### 2024-04-26 01_Example/Manuscript.qmd
created minimal example

---
title: "A Study on Penguins: A Minimal Reproducible Example"
author:
  - name: "Kan Garoo"
    affiliation: "University of Melbourne"
    email: "kangaroo@email.com"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  quarto::apa_document
bibliography: Bibliography.bib
---

# Introduction

Penguins are fascinating creatures that inhabit various regions of the Southern Hemisphere, including Antarctica and surrounding islands.
The study of penguins provides valuable insights into ecosystem dynamics, climate change impacts, and evolutionary biology [@smith2020; @jones2018].

This manuscript presents a minimal reproducible example utilizing the penguins dataset to demonstrate scientific workflows in R.

```{r t_test, echo=FALSE}

# Load penguins data set
data <- read.csv("Data.csv")

# Remove rows with NA
data_clean <- data[complete.cases(data),]

# Perform t-test
t_test_result <- t.test(bill_length_mm ~ sex, data = data_clean)

```

# Methods

We conducted a two-sample t-test to compare the average bill lengths between male and female penguins.
The null hypothesis ($H_0$) states that there is no difference in bill lengths between male and female penguins, while the alternative hypothesis ($H_1$) suggests a significant difference.

The t-test was performed using the `t.test()` function in R, with a significance level of 0.05.

# Results

Descriptive statistics of the data set are displayed in Table 1.

```{r descriptive_statistics, echo=FALSE, results='asis'}

# Call descriptive table with the vtable package
vtable::sumtable(data_clean, group = "species")
```

The t-test revealed a significant difference in bill lengths between male and female penguins ($t($`{r} round(t_test_result$parameter, digits = 2)` $)=$ `{r} round(t_test_result$statistic, digits = 2)`, $p$ `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`).
Female penguins ($M=$ `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)`mm, $SD=$ `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` mm) exhibited shorter bill lengths compared to male penguins ($M=$ `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)`mm, $SD=$ `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` mm).

```{r bill_length_comparison, echo=FALSE, fig.cap="Histogram with Scatter Plot of Bill Lengths by Sex"}

# Create ggplot object
ggplot2::ggplot(data_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
  ggplot2::geom_boxplot(width = 0.5, show.legend = FALSE) +
  ggplot2::geom_jitter(width = 0.15, show.legend = FALSE) +
  ggplot2::labs(x = "Sex", y = "Bill length in mm") +
  ggplot2::theme_minimal()
```

# Discussion

The significant difference in bill lengths between male and female penguins suggests potential sexual dimorphism in this trait.
This finding aligns with previous research indicating differential foraging strategies and resource partitioning between male and female penguins [@brown2015; @wilson2019].

Understanding the factors influencing bill morphology in penguins is crucial for conservation efforts and ecosystem management, particularly in the face of ongoing environmental challenges.

# References

::: {#refs}
:::

### 2024-04-29 01_Example/Manuscript.qmd
created pdf

@@
 date: "`r format(Sys.Date(), '%B %d, %Y')`"
-output:
-  quarto::apa_document
 bibliography: Bibliography.bib
+format:
+  #apaquarto-docx: default
+  #apaquarto-html: default
+  apaquarto-pdf:
+    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
+    documentmode: man
+    keep-tex: true
 ---

### 2024-04-29 01_Example/Manuscript.tex
created pdf

\documentclass[
  man,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
% \setlength\LTleft{0pt}
\RequirePackage{threeparttablex}

% % % \RequirePackage{rotating}
% \RequirePackage{threeparttablex}
% \DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
% \DeclareDelayedFloatFlavor{sidewaystable}{table}
% \DeclareDelayedFloatFlavor{longtable}{table}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}

% %

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{A Study on Penguins: A Minimal Reproducible Example}
\shorttitle{A Study on Penguins: A Minimal Reproducible Example}

\usepackage{etoolbox}

\author{Kan Garoo}

\affiliation{
{University of Melbourne}}

\leftheader{Garoo}

\date{2024-04-29}

%

\authornote{
\par{ }
\par{       }
\par{Correspondence concerning this article should be addressed to }
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother
\RequirePackage{longtable}
\DeclareDelayedFloatFlavor{longtable}{table}
% \RequirePackage{threeparttablex}
% \DeclareDelayedFloatFlavor{ThreePartTable}{table}

\urlstyle{same}

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Penguins are fascinating creatures that inhabit various regions of the
Southern Hemisphere, including Antarctica and surrounding islands. The
study of penguins provides valuable insights into ecosystem dynamics,
climate change impacts, and evolutionary biology (Jones, 2018; Smith,
2020).

This manuscript presents a minimal reproducible example utilizing the
penguins dataset to demonstrate scientific workflows in R.

\section{Methods}\label{methods}

We conducted a two-sample t-test to compare the average bill lengths
between male and female penguins. The null hypothesis (\(H_0\)) states
that there is no difference in bill lengths between male and female
penguins, while the alternative hypothesis (\(H_1\)) suggests a
significant difference.

The t-test was performed using the \texttt{t.test()} function in R, with
a significance level of 0.05.

\section{Results}\label{results}

Descriptive statistics of the data set are displayed in Table 1.

\begin{table}

\caption{\label{tab:descriptive_statistics}Summary Statistics}
\centering
\begin{tabular}[t]{llllllllll}
\toprule
\multicolumn{1}{c}{species} & \multicolumn{3}{c}{Adelie} & \multicolumn{3}{c}{Chinstrap} & \multicolumn{3}{c}{Gentoo} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
Variable & N & Mean & SD & N & Mean & SD & N & Mean & SD\\
\midrule
island & 146 &  &  & 68 &  &  & 119 &  & \\
... Biscoe & 44 & 30\% &  & 0 & 0\% &  & 119 & 100\% & \\
... Dream & 55 & 38\% &  & 68 & 100\% &  & 0 & 0\% & \\
... Torgersen & 47 & 32\% &  & 0 & 0\% &  & 0 & 0\% & \\
bill\_length\_mm & 146 & 39 & 2.7 & 68 & 49 & 3.3 & 119 & 48 & 3.1\\
\addlinespace
bill\_depth\_mm & 146 & 18 & 1.2 & 68 & 18 & 1.1 & 119 & 15 & 0.99\\
flipper\_length\_mm & 146 & 190 & 6.5 & 68 & 196 & 7.1 & 119 & 217 & 6.6\\
body\_mass\_g & 146 & 3706 & 459 & 68 & 3733 & 384 & 119 & 5092 & 501\\
sex & 146 &  &  & 68 &  &  & 119 &  & \\
... female & 73 & 50\% &  & 34 & 50\% &  & 58 & 49\% & \\
\addlinespace
... male & 73 & 50\% &  & 34 & 50\% &  & 61 & 51\% & \\
year & 146 & 2008 & 0.81 & 68 & 2008 & 0.86 & 119 & 2008 & 0.79\\
\bottomrule
\end{tabular}
\end{table}

The t-test revealed a significant difference in bill lengths between
male and female penguins (\$t(\$329.29 \()=\) -6.67, \(p\) \textless{}
0.001). Female penguins (\(M=\) 42.1mm, \(SD=\) 4.9 mm) exhibited
shorter bill lengths compared to male penguins (\(M=\) 45.85mm, \(SD=\)
5.37 mm).

\begin{figure}[H]

\caption{Histogram with Scatter Plot of Bill Lengths by Sex}

{\centering \includegraphics{Manuscript_files/figure-pdf/bill_length_comparison-1.pdf}

}

\end{figure}%

\section{Discussion}\label{discussion}

The significant difference in bill lengths between male and female
penguins suggests potential sexual dimorphism in this trait. This
finding aligns with previous research indicating differential foraging
strategies and resource partitioning between male and female penguins
(Brown, 2015; Wilson, 2019).

Understanding the factors influencing bill morphology in penguins is
crucial for conservation efforts and ecosystem management, particularly
in the face of ongoing environmental challenges.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-brown2015}
Brown, E. F. (2015). Sexual dimorphism in bill lengths of penguins.
\emph{Journal of Ornithology}, \emph{20}, 67--79.

\bibitem[\citeproctext]{ref-jones2018}
Jones, C. D. (2018). Penguins and climate change: An overview.
\emph{Environmental Science Review}, \emph{8}, 45--58.

\bibitem[\citeproctext]{ref-smith2020}
Smith, A. B. (2020). Penguin behavior: A comprehensive review.
\emph{Journal of Penguin Studies}, \emph{15}, 123--135.

\bibitem[\citeproctext]{ref-wilson2019}
Wilson, G. H. (2019). Foraging strategies in male and female penguins.
\emph{Behavioral Ecology}, \emph{25}, 102--115.

\end{CSLReferences}

\end{document}

### 2024-04-29 01_Example/Manuscript.qmd
minimal exampl journal

--- 2024-04-26 01_Example/Manuscript.qmd
@@
 author:
-  - name: "Kan Garoo"
-    affiliation: "University of Melbourne"
-    email: "kangaroo@email.com"
+  - name: "Josephine Zerna"
+    email: "josephine.zerna@tu-dresden.de"
+    orcid: 0000-0003-2892-884X
+    roles:
+      - conceptualization
+      - writing
+    affiliations:
+      - id: id1
+        name: "Technische Universität Dresden"
+        group: Chair of Differential and Personality Psychology
+        department: Faculty of Psychology
+        city: Dresden
+        postal-code: 01069
+  - name: "Christoph Scheffel"
+    affiliation: 
+      - id: id1
+    email: "christoph_scheffel@tu-dresden.de"
+    orcid: 0000-0001-5963-9229
+    corresponding: true
+    roles:
+      - project administration
+      - methodology
 date: "`r format(Sys.Date(), '%B %d, %Y')`"
-output:
-  quarto::apa_document
+abstract: "This document is a minimal, reproducible manuscript using the penguins data set as an example."
+keywords: [penguins, reproducibility, minimal, example]
+floatsintext: true
 bibliography: Bibliography.bib
+format:
+  #apaquarto-docx: default
+  #apaquarto-html: default
+  apaquarto-pdf:
+    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
+    documentmode: jou
+    keep-tex: true
 ---

### 2024-04-29 01_Example/Manuscript.tex
minimal exampl journal

@@
 \documentclass[
-  man,
+  jou,
+  floatsintext,
   longtable,
@@
 \RequirePackage{threeparttablex}
-
-% % % \RequirePackage{rotating}
-% \RequirePackage{threeparttablex}
-% \DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
-% \DeclareDelayedFloatFlavor{sidewaystable}{table}
-% \DeclareDelayedFloatFlavor{longtable}{table}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
 
@@
 
-\author{Kan Garoo}
+\authorsnames{Josephine Zerna,Christoph Scheffel}
 
-\affiliation{
-{University of Melbourne}}
+\authorsaffiliations{
+{Faculty of Psychology, Technische Universität Dresden},{}}
 
-\leftheader{Garoo}
+\leftheader{Zerna and Scheffel}
 
@@
 
+\abstract{This document is a minimal, reproducible manuscript using the
+penguins data set as an example.}
 % 
+\keywords{penguins, reproducibility, minimal, example}
 
-\authornote{
+\authornote{\par{\addORCIDlink{Josephine
+Zerna}{0000-0003-2892-884X}}\par{\addORCIDlink{Christoph
+Scheffel}{0000-0001-5963-9229}}
 \par{ }
-\par{       }
-\par{Correspondence concerning this article should be addressed to }
+\par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Josephine
+Zerna:   conceptualization, writing; Christoph Scheffel:   project
+administration, methodology}
+\par{Correspondence concerning this article should be addressed
+to Christoph Scheffel, Email: christoph\_scheffel@tu-dresden.de}
 }
+
+\usepackage{pbalance} 
+\usepackage{float}
+\makeatletter
+\let\oldtpt\ThreePartTable
+\let\endoldtpt\endThreePartTable
+\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
+\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
+\onecolumn
+\begin{minipage}{0.5\textwidth}
+\oldtpt[#1]
+}
+\def\ThreePartTable@ii{\begin{figure}[!htbp]
+\onecolumn
+\begin{minipage}{0.5\textwidth}
+\oldtpt
+}
+\def\endThreePartTable{
+\endoldtpt
+\end{minipage}
+\twocolumn
+\end{figure}}
+\makeatother
 
@@
 \hline
-\endoldlt
-}
+\endoldlt}
 \makeatother
-\RequirePackage{longtable}
-\DeclareDelayedFloatFlavor{longtable}{table}
-% \RequirePackage{threeparttablex}
-% \DeclareDelayedFloatFlavor{ThreePartTable}{table}
+
+\newenvironment{twocolumntable}% environment name
+{% begin code
+\begin{table*}[!htbp]%
+\onecolumn%
+}%
+{%
+\twocolumn%
+\end{table*}%
+}% end code
 

### 2024-05-16 03_Supplement/Checklist.qmd
restructured folders

---
title: "Checklist"
format: html
---

# Workshop Checklist for Reproducible Manuscripts

You can use this checklist for the project you worked on today or for any future projects that you make reproducible. The more points you can check off, the more reproducible your project is.

You will encounter some points that were not part of this workshop, but don't worry, we provide additional resources so you can seek more information about them if you wish to.

1.  Project Setup
  + Independence
    -   Are all files contained in an .Rproj environment?
    -   Are all files being accessed with relative paths using the 'here' package?
  + Stability
    -   Are package dependencies being stabilized with the 'renv' package?
    -   Is the environment encapsulated with Binder?
    -   Does the online version have a persistent identifier (DOI)?
    -   If you are the only author, does a coworker with a longer contract have access as well?
  + Version control
    -   Is the project being controlled through a Git integration?
    -   Are all commits small and meaningfully labelled?
    -   Does the repository have a license file?
    -   Does the repository have a short but meaningful README file?

2.  Data Management
  + Structure
    -   Are all files clearly named with consistent conventions?
    -   Is the folder structure intuitive? There should be ≤3 levels with ≤7 folders each.
  + Resolution level
    -   Have sensitive data been anonymized/pseudonymized?
    -   Are raw or preprocessed data being published? Clearly label the data as one or the other and describe any preprocessing steps in a README file.
    -   Are there any legal restrictions regarding the data? If your work involves patents, contracts, patient data, etc., contact your data protection officer for guidelines.
  + Accessibility
    -   Where will the data be stored? Small datasets can be stored in Github, larger ones on the servers of storage providers, e.g. OSF or Figshare.
    -   Where will the data be archived? On your institution's servers and/or publicly.
    -   Under which license/restrictions is the data being published?

3.  Analyses and Documentation
  + Informativity
    -   Is all code meaningfully commented?
    -   Is there a codebook?
    -   Are all analyses described in sufficient detail?
  + Ease
    -   Is there metadata provided to increase discoverability?
    -   Are the analyses executable with one click?
    -   If multiple steps are necessary, are they clearly instructed?

4.  Manuscript
  + References
    -   Does the provided .bib-file contain all relevant and correct citations?
    -   Are the citation keys meaningful?
  + Tables
    -   Can the tables be reproduced by running the applicable code section?
  + Figures
    -   Can the figures be reproduced by running the applicable code section?
    -   If multiple steps are necessary, are they clearly instructed?

5.  Review
  + Internally
    -   Has a coworker attempted to reproduce your project? Bonus points, if it's on a different operating system.
  + Externally
    -   Is the repository accessible to reviewers? Some repositories provide anonymized links for peer-review.
    -   Does the repository link to the preprint/final manuscript?

### 2024-05-16 Instructions.qmd
created instructions file

---
title: "Workshop Instructions: Reproducible Scientific Manuscripts"
---

# Workshop Instructions: Reproducible Scientific Manuscripts

## Step 1: Clone the Repository

1. Open your terminal or command prompt.
2. Navigate to the directory where you want to clone the repository.
3. Execute the following command:

```{r eval=FALSE}
git clone <repository_url>
```

Replace `<repository_url>` with the URL of the GitHub repository provided to you.

## Step 2: Execute the Minimal Example

1. Navigate to the "01_Example" folder in the cloned repository.
2. Open the "Example.Rproj" file in RStudio.
3. Run the provided script, "Manuscript.qmd", to execute the minimal example. Data, bibliography and citation style file are already provided in the folder, there is no need to change or add anything.
4. Review the generated manuscript files ("Manuscript.docx", "Manuscript.html", etc.) to ensure they were created successfully.

## Step 3: Rebuild the Minimal Example as Practise

1. Open your terminal or command prompt.
2. Navigate to the "02_Practise" folder in the cloned repository.
3. Create a new R Project in this folder and initialize the *renv* environment:

```{r eval=FALSE}
renv::init()
```

4. Open the newly created ".Rproj" file in RStudio.
5. Within the R Project, create a new Quarto document:
- Click on File > New File > Quarto Document...
6. Replace the content of the new Quarto document with the provided text.
7. Modify the fiel paths to ensure that all provided files (data, bibliography, etc.) are correctly imported and referenced in your Quarto document.
8. Execute the Quarto document to generate your practice manuscript.

## Step 4: Apply the Methods to Your Own Project/Data

1. Navigate to your own project directory or create a new one if needed.
2. Create a Quarto document and copy text from your existing manuscript into the document.
3. Modify the formatting.
4. Copy a figure into the project folder and insert it into the document.
5. Create a table, either by hand or by importing your dataset into the project folder and using a package to generate a table.
6. Add your bibliography as a .bib file.
7. Choose a citation style and download the appropriate .csl file from the web.
8. Execute your modified scripts and review the generated manuscript files.
9. Add a license file.
10. Add a README file to let others know what the project contains, how to execute it, and who to contact in case of trouble.

## Step 5: Use the Checklist

1. Navigate to the "03_Supplement" folder in the cloned repository.
2. Open the "Checklist.qmd" in RStudio or the "Checklist.html" file in your web browser.
3. Review each item on the checklist to ensure that your work meets the reproducibility criteria.
4. Tick off each item as you verify its completion in your project.

### 2024-05-16 01_Example/Manuscript.qmd
updated minimal exmple

--- 2024-04-26 01_Example/Manuscript.qmd
@@
 author:
-  - name: "Kan Garoo"
-    affiliation: "University of Melbourne"
-    email: "kangaroo@email.com"
+  - name: "Josephine Zerna"
+    email: "josephine.zerna@tu-dresden.de"
+    orcid: 0000-0003-2892-884X
+    roles:
+      - conceptualization
+      - writing
+    affiliations:
+      - id: id1
+        name: "Technische Universität Dresden"
+        group: Chair of Differential and Personality Psychology
+        department: Faculty of Psychology
+        city: Dresden
+        postal-code: 01069
+  - name: "Christoph Scheffel"
+    affiliation: 
+      - id: id1
+    email: "christoph_scheffel@tu-dresden.de"
+    orcid: 0000-0001-5963-9229
+    corresponding: true
+    roles:
+      - project administration
+      - methodology
 date: "`r format(Sys.Date(), '%B %d, %Y')`"
-output:
-  quarto::apa_document
+abstract: "This document is a minimal, reproducible manuscript using the penguins data set as an example."
+keywords: [penguins, reproducibility, minimal, example]
+floatsintext: false
 bibliography: Bibliography.bib
+format:
+  #apaquarto-docx: default
+  #apaquarto-html: default
+  apaquarto-pdf:
+    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
+    documentmode: man
+    keep-tex: true
+
 ---
@@
 vtable::sumtable(data_clean, group = "species")
+
 ```

### 2024-05-16 01_Example/Manuscript.tex
updated minimal exmple

--- 2024-04-29 01_Example/Manuscript.tex
@@
 
-\author{Kan Garoo}
+\authorsnames{Josephine Zerna,Christoph Scheffel}
 
-\affiliation{
-{University of Melbourne}}
+\authorsaffiliations{
+{Faculty of Psychology, Technische Universität Dresden},{}}
 
-\leftheader{Garoo}
+\leftheader{Zerna and Scheffel}
 
-\date{2024-04-29}
+\date{Invalid Date}
 
+\abstract{This document is a minimal, reproducible manuscript using the
+penguins data set as an example.}
 % 
+\keywords{penguins, reproducibility, minimal, example}
 
-\authornote{
+\authornote{\par{\addORCIDlink{Josephine
+Zerna}{0000-0003-2892-884X}}\par{\addORCIDlink{Christoph
+Scheffel}{0000-0001-5963-9229}}
 \par{ }
-\par{       }
-\par{Correspondence concerning this article should be addressed to }
+\par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Josephine
+Zerna:   conceptualization, writing; Christoph Scheffel:   project
+administration, methodology}
+\par{Correspondence concerning this article should be addressed
+to Christoph Scheffel, Email: christoph\_scheffel@tu-dresden.de}
 }

### 2024-05-16 Instructions.qmd
Update Instructions.qmd

@@
 # Workshop Instructions: Reproducible Scientific Manuscripts
+
+In this part of the workshop, you will integrate all of the skills, you acquired in the previous lessons.
+You will firstly execute a self-contained minimal example that is provided in this repository.
+Secondly, you will recreate the minimal example using all the components provided in the folder "02_practice".
+Thirdly, you will create your own first reproducible manuscript based on a conventional manuscript you already have.
+And finally, you can use the checklist provided in the "03_supplement" folder to go over your project and see how many reproducibility measure you applied.
+
+<!-- Please insert links to the appropriate workshop components and folders -->
 

### 2024-05-16 Instructions.qmd
created readme files

@@
 1. Navigate to the "01_Example" folder in the cloned repository.
-2. Open the "Example.Rproj" file in RStudio.
-3. Run the provided script, "Manuscript.qmd", to execute the minimal example. Data, bibliography and citation style file are already provided in the folder, there is no need to change or add anything.
-4. Review the generated manuscript files ("Manuscript.docx", "Manuscript.html", etc.) to ensure they were created successfully.
+2. Open the README.md file. 
+Every repository should contain a README file describing the content of the repository, how to execute the files, and who to contact in case of issues. 
+Therefore, you will find the instructions on how to execute the Minimal Example within this README file.
 

### 2024-05-16 Instructions.qmd
added readme instruction

@@
 You will firstly execute a self-contained minimal example that is provided in this repository.
-Secondly, you will recreate the minimal example using all the components provided in the folder "02_practice".
+Secondly, you will recreate the minimal example using all the components provided in the folder "02_Practice".
 Thirdly, you will create your own first reproducible manuscript based on a conventional manuscript you already have.
-And finally, you can use the checklist provided in the "03_supplement" folder to go over your project and see how many reproducibility measure you applied.
+And finally, you can use the checklist provided in the "03_Supplement" folder to go over your project and see how many reproducibility measures you have applied.
 
@@
 8. Execute the Quarto document to generate your practice manuscript.
+9. Add a README.md file to the "02_Practise" folder and describe, which steps users have to to in order to execute your Minimal Example.
 

### 2024-08-21 01_Example/Manuscript.qmd
Fix whitespace

--- 2024-04-29 01_Example/Manuscript.qmd
@@
 author:
-  - name: "Kan Garoo"
-    affiliation: "University of Melbourne"
-    email: "kangaroo@email.com"
+  - name: "Josephine Zerna"
+    email: "josephine.zerna@tu-dresden.de"
+    orcid: 0000-0003-2892-884X
+    roles:
+      - conceptualization
+      - writing
+    affiliations:
+      - id: id1
+        name: "Technische Universität Dresden"
+        group: Chair of Differential and Personality Psychology
+        department: Faculty of Psychology
+        city: Dresden
+        postal-code: 01069
+  - name: "Christoph Scheffel"
+    affiliation: 
+      - id: id1
+    email: "christoph_scheffel@tu-dresden.de"
+    orcid: 0000-0001-5963-9229
+    corresponding: true
+    roles:
+      - project administration
+      - methodology
 date: "`r format(Sys.Date(), '%B %d, %Y')`"
+abstract: "This document is a minimal, reproducible manuscript using the penguins data set as an example."
+keywords: [penguins, reproducibility, minimal, example]
+floatsintext: false
 bibliography: Bibliography.bib

### 2024-08-21 01_Example/Manuscript.qmd
Use YAML chunk options and avoid labels with underscores
Regarding underscores, see https://quarto.org/docs/authoring/cross-references.html

@@
 
-```{r t_test, echo=FALSE}
+```{r}
+#| label: "t-test"
+#| echo: false
 
@@
 
-```{r descriptive_statistics, echo=FALSE, results='asis'}
+```{r}
+#| label: "descriptive-statistics"
+#| echo: false
+#| results: "asis"
 
@@
 
-```{r bill_length_comparison, echo=FALSE, fig.cap="Histogram with Scatter Plot of Bill Lengths by Sex"}
+```{r}
+#| label: "bill-length-comparison"
+#| echo: false
+#| fig.cap: "Histogram with Scatter Plot of Bill Lengths by Sex"
 

### 2024-08-21 01_Example/Manuscript.qmd
Use violin plot rather than box plot
Friends don't let friends use box plots

@@
 #| echo: false
-#| fig.cap: "Histogram with Scatter Plot of Bill Lengths by Sex"
+#| fig-cap: "Scatter Plot of Bill Lengths by Sex"
+#| fig-subcap: "Violin Plot Displays Quartiles"
 
@@
 ggplot2::ggplot(data_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
-  ggplot2::geom_boxplot(width = 0.5, show.legend = FALSE) +
+  ggplot2::geom_violin(
+    draw_quantiles = c(0.25, 0.5, 0.75),
+    show.legend = FALSE
+  ) +
   ggplot2::geom_jitter(width = 0.15, show.legend = FALSE) +

### 2024-08-21 01_Example/Manuscript.qmd
Only use cap
Use meaningful affiliation id

@@
     affiliations:
-      - id: id1
+      - id: "tu-dresden"
         name: "Technische Universität Dresden"
@@
     affiliation:
-      - id: id1
+      - id: "tu-dresden"
     email: "christoph_scheffel@tu-dresden.de"
@@
 #| echo: false
-#| fig-cap: "Scatter Plot of Bill Lengths by Sex"
-#| fig-subcap: "Violin Plot Displays Quartiles"
+#| fig-cap: "Scatter Plot of Bill Lengths by Sex With Violin Plot Showing Quartiles"
 

### 2024-08-22 01_Example/Manuscript.qmd
Fix LaTeX math
Wrap everything inside dollars

@@
 
-The t-test revealed a significant difference in bill lengths between male and female penguins ($t($`{r} round(t_test_result$parameter, digits = 2)` $)=$ `{r} round(t_test_result$statistic, digits = 2)`, $p$ `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`).
-Female penguins ($M=$ `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)`mm, $SD=$ `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` mm) exhibited shorter bill lengths compared to male penguins ($M=$ `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)`mm, $SD=$ `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` mm).
+The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`{r} round(t_test_result$parameter, digits = 2)`) = `{r} round(t_test_result$statistic, digits = 2)`, p `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
+Female penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}$).
 

### 2024-08-22 01_Example/Manuscript.qmd
Add proper prefixes for table and figure

@@
 ```{r}
-#| label: "descriptive-statistics"
+#| label: "tbl-descriptive-statistics"
 #| echo: false
@@
 ```{r}
-#| label: "bill-length-comparison"
+#| label: "fig-bill-length-comparison"
 #| echo: false

### 2024-08-22 01_Example/Manuscript.qmd
Use {modelsummary} rather than {vtable}
Because {modelsummary} has fewer dependencies

@@
 ```{r}
-#| label: "tbl-descriptive-statistics"
+#| label: "descriptive-statistics"
 #| echo: false
 #| results: "asis"
+#| tbl-cap: "Summary Statistics"
 
-# Call descriptive table with the vtable package
-vtable::sumtable(data_clean, group = "species")
+options(modelsummary_warning_latex_siunitx_preamble = FALSE) # Hide LaTeX compilation tips
+modelsummary::datasummary_balance(formula = ~ species, data = data_clean, output = "latex", dinm = FALSE)
 

### 2024-08-22 01_Example/Manuscript.qmd
Add table prefix (again)
Remove prefix again due to display issues and don't assume output format
See https://github.com/vincentarelbundock/tinytable/issues/327

--- 2024-08-22 01_Example/Manuscript.qmd
@@
 ```{r}
-#| label: "tbl-descriptive-statistics"
+#| label: "descriptive-statistics"
 #| echo: false
-#| results: "asis"
+#| tbl-cap: "Summary Statistics"
 
-# Call descriptive table with the vtable package
-vtable::sumtable(data_clean, group = "species")
+modelsummary::datasummary_balance(formula = ~ species, data = data_clean, dinm = FALSE)
 

### 2024-08-22 01_Example/Manuscript.qmd
Add document language explicitly

@@
 date: "`r format(Sys.Date(), '%B %d, %Y')`"
+lang: en
 abstract: "This document is a minimal, reproducible manuscript using the penguins data set as an example."

### 2024-08-22 01_Example/Manuscript.qmd
Add small space between number and unit

@@
 The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`{r} round(t_test_result$parameter, digits = 2)`) = `{r} round(t_test_result$statistic, digits = 2)`, p `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
-Female penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}$).
+Female penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)`\,\mathrm{mm}$).
 

### 2024-08-22 01_Example/Manuscript.tex
Recreate output

@@
 
+\usepackage[bidi=default]{babel}
+\babelprovide[main,import]{english}
+
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+
 \RequirePackage{longtable}
-% \setlength\LTleft{0pt}
 \RequirePackage{threeparttablex}
-
-% % % \RequirePackage{rotating}
-% \RequirePackage{threeparttablex}
-% \DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
-% \DeclareDelayedFloatFlavor{sidewaystable}{table}
-% \DeclareDelayedFloatFlavor{longtable}{table}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-
-% %
 
@@
 
-\newlength\Oldarrayrulewidth
-\newlength\Oldtabcolsep
+%\newlength\Oldarrayrulewidth
+%\newlength\Oldtabcolsep
 
@@
 \title{A Study on Penguins: A Minimal Reproducible Example}
+
 \shorttitle{A Study on Penguins: A Minimal Reproducible Example}
@@
 
-\date{Invalid Date}
+\date{2024-08-22}
 
@@
 penguins data set as an example.}
-% 
+
 \keywords{penguins, reproducibility, minimal, example}
@@
 Scheffel}{0000-0001-5963-9229}}
-\par{ }
+
 \par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Josephine
@@
 \DeclareDelayedFloatFlavor{longtable}{table}
-% \RequirePackage{threeparttablex}
-% \DeclareDelayedFloatFlavor{ThreePartTable}{table}
 
 \urlstyle{same}
+
+\usepackage{float}
+\usepackage{tabularray}
+\usepackage[normalem]{ulem}
+\usepackage{graphicx}
+\UseTblrLibrary{booktabs}
+\UseTblrLibrary{siunitx}
+\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
+\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
+\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
 
@@
 study of penguins provides valuable insights into ecosystem dynamics,
-climate change impacts, and evolutionary biology (Jones, 2018; Smith,
-2020).
+climate change impacts, and evolutionary biology
+(\citeproc{ref-jones2018}{Jones, 2018}; \citeproc{ref-smith2020}{Smith,
+2020}).
 
@@
 \begin{table}
+\caption{Summary Statistics}\tabularnewline
 
-\caption{\label{tab:descriptive_statistics}Summary Statistics}
 \centering
-\begin{tabular}[t]{llllllllll}
+\begin{tblr}[         %% tabularray outer open
+]                     %% tabularray outer close
+{                     %% tabularray inner open
+colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
+cell{1}{3}={c=2,}{halign=c,},
+cell{1}{5}={c=2,}{halign=c,},
+cell{1}{7}={c=2,}{halign=c,},
+column{1}={halign=l,},
+column{2}={halign=l,},
+column{3}={halign=r,},
+column{4}={halign=r,},
+column{5}={halign=r,},
+column{6}={halign=r,},
+column{7}={halign=r,},
+column{8}={halign=r,},
+row{1}={halign=c,},
+hline{8}={1,2,3,4,5,6,7,8}{solid, 0.05em, black},
+}                     %% tabularray inner close
 \toprule
-\multicolumn{1}{c}{species} & \multicolumn{3}{c}{Adelie} & \multicolumn{3}{c}{Chinstrap} & \multicolumn{3}{c}{Gentoo} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
-Variable & N & Mean & SD & N & Mean & SD & N & Mean & SD\\
-\midrule
-island & 146 &  &  & 68 &  &  & 119 &  & \\
-... Biscoe & 44 & 30\% &  & 0 & 0\% &  & 119 & 100\% & \\
-... Dream & 55 & 38\% &  & 68 & 100\% &  & 0 & 0\% & \\
-... Torgersen & 47 & 32\% &  & 0 & 0\% &  & 0 & 0\% & \\
-bill\_length\_mm & 146 & 39 & 2.7 & 68 & 49 & 3.3 & 119 & 48 & 3.1\\
-\addlinespace
-bill\_depth\_mm & 146 & 18 & 1.2 & 68 & 18 & 1.1 & 119 & 15 & 0.99\\
-flipper\_length\_mm & 146 & 190 & 6.5 & 68 & 196 & 7.1 & 119 & 217 & 6.6\\
-body\_mass\_g & 146 & 3706 & 459 & 68 & 3733 & 384 & 119 & 5092 & 501\\
-sex & 146 &  &  & 68 &  &  & 119 &  & \\
-... female & 73 & 50\% &  & 34 & 50\% &  & 58 & 49\% & \\
-\addlinespace
-... male & 73 & 50\% &  & 34 & 50\% &  & 61 & 51\% & \\
-year & 146 & 2008 & 0.81 & 68 & 2008 & 0.86 & 119 & 2008 & 0.79\\
+&  & Adelie (N=146) &  & Chinstrap (N=68) &  & Gentoo (N=119) &  \\ \cmidrule[lr]{3-4}\cmidrule[lr]{5-6}\cmidrule[lr]{7-8}
+&    & Mean & Std. Dev. & Mean & Std. Dev. & Mean & Std. Dev. \\ \midrule %% TinyTableHeader
+bill\_length\_mm    &           & \num{38.8}   & \num{2.7}   & \num{48.8}   & \num{3.3}   & \num{47.6}   & \num{3.1}   \\
+bill\_depth\_mm     &           & \num{18.3}   & \num{1.2}   & \num{18.4}   & \num{1.1}   & \num{15.0}   & \num{1.0}   \\
+flipper\_length\_mm &           & \num{190.1}  & \num{6.5}   & \num{195.8}  & \num{7.1}   & \num{217.2}  & \num{6.6}   \\
+body\_mass\_g       &           & \num{3706.2} & \num{458.6} & \num{3733.1} & \num{384.3} & \num{5092.4} & \num{501.5} \\
+year                  &           & \num{2008.1} & \num{0.8}   & \num{2008.0} & \num{0.9}   & \num{2008.1} & \num{0.8}   \\
+&           & N             & Pct.         & N             & Pct.         & N             & Pct.         \\
+island                & Biscoe    & \num{44}     & \num{30.1}  & \num{0}      & \num{0.0}   & \num{119}    & \num{100.0} \\
+& Dream     & \num{55}     & \num{37.7}  & \num{68}     & \num{100.0} & \num{0}      & \num{0.0}   \\
+& Torgersen & \num{47}     & \num{32.2}  & \num{0}      & \num{0.0}   & \num{0}      & \num{0.0}   \\
+sex                   & female    & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{58}     & \num{48.7}  \\
+& male      & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{61}     & \num{51.3}  \\
 \bottomrule
-\end{tabular}
+\end{tblr}
 \end{table}
@@
 The t-test revealed a significant difference in bill lengths between
-male and female penguins (\$t(\$329.29 \()=\) -6.67, \(p\) \textless{}
-0.001). Female penguins (\(M=\) 42.1mm, \(SD=\) 4.9 mm) exhibited
-shorter bill lengths compared to male penguins (\(M=\) 45.85mm, \(SD=\)
-5.37 mm).
+male and female penguins (\(t(329.29) = -6.67, p < 0\.001\)). Female
+penguins (\(M = 42.1\,\mathrm{mm}, SD = 4.9\,\mathrm{mm}\)) exhibited
+shorter bill lengths compared to male penguins
+(\(M = 45.85\,\mathrm{mm}, SD = 5.37\,\mathrm{mm}\)).
 
-\begin{figure}[H]
+\begin{figure}
 
-\caption{Histogram with Scatter Plot of Bill Lengths by Sex}
+\caption{\label{fig-bill-length-comparison}Scatter Plot of Bill Lengths
+by Sex With Violin Plot Showing Quartiles}
 
-{\centering \includegraphics{Manuscript_files/figure-pdf/bill_length_comparison-1.pdf}
+\centering{
+
+\includegraphics{Manuscript_files/figure-pdf/fig-bill-length-comparison-1.pdf}
 
@@
 strategies and resource partitioning between male and female penguins
-(Brown, 2015; Wilson, 2019).
+(\citeproc{ref-brown2015}{Brown, 2015};
+\citeproc{ref-wilson2019}{Wilson, 2019}).
 

### 2024-08-23 01_Example/Manuscript.qmd
Use variables for inline R

--- 2024-08-22 01_Example/Manuscript.qmd
@@
 
+```{r}
+#| label: "variables"
+#| echo: false
+
+blm_female <- data_clean$bill_length_mm[data_clean$sex == "female"]
+blm_male <- data_clean$bill_length_mm[data_clean$sex == "male"]
+```
+
 The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`{r} round(t_test_result$parameter, digits = 2)`) = `{r} round(t_test_result$statistic, digits = 2)`, p `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
-Female penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "female"], na.rm = TRUE), digits = 2)` \mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}, SD = `{r} round(sd(data$bill_length_mm[data$sex == "male"], na.rm = TRUE), digits = 2)` \mathrm{mm}$).
+Female penguins ($M = `{r} round(mean(blm_female), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(blm_male), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).
 

### 2024-08-23 01_Example/Manuscript.qmd
Show summary statistics by sex rather than species
Fix missing decimal point
See https://github.com/quarto-dev/quarto-cli/issues/10595

@@
 
-modelsummary::datasummary_balance(formula = ~ species, data = data_clean, dinm = FALSE)
+modelsummary::datasummary_balance(formula = ~ sex, data = data_clean, dinm = FALSE)
 
@@
 
-The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`{r} round(t_test_result$parameter, digits = 2)`) = `{r} round(t_test_result$statistic, digits = 2)`, p `{r} format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
-Female penguins ($M = `{r} round(mean(blm_female), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `{r} round(mean(blm_male), digits = 2)`\,\mathrm{mm}, SD = `{r} round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).
+The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`r round(t_test_result$parameter, digits = 2)`) = `r round(t_test_result$statistic, digits = 2)`, p `r format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
+Female penguins ($M = `r round(mean(blm_female), digits = 2)`\,\mathrm{mm}, SD = `r round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `r round(mean(blm_male), digits = 2)`\,\mathrm{mm}, SD = `r round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).
 

### 2024-08-23 01_Example/Manuscript.qmd
Use "dat" rather than "data"
Because "data()" is already a function

@@
 # Load penguins data set
-data <- read.csv("Data.csv")
+dat <- read.csv("Data.csv")
 
 # Remove rows with NA
-data_clean <- data[complete.cases(data),]
+dat_clean <- dat[complete.cases(dat),]
 
 # Perform t-test
-t_test_result <- t.test(bill_length_mm ~ sex, data = data_clean)
+t_test_result <- t.test(bill_length_mm ~ sex, data = dat_clean)
 
@@
 
-modelsummary::datasummary_balance(formula = ~ sex, data = data_clean, dinm = FALSE)
+modelsummary::datasummary_balance(formula = ~ sex, data = dat_clean, dinm = FALSE)
 
@@
 
-blm_female <- data_clean$bill_length_mm[data_clean$sex == "female"]
-blm_male <- data_clean$bill_length_mm[data_clean$sex == "male"]
+blm_female <- dat_clean$bill_length_mm[dat_clean$sex == "female"]
+blm_male <- dat_clean$bill_length_mm[dat_clean$sex == "male"]
 ```
@@
 # Create ggplot object
-ggplot2::ggplot(data_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
+ggplot2::ggplot(dat_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
   ggplot2::geom_violin(

### 2024-08-23 01_Example/Manuscript.qmd
Show missings in descriptive statistics

@@
 
-modelsummary::datasummary_balance(formula = ~ sex, data = dat_clean, dinm = FALSE)
+within(dat, {
+  sex[is.na(sex)] <- "unknown"
+}) |>
+  (\(x) { # Add number of non-missings
+    idx <- which(sapply(x, is.numeric))
+    n <- sapply(x, \(v) sum(!is.na(v)))
+    names(x) <- names(x) |>
+      replace(
+        list = idx,
+        values = paste0(names(x), " (N=", n, ")")[idx]
+      )
+    x
+  })() |>
+  modelsummary::datasummary_balance(formula = ~ sex, dinm = FALSE) |>
+  tinytable::style_tt(
+    # Make table more compact to fit on page
+    tabularray_inner = "rowsep={0.2em}, colsep = {0.2em}"
+  )
 

### 2024-08-23 01_Example/Manuscript.qmd
Treat year as categorical variable

--- 2024-08-23 01_Example/Manuscript.qmd
@@
 
-modelsummary::datasummary_balance(formula = ~ sex, data = dat_clean, dinm = FALSE)
+within(dat, {
+  sex[is.na(sex)] <- "unknown"
+  year <- as.factor(year) # Year is more interesting if analyzed as factor
+}) |>
+  (\(x) { # Add number of non-missings
+    idx <- which(sapply(x, is.numeric))
+    n <- sapply(x, \(v) sum(!is.na(v)))
+    names(x) <- names(x) |>
+      replace(
+        list = idx,
+        values = paste0(names(x), " (N=", n, ")")[idx]
+      )
+    x
+  })() |>
+  modelsummary::datasummary_balance(formula = ~ sex, dinm = FALSE) |>
+  tinytable::style_tt(
+    # Make table more compact to fit on page
+    tabularray_inner = "rowsep={0.2em}, colsep = {0.2em}"
+  )
 

### 2024-08-23 01_Example/Manuscript.qmd
Rename table to match text
Improve readibility

--- 2024-08-23 01_Example/Manuscript.qmd
@@
 #| echo: false
-#| tbl-cap: "Summary Statistics"
+#| tbl-cap: "Descriptive Statistics"
 
-modelsummary::datasummary_balance(formula = ~ sex, data = dat_clean, dinm = FALSE)
+within(dat, {
+  sex[is.na(sex)] <- "unknown"
+  year <- as.factor(year) # Year is more interesting if analyzed as factor
+}) |>
+  (\(x) { # Add number of non-missings
+    n <- sapply(x, \(v) sum(!is.na(v)))
+    labels <- paste0(names(x), " (N=", n, ")")
+    idx <- which(sapply(x, is.numeric))
+    names(x) <- names(x) |>
+      replace(
+        list = idx,
+        values = labels[idx]
+      )
+    x
+  })() |>
+  modelsummary::datasummary_balance(formula = ~ sex, dinm = FALSE) |>
+  tinytable::style_tt(
+    # Make table more compact to fit on page
+    tabularray_inner = "rowsep={0.2em}, colsep = {0.2em}"
+  )
 

### 2024-08-23 01_Example/Manuscript.qmd
Make Welch t-test explicit

@@
 # Perform t-test
-t_test_result <- t.test(bill_length_mm ~ sex, data = dat_clean)
+t_test_result <- t.test(bill_length_mm ~ sex, data = dat_clean, var.equal = FALSE)
 
@@
 
-We conducted a two-sample t-test to compare the average bill lengths between male and female penguins.
+We conducted a two-sample Welch t-test to compare the average bill lengths between male and female penguins.
 The null hypothesis ($H_0$) states that there is no difference in bill lengths between male and female penguins, while the alternative hypothesis ($H_1$) suggests a significant difference.

### 2024-08-23 01_Example/Manuscript.qmd
Add spacing for DOCX
See https://tex.stackexchange.com/questions/303110/avoid-space-after-commas-used-as-thousands-separator-in-math-mode/303127#303127
and https://tex.stackexchange.com/questions/74353/what-commands-are-there-for-horizontal-spacing/74354#74354

@@
 
-The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`r round(t_test_result$parameter, digits = 2)`) = `r round(t_test_result$statistic, digits = 2)`, p `r format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
-Female penguins ($M = `r round(mean(blm_female), digits = 2)`\,\mathrm{mm}, SD = `r round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `r round(mean(blm_male), digits = 2)`\,\mathrm{mm}, SD = `r round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).
+The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`r round(t_test_result$parameter, digits = 2)`) = `r round(t_test_result$statistic, digits = 2)`{,}\> p `r format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
+Female penguins ($M = `r round(mean(blm_female), digits = 2)`\,\mathrm{mm}{,}\> SD = `r round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `r round(mean(blm_male), digits = 2)`\,\mathrm{mm}{,}\> SD = `r round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).
 

### 2024-08-23 01_Example/Manuscript.qmd
Use A4 format

@@
     keep-tex: true
+    a4paper: true
 ---

### 2024-08-23 01_Example/Manuscript.qmd
Use flextable rather than tinytable
to allow referencing tables
see https://github.com/wjschne/apaquarto/issues/83

@@
 
-Descriptive statistics of the data set are displayed in Table 1.
+Descriptive statistics of the data set are given in @tbl-descriptive-statistics and individual bill lengths are displayed in @fig-bill-length-comparison.
 
 ```{r}
-#| label: "descriptive-statistics"
+#| label: "tbl-descriptive-statistics"
 #| echo: false
 #| tbl-cap: "Descriptive Statistics"
+#| apa-twocolumn: true
 
@@
   })() |>
-  modelsummary::datasummary_balance(formula = ~ sex, dinm = FALSE) |>
-  tinytable::style_tt(
-    # Make table more compact to fit on page
-    tabularray_inner = "rowsep={0.2em}, colsep = {0.2em}"
-  )
+  modelsummary::datasummary_balance(formula = ~ sex, output = "flextable", dinm = FALSE) |>
+  flextable::theme_apa() |>
+  flextable::font(fontname = "TeX Gyre Termes", part = "all")
 

### 2024-08-23 01_Example/Manuscript.qmd
Emphasize t in t-test

@@
 
-We conducted a two-sample Welch t-test to compare the average bill lengths between male and female penguins.
+We conducted a two-sample Welch *t*-test to compare the average bill lengths between male and female penguins.
 The null hypothesis ($H_0$) states that there is no difference in bill lengths between male and female penguins, while the alternative hypothesis ($H_1$) suggests a significant difference.
 
-The t-test was performed using the `t.test()` function in R, with a significance level of 0.05.
+The *t*-test was performed using the `t.test()` function in R, with a significance level of 0.05.
 
@@
 
-The t-test revealed a significant difference in bill lengths between male and female penguins ($t(`r round(t_test_result$parameter, digits = 2)`) = `r round(t_test_result$statistic, digits = 2)`{,}\> p `r format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
+The *t*-test revealed a significant difference in bill lengths between male and female penguins ($t(`r round(t_test_result$parameter, digits = 2)`) = `r round(t_test_result$statistic, digits = 2)`{,}\> p `r format.pval(t_test_result$p.value, eps = 0.001, scientific = FALSE)`$).
 Female penguins ($M = `r round(mean(blm_female), digits = 2)`\,\mathrm{mm}{,}\> SD = `r round(sd(blm_female), digits = 2)`\,\mathrm{mm}$) exhibited shorter bill lengths compared to male penguins ($M = `r round(mean(blm_male), digits = 2)`\,\mathrm{mm}{,}\> SD = `r round(sd(blm_male), digits = 2)`\,\mathrm{mm}$).

### 2024-08-23 01_Example/Manuscript.tex
Update output files

@@
   longtable,
+  a4paper,
   nolmodern,
@@
 
-\date{2024-08-22}
+\date{2024-08-23}
 
@@
 
+\usepackage{fontspec}
+\usepackage{multirow}
+\usepackage{multicol}
+\usepackage{colortbl}
+\usepackage{hhline}
+\newlength\Oldarrayrulewidth
+\newlength\Oldtabcolsep
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{hyperref}
 \usepackage{float}
-\usepackage{tabularray}
-\usepackage[normalem]{ulem}
-\usepackage{graphicx}
-\UseTblrLibrary{booktabs}
-\UseTblrLibrary{siunitx}
-\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
-\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
-\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
+\usepackage{wrapfig}
 \makeatletter
@@
 
-We conducted a two-sample t-test to compare the average bill lengths
-between male and female penguins. The null hypothesis (\(H_0\)) states
-that there is no difference in bill lengths between male and female
-penguins, while the alternative hypothesis (\(H_1\)) suggests a
-significant difference.
+We conducted a two-sample Welch \emph{t}-test to compare the average
+bill lengths between male and female penguins. The null hypothesis
+(\(H_0\)) states that there is no difference in bill lengths between
+male and female penguins, while the alternative hypothesis (\(H_1\))
+suggests a significant difference.
 
-The t-test was performed using the \texttt{t.test()} function in R, with
-a significance level of 0.05.
+The \emph{t}-test was performed using the \texttt{t.test()} function in
+R, with a significance level of 0.05.
 
@@
 
-Descriptive statistics of the data set are displayed in Table 1.
+Descriptive statistics of the data set are given in
+Table~\ref{tbl-descriptive-statistics} and individual bill lengths are
+displayed in Figure~\ref{fig-bill-length-comparison}.
 
 \begin{table}
-\caption{Summary Statistics}\tabularnewline
 
-\centering
-\begin{tblr}[         %% tabularray outer open
-]                     %% tabularray outer close
-{                     %% tabularray inner open
-colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
-cell{1}{3}={c=2,}{halign=c,},
-cell{1}{5}={c=2,}{halign=c,},
-cell{1}{7}={c=2,}{halign=c,},
-column{1}={halign=l,},
-column{2}={halign=l,},
-column{3}={halign=r,},
-column{4}={halign=r,},
-column{5}={halign=r,},
-column{6}={halign=r,},
-column{7}={halign=r,},
-column{8}={halign=r,},
-row{1}={halign=c,},
-hline{8}={1,2,3,4,5,6,7,8}{solid, 0.05em, black},
-}                     %% tabularray inner close
-\toprule
-&  & Adelie (N=146) &  & Chinstrap (N=68) &  & Gentoo (N=119) &  \\ \cmidrule[lr]{3-4}\cmidrule[lr]{5-6}\cmidrule[lr]{7-8}
-&    & Mean & Std. Dev. & Mean & Std. Dev. & Mean & Std. Dev. \\ \midrule %% TinyTableHeader
-bill\_length\_mm    &           & \num{38.8}   & \num{2.7}   & \num{48.8}   & \num{3.3}   & \num{47.6}   & \num{3.1}   \\
-bill\_depth\_mm     &           & \num{18.3}   & \num{1.2}   & \num{18.4}   & \num{1.1}   & \num{15.0}   & \num{1.0}   \\
-flipper\_length\_mm &           & \num{190.1}  & \num{6.5}   & \num{195.8}  & \num{7.1}   & \num{217.2}  & \num{6.6}   \\
-body\_mass\_g       &           & \num{3706.2} & \num{458.6} & \num{3733.1} & \num{384.3} & \num{5092.4} & \num{501.5} \\
-year                  &           & \num{2008.1} & \num{0.8}   & \num{2008.0} & \num{0.9}   & \num{2008.1} & \num{0.8}   \\
-&           & N             & Pct.         & N             & Pct.         & N             & Pct.         \\
-island                & Biscoe    & \num{44}     & \num{30.1}  & \num{0}      & \num{0.0}   & \num{119}    & \num{100.0} \\
-& Dream     & \num{55}     & \num{37.7}  & \num{68}     & \num{100.0} & \num{0}      & \num{0.0}   \\
-& Torgersen & \num{47}     & \num{32.2}  & \num{0}      & \num{0.0}   & \num{0}      & \num{0.0}   \\
-sex                   & female    & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{58}     & \num{48.7}  \\
-& male      & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{61}     & \num{51.3}  \\
-\bottomrule
-\end{tblr}
+{\caption{{Descriptive Statistics}{\label{tbl-descriptive-statistics}}}
+\vspace{-20pt}}
+
+\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}
+
+\global\setlength{\Oldtabcolsep}{\tabcolsep}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\begin{longtable*}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\ascline{0.75pt}{000000}{1-8}
+
+\multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ }}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{female\ (N=165)}}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{male\ (N=168)}}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{unknown\ (N=11)}}}} \\
+
+\ascline{1.5pt}{666666}{1-8}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ \ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean\ \ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.\ \ }}}} \\
+
+\ascline{0.75pt}{000000}{1-8}\endfirsthead
+
+\ascline{0.75pt}{000000}{1-8}
+
+\multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ }}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{female\ (N=165)}}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{male\ (N=168)}}}} & \multicolumn{2}{>{\centering}m{\dimexpr 1.5in+2\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{unknown\ (N=11)}}}} \\
+
+\ascline{1.5pt}{666666}{1-8}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{\ \ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.\ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Mean\ \ }}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Std.\ Dev.\ \ }}}} \\
+
+\ascline{0.75pt}{000000}{1-8}\endhead
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{bill\_length\_mm\ (N=342)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{42.1}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{4.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{45.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{5.4}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{41.3}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{4.6}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{bill\_depth\_mm\ (N=342)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{16.4}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{1.8}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{17.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{1.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{16.6}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{2.2}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{flipper\_length\_mm\ (N=342)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{197.4}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{12.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{204.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{14.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{199.0}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{16.5}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{body\_mass\_g\ (N=342)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{3862.3}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{666.2}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{4545.7}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{787.6}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{4005.6}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{679.4}}}} \\
+
+\ascline{1pt}{000000}{1-8}
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{N}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Pct.}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{N}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Pct.}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{N}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Pct.}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{species}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Adelie}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{73}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{44.2}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{73}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{43.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{6}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{54.5}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Chinstrap}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{34}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{20.6}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{34}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{20.2}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{0}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{0.0}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Gentoo}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{58}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{35.2}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{61}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{36.3}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{45.5}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{island}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Biscoe}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{80}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{48.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{83}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{49.4}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{45.5}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Dream}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{61}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{37.0}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{62}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{36.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{1}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{9.1}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{Torgersen}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{24}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{14.5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{23}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{13.7}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{5}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{45.5}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{year}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{2007}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{51}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{30.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{52}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{31.0}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{7}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{63.6}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{2008}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{56}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{33.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{57}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{33.9}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{1}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{9.1}}}} \\
+
+\multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{2009}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{58}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{35.2}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{59}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{35.1}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{3}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{TeX Gyre Termes}{27.3}}}} \\
+
+\ascline{0.75pt}{000000}{1-8}
+
+\end{longtable*}
+
+\arrayrulecolor[HTML]{000000}
+
+\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}
+
+\global\setlength{\tabcolsep}{\Oldtabcolsep}
+
+\renewcommand*{\arraystretch}{1}
+
 \end{table}
 
-The t-test revealed a significant difference in bill lengths between
-male and female penguins (\(t(329.29) = -6.67, p < 0\.001\)). Female
-penguins (\(M = 42.1\,\mathrm{mm}, SD = 4.9\,\mathrm{mm}\)) exhibited
-shorter bill lengths compared to male penguins
-(\(M = 45.85\,\mathrm{mm}, SD = 5.37\,\mathrm{mm}\)).
+The \emph{t}-test revealed a significant difference in bill lengths
+between male and female penguins (\(t(329.29) = -6.67{,}\> p < 0.001\)).
+Female penguins (\(M = 42.1\,\mathrm{mm}{,}\> SD = 4.9\,\mathrm{mm}\))
+exhibited shorter bill lengths compared to male penguins
+(\(M = 45.85\,\mathrm{mm}{,}\> SD = 5.37\,\mathrm{mm}\)).
 

### 2024-08-23 01_Example/Manuscript.qmd
Use LuaLaTeX
to make TeX Gyre Termes work
(a Times New Roman replacement)

@@
     keep-tex: true
+    pdf-engine: lualatex
     a4paper: true

### 2024-08-23 01_Example/Manuscript.qmd
Add more comments

@@
 }) |>
-  (\(x) { # Add number of non-missings
-    n <- sapply(x, \(v) sum(!is.na(v)))
-    labels <- paste0(names(x), " (N=", n, ")")
-    idx <- which(sapply(x, is.numeric))
-    names(x) <- names(x) |>
-      replace(
-        list = idx,
-        values = labels[idx]
-      )
+  (\(x) { # Add number of non-missings to labels
+    n <- sapply(x, \(v) sum(!is.na(v))) # Calculate number of non-missings
+    labels <- paste0(names(x), " (N=", n, ")") # Paste labels and numbers together
+    idx <- which(sapply(x, is.numeric)) # Get numeric variables
+    names(x)[idx] <- labels[idx] # Change the labels
     x

### 2024-08-23 01_Example/Manuscript.qmd
Simplify descriptive statistics

--- 2024-08-23 01_Example/Manuscript.qmd
@@
 
-within(dat, {
-  sex[is.na(sex)] <- "unknown"
-  year <- as.factor(year) # Year is more interesting if analyzed as factor
-}) |>
-  (\(x) { # Add number of non-missings
-    n <- sapply(x, \(v) sum(!is.na(v)))
-    labels <- paste0(names(x), " (N=", n, ")")
-    idx <- which(sapply(x, is.numeric))
-    names(x) <- names(x) |>
-      replace(
-        list = idx,
-        values = labels[idx]
-      )
-    x
-  })() |>
-  modelsummary::datasummary_balance(formula = ~ sex, output = "flextable", dinm = FALSE) |>
+dat_desc <- dat
+
+dat_desc$sex[is.na(dat_desc$sex)] <- "unknown" # Convert NA to "unknown"
+dat_desc$year <- as.factor(dat_desc$year) # Convert year to factor
+
+# Add number of non-missings to names of numeric variables
+n <- sapply(dat_desc, \(v) sum(!is.na(v))) # Calculate number of non-missings
+labels <- paste0(names(dat_desc), " (N=", n, ")") # Paste variable names and numbers together
+idx <- which(sapply(dat_desc, is.numeric)) # Identify numeric variables
+names(dat_desc)[idx] <- labels[idx] # Change the variable names
+
+modelsummary::datasummary_balance(
+  formula = ~ sex,
+  data = dat_desc,
+  output = "flextable",
+  dinm = FALSE
+) |>
   flextable::theme_apa() |>
@@
 ```{r}
-#| label: "variables"
+#| label: "bill-lengths"
 #| echo: false

### 2024-08-23 01_Example/Manuscript.tex
Radically simplify analysis

--- 2024-08-22 01_Example/Manuscript.tex
@@
 
-\usepackage{float}
-\usepackage{tabularray}
-\usepackage[normalem]{ulem}
-\usepackage{graphicx}
-\UseTblrLibrary{booktabs}
-\UseTblrLibrary{siunitx}
-\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
-\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
-\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
 \makeatletter
@@
 
-We conducted a two-sample t-test to compare the average bill lengths
-between male and female penguins. The null hypothesis (\(H_0\)) states
-that there is no difference in bill lengths between male and female
-penguins, while the alternative hypothesis (\(H_1\)) suggests a
-significant difference.
+We conducted a Welch two-sample \emph{t}-test to compare the average
+bill lengths between male and female penguins. The null hypothesis
+(\(H_0\)) states that there is no difference in bill lengths between
+male and female penguins, while the alternative hypothesis (\(H_1\))
+suggests a significant difference.
 
-The t-test was performed using the \texttt{t.test()} function in R, with
-a significance level of 0.05.
+The \emph{t}-test was performed using the \texttt{t.test()} function in
+R, with a significance level of 0.05.
 
@@
 
-Descriptive statistics of the data set are displayed in Table 1.
+Descriptive statistics of the data set are given in
+Table~\ref{tbl-descriptive-statistics} and individual bill lengths are
+displayed in Figure~\ref{fig-bill-length-comparison}.
+
+The Welch Two Sample t-test testing the difference of bill\_length\_mm
+by sex (mean in group female = 42.10, mean in group male = 45.85)
+suggests that the effect is negative, statistically significant, and
+large (difference = -3.76, 95\% CI {[}-4.87, -2.65{]}, t(329.29) =
+-6.67, p \textless{} .001, Cohen's d = -0.73)
 
 \begin{table}
-\caption{Summary Statistics}\tabularnewline
 
-\centering
-\begin{tblr}[         %% tabularray outer open
-]                     %% tabularray outer close
-{                     %% tabularray inner open
-colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
-cell{1}{3}={c=2,}{halign=c,},
-cell{1}{5}={c=2,}{halign=c,},
-cell{1}{7}={c=2,}{halign=c,},
-column{1}={halign=l,},
-column{2}={halign=l,},
-column{3}={halign=r,},
-column{4}={halign=r,},
-column{5}={halign=r,},
-column{6}={halign=r,},
-column{7}={halign=r,},
-column{8}={halign=r,},
-row{1}={halign=c,},
-hline{8}={1,2,3,4,5,6,7,8}{solid, 0.05em, black},
-}                     %% tabularray inner close
-\toprule
-&  & Adelie (N=146) &  & Chinstrap (N=68) &  & Gentoo (N=119) &  \\ \cmidrule[lr]{3-4}\cmidrule[lr]{5-6}\cmidrule[lr]{7-8}
-&    & Mean & Std. Dev. & Mean & Std. Dev. & Mean & Std. Dev. \\ \midrule %% TinyTableHeader
-bill\_length\_mm    &           & \num{38.8}   & \num{2.7}   & \num{48.8}   & \num{3.3}   & \num{47.6}   & \num{3.1}   \\
-bill\_depth\_mm     &           & \num{18.3}   & \num{1.2}   & \num{18.4}   & \num{1.1}   & \num{15.0}   & \num{1.0}   \\
-flipper\_length\_mm &           & \num{190.1}  & \num{6.5}   & \num{195.8}  & \num{7.1}   & \num{217.2}  & \num{6.6}   \\
-body\_mass\_g       &           & \num{3706.2} & \num{458.6} & \num{3733.1} & \num{384.3} & \num{5092.4} & \num{501.5} \\
-year                  &           & \num{2008.1} & \num{0.8}   & \num{2008.0} & \num{0.9}   & \num{2008.1} & \num{0.8}   \\
-&           & N             & Pct.         & N             & Pct.         & N             & Pct.         \\
-island                & Biscoe    & \num{44}     & \num{30.1}  & \num{0}      & \num{0.0}   & \num{119}    & \num{100.0} \\
-& Dream     & \num{55}     & \num{37.7}  & \num{68}     & \num{100.0} & \num{0}      & \num{0.0}   \\
-& Torgersen & \num{47}     & \num{32.2}  & \num{0}      & \num{0.0}   & \num{0}      & \num{0.0}   \\
-sex                   & female    & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{58}     & \num{48.7}  \\
-& male      & \num{73}     & \num{50.0}  & \num{34}     & \num{50.0}  & \num{61}     & \num{51.3}  \\
-\bottomrule
-\end{tblr}
+{\caption{{Descriptive Statistics}{\label{tbl-descriptive-statistics}}}
+\vspace{-20pt}}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule\noalign{}
+Variable & female (n=165) & male (n=168) (n=333) \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+species {[}Adelie{]}, \% & 44.2 & 43.5 \\
+species {[}Chinstrap{]}, \% & 20.6 & 20.2 \\
+species {[}Gentoo{]}, \% & 35.2 & 36.3 \\
+island {[}Biscoe{]}, \% & 48.5 & 49.4 \\
+island {[}Dream{]}, \% & 37.0 & 36.9 \\
+island {[}Torgersen{]}, \% & 14.5 & 13.7 \\
+Mean bill\_length\_mm (SD) & 42.10 (4.90) & 45.85 (5.37) \\
+Mean bill\_depth\_mm (SD) & 16.43 (1.80) & 17.89 (1.86) \\
+Mean flipper\_length\_mm (SD) & 197.36 (12.50) & 204.51 (14.55) \\
+Mean body\_mass\_g (SD) & 3862.27 (666.17) & 4545.68 (787.63) \\
+year {[}2007{]}, \% & 30.9 & 31.0 \\
+year {[}2008{]}, \% & 33.9 & 33.9 \\
+year {[}2009{]}, \% & 35.2 & 35.1 \\
+\end{longtable}
+
 \end{table}
-
-The t-test revealed a significant difference in bill lengths between
-male and female penguins (\(t(329.29) = -6.67, p < 0\.001\)). Female
-penguins (\(M = 42.1\,\mathrm{mm}, SD = 4.9\,\mathrm{mm}\)) exhibited
-shorter bill lengths compared to male penguins
-(\(M = 45.85\,\mathrm{mm}, SD = 5.37\,\mathrm{mm}\)).
 

### 2024-08-26 index.qmd
Add Quarto site

---
title: "Code publishing tutorial"
---

This is a Quarto website.

To learn more about Quarto websites visit <https://quarto.org/docs/websites>.

### 2024-09-04 about.qmd
Add content pages

---
title: "About"
bibliography: grateful-refs.bib
---

```{r}
#| echo: false

grateful::cite_packages(
  output = "paragraph",
  out.dir = ".",
  pkgs = unique(renv::dependencies(progress = FALSE)$Package),
  passive.voice = TRUE
) |>
  withr::with_options(new = list(warnPartialMatchDollar = FALSE))
```

```{r}
#| output: "asis"
#| echo: false

cat(c("Quarto:", as.character(quarto::quarto_version())))
```

```{r}
#| echo: false

sessioninfo::session_info()
```

### 2024-09-04 add_readme.qmd
Add content pages

---
title: "Add a README"
lang: en
engine: knitr
---

## Creating a README

Once you have prepared your project and settled on a license, it is time to add a final touch. Imagine you are coming back to your project in five years, having forgotten most of the details of what you did exactly. What would be useful to know in order to quickly understand what is going on in the project? This is what needs to be described in the README. While you could just start writing along, most READMEs have some sections in common which we describe below.

Name

:   How is the project called?

Badges and project status

:   Badges typically report quick facts, like [whether the project is actively maintained](https://www.repostatus.org/), [how many dependencies it has](https://tinyverse.netlify.app/), or [whether it has been published in a package repository](https://www.r-pkg.org/services#badges). Sometimes, the license, any associated DOIs, or quality metrics like code coverage by unit tests are communicated via badges as well.

Description

:   What is the project about? What are its features? Is there a published article associated with it?

Visuals

:   If there is anything you can show that demonstrates how the project can be used? Screenshots or other visuals can make the README more appealing.

Installation/Dependencies

:   What steps need to be taken to run the project? What software needs to be installed? Mention all dependencies here that are not explicitly managed by `renv`, such as the system dependencies of R packages as well as the version of Quarto.

    An overview over the system dependencies of one more multiple R packages can be created using the function `pkg_sysreqs()` from the R package `pak`. In combination with `renv`, we can obtain the system dependencies of all R packages in the current project:

    ```{.r filename="Console"}
    pak::pkg_sysreqs(renv::dependencies()$Package)
    ```

    The output may look like the following:

    ```
    ── Install scripts ─────────────────────── 
    dnf install -y make pandoc

    ── Packages and their system dependencies 
    fs        – make
    knitr     – pandoc
    rmarkdown – pandoc
    sass      – make
    ```

    We can see that `make` and `pandoc` were identified as system dependencies. One can obtain their version by running them with the `--version` argument:

    ```{.bash filename="Terminal"}
    make --version
    pandoc --version
    ```

    We also know that we need Quarto to create the PDF and that we installed the `apaquarto` extension, so let's find out these versions as well:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]

    ```{.bash filename="Terminal"}
    quarto --version
    quarto list extensions
    ```

    Finally, we know that we installed a $\TeX$ distribution to create the PDF, so let's find out its version by running:

    ```{.bash filename="Terminal"}
    quarto check
    ```

    The output is quite long and it might look slightly different for you, but the relevant sections are the following:

    ```
    [✓] Checking tools....................OK
          TinyTeX: (not installed)
          Chromium: (not installed)

    [✓] Checking LaTeX....................OK
          Using: Installation From Path
          Path: /usr/local/texlive/2024/bin/x86_64-linux
          Version: 2024
    ```

    We put all the information obtained about system dependencies and their versions in the README. Sharing the install scripts that were provided by `pak` is an additional courtesy to your readers.

Usage

:   Which files does the project contain? How are they organized? How can one run the project -- is there a master script or a particular order in which any scripts need to be executed? How long does it take to run all scripts? Is there additional documentation available?

Support

:   Do you offer support or help, for example, via GitHub discussions or a mailing list?

Contributing

:   If the project is active: Can other people contribute? How? Do you accept contributions? Do you review issues? This section is sometimes outsourced into a file called `CONTRIBUTING.md`.

Authors

:   Who was involved in creating this project? This involves you, your co-authors, and anybody you accepted contributions from.

Citation

:   Is there a recommended way to cite this project? We recommend to increase citability by using the [CodeMeta generator](https://codemeta.github.io/codemeta-generator/) to create the file `codemeta.json`. Alternatively, one can also use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create the file.

License

:   Under which license are the contents in this project available?

Create your README now as the file `README.md`.

### 2024-09-04 archive.qmd
Add content pages

---
title: "Archive"
lang: "en"
bibliography: literature.bib
---

The project is now ready to be deposited for long-term archival. When considering an appropriate repository, there are a few questions to ponder [@Jackson2018]:

> __Does the digital repository give you a unique persistent digital identifier for your deposit?__
> 
> Persistent digital identifiers are identifiers that are provided under the assumption that they, and the digital objects they refer to, will remain available long term. Examples of unique persistent digital identifiers include DOIs and ARKs. The use of persistent digital identifiers for research outputs, including papers, data and software, and the use of these for citation purposes, is becoming prevalent in research.
> 
> __Can the digital repository accommodate the size of your deposit?__
> 
> Digital repositories may have constraints on the size of deposits they accept. For example, Zenodo has a limit of 50 GB and the University of Edinburgh DataShare has a limit of 20 GB for deposits done via a browser and 100 GB otherwise.
> 
> __Has your funder or publisher recommended or mandated a digital repository to use?__
> 
> If both a funder and publisher mandate two different digital repositories you could deposit within them both, or discuss with each to see if they will agree for you to use one only, even if it is not the one they mandate.
> 
> __Are the policies of the digital repository acceptable to you?__
> 
> These can include terms and conditions of use; availability of the service (e.g. does the service regularly go down for updates or does it run 24/7); how resilient the service is to breakdowns; where deposits are stored; how they are stored; what are the plans for stored deposits if the repository goes out of business; how secure the storage is; how regularly it is backed up; how long backups are kept for; and what disaster recovery plans are in place. Knowing the quality of service can help you judge whether your deposit will be available to others, and yourself, even in the face of disaster. It also allows you, and your institution, to develop contingency plans should such drastic circumstances arise.
> 
> __Is the longevity of the digital repository acceptable to you?__
> 
> You want to know that the digital repository will host your deposit for as long as you need it too, not just 6 months or so. The policies or terms and conditions of the digital repository should state this. For example, Zenodo states that their lifetime is currently 20 years and figshare guarantees 10 years of persistent availability.
> 
> __Is the digital repository free or do you have to pay a fee?__
> __If there is a fee, is this a one-off payment and can you afford it?__
> 
> If a digital repository expects a one-off payment, to subsidise its hosting, you may be able to pay for this out of your project budget or your institution may pay this for you. It is recommended that you avoid any digital repository that expects a regular fee for the duration for which they hold the deposit, since it would most likely be unsustainable for your institution to commit to ever-increasing fees over a long term. There are many free digital repositories available.
> 
> __Is the digital repository accredited or certified?__
> 
> The digital repository may be accredited or certified by an independent agency that reviews digital repositories according to various criteria related to its policies and processes. The presence of an accreditation or certification may provide you with an additional degree of confidence that the digital repository meets your requirements. For example, the Data Seal of Approval can be awarded to digital repositories who satisfy a set of quality guidelines relating to how they manage, preserve and make available the deposits they hold. The presence of a Data Seal of Approval could give you additional confidence that a digital repository is suitable for holding your deposit.

Additional guidance is provided by @Daquin2023 and @Garijo2022. We will cover the archival using the Software Heritage archive and Zenodo.

For further information, you may be interested in [Software Deposit Guidance for Researchers](https://softwaresaved.github.io/software-deposit-guidance/).

### 2024-09-04 choose_license.qmd
Add content pages

---
title: "Choose a License"
lang: en
bibliography: literature.bib
format:
  html:
    mermaid:
      theme: neutral
engine: knitr
---

## A Primer on Licenses

Whenever you create a literary or artistic work (such as a text, image, video, or software), the copyright law in most countries limits other people from copying, modifying, and sharing it without your express permission. [If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it. This legal default of "all rights reserved" runs counter to many cultural and scientific processes. Copyright licenses enable authors to free up their works for reuse by others.

::: {.column-margin}
For the purpose of this tutorial, by _license_ we mean _copyright license_.
:::

A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights. And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others. By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]

::: {#cau-exclusive-rights .callout-caution}
### Publishing Without Exclusive Rights

When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license that goes further than the open access license available to readers of the journal [@Rumsey2022]. However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide by @Elson2016 to learn how to do that.

If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) and [Dissemin](https://dissem.in/) for legal options to still make it available free of charge to readers.
:::

The licenses we discuss here mostly regard copyright. Therefore, recipients may lack other rights such as publicity, privacy, moral, patent,^[In fact, the software licenses we recommend on this page have been specifically selected to provide an express grant of patent rights.] or trademark rights. For example, sharing photos that depict people is not only a matter of copyright, but also of privacy rights. Conversely, the licenses do not apply if recipients are allowed to use the works for other reasons such as fair use, the right to quote, or because they made a different arrangement with the author.

## Which License to Choose?

Many boilerplate licenses are available to apply to your work. Which license is appropriate depends on several factors, including existing licenses in place and the type of work, but also your personal considerations. We strongly recommend to apply a _free/open_ license to your work, which means that the work "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015]. Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes. By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that modifications are only shared under the same license, and that no further restrictions are imposed on the work. Software licenses may additionally require to make the source code available to everybody the software is shared with. Because there are many free/open licenses available, the following decision tree only contains a recommended subset.

::: {.column-margin}
If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos). For software that's almost universally the [GPLv3](https://www.gnu.org/licenses/license-list.en.html#GPLCompatibleLicenses) and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/compatiblelicenses). For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and CC\ BY-SA\ 4.0 are acceptable.
:::

::: {#nte-terminology .callout-note collapse="true"}
### A Note on Terminology

The terms _free_ and _open_, especially with regard to software, come with a history. The Free Software Foundation (FSF) was founded in 1985 to protect "four essential freedoms" [-@FSF2024] of a program's user. These are the freedoms to use, study, share, and improve a program. Software whose users legally and practically have these freedoms (because, among other things, they have access to its source code) is considered _free_. The four freedoms are seen as vital for a society as a whole in the sense that they enable sharing, cooperation and ultimately freedom in general. The FSF maintains [a list of software licenses](https://www.gnu.org/licenses/license-list.en.html#SoftwareLicenses) that it considers to be protecting the four freedoms. Sometimes the term _libre_ (Spanish and French for _free_) is used to make a distinction from _gratis_ software. You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).

The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach. It is concerned with developing high-quality software, for which everyone's ability to obtain, modify and contribute back the source code is considered beneficial. Access to the source code is one out of multiple conditions for software to be considered _open source_ by the OSI [-@OSI2007], which equally maintains [a list of approved licenses](https://opensource.org/licenses).

Access to the source code is a necessary, but not sufficient requirement both for _free software_ and for _open source software_. Conversely, both can (and frequently are) sold for money, as their respective criteria only apply once one has access to the software. Throughout this tutorial, we write "free/open license" to mean a license that is approved by either the FSF or the OSI. Software which is neither but makes available its source code is sometimes referred to as _source-available software_.

Apart from software, the term _open access_ has often been used for works that are available at no cost. For example, this is the commonality of bronze, green, hybrid, gold, and diamond/platinum open access _articles_, which otherwise vary in the rights that are granted to readers. In 2002, the Budapest Open Access Initiative declared that _open access_ additionally includes the right to use articles for any purpose, and in 2003, the Bethesda Statement and the Berlin Declaration added the right to make derivative works.

Two other notable definitions include the _Open Definition_ [@OKFN2016], which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015], for which the open editing phase began in 2006. They are largely viewed as compatible with one another.

:::: {.column-margin}
![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
::::
:::

### Existing License?

First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications. The same applies if there are strong community norms to use a particular free/open license.^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.

### Work Type?

If you create a new work and no strong community norms suggest a particular license, you need to choose the license yourself. Which license to choose depends on the type of work you create. Software licenses, for example, may consider that the source code is the preferred form for making modifications, while licenses for data can differentiate between the database and any works produced from it. We have created a flowchart that covers the most likely types of works you will create as a researcher: software, writing (i.e., text), images, audio, video, and data (see @fig-flowchart-simple). The flowchart always recommends the most permissive license possible to maximize reuse. Click on the name of a license to learn more about it.

Note, that for data you have, at least in principle, the option to apply separate licenses to the individual entries and the collective database. For example, if you were to create a database of artworks by others, those artworks would be licensed individually as chosen by the artists, but the license for the database as a whole could be chosen by you. The latter includes the structure of the database (e.g., the selection of entries and field names) and any sui generis database rights. However, if the content was created by you, we recommend you to choose the same license for both content and database. Metadata, in particular, should always be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

::: {#fig-flowchart-simple}
```{mermaid}
flowchart TB
  start("We want<br>to choose<br>a license.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
  start --"We created the work<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")

  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
  norm --"Does not<br>exist"--> type("Work type?")

  type --"Software"--> apache["Apache 2.0"]
  type --"Writing, image, audio, video"--> cc0["CC0 1.0"]
  type --"Data(base)"--> existing_license_data("Adapting individual<br>data entries by others?")

  existing_license_data --"No, we created them<br>entirely by ourselves."--> cc0_data["CC0 1.0 <em>for database<br>and its content</em>"]
  existing_license_data --"Yes, they were<br>shared under a<br>free/open license."--> use_existing_license_data["<em>Use their license<br>for content and</em><br>CC0 1.0 <em>for<br>the database</em>"]

  click apache href "https://choosealicense.com/licenses/apache-2.0/"
  click cc0 href "https://creativecommons.org/publicdomain/zero/1.0/"
  click cc0_data href "https://creativecommons.org/publicdomain/zero/1.0/"
```

Flowchart for Choosing a License
:::

::: {#nte-other-work-types .callout-note collapse="true"}
#### Other Work Types

There are specific licenses available (and recommended) for fonts and hardware:

- fonts: [SIL Open Font License 1.1](https://openfontlicense.org/)
- hardware: [CERN Open Hardware Licence](https://cern-ohl.web.cern.ch/)

Furthermore, one should be cautious about the restrictions of licenses applied to the following types of works:

- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/gpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016]. And if a document contained source code covered by the GPLv3, the same license would also apply to the document [@MadHatter2022].
- __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license. For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access if shared [see @Poole2017]. Similarly, following an example from @Matt2009 if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared. If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.

Finally, works in the public domain should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
:::

### Copyleft?

The consequence of applying a permissive license to your work is that others may distribute their modifications to it under stricter terms, possibly without a free/open license. You may prefer that everybody who receives somebody else's adaptations of your work enjoys the same rights that you granted to them. This can be achieved by _copyleft_ licenses.^[So called because copyleft licenses make content permanently free, thus turning copyright around: © → 🄯] For software, they come in two flavours:

- __Weak copyleft licenses__ only require that modifications to the software itself are licensed under the same or a compatible license if shared. For example, if I create and publish an R package under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license. However, people who merely _use_ my R package in their scripts can choose any license for their scripts.
- __Strong copyleft licenses,__ on the other side, insist that any larger works that use the copyleft-licensed software must also be licensed under the same or a compatible license if shared. For example, if my R package were to be put under a strong copyleft license, everybody publishing a data analysis script that uses my package would need to put it under the same license. Similarly, a Quarto manuscript using such a package would need to disclose its source code if shared.

Note, however, that the copyleft licenses we discuss here do not mandate sharing. Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015]. It is also worth reiterating that these licenses do not restrict the original author(s): They are still permitted to distribute their work under a different license and without sharing the source code.

We have prepared two advanced license flowcharts, one for software, writing, images, audio, and video in @fig-flowchart-non-data and one for data in @fig-flowchart-data where you can make additional choices. Note, however, that especially the advanced flowchart for licensing data is quite complex and we recommend you to seek legal counsel if you want to be sure.

::: {#cau-license-versions .callout-caution}
### License Versions Are Important

You may have noticed that we mostly refer to licenses using a name _and_ a version number. This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses. For example, the Creative Commons licenses (that start with `CC`) were first published in 2002. Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and sui generis database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)). There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.

For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022].
:::

### Attribution and Anti-DRM?

The advanced license flowcharts also allow you to make two additional decisions for Creative Commons licenses (i.e., when deciding between CC0\ 1.0, CC\ BY\ 4.0, and CC\ BY-SA\ 4.0):

- __Attribution__ means whether recipients of your work are required to provide attribution to you and indicate if changes were made. Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
- __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.

All software licenses in @fig-flowchart-non-data require providing attribution, although only [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) and [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/) require others to indicate changes. Of the software licenses listed, only the AGPLv3 also has an anti-DRM provision. For data, the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) allows for technological measures that restrict the work only if a version of the database is provided in parallel without such measures.

::: {#nte-silos .callout-note collapse="true"}
### Other Restrictions

As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use. There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015], but we would like to highlight one reason in particular.

With the rights provided by free/open licenses comes the possibility to build on and combine multiple works by different authors, which is essential for any cultural and scientific activity. However, this is only possible if the various licenses involved are compatible with each other. For example, one is allowed to remix two figures if the first is licensed under CC\ BY\ 4.0 and the second under CC\ BY-SA\ 4.0 because the licenses were written to be compatible with each other [compare @CC2024FAQ]. Also note that CC\ BY-SA\ 4.0 is [one-way compatible](https://creativecommons.org/compatiblelicenses) with the GPLv3, which in turn is compatible with the AGPLv3.

However, applying a restriction such as only permitting non-commercial use or not allowing derivative works creates _silos_ of works which are mutually incompatible with each other. Put differently, one cannot share a remix of two works where one is licensed under CC\ BY-SA\ 4.0 and the other only allows non-commercial use. In order to avoid silos, one should only choose license which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014]. If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012. The following diagram provides an overview of the compatibility of various licenses:

!["Open data-capable license interoperability" by Robbie Morrison licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/). Taken without modification from @Morrison2024](images/Morrison2024.jpg){.lightbox}
:::

## Applying the License

Having selected the licenses of your choice -- you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed in "[Add a README](add_readme.qmd)".

For example, the `apaquarto` extension that you included in your project contains works by others. You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with its terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice. In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.

::: {#tip-license-help .callout-tip}
### Follow the Help Provided by the License Authors

For all the licenses recommended in this tutorial, the organizations that created these licenses provide more information on how to apply them to your work:

::: {layout-ncol=2}

:::: {}
- [Apache\ 2.0](https://www.apache.org/licenses/LICENSE-2.0#apply)
- [MPL\ 2.0](https://www.mozilla.org/MPL/2.0/FAQ/#license-use)
- [AGPLv3](https://www.gnu.org/licenses/gpl-howto.html)
::::

:::: {}
- [CC licenses](https://creativecommons.org/faq/#how-do-i-apply-a-creative-commons-license-to-my-material)
- [ODbL](https://opendatacommons.org/licenses/odbl/)
::::
:::

Creative Commons even provides a range of considerations for licensors and licensees [@CC2013] and an [interactive chooser](https://creativecommons.org/choose/) which you can use to create snippet that you can copy and paste to the desired location.
:::

While it is common to state the chosen license(s) in the README, usually one of the following actions is taken in addition:

In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim. This is a practice propagated by GitHub, who provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). However, if the project is not completely covered by one single license, this practice may become unwieldy. For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.

Individual programming languages also have their own way of stating which license a package is distributed under. For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].

Finally, @nte-reuse explains how to use the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable. This is the approach we recommend taking.

::: {#nte-reuse .callout-note collapse="true"}
### Using REUSE to Record Licenses

Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/) which allows communicating the license choice unequivocally. We will be using that to indicate the license for every file in your project, along with the year of publication and the copyright holder. To do this, we add a comment to the beginning of every file and add the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`. How this works depends on the file type, as the syntax for a comment varies.

For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name:

```{.r filename="create_data_dictionary.R"}
# SPDX-FileCopyrightText: <YEAR> <NAME>
#
# SPDX-License-Identifier: CC0-1.0
```

You need to use `#` to start the comment because this is the symbol that starts comment lines in R scripts. Alternatively, you can use the [reuse tool](https://github.com/fsfe/reuse-tool) to add these information for you. After installing it with...

```{.bash filename="Terminal"}
pip install reuse
```

...you can add the copyright information using the following command -- the current year will be added automatically:

```{.bash .code-overflow-wrap filename="Terminal"}
reuse annotate --copyright="<NAME>" --license="CC0-1.0" create_data_dictionary.R
```

In many cases, the reuse tool will figure out the appropriate comment style for you. If this is not the case, as currently with Quarto files, you can tell it directly which comment style to use (`html` in this case):

```{.bash .code-overflow-wrap filename="Terminal"}
reuse annotate --copyright="Josephine Zerna <josephine.zerna@tu-dresden.de>" --copyright="Christoph Scheffel <christoph_scheffel@tu-dresden.de>" --copyright="Florian Kohrt" --license="CC-BY-4.0" --style=html Manuscript.qmd
```

This adds the following header to `Manuscript.qmd`:

```{.html filename="Manuscript.qmd"}
<!--
SPDX-FileCopyrightText: 2024 Christoph Scheffel <christoph_scheffel@tu-dresden.de>
SPDX-FileCopyrightText: 2024 Florian Kohrt
SPDX-FileCopyrightText: 2024 Josephine Zerna <josephine.zerna@tu-dresden.de>

SPDX-License-Identifier: CC-BY-4.0
-->
```

Note that `Manuscript.qmd` was provided to you under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), which is what you indicate with the previous comment. If you edited the file, you may also add yourself.

Sometimes, there are file types which do not allow for adding the license information inside them, such as PDF and CSV files. For these, a corresponding `.license` file can be created. Try the following command which indicates that the data were published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):

```{.bash .code-overflow-wrap filename="Terminal"}
reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" Data.csv
```

You will notice that this creates another file called `Data.csv.license` containing the relevant information:

```{.yml filename="Data.csv.license"}
SPDX-FileCopyrightText: 2024 Kristen Gorman

SPDX-License-Identifier: CC0-1.0
```

If you want to indicate the license for all files in a particular folder, you can create a filed called `REUSE.toml` and add an `[[annotations]]` table for them:

```{.toml filename="REUSE.toml"}
version = 1

# apaquarto extension from https://github.com/wjschne/apaquarto
[[annotations]]
path = "_extensions/wjschne/apaquarto/*"
SPDX-FileCopyrightText = "2024 William Joel Schneider <w.joel.schneider@gmail.com>"
SPDX-License-Identifier = "CC0-1.0"
```

Finally, there may be some minor files which are build artifacts. You can either add them to your `.gitignore` file or use the CC0\ 1.0 license/waiver with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder. For more information, also discussing other corner cases, you can read their [Frequently Asked Questions](https://reuse.software/faq/).

Once you are done, you can download the texts of all indicated licenses using...

```{.bash filename="Terminal"}
reuse download --all
```

...and verify that you did not miss a file by running...

```{.bash filename="Terminal"}
reuse lint
```
:::

Regardless of how exactly the licenses are added to the project, this is a good opportunity to verify one last time that all third party content is provided to you under a free/open license and that you comply with it. Please add a license to your project now, either creating a file `LICENSE.txt` or following the REUSE standard.

## Wrap-up

If you would like to learn more about copyright and licenses you might find the following resources interesting:

- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2014
- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004

## Additional Figures {.appendix}

::: {#fig-flowchart-non-data}
```{mermaid}
flowchart TB
  start("We want to choose a<br>license for software,<br>writing, image, audio,<br>or video.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
  start --"We created the work<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")

  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
  norm --"Does not<br>exist"--> type("Work type?")

  type --"Software"--> code_sa(Copyleft?)
  type --"Writing, image, audio, video"--> nocode_cc("Anti-DRM?<br>Attribution?<br>Copyleft?")

  code_sa --"No (permissive)"--> apache["Apache 2.0"]
  code_sa --"Yes (weak)"--> mpl["MPL 2.0"]
  code_sa --"Yes (strong)"--> agpl["AGPLv3"]

  nocode_cc --"Neither"--> cc0["CC0 1.0"]
  nocode_cc --"Anti-DRM &<br>Attribution"--> cc_by["CC BY 4.0"]
  nocode_cc --"Anti-DRM &<br>Attribution &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]

  click apache href "https://choosealicense.com/licenses/apache-2.0/"
  click mpl href "https://choosealicense.com/licenses/mpl-2.0/"
  click agpl href "https://choosealicense.com/licenses/agpl-3.0/"
  click cc0 href "https://creativecommons.org/publicdomain/zero/1.0/"
  click cc_by href "https://creativecommons.org/licenses/by/4.0/"
  click cc_by_sa href "https://creativecommons.org/licenses/by-sa/4.0/"
```

Advanced License Flowchart for Software, Writing, Images, Audio, and Video
:::

::: {#fig-flowchart-data}
```{mermaid}
flowchart TB
  start("We want to choose a<br>license for data.") --"We adapted a database by<br>others shared under a<br>free/open license."--> use_existing_license_db["<em>Use its license(s)<br>for content and database</em>"]
  start --"We created a database<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")

  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
  norm --"Does not<br>exist"--> existing_license_content("Adapting content<br> by others?")

  subgraph "License for individual data entries (content)"
  existing_license_content --"No, we created the content<br>entirely by ourselves."--> metadata("Is metadata?")
  existing_license_content --"Yes, it was<br>shared under a<br>free/open license."--> use_existing_license_content["<em>Use that license</em>"]
  metadata --"Yes"--> cc0_content_metadata["CC0 1.0"]
  metadata --"No"--> choose_license["<em>Consult flowchart for<br>software, writing,<br>image, audio, and video</em>"]
  end

  subgraph "License for combination of data (database)"
  choose_license --> switch_license["<em>Depending on content license</em>"]
  use_existing_license_content --> switch_license

  cc0_content_metadata --> cc0_db["CC0 1.0"]
  switch_license --"CC0 or<br>non-CC license"--> sa(Copyleft?)
  switch_license --"CC BY or<br>CC BY-SA"--> same["<em>Same license for DB</em>"]

  sa --"No (permissive)"--> cc0_db
  sa --"Yes"--> odbl["ODbL 1.0"]

  %% the following link is only added to have terminal nodes on the same level
  sa ~~~ same
  end

  click cc0_content_metadata href "https://creativecommons.org/publicdomain/zero/1.0/"
  click cc0_db href "https://creativecommons.org/publicdomain/zero/1.0/"
  click odbl href "https://opendatacommons.org/licenses/odbl/summary/"
```

Advanced License Flowchart for Data(base)
:::

### 2024-09-04 index.qmd
Add content pages

@@
 
-This is a Quarto website.
+## Content
 
-To learn more about Quarto websites visit <https://quarto.org/docs/websites>.
+This tutorial covers how to make your code available to others.
+While the general principles taught in this tutorial apply to most researchers publishing code, we will make very specific suggestions what tools to use.
+
+We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:
+
+- removing legal barriers: adding a appropriate license
+- high-level documentation: adding a README file
+- archival: uploading to a repository
+
+## Prerequisites
+
+We will assume you have installed [Git](https://git-scm.com/), [R](https://www.r-project.org/), and [RStudio](https://posit.co/download/rstudio-desktop/).
+
+If you are really new to R, we recommend looking at the tutorial "[Introduction to R](https://lmu-osc.github.io/introduction-to-R/)".
+If you have not used Quarto before, we have also prepared a "[Quarto Workshop](https://lmu-osc.github.io/introduction-to-Quarto/)".
+
+[Get started](intro.qmd){.btn-action-primary .btn-action .btn .btn-success .btn-lg}

### 2024-09-04 intro.qmd
Add content pages

---
title: "Introduction"
lang: en
bibliography: literature.bib
---

Together, we will create a research compendium whilst taking the following ten things into account [@Arguillas2022]:

> __Does the research compendium contain everything needed to reproduce a predefined outcome in an organized and parsimonious way?__
> 
> 1. __Completeness:__ The research compendium contains all of the objects needed to reproduce a predefined outcome.
> 2. __Organization:__ It is easy to understand and keep track of the various objects in the research compendium and their relationship over time.
> 3. __Economy:__ Fewer extraneous objects in the compendium mean fewer things that can break and require less maintenance over time.
> 
> __Is descriptive information about the research compendium and its components available and easy to understand?__
> 
> 4. __Transparency:__ The research compendium provides full disclosure of the research process that produced the scientific claim.
> 5. __Documentation:__ Information describing compendium objects is provided in enough detail to enable independent understanding and use of the compendium.
> 
> __Is information about how the research compendium and its components can be used available and easy to understand?__
> 
> 6. __Access:__ It is clear who can use what, how, and under what conditions, with open access preferred.
> 7. __Provenance:__ The origin of the components of the research compendium and how each has changed over time is evident.
> 
> __Is information about the research compendium and its components embedded in code?__
> 
> 8. __Metadata:__ Information about the research compendium and its components is embedded in a standardized, machine-readable code.
> 9. __Automation:__ As much as possible, the computational workflow is script- or workflow-based so that the workflow can be re-executed using minimal actions.
> 
> __Is there a plan for reviewing the research compendium for FAIR and computational reproducibility standards over time?__
> 
> 10. __Review:__ A series of managed activities are needed to ensure continued access to and functionality of the research compendium and its components for as long as necessary.

### 2024-09-04 setup.qmd
Add content pages

---
title: "Project Setup"
lang: en
bibliography: literature.bib
engine: knitr
---

We will start by setting up a simple example of a reproducible report.

## Create Quarto Project

First, we will need to create a new Quarto project.

If you haven't already, open RStudio. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.

![](images/setup_01.png){width=500px fig-align="center"}

Here, select _New Directory_

![](images/setup_02.png){width=600px fig-align="center"}

And choose the project type _Quarto Project_.

![](images/setup_03.png){width=600px fig-align="center"}

Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.

As we will use Git to track the version history of files, be sure to check _Create a git repository_. If you don't know what Git is, have a look at the tutorial "[Introduction to version Control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".

::: {.column-margin}
![renv: A dependency management toolkit for R](images/renv.svg){width=250px}
:::

Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on. Therefore make sure that the box _Use renv with this project_ is checked. Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".

If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.

![](images/setup_04.png){width=600px fig-align="center"}

Click on _Create Project_. Your RStudio window should now look similar to this:

![](images/setup_05.png){.lightbox fig-align="center" alt="The project `code-publishing-exercise` opened in RStudio. The source pane to the top left has a Quarto file open called \"code-publishing-exercise.qmd\". The console pane to the bottom left indicates by its output that renv is active in the current project. The environment pane to the top right indicates that the environment is currently empty. The output pane to the bottom right shows the files in the current project."}

If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.

Make sure that your project is in a consistent state according to `renv` by running:

```{.r filename="Console"}
renv::status()
```

If it reports packages that are not used, synchronize the lock file using:

```{.r filename="Console"}
renv::snapshot()
```

You are now ready to stage and commit your files. You can either stage files separately or the whole project folder at once. If you do the latter, we recommend you to inspect the untracked changes before staging all of them:

::: {.column-margin}
In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
:::

```{.bash filename="Terminal"}
git status
```

Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file. If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:

```{.bash filename="Terminal"}
git add .
git commit -m "Initial commit"
```

If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file. If you don't check your changes before committing, you might accidentally commit something you'd rather not.

::: {#tip-author-identity-unknown .callout-tip}
If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:

```{.bash filename="Terminal"}
git config user.name "YOUR NAME"
git config user.email "YOUR EMAIL ADDRESS"
```

Then, commit again.
:::

## Decide on Structure

Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it. In general, the directory structure should facilitate understanding a project by breaking it into logical chunks. There is no single best solution, as a good structure depends on where a project's complexity lies. However, it is usually helpful if the different files and folders reflect the execution order. For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders. Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.

Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021]. We recommend you to follow the standards of your field.

For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.

## Add Data

You can now download the data set we have prepared for you and put into your project folder: [`Data.csv`](_01_Example/Data.csv){download=""}

::: {.column-margin}
![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
:::

The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica. It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

::: {#wrn-legal-restrictions .callout-warning}
### Consider Legal Restrictions Before Sharing

Everything you put into the project folder will be shared publicly. For reasons of reproducibility, this should include the data you analyze. Of course, you should only share them to the extent you are allowed to, taking into account:

1. applicable privacy laws (e.g., the GDPR for European citizens),
2. contractual obligations (e.g., with your data provider),
3. copyright of the data and their particular structure, and
4. any sui generis database right.

Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic dataset^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from. For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).

Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are. Additionally, in some jurisdictions data can be subject to _sui generis database rights_ which prevent extracting substantial parts of a database. As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
:::

When publishing a data set, it is important to document the meaning (e.g., units) and possible values of its variables. This is typically done with a _data dictionary_ (also called a _codebook_). In the following, we will demonstrate how to create a simple data dictionary using the R packages [`tinylabels`](https://cran.r-project.org/package=tinylabels), [`datawizard`](https://easystats.github.io/datawizard/), and [`tinytable`](https://vincentarelbundock.github.io/tinytable/). You can install them now using:

```{.r filename="Console"}
renv::install(c(
  "tinylabels",
  "datawizard",
  "tinytable"
))
```

You can put the code that follows for creating the data dictionary into a new file called `create_data_dictionary.R`.

Using `tinylabels` we can add labels to the variables of a `data.frame` in R:^[Note that the code provided does not alter the data file -- no labels will be added to `Data.csv`. The labels are only added to a (temporary) copy of the data set within R in order to create the data dictionary.]

```{r, echo = -1, eval = -2, filename="create_data_dictionary.R"}
dat <- read.csv(file.path("_01_Example", "Data.csv"))
dat <- read.csv("Data.csv")

descriptions <- c(
  species = "a character string denoting penguin species",
  island = "a character string denoting island in Palmer Archipelago, Antarctica",
  bill_length_mm = "a number denoting bill length (millimeters)",
  bill_depth_mm = "a number denoting bill depth (millimeters)",
  flipper_length_mm = "an integer denoting flipper length (millimeters)",
  body_mass_g = "an integer denoting body mass (grams)",
  sex = "a character string denoting penguin sex",
  year = "an integer denoting the study year"
)
tinylabels::variable_label(dat) <- descriptions
```

Subsequently, `datawizard` can be employed to create the data dictionary containing the name and label, but also some other information about each variable:

::: {.column-margin}
![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
:::

```{r, filename="create_data_dictionary.R"}
(dict <- datawizard::data_codebook(dat) |>
  subset(select = -.row_id) |>
  tinytable::tt())
```

Finally, we can store the data dictionary inside an HTML file using the R package `tinytable` and put the HTML file into the project folder as well.

::: {.column-margin}
![tinytable: Simple and Configurable Tables](images/tinytable.svg){width=250px}
:::

```{.r filename="create_data_dictionary.R"}
tinytable::save_tt(dict, output = "data_dictionary.html")
```

A human-readable data dictionary is necessary for making one's research reproducible and the example we provided demonstrates only the bare minimum. A full data documentation including measurement instruments, sampling procedures, appropriate weighting, contact information, and more information about the study can be created with the R package [`pointblank`](https://rstudio.github.io/pointblank/articles/INFO-1.html). And one could go even further by making the information machine-readable in a standardized way. We provide an optional example of that in @nte-frictionless. If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Data management](https://github.com/MPDL/FAIR-Data-Management)".

::: {#nte-frictionless .callout-note collapse="true"}
### Create Machine-Readable Variable Documentation

This example demonstrates how the title and description of the data set, the description of the variables and their possible values are stored in a machine-readable way.

```r
table_info <- c(
  title = "penguins dataset",
  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
)
descriptions <- c(
  species = "a character string denoting penguin species",
  island = "a character string denoting island in Palmer Archipelago, Antarctica",
  bill_length_mm = "a number denoting bill length (millimeters)",
  bill_depth_mm = "a number denoting bill depth (millimeters)",
  flipper_length_mm = "an integer denoting flipper length (millimeters)",
  body_mass_g = "an integer denoting body mass (grams)",
  sex = "a character string denoting penguin sex",
  year = "an integer denoting the study year"
)
vals <- list(
  species = c("Adelie", "Gentoo", "Chinstrap"),
  island = c("Torgersen", "Biscoe", "Dream"),
  sex = c("male", "female"),
  year = c(2007, 2008, 2009)
)
```

Generally, metadata are either stored embedded into the data or externally, for example, in a separate file. We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately.

Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data. For the purpose of the following code, it is just a nested list that we edit to include our own information. We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/). Finally, the package is used to create a metadata file that contains the schema.

```r
# Read data and create schema
dat_filename <- "Data.csv"
dat <- read.csv(dat_filename)
dat_schema <- frictionless::create_schema(dat)

# Add descriptions to the fields
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  c(x, description = descriptions[[x$name]])
})

# Record possible values
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  if (x$name %in% names(vals)) {
    modifyList(x, list(constraints = list(enum = vals[[x$name]])))
  } else {
    x
  }
})

# Define missing values
dat_schema$missingValues <- c("", "NA")

# Create package with license info and write it
dat_package <- frictionless::create_package() |>
  frictionless::add_resource(
    resource_name = "penguins",
    data = dat_filename,
    schema = dat_schema,
    title = table_info[["title"]],
    description = table_info[["description"]],
    licenses = list(list(
      name = "CC0-1.0",
      path = "https://creativecommons.org/publicdomain/zero/1.0/",
      title = "CC0 1.0 Universal"
    ))
  )
frictionless::write_package(dat_package, directory = ".")
```

This creates the metadata file `datapackage.json` in the current directory. Make sure it is located in the same folder as `Data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
:::

Having added the data and its documentation, one can view and record the utilized packages with `renv`...

```{.r filename="Console"}
renv::status()
renv::snapshot()
```

...and go through the commit routine:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add data"
```

## Add Code

In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file. The manuscript contains code together with a written narrative. Download the two files to your computer and put them into your project folder.

- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}

The manuscript explores differences in bill length between male and female penguins, feel free to read through it.

::: {#wrn-take-copyright-seriously .callout-warning}
### Take Copyright Seriously

When you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so. Only build on existing work for which you are given an express grant of relevant rights. How do you know you are allowed to copy, edit, and share the two files linked above?
:::

As the manuscript uses some new packages, install them with:

```{.r filename="Console"}
renv::install()
```

The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which needs to be installed in the project before it can be rendered:

```{.bash filename="Terminal"}
quarto add --no-prompt wjschne/apaquarto
```

Also, you need to have a $\TeX$ distribution installed on your computer. A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:

```{.bash filename="Terminal"}
quarto install tinytex
```

You should now be able to render the document using Quarto:

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```

This should create a PDF file called `Manuscript.pdf` in your project folder.

::: {#tip-update-quarto .callout-tip}
If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however `apaquarto` sometimes requires more recent versions.
:::

With the code being added, one can use `renv` again to view and record the new packages:

```{.r filename="Console"}
renv::status()
renv::snapshot()
```

::: {#tip-renv-status .callout-tip}
Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
:::

Finally, make your changes known to Git:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add manuscript"
```

::: {#wrn-credentials .callout-warning}
### Beware of Credentials

Sometimes, a data analysis requires the interaction with online services:

- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
- an analysis may be conducted with the help of AI providers.

In these cases, make sure that the code you check in to Git does not contain any credentials that are required for accessing these services. Instead, make use of environment variables which are defined in a location that is excluded from version control. When programming with R, you can define them in a file called `.Renviron` in the root of your project folder:

```{.ini filename=".Renviron"}
MY_FIRST_KEY="your_api_key_here"
MY_SECOND_KEY="your_api_key_here"
```

When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:

```r
query_api(..., api_key = Sys.getenv("MY_FIRST_KEY"))
```

Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that you found contains credentials, you can follow @Chacon2024.
:::

### Coding Best Practices

Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.

- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root. For example, don't write `C:\Users\Public\Documents\my_project\images\result.png`, instead write `images\result.png`.^[In R, use `file.path()` for a platform-independent way to concatenate folders to build a path.]

- __Keep it simple.__ Add complexity only when you must. Whenever there's a boring way to do something and a clever way, go for the boring way. If the code grows increasingly complex, refactor it into separate functions and files.

- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.

- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.

- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).

- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future. Prefer packages that are well-maintained and light on dependencies^[You can use the [R package dependency badges](https://tinyverse.netlify.app/) to count the direct and total dependencies.]. We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.

- __Fail early, often and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.

- __Test your code.__ Test your code with scenarios where you know what the result should be. Turn bugs you discovered into test cases. Use linting tools, for example, the R package [`lintr`](https://lintr.r-lib.org/).

- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality. R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/). Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013. And even if you don't follow a style guide, be consistent.

This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:

- "Tidy design principles" [@Wickham2023Design]
- "The Good Research Code Handbook" [@Mineault2021]
- "The Art of UNIX Programming" [@Raymond2003]

> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
> 
> --- Martin Fowler, British software engineer

## The Last Mile

`renv` only records the versions of R packages and of R itself. This means that potential system dependencies of R packages and other tools utilized in the project are not documented anywhere, including Quarto.^[As of August 2024, a proposal to record the version of Quarto and Pandoc has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] We will manually write them down when [creating a README](add_readme.qmd). For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies). Do run the following:

```{.bash filename="Terminal"}
quarto use binder
```

This will create a few additional files which facilitate reconstructing the computational environment in the future.^[Either using [repo2docker](https://repo2docker.readthedocs.io/) or the public [binder](https://mybinder.org/) service.] As always, commit your changes:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add repo2docker config"
```

You are now all set up to prepare your project for sharing!

### 2024-09-07 about.qmd
Add authors, license and funding note

@@
 title: "About"
+author:
+  - name: "Florian Kohrt"
+    orcid: "0000-0003-0374-5625"
+  - name: "Josephine Zerna"
+    email: "josephine.zerna@tu-dresden.de"
+    orcid: "0000-0003-2892-884X"
+  - name: "Christoph Scheffel"
+    email: "christoph_scheffel@tu-dresden.de"
+    orcid: "0000-0001-5963-9229"
 bibliography: grateful-refs.bib
 ---
+
+# Utilized Software
 
@@
 ```{r}
-#| echo: false
+#| class.output: "txt code-overflow-scroll"
 
@@
 ```
+
+# License and Disclaimer
+
+If not indicated otherwise, this tutorial is licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/). None of the discussion in this tutorial constitutes legal advice.
+
+Any screenshots of [RStudio](https://github.com/rstudio/rstudio) are released under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/gpl-3.0.html).
+
+```{.txt .code-overflow-wrap}
+Copyright (C) 2024 Posit Software, PBC
+
+This program is free software: you can redistribute it and/or modify it under the terms of version 3 of the GNU Affero General Public License as published by the Free Software Foundation.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU Affero General Public License for more details.
+
+You should have received a copy of the GNU Affero General Public License along with this program.  If not, see <https://www.gnu.org/licenses/>.
+```
+
+# Funding
+
+::: {.column-margin}
+![](images/bitss.png){width=250px}
+:::
+
+This work was partly funded by the Berkeley Initiative for Transparency in the Social Sciences (BITSS), managed by the Center for Effective Global Action (CEGA).

### 2024-09-07 index.qmd
Fix typo and add RTools and Quarto

@@
 
-- removing legal barriers: adding a appropriate license
+- removing legal barriers: adding an appropriate license
 - high-level documentation: adding a README file
-- archival: uploading to a repository
+- archival: uploading it to a repository
 
@@
 
-We will assume you have installed [Git](https://git-scm.com/), [R](https://www.r-project.org/), and [RStudio](https://posit.co/download/rstudio-desktop/).
+We will assume you have installed [Git](https://git-scm.com/download/), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/).
+Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).
 

### 2024-09-07 add_readme.qmd
Fix typo and format raw output nicely

@@
 
-    An overview over the system dependencies of one more multiple R packages can be created using the function `pkg_sysreqs()` from the R package `pak`.
+    An overview over the system dependencies of R packages can be created using the function `pak::pkg_sysreqs()`.
 In combination with `renv`, we can obtain the system dependencies of all R packages in the current project:
@@
 
-    ```
+    ```txt
     ── Install scripts ─────────────────────── 
@@
 
-    ```
+    ```txt
     [✓] Checking tools....................OK

### 2024-09-07 archive.qmd
Add appropriate attribution and license info

@@
 
-The project is now ready to be deposited for long-term archival. When considering an appropriate repository, there are a few questions to ponder [@Jackson2018]:
+The project is now ready to be deposited for long-term archival.
+When considering an appropriate repository, there are a few questions to ponder [excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 

### 2024-09-07 intro.qmd
Add appropriate attribution and license info

@@
 
-Together, we will create a research compendium whilst taking the following ten things into account [@Arguillas2022]:
+Together, we will create a research compendium whilst taking the following ten things into account [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 

### 2024-09-07 choose_license.qmd
Rework section on licenses
Fix factual errors
Provide specific advise for R users
Make flowcharts more explicit

@@
 Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes.
-By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that modifications are only shared under the same license, and that no further restrictions are imposed on the work.
+By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work.
 Software licenses may additionally require to make the source code available to everybody the software is shared with.
-Because there are many free/open licenses available, the following decision tree only contains a recommended subset.
+Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.
 
@@
 We have created a flowchart that covers the most likely types of works you will create as a researcher: software, writing (i.e., text), images, audio, video, and data (see @fig-flowchart-simple).
-The flowchart always recommends the most permissive license possible to maximize reuse.
+This flowchart always recommends the most permissive license possible to maximize reuse -- below we provide two additional flowcharts that allow for more choices.
 Click on the name of a license to learn more about it.
@@
 
-There are specific licenses available (and recommended) for fonts and hardware:
+One should be cautious about the restrictions of licenses applied to the following types of works:
 
-- fonts: [SIL Open Font License 1.1](https://openfontlicense.org/)
-- hardware: [CERN Open Hardware Licence](https://cern-ohl.web.cern.ch/)
-
-Furthermore, one should be cautious about the restrictions of licenses applied to the following types of works:
-
+- __fonts:__ Copyleft licenses applied to fonts can a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
+If the intent is that only derived fonts, if published, have to be put under the same license, the [SIL Open Font License\ 1.1](https://openfontlicense.org/) is an appropriate choice.
+Note, however, that it doesn't require attribution for usage.
+If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
 - __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/gpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
@@
 If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
-
-Finally, works in the public domain should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
+Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
+- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
 :::
@@
 You may prefer that everybody who receives somebody else's adaptations of your work enjoys the same rights that you granted to them.
-This can be achieved by _copyleft_ licenses.^[So called because copyleft licenses make content permanently free, thus turning copyright around: © → 🄯] For software, they come in two flavours:
+This can be achieved by _copyleft_ licenses.^[So called because copyleft licenses make content permanently free, thus turning copyright around: © → 🄯.
+In the realm of Creative Commons licenses, copyleft is also called _share alike_.] For software, they come in two flavors:
 
 - __Weak copyleft licenses__ only require that modifications to the software itself are licensed under the same or a compatible license if shared.
-For example, if I create and publish an R package under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license.
-However, people who merely _use_ my R package in their scripts can choose any license for their scripts.
+For example, if you create and publish software under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license.
+However, people who merely _use_ your software in their own work which they make publicly available can choose any license for it.
 - __Strong copyleft licenses,__ on the other side, insist that any larger works that use the copyleft-licensed software must also be licensed under the same or a compatible license if shared.
-For example, if my R package were to be put under a strong copyleft license, everybody publishing a data analysis script that uses my package would need to put it under the same license.
-Similarly, a Quarto manuscript using such a package would need to disclose its source code if shared.
+For example, if your software were to be put under a strong copyleft license, everybody publishing software that uses your software would need to put it under the same license.
+Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024].
 
@@
 
+::: {#tip-license-r .callout-tip}
+#### Projects Involving R Code
+
+In most cases, the output of software, like images or tables, does not depend on its license.
+Therefore, if you use an R package under a copyleft license to create content, you are likely the copyright owner.
+However, if the output is based in data, it can be considered a derivative work and the license of the data also applies.
+
+If a Quarto document contains both R code and writing, one may wonder what type of license to apply.
+It is possible to make such a file simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
+This is called dual-licensing and makes it easier to reuse both rendered PDF as well as the code.
+For example, one could write:
+
+> The Quarto files in this project are made available under CC\ BY-SA\ 4.0, or (at your option) under the terms of the Apache Software License\ 2.0.
+
+It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license.
+Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense].
+You can learn which license an installed package uses via `packageDescription("<PACKAGE_NAME>", fields = "License")`.
+And to identify which licenses are being used by the R packages your project depends on, you can use the following code:
+
+```r
+direct_deps <- unique(renv::dependencies()$Package)
+all_deps <- tools::package_dependencies(
+  packages = direct_deps,
+  which = "strong",
+  recursive = TRUE
+) |>
+  unlist() |>
+  append(deps) |>
+  unique()
+unique(installed.packages(fields = "License")[pkgs, "License"])
+```
+:::
+
 We have prepared two advanced license flowcharts, one for software, writing, images, audio, and video in @fig-flowchart-non-data and one for data in @fig-flowchart-data where you can make additional choices.
@@
 
-::: {#cau-license-versions .callout-caution}
-### License Versions Are Important
+### Attribution, State Changes, and Anti-DRM?
 
-You may have noticed that we mostly refer to licenses using a name _and_ a version number.
-This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses.
-For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and sui generis database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
-There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
+The advanced license flowcharts also allow you to make additional decisions:
 
-For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022].
-:::
+- __Attribution__ means whether recipients of your work are required to provide attribution to you.
+For software licenses, this is called a copyright notice.
+Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
+- __State changes__ means that recipients need to indicate if changes were made.
+- __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
+DRM is an abbreviation for **d**igital **r**ights **m**anagement.
 
-### Attribution and Anti-DRM?
-
-The advanced license flowcharts also allow you to make two additional decisions for Creative Commons licenses (i.e., when deciding between CC0\ 1.0, CC\ BY\ 4.0, and CC\ BY-SA\ 4.0):
-
-- __Attribution__ means whether recipients of your work are required to provide attribution to you and indicate if changes were made.
-Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
-- __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
-
+From the Creative Commons licenses, only [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) does not require providing attribution.
 All software licenses in @fig-flowchart-non-data require providing attribution, although only [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) and [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/) require others to indicate changes.
-Of the software licenses listed, only the AGPLv3 also has an anti-DRM provision.
 For data, the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) allows for technological measures that restrict the work only if a version of the database is provided in parallel without such measures.
@@
 
+::: {#cau-license-versions .callout-caution}
+### License Versions Are Important
+
+You may have noticed that we mostly refer to licenses using a name _and_ a version number.
+This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses.
+For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
+Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and sui generis database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
+There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
+
+For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022].
+:::
+
 For example, the `apaquarto` extension that you included in your project contains works by others.
@@
 
-Creative Commons even provides a range of considerations for licensors and licensees [@CC2013] and an [interactive chooser](https://creativecommons.org/choose/) which you can use to create snippet that you can copy and paste to the desired location.
+Creative Commons even provides a range of considerations for licensors and licensees [@CC2013] and an [interactive chooser](https://creativecommons.org/choose/) which you can use to create text snippet that you can copy and paste to the desired location.
 :::
@@
 
-  type --"Software"--> code_sa(Copyleft?)
-  type --"Writing, image, audio, video"--> nocode_cc("Anti-DRM?<br>Attribution?<br>Copyleft?")
+  type --"Software"--> code_sa("Attribution?<br>State changes?<br>Copyleft?")
+  type --"Writing, image, audio, video"--> nocode_cc("Anti-DRM?<br>Attribution?<br>State changes?<br>Copyleft?")
 
-  code_sa --"No (permissive)"--> apache["Apache 2.0"]
-  code_sa --"Yes (weak)"--> mpl["MPL 2.0"]
-  code_sa --"Yes (strong)"--> agpl["AGPLv3"]
+  code_sa --"Attribution &<br>State changes"--> apache["Apache 2.0"]
+  code_sa --"Attribution &<br>Weak copyleft"--> mpl["MPL 2.0"]
+  code_sa --"Attribution &<br>State changes &<br>Strong copyleft"--> agpl["AGPLv3"]
 
   nocode_cc --"Neither"--> cc0["CC0 1.0"]
-  nocode_cc --"Anti-DRM &<br>Attribution"--> cc_by["CC BY 4.0"]
-  nocode_cc --"Anti-DRM &<br>Attribution &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
+  nocode_cc --"Anti-DRM &<br>Attribution &<br>State changes"--> cc_by["CC BY 4.0"]
+  nocode_cc --"Anti-DRM &<br>Attribution &<br>State changes &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
 
@@
 ```
+
+DRM = digital rights management
 

### 2024-09-07 setup.qmd
Use OSC link for FAIR data workshop and remove note about file.path()
Windows users are already encouraged to use
forward slashes rather than backward slashes,
so using file.path() probably doesn't add
too much value.

@@
 We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Data management](https://github.com/MPDL/FAIR-Data-Management)".
+If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
 
@@
 - __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root.
-For example, don't write `C:\Users\Public\Documents\my_project\images\result.png`, instead write `images\result.png`.^[In R, use `file.path()` for a platform-independent way to concatenate folders to build a path.]
+For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
 

### 2024-09-07 choose_license.qmd
Make data flowchart more explicit

@@
   type --"Software"--> code_sa("Attribution?<br>State changes?<br>Copyleft?")
-  type --"Writing, image, audio, video"--> nocode_cc("Anti-DRM?<br>Attribution?<br>State changes?<br>Copyleft?")
+  type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>State changes?<br>Anti-DRM?<br>Copyleft?")
 
@@
   nocode_cc --"Neither"--> cc0["CC0 1.0"]
-  nocode_cc --"Anti-DRM &<br>Attribution &<br>State changes"--> cc_by["CC BY 4.0"]
-  nocode_cc --"Anti-DRM &<br>Attribution &<br>State changes &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
+  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM"--> cc_by["CC BY 4.0"]
+  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
 
@@
 
-DRM = digital rights management
+Note: DRM = digital rights management
 
@@
   cc0_content_metadata --> cc0_db["CC0 1.0"]
-  switch_license --"CC0 or<br>non-CC license"--> sa(Copyleft?)
+  switch_license --"CC0 or<br>non-CC license"--> sa("Attribution?<br>Anti-DRM?<br>Copyleft?")
   switch_license --"CC BY or<br>CC BY-SA"--> same["<em>Same license for DB</em>"]
 
-  sa --"No (permissive)"--> cc0_db
-  sa --"Yes"--> odbl["ODbL 1.0"]
+  sa --"Neither"--> cc0_db
+  sa --"Attribution &<br>Anti-DRM &<br>Copyleft"--> odbl["ODbL 1.0"]
 

### 2024-09-08 choose_license.qmd
Fix typos

@@
 
-First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
+First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
 The same applies if there are strong community norms to use a particular free/open license.^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.

### 2024-09-08 make_readme.qmd
Rename "Add a README" to "Make a README"
Because it fits with the corresponding website:
https://www.makeareadme.com/
Fix typos

--- 2024-09-07 add_readme.qmd
@@
 ---
-title: "Add a README"
+title: "Make a README"
 lang: en
@@
 Once you have prepared your project and settled on a license, it is time to add a final touch.
-Imagine you are coming back to your project in five years, having forgotten most of the details of what you did exactly.
+Imagine returning to your project in five years, having forgotten most of the details of what you did exactly.
 What would be useful to know in order to quickly understand what is going on in the project?
@@
 
-:   If there is anything you can show that demonstrates how the project can be used? Screenshots or other visuals can make the README more appealing.
+:   Is there anything you can show that demonstrates how the project can be used? Screenshots or other visuals can make the README more appealing.
 
@@
 :   Is there a recommended way to cite this project?
-We recommend to increase citability by using the [CodeMeta generator](https://codemeta.github.io/codemeta-generator/) to create the file `codemeta.json`.
-Alternatively, one can also use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create the file.
+We recommend to increase citability by using the [CodeMeta generator](https://codemeta.github.io/codemeta-generator/) to create a file called `codemeta.json`.
+Alternatively, one can also use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
+Then, put it into the project folder.
 

### 2024-09-08 setup.qmd
Fix typos

@@
 As we will use Git to track the version history of files, be sure to check _Create a git repository_.
-If you don't know what Git is, have a look at the tutorial "[Introduction to version Control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".
+If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".
 
@@
 
-Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository.
-If you already committed a file that you found contains credentials, you can follow @Chacon2024.
+Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that contains credentials, you can follow @Chacon2024.
 :::

### 2024-09-09 choose_license.qmd
Fix code

@@
   unlist() |>
-  append(deps) |>
+  append(direct_deps) |>
   unique()
-unique(installed.packages(fields = "License")[pkgs, "License"])
+unique(installed.packages(fields = "License")[all_deps, "License"])
 ```

### 2024-09-09 choose_license.qmd
Use `pak` to calculate the dependencies

--- 2024-09-08 choose_license.qmd
@@
 ```r
-direct_deps <- unique(renv::dependencies()$Package)
-all_deps <- tools::package_dependencies(
-  packages = direct_deps,
-  which = "strong",
-  recursive = TRUE
-) |>
-  unlist() |>
-  append(deps) |>
-  unique()
-unique(installed.packages(fields = "License")[pkgs, "License"])
+deps <- renv::dependencies()$Package |>
+  unique() |>
+  pak::pkg_deps(dependencies = NA) |>
+  getElement("package")
+unique(installed.packages(fields="License")[deps, "License"])
 ```

### 2024-09-09 setup.qmd
Use `pak` to calculate the dependencies

@@
 - __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
-Prefer packages that are well-maintained and light on dependencies^[You can use the [R package dependency badges](https://tinyverse.netlify.app/) to count the direct and total dependencies.].
+Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
 We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.

### 2024-09-12 choose_license.qmd
Minor stylistic improvements

@@
 ::: {#cau-exclusive-rights .callout-caution}
-### Publishing Without Exclusive Rights
+### Giving Publishers Exclusive Rights
 
@@
 However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice.
-If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide by @Elson2016 to learn how to do that.
+If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
 
@@
 
-For example, the `apaquarto` extension that you included in your project contains works by others.
-You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with its terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice.
+For example, the `apaquarto` extension that you included in your project is a work by others.^[Unless you are the author of the `apaquarto` extension.] You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with any terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice.
 In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.
@@
 - [CC licenses](https://creativecommons.org/faq/#how-do-i-apply-a-creative-commons-license-to-my-material)
-- [ODbL](https://opendatacommons.org/licenses/odbl/)
+- [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/)
 ::::
@@
 In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim.
-This is a practice propagated by GitHub, who provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
+This is a practice propagated by GitHub, which provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
 However, if the project is not completely covered by one single license, this practice may become unwieldy.
@@
 We will be using that to indicate the license for every file in your project, along with the year of publication and the copyright holder.
-To do this, we add a comment to the beginning of every file and add the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`.
+To do this, we add a comment to the beginning of every file and include the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`.
 How this works depends on the file type, as the syntax for a comment varies.
 
-For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name:
+For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name -- of course, you can also choose a different license:
 
@@
 Note that `Manuscript.qmd` was provided to you under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), which is what you indicate with the previous comment.
-If you edited the file, you may also add yourself.
+If you edited the file, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them, although you do not need to provide your real name.]
 

### 2024-09-12 index.qmd
Rename "Add a README" to "Make a README"
Because it fits with the corresponding website:
https://www.makeareadme.com/
Mention other tutorials in one place

--- 2024-08-26 index.qmd
@@
 ---
-title: "Code publishing tutorial"
+title: "Code Publishing Tutorial"
 ---
 
-This is a Quarto website.
+This tutorial covers how to make your code available to others. It wraps up a series of tutorials which we recommend to go through in advance:
 
-To learn more about Quarto websites visit <https://quarto.org/docs/websites>.
+- [Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/): This tutorial explains how to set up a simple project with version control.
+- [Collaborative coding with GitHub and RStudio](https://lmu-osc.github.io/Collaborative-RStudio-GitHub/): Here you can learn how to collaborate with others on code projects using GitHub.
+- [Introduction to R](https://lmu-osc.github.io/introduction-to-R/): This is an introduction to the R programming language.
+- [Introduction to `renv`](https://lmu-osc.github.io/introduction-to-renv/): Here you can learn how to manage the dependencies of an R project.
+- [Introduction to Quarto](https://lmu-osc.github.io/introduction-to-Quarto/): Learn how to combine prose and code to make documents more reproducible.
+- [Introduction to Zotero](https://lmu-osc.github.io/introduction-to-zotero/): Get introduced to the reference management software Zotero.
+- [FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/): Learn how to make your data findable, accessible, interoperable, and reusable.
+
+While the general principles taught in this tutorial apply to most researchers publishing code, we will make very specific suggestions what tools to use.
+
+In particular, we will assume you have installed [Git](https://git-scm.com/download/), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/).
+Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).
+
+We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:
+
+- removing legal barriers: adding an appropriate license
+- high-level documentation: creating a README file
+- archival: uploading it to a repository
+
+[Get started](intro.qmd){.btn-action-primary .btn-action .btn .btn-success .btn-lg}

### 2024-09-12 setup.qmd
Implement Malika's feedback on the setup

@@
 
-If you haven't already, open RStudio. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
+If you haven't already, open RStudio -- see @nte-terminal for an alternative to RStudio. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
 
@@
 Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on.
+Using it makes it easier for others to view and obtain them at a later point in time.
 Therefore make sure that the box _Use renv with this project_ is checked.
@@
 ```
+
+::: {#nte-terminal .callout-note collapse="true"}
+### Without RStudio
+
+Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:
+
+```{.bash filename="Terminal"}
+quarto create project default code-publishing-exercise
+cd code-publishing-exercise/
+rm code-publishing-exercise.qmd
+git init
+git checkout -b main
+```
+
+Then, one can open an R session, by simply typing `R` into the terminal.
+Next, sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
+If not, set it using `setwd("code-publishing-exercise")`.
+Then, initialize `renv`:
+
+```{.r filename="Console"}
+renv::init()
+```
+:::
 
@@
 We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
+If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Research Data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
 
@@
 
-The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which needs to be installed in the project before it can be rendered:
+The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the @APA2020.
+It can be installed in the project using the following command:
 
@@
 
-Also, you need to have a $\TeX$ distribution installed on your computer. A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
+::: {#tip-apaquarto .callout-tip}
+#### Not a psychologist?
+
+If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.
+
+Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:
+
+```{.yml filename="Manuscript.qmd"}
+format:
+  pdf:
+    pdf-engine: lualatex
+    documentclass: scrartcl
+    papersize: a4
+```
+:::
+
+Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents.
+A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
 
@@
 Turn bugs you discovered into test cases.
-Use linting tools, for example, the R package [`lintr`](https://lintr.r-lib.org/).
+Use linting tools^[A linting tool analyzes your code without actually running it.
+This process is called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
 
@@
 `renv` only records the versions of R packages and of R itself.
-This means that potential system dependencies of R packages and other tools utilized in the project are not documented anywhere, including Quarto.^[As of August 2024, a proposal to record the version of Quarto and Pandoc has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] We will manually write them down when [creating a README](add_readme.qmd).
+This means that potential system dependencies of R packages and other tools utilized in the project are not documented anywhere, including Quarto.^[As of August 2024, a proposal to record the version of Quarto has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] We will manually write them down when [creating a README](make_readme.qmd).
 For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).

### 2024-09-13 choose_license.qmd
Add a TL;DR to the chapter on licenses

@@
 ---
+
+## TL;DR
+
+Every expression of an idea, such as literary or artistic works, are automatically covered by copyright.
+This means that nobody else than the copyright holder is allowed to copy, modify, or share it.
+Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
+
+Researchers are only allowed to share material they have the necessary rights to, for example, through a free/open license.
+These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.
+In addition, researchers should share any work by their own under such a license.
+
+"Choosing a License" involves:
+
+1. Recording the copyright notices and licenses for files by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
+2.
+In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously (at the choice of the recipient) to all own files (for example, again by recording the license in the file `LICENSE.txt`).
 
@@
 
-## Which License to Choose?
+## Which License to Choose for a Work?
 
@@
 In most cases, the output of software, like images or tables, does not depend on its license.
-Therefore, if you use an R package under a copyleft license to create content, you are likely the copyright owner.
-However, if the output is based in data, it can be considered a derivative work and the license of the data also applies.
+Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
+However, if the output is based on data, it can be considered a derivative work and the license of the data also applies.
+For example, maps may be considered as a derivative of the geographic data they are based on.
 
@@
 
+## Which Licenses to Choose for a Project?
+
+So far, we only discussed how to choose a license for works _of one type_.
+But what if you want to share a project with all kinds of files?
+For example, the project from this tutorial (among other things) includes a data file, a manuscript file with intermingled code and writing, and an R file.
+And what if you also want to share files by others, as is the case with `apaquarto` which you may have installed in this project?
+
+The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
+The easiest approach is to make note of every foreign work included in your project and record its license.
+Then, dual-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.
+
 ## Applying the License
 
-Having selected the licenses of your choice -- you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect.
+Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect.
 Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed in "[Add a README](add_readme.qmd)".
+Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed "[later](make_readme.qmd)".
 

### 2024-09-13 archive.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

@@
 ---
-title: "Archive"
+title: "Archive Your Project"
 lang: "en"
@@
 ---
+
+## Where to Deposit Software
 
@@
 
-Additional guidance is provided by @Daquin2023 and @Garijo2022. We will cover the archival using the Software Heritage archive and Zenodo.
+Additional guidance is provided by @Daquin2023 and @Garijo2022.
+
+## Using Zenodo to Archive Your Project
+
+In the following we will demonstrate how to use Zenodo to archive one's project and obtain a DOI.
+
+::: {#cau-sandbox .callout-caution}
+### Zenodo Sandbox
+
+Note that during this tutorial, we will be using Zenodo's sandbox instance located at [`sandbox.zenodo.org`](https://sandbox.zenodo.org/) in order to learn how to use it while not wasting Zenodo's resources.
+For real projects, use the production instance at [`zenodo.org`](https://zenodo.org/), which requires a separate account.
+:::
+
+First, create an account on the Zenodo Sandbox by clicking on _Sign up_ in the upper right corner and submitting the provided form.
+Then, confirm your email address (also check your spam folder if you cannot find the confirmation email).
+
+Now you can follow the guide "[Create new upload](https://help.zenodo.org/docs/deposit/create-new-upload/)", taking into account the following notes specific to each step:
+
+- __Add/remove files__: To circumvent the file limit on Zenodo, you can bundle your project folder in one ZIP file and upload that instead.
+- __Fill in metadata__
+  - __Digital Object Identifier__ (DOI): Usually, you will want a DOI, therefore you need to choose _No_ for the question whether you already have a DOI for this upload, and click on _Get a DOI now!_.
+  - __Resource type__: Select the most appropriate one, for example, _Publication / Preprint_.
+  - __Licenses__: If your project only contains content by you and if you previously only chose one license, you can select that by choosing _Add standard_.
+Otherwise, click on _Add custom_ and describe how each file is licensed.
+If you have been using REUSE to record license information, you can run enter `SPDX Document` as title and run the reuse tool with `reuse spdx -o reuse.spdx` to create the file `reuse.spdx`.
+Then you can paste the content of `reuse.spdx` into the description field of the custom license as it contains information about the license of each file.
+
+If you previously uploaded your project to a public Git repository (which we didn't cover during this tutorial), you could also add the new DOI to your README, for example, under the section "Citation".
+
+## Final Remarks
+
+There are situations in which you are not allowed to share your data publicly.
+In this case, consider using a repository specifically for research data, which later grants access for valid uses individually.
 

### 2024-09-13 make_readme.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

@@
 
-## Creating a README
+## What does a README contain?
 
@@
 
-Badges and project status
+Badges/Project Status
 
@@
 Mention all dependencies here that are not explicitly managed by `renv`, such as the system dependencies of R packages as well as the version of Quarto.
-
-    An overview over the system dependencies of R packages can be created using the function `pak::pkg_sysreqs()`.
-In combination with `renv`, we can obtain the system dependencies of all R packages in the current project:
-
-    ```{.r filename="Console"}
-    pak::pkg_sysreqs(renv::dependencies()$Package)
-    ```
-
-    The output may look like the following:
-
-    ```txt
-    ── Install scripts ─────────────────────── 
-    dnf install -y make pandoc
-
-    ── Packages and their system dependencies 
-    fs        – make
-    knitr     – pandoc
-    rmarkdown – pandoc
-    sass      – make
-    ```
-
-    We can see that `make` and `pandoc` were identified as system dependencies. One can obtain their version by running them with the `--version` argument:
-
-    ```{.bash filename="Terminal"}
-    make --version
-    pandoc --version
-    ```
-
-    We also know that we need Quarto to create the PDF and that we installed the `apaquarto` extension, so let's find out these versions as well:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]
-
-    ```{.bash filename="Terminal"}
-    quarto --version
-    quarto list extensions
-    ```
-
-    Finally, we know that we installed a $\TeX$ distribution to create the PDF, so let's find out its version by running:
-
-    ```{.bash filename="Terminal"}
-    quarto check
-    ```
-
-    The output is quite long and it might look slightly different for you, but the relevant sections are the following:
-
-    ```txt
-    [✓] Checking tools....................OK
-          TinyTeX: (not installed)
-          Chromium: (not installed)
-
-    [✓] Checking LaTeX....................OK
-          Using: Installation From Path
-          Path: /usr/local/texlive/2024/bin/x86_64-linux
-          Version: 2024
-    ```
-
-    We put all the information obtained about system dependencies and their versions in the README. Sharing the install scripts that were provided by `pak` is an additional courtesy to your readers.
+An R package's system dependency is any additional software that you need to install on your computer in order to use a particular R package.
+For example, the R package [`reticulate`](https://rstudio.github.io/reticulate/) allows to run Python code from within R.
+However, in order to actually use it, one has to additionally install Python itself as it does not come together with `reticulate` -- rather, it is a system dependency.
+See @sec-dependencies for additional information.
 
@@
 
-Create your README now as the file `README.md`.
+## Installation/Dependencies {#sec-dependencies}
+
+An overview over the system dependencies of R packages can be created using the function `pak::pkg_sysreqs()`.
+In combination with `renv`, we can obtain the system dependencies of all R packages in the current project:
+
+```{.r filename="Console"}
+pak::pkg_sysreqs(renv::dependencies()$Package)
+```
+
+The output may look like the following:
+
+```txt
+── Install scripts ─────────────────────── 
+dnf install -y make pandoc
+
+── Packages and their system dependencies 
+fs        – make
+knitr     – pandoc
+rmarkdown – pandoc
+sass      – make
+```
+
+We can see that `make` and `pandoc` were identified as system dependencies. One can obtain their version by running them with the `--version` argument:
+
+```{.bash filename="Terminal"}
+make --version
+pandoc --version
+```
+
+We also know that we need Quarto to create the PDF, so let's find out its version as well:
+
+```{.bash filename="Terminal"}
+quarto --version
+```
+
+If you installed `apaquarto` or any other Quarto extension, one can query their versions as follows:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]
+
+```{.bash filename="Terminal"}
+quarto list extensions
+```
+
+Finally, we know that we installed a $\TeX$ distribution to create the PDF, so let's find out its version by running:
+
+```{.bash filename="Terminal"}
+quarto check
+```
+
+The output is quite long and it might look slightly different for you, but the relevant sections are the following:
+
+```txt
+[✓] Checking tools....................OK
+      TinyTeX: v2024.09
+      Chromium: (not installed)
+
+[✓] Checking LaTeX....................OK
+      Using: TinyTex
+      Path: /home/r155953/.TinyTeX/bin/x86_64-linux
+      Version: 2024
+```
+
+## Create It!
+
+Create your README now as the file `README.md`. If you feel stuck, you can have a look at the following examples:
+
+::: {#tip-name .callout-tip collapse="true"}
+### Name
+
+```{.md .code-overflow-wrap filename="README.md"}
+# Penguin Paper
+
+```
+:::
+
+::: {#tip-project-status .callout-tip collapse="true"}
+### Badges/Project Status
+
+```{.md .code-overflow-wrap filename="README.md"}
+[![Project Status: Unsupported – The project has reached a stable, usable state but the author(s) have ceased all work on it.
+A new maintainer may be desired.](https://www.repostatus.org/badges/latest/unsupported.svg)](https://www.repostatus.org/#unsupported)
+```
+:::
+
+::: {#tip-description .callout-tip collapse="true"}
+### Description
+
+```{.md .code-overflow-wrap filename="README.md"}
+This project contains the Quarto manuscript of our study on penguins. It is written in R and uses `renv` to track its dependencies.
+```
+:::
+
+::: {#tip-installation .callout-tip collapse="true"}
+### Installation/Dependencies
+
+``````{.md .code-overflow-wrap filename="README.md"}
+## Dependencies
+
+This manuscript requires the following system software to be installed. In addition, we provide the version numbers this manuscript has last been run with:
+
+- [Quarto](https://quarto.org/docs/download/) 1.6.9
+- GNU Make 4.4.1
+- Pandoc 3.3
+- TinyTeX 2024.09
+- [R](https://cloud.r-project.org/) 4.4.1
+
+On Fedora Linux, Make and Pandoc can be installed as follows:
+
+```bash
+dnf install -y make pandoc
+```
+
+Quarto and R can be installed using the links provided. TinyTeX can be installed using Quarto by entering the following into the terminal:
+
+```bash
+quarto install tinytex
+```
+
+All R packages that this project requires are managed using [`renv`](https://cran.r-project.org/package=renv).
+Therefore, `renv` needs to be installed first, by entering the following in the R console:
+
+```r
+install.packages("renv")
+```
+
+Next, one can open a new R session in the root directory of this project and run the following, which should install all required R packages at their recorded versions:
+
+```r
+renv::restore()
+```
+``````
+:::
+
+::: {#tip-usage .callout-tip collapse="true"}
+### Usage
+
+`````{.md .code-overflow-wrap filename="README.md"}
+## Usage
+
+The most important file in this project is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:
+
+- `Bibliography.bib`: bibliographic references used in the manuscript
+- `Data.csv`: a data set containing the simplified `palmerpenguins` data created using `write.csv(palmerpenguins::penguins, "dat.csv", row.names = FALSE)`
+- `data_dictionary.html`: a dictionary to the data file, created using `create_data_dictionary.R`
+
+The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.
+
+The manuscript can be rendered to PDF using the following command:
+
+```bash
+quarto render Manuscript.qmd
+```
+``````
+:::
+
+::: {#tip-citation .callout-tip collapse="true"}
+### Citation
+
+```{.md .code-overflow-wrap filename="README.md"}
+## Citation
+
+Please cite this draft as follows:
+
+> Zerna & Scheffel (2024): "A Study on Penguins: A Minimal Reproducible Example". Unpublished manuscript.
+```
+:::
+
+::: {#tip-citation .callout-tip collapse="true"}
+### License
+
+```{.md .code-overflow-wrap filename="README.md"}
+## License
+
+The manuscript file and the bibliograhpic references are made available by us under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/).
+The data set by Kristen Gorman has been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+`apaquarto` has also been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+All other files in this project are available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+```
+:::

### 2024-09-13 setup.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

@@
 
-The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the @APA2020.
+The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
 It can be installed in the project using the following command:

### 2024-09-13 choose_license.qmd
Rework setup and licensing chapters
Following suggestions by Jonas

@@
 For example, if you were to create a database of artworks by others, those artworks would be licensed individually as chosen by the artists, but the license for the database as a whole could be chosen by you.
-The latter includes the structure of the database (e.g., the selection of entries and field names) and any sui generis database rights.
+The latter includes the structure of the database (e.g., the selection of entries and field names) and any _sui generis_ database rights.
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
@@
 Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015].
+This means that if you only use a work internally, you do not need to share your derivative works.
 It is also worth reiterating that these licenses do not restrict the original author(s): They are still permitted to distribute their work under a different license and without sharing the source code.
@@
 For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and sui generis database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
+Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
 There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
@@
 ```{.bash filename="Terminal"}
-pip install reuse
+pipx install reuse
 ```
@@
 
-If you want to indicate the license for all files in a particular folder, you can create a filed called `REUSE.toml` and add an `[[annotations]]` table for them:
+If you want to indicate the license for all files in a particular folder, you can create a file called `REUSE.toml` and add an `[[annotations]]` table for them:
 

### 2024-09-13 setup.qmd
Rework setup and licensing chapters
Following suggestions by Jonas

@@
 3. copyright of the data and their particular structure, and
-4. any sui generis database right.
+4. any _sui generis_ database right.
 
@@
 Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are.
-Additionally, in some jurisdictions data can be subject to _sui generis database rights_ which prevent extracting substantial parts of a database.
+Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.
 As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
@@
 
+:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
+### Hint for "Take Copyright Seriously"
+
+Have a look at the [about page](about.qmd).
+::::
+
 As the manuscript uses some new packages, install them with:

### 2024-09-18 choose_license.qmd
Fix typo
Provide more details for TL;DR

@@
 
-Every expression of an idea, such as literary or artistic works, are automatically covered by copyright.
+Every expression of an idea, such as a literary or an artistic works, is automatically covered by copyright.
 This means that nobody else than the copyright holder is allowed to copy, modify, or share it.
@@
 2.
-In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously (at the choice of the recipient) to all own files (for example, again by recording the license in the file `LICENSE.txt`).
+In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously to all own files (for example, again by recording the license in the file `LICENSE.txt`).
+Recipients are then free to choose under which of the two licenses to use the material.
+
+We recommend to choose [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) for all own works.
+Take your time to read the summaries behind the links and/or the actual license text to understand its legal effect.
+
+You can adapt the following wording to your use case:
+
+> Except otherwise noted, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
 
@@
 
-> The Quarto files in this project are made available under CC\ BY-SA\ 4.0, or (at your option) under the terms of the Apache Software License\ 2.0.
+> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
 

### 2024-09-18 setup.qmd
Fix typos

@@
 
-If you haven't already, open RStudio -- see @nte-terminal for an alternative to RStudio. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
+If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
 
@@
 Then, one can open an R session, by simply typing `R` into the terminal.
-Next, sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
+Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
 If not, set it using `setwd("code-publishing-exercise")`.
@@
 We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR Research Data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
+If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
 

### 2024-09-18 choose_license.qmd
Fix typo

@@
 
-> Except otherwise noted, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
+> Except where noted otherwise, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
 

### 2024-09-18 choose_license.qmd
Create separate section on dual-licensing

@@
 
-"Choosing a License" involves:
+_Choosing a license_ involves:
 
-1. Recording the copyright notices and licenses for files by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
+1. Recording the copyright notices and licenses for content by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
 2.
-In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously to all own files (for example, again by recording the license in the file `LICENSE.txt`).
+In the simplest case, choosing two licenses (one suitable for code, one for everything else) which to apply  simultaneously to all own content (for example, again by recording the license in the file `LICENSE.txt`).
 Recipients are then free to choose under which of the two licenses to use the material.
@@
 
+::: {#tip-dual-licensing .callout-tip}
+#### Dual-licensing
+
+Sometimes, the type of a work is not obvious. For example, a Quarto document...
+
+- ...contains both R code and writing, and
+- ...may be distributed in the source format or as rendered document, possibly including images.
+
+One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to keep the license notice intact]
+
+One solution is to make such a work simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
+This is called dual-licensing and makes it easier to reuse both the rendered document as well as the code.
+For example, one could write:
+
+> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
+:::
+
 Note, that for data you have, at least in principle, the option to apply separate licenses to the individual entries and the collective database.
@@
 
-If a Quarto document contains both R code and writing, one may wonder what type of license to apply.
-It is possible to make such a file simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
-This is called dual-licensing and makes it easier to reuse both rendered PDF as well as the code.
-For example, one could write:
-
-> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
-
 It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license.
 Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense].
+
 You can learn which license an installed package uses via `packageDescription("<PACKAGE_NAME>", fields = "License")`.
@@
 Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed "[later](make_readme.qmd)".
+Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
 
@@
 In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim.
-This is a practice propagated by GitHub, which provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
+This is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
 However, if the project is not completely covered by one single license, this practice may become unwieldy.

### 2024-09-18 about.qmd
Fix heading level
The title is already level 1, hence
all headings need to start with level 2.
Remove email address by Josephine
As it is outdated

@@
   - name: "Josephine Zerna"
-    email: "josephine.zerna@tu-dresden.de"
     orcid: "0000-0003-2892-884X"
@@
 
-# Utilized Software
+## Utilized Software
 
@@
 
-# License and Disclaimer
+## License and Disclaimer
 
@@
 
-# Funding
+## Funding
 

### 2024-09-21 make_readme.qmd
Mention CFF
Published articles are mostly a matter of citation

@@
 
-:   What is the project about? What are its features? Is there a published article associated with it?
+:   What is the project about? What are its features? Why was it created?
 
@@
 :   Is there a recommended way to cite this project?
+Is there a published article associated with it that you would like to have cited?
 We recommend to increase citability by using the [CodeMeta generator](https://codemeta.github.io/codemeta-generator/) to create a file called `codemeta.json`.
@@
 Then, put it into the project folder.
+Another common standard is the [Citation File Format](https://citation-file-format.github.io/) (CFF), which is an equally valid option.
 
@@
 
-:   Under which license are the contents in this project available?
+:   Under which licenses are the works in this project available?
 

### 2024-09-21 choose_license.qmd
A few improvements to license infos

@@
 [If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it.
-This legal default of "all rights reserved" runs counter to many cultural and scientific processes.
+This legal default of "all rights reserved" was created to benefit publishers, not authors, [@Fogel2006] and runs counter to many cultural and scientific processes.
 Copyright licenses enable authors to free up their works for reuse by others.
@@
 
-When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license that goes further than the open access license available to readers of the journal [@Rumsey2022].
-However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice.
+When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
+This would give them more rights than the readers of the article have through the chosen open access license [@Rumsey2022], and exceeds exceeds by far what is necessary to make publication possible [@Suber2022].
+Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
 If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
@@
 By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work.
-Software licenses may additionally require to make the source code available to everybody the software is shared with.
+Software licenses may additionally require to make the source code available to everybody the software is shared with and often require to display the full text of the license upon usage.
 Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.
@@
 
-::: {#tip-dual-licensing .callout-tip}
-#### Dual-licensing
+::: {#tip-multi-licensing .callout-tip}
+#### Multi-licensing
 
@@
 
-One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to keep the license notice intact]
+One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license]
 
-One solution is to make such a work simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
-This is called dual-licensing and makes it easier to reuse both the rendered document as well as the code.
+One solution is to make such a work simultaneously available under two (or more) licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
+This is called multi-licensing and makes it easier to reuse both the rendered document as well as the code.
 For example, one could write:
@@
 As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
-There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015], but we would like to highlight one reason in particular.
+There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023], but we would like to highlight one reason in particular.
 
@@
 The easiest approach is to make note of every foreign work included in your project and record its license.
-Then, dual-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.
+Then, multi-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.
 
@@
 
-For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022].
+For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
 :::

### 2024-09-21 choose_license.qmd
Fix typo

@@
 [If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it.
-This legal default of "all rights reserved" was created to benefit publishers, not authors, [@Fogel2006] and runs counter to many cultural and scientific processes.
+This legal default of "all rights reserved" was created to benefit publishers, not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
 Copyright licenses enable authors to free up their works for reuse by others.

### 2024-09-21 choose_license.qmd
Fix typo

@@
 When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
-This would give them more rights than the readers of the article have through the chosen open access license [@Rumsey2022], and exceeds exceeds by far what is necessary to make publication possible [@Suber2022].
+This would give them more rights than the readers of the article have through the chosen open access license [@Rumsey2022] and exceeds exceeds by far what is necessary to make publication possible [@Suber2022].
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.

### 2024-11-26 index.qmd
Mention alternatives to Quarto

@@
 
+::: {#nte-quarto-alternatives .callout-note collapse="true"}
+## Alternatives to Quarto
+
+Quarto is a tool to create reproducible documents, but not the only one.
+You may also use [RMarkdown](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [Manubot](https://manubot.org/), [showyourwork!](https://show-your.work/), or one of the countless others.
+:::
+
 We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:

### 2024-11-26 choose_license.qmd
Improve wording
Improve on license choosing

@@
 Every expression of an idea, such as a literary or an artistic works, is automatically covered by copyright.
-This means that nobody else than the copyright holder is allowed to copy, modify, or share it.
+Put simply, this means that nobody else than the copyright holder is allowed to copy, modify, or share it.
 Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
+Certain exceptions and limitations to copyright (such as fair use or the right to quote) may obviate the need for a license.
 
@@
 
-When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
-This would give them more rights than the readers of the article have through the chosen open access license [@Rumsey2022] and exceeds exceeds by far what is necessary to make publication possible [@Suber2022].
+When you have your article published under a Creative Commons license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
+This would give them more rights than the readers of the article have through the chosen Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
@@
 We have created a flowchart that covers the most likely types of works you will create as a researcher: software, writing (i.e., text), images, audio, video, and data (see @fig-flowchart-simple).
-This flowchart always recommends the most permissive license possible to maximize reuse -- below we provide two additional flowcharts that allow for more choices.
+This flowchart always recommends the most permissive license possible to maximize reuse -- though we provide two additional flowcharts below that allow for more choices.
 Click on the name of a license to learn more about it.
@@
 
-Note, that for data you have, at least in principle, the option to apply separate licenses to the individual entries and the collective database.
+Note, that for data you have, at least in principle, the option to apply different licenses to the individual entries and the collective database.
 For example, if you were to create a database of artworks by others, those artworks would be licensed individually as chosen by the artists, but the license for the database as a whole could be chosen by you.
@@
 
-- __fonts:__ Copyleft licenses applied to fonts can a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
+- __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
 If the intent is that only derived fonts, if published, have to be put under the same license, the [SIL Open Font License\ 1.1](https://openfontlicense.org/) is an appropriate choice.
@@
 For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access if shared [see @Poole2017].
-Similarly, following an example from @Matt2009 if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
+Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
 If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
@@
 
-The consequence of applying a permissive license to your work is that others may distribute their modifications to it under stricter terms, possibly without a free/open license.
+The consequence of applying a permissive license to your work is that others may distribute their adaptations from it under stricter terms, possibly without a free/open license.
 You may prefer that everybody who receives somebody else's adaptations of your work enjoys the same rights that you granted to them.
@@
 The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
+Sometimes, even different parts of a file might be subject to different licenses.
 The easiest approach is to make note of every foreign work included in your project and record its license.

### 2024-11-26 archive.qmd
Add reference
Provide some advice on finding a suitable repository

@@
 
-Additional guidance is provided by @Daquin2023 and @Garijo2022.
+Additional guidance is provided by @Daquin2023, @Garijo2022, and @ScienceEurope2021.
+You can use the registry of research data repositories [re3data.org](https://www.re3data.org/) to find one that matches your criteria.
+We think that the [Software Heritage code archive](https://www.softwareheritage.org/) and [Zenodo](https://zenodo.org/) are good choices, but don't forget to also check your local institutions for recommendations.
 

### 2024-11-26 setup.qmd
Provide additional references

@@
 We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately.
+Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).
 
@@
 - "The Good Research Code Handbook" [@Mineault2021]
+- "Quality assurance of code for analysis and research" [@UKGovernment2020]
 - "The Art of UNIX Programming" [@Raymond2003]

### 2024-11-26 index.qmd
Remove Manubot as it's not primarily for executing code

--- 2024-09-12 index.qmd
@@
 
+::: {#nte-quarto-alternatives .callout-note collapse="true"}
+## Alternatives to Quarto
+
+Quarto is a tool to create reproducible documents, but not the only one.
+You may also use [RMarkdown](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), or one of the countless others.
+:::
+
 We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:

### 2024-11-27 choose_license.qmd
Remove Dissemin and make database parts more clear

@@
 
-If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) and [Dissemin](https://dissem.in/) for legal options to still make it available free of charge to readers.
+If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
 :::
@@
 If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos).
-For software that's almost universally the [GPLv3](https://www.gnu.org/licenses/license-list.en.html#GPLCompatibleLicenses) and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/compatiblelicenses).
+For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the CC\ BY-SA\ 4.0.
 For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and CC\ BY-SA\ 4.0 are acceptable.
@@
 
-Note, that for data you have, at least in principle, the option to apply different licenses to the individual entries and the collective database.
+Note, that for data there is, at least in principle, the possibility that a different license applies to the individual entries than to the collective database.
 For example, if you were to create a database of artworks by others, those artworks would be licensed individually as chosen by the artists, but the license for the database as a whole could be chosen by you.
-The latter includes the structure of the database (e.g., the selection of entries and field names) and any _sui generis_ database rights.
+The latter includes the selection of entries, field names and any _sui generis_ database rights.
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Metadata, in particular, should always be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Dryad2011; @ScienceCommons2018] -- otherwise, see @nte-other-work-types for some caveats.
 
@@
 - __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
-For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access if shared [see @Poole2017].
+For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017].
 Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
@@
 
-In most cases, the output of software, like images or tables, does not depend on its license.
+In most cases, the output of software, like images or tables, does not depend on the software's license.
 Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
@@
 
-Note: DRM = digital rights management
+Note. DRM = digital rights management
 
@@
   subgraph "License for individual data entries (content)"
-  existing_license_content --"No, we created the content<br>entirely by ourselves."--> metadata("Is metadata?")
+  existing_license_content --"No, we created the content<br>entirely by ourselves."--> metadata("Entries are facts<br>(like measurements<br>or metadata)?")
   existing_license_content --"Yes, it was<br>shared under a<br>free/open license."--> use_existing_license_content["<em>Use that license</em>"]

### 2024-11-27 choose_license.qmd
Remove references to Science Commons
Their argument is likely not upheld anymore as they
advertise other licenses than CC0 for data.

@@
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Dryad2011; @ScienceCommons2018] -- otherwise, see @nte-other-work-types for some caveats.
+Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) -- otherwise, consult @nte-other-work-types for some caveats.
 

### 2024-11-27 setup.qmd
Use APA citation style
Mention v2 of Data Package standard

@@
 lang: en
-bibliography: literature.bib
 engine: knitr
@@
 
-Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.
-For the purpose of the following code, it is just a nested list that we edit to include our own information.
+Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released.
+As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information.
 We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

### 2024-11-27 index.qmd
Fix link

@@
 
-In particular, we will assume you have installed [Git](https://git-scm.com/download/), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/).
+In particular, we will assume you have installed [Git](https://git-scm.com/downloads), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/).
 Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).

### 2024-11-27 index.qmd
Mention Org Mode and StatTag

@@
 Quarto is a tool to create reproducible documents, but not the only one.
-You may also use [RMarkdown](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), or one of the countless others.
+You may also use [RMarkdown](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), [Org Mode](https://orgmode.org/), [StatTag](https://sites.northwestern.edu/stattag/), or one of the countless others.
 :::

### 2024-11-27 intro.qmd
Use APA citation style
Add a background on research compendia to introduction

--- 2024-09-04 intro.qmd
@@
 lang: en
-bibliography: literature.bib
 ---
 
-Together, we will create a research compendium whilst taking the following ten things into account [@Arguillas2022]:
+In the following, we will provide a brief introduction to the concept of _research compendia_.
+
+## The importance of sharing
+
+Suppose you are reading an article about a new imaging method to turn seismological data into subsurface images.
+The article describes the ideas that went into developing this method and presents a few examples to illustrate its superiority over previous approaches.
+You got interested and would like to apply this method to your own data.
+However, with only the article available, it could take months to come up with a working solution, if possible at all.
+This situation has been put aptly by @Buckheit1995, distilling an idea by Jon Claerbout:
+
+> "An article about computational science in a scientific publication is not the scholarship itself, it is merely advertising of the scholarship.
+The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."
+
+Even for articles that are use computers to apply existing methods (rather than reporting on a new method), sharing the source code and being transparent about the computational environment is imperative to making research reproducible [@Ince2012].
+By reproducibility, we mean "obtaining consistent results using the same input data; computational steps, methods, and code; and conditions of analysis" [@NASEM2019, p.
+46].
+
+## Linking results and computations
+
+This tutorial not only covers sharing the source code, but also connecting it to the results through the creation of dynamic documents.
+Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article.
+Dynamic documents bundled together with any necessary data and auxiliary software are called a [research compendium](https://research-compendium.science/) [@Gentleman2007].
+
+The practice to interleave narrative text with code has its roots in the paradigm of literate programming, where documentation and source code are treated as equals and are arranged in a way to maximize understanding [@Knuth1984].
+Alternating text and code can be also found in notebook interfaces for exploratory programming, such as provided by Wolfram Mathematica or Jupyter Notebooks [@Kluyver2016], with the added capability of executing the code and embedding its output.
+With _Sweave_ [@Leisch2002], ideas from both worlds -- literate programming and embedding program output -- were combined into one tool for rendering dynamic documents using the R programming language.
+It is the predecessor of the R package `knitr` [@Xie2015] which is being used under the hood in this tutorial.^[Specifically, Quarto employs `knitr` to execute chunks of R code.]
+
+Linking results with their computations has benefits for authors and readers.
+For the author, articles always contain the most recent version of figures, as they are updated automatically when the computation changes.
+For the readers, it enables understanding exactly how a particular result was obtained if they get access to the underlying research compendium.
+
+## Best practices
+
+When creating the research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 
@@
 > 10. __Review:__ A series of managed activities are needed to ensure continued access to and functionality of the research compendium and its components for as long as necessary.
+
+Although this tutorial guides you through the creation of a research compendium, you are invited to revisit these questions after completion and check whether and how each point was addressed (or not).
+Further, you can consult them as a checklist for future projects.

### 2024-11-28 choose_license.qmd
Use APA citation style
Mention regulation by funder

@@
 lang: en
-bibliography: literature.bib
 format:
@@
 
+A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
+And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
+By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
+
 ::: {.column-margin}
@@
 :::
-
-A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
-And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
-By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
 
@@
 First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
-The same applies if there are strong community norms to use a particular free/open license.^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.
+The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder.
+Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.
 
@@
   start("We want<br>to choose<br>a license.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 
-  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
-  norm --"Does not<br>exist"--> type("Work type?")
+  regulation --"Exists"--> follow_existing_norms["<em>Follow that</em>"]
+  regulation --"Does not<br>exist"--> type("Work type?")
 
@@
 If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
-- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/gpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
+- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
 And if a document contained source code covered by the GPLv3, the same license would also apply to the document [@MadHatter2022].
@@
   start("We want to choose a<br>license for software,<br>writing, image, audio,<br>or video.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 
-  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
-  norm --"Does not<br>exist"--> type("Work type?")
+  regulation --"Exists"--> follow_existing_norms["<em>Follow that</em>"]
+  regulation --"Does not<br>exist"--> type("Work type?")
 
@@
   start("We want to choose a<br>license for data.") --"We adapted a database by<br>others shared under a<br>free/open license."--> use_existing_license_db["<em>Use its license(s)<br>for content and database</em>"]
-  start --"We created a database<br>entirely by ourselves."--> norm("Community norm<br>regarding license?")
+  start --"We created a database<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 
-  norm --"Exists"--> follow_existing_norms["<em>Follow that norm</em>"]
-  norm --"Does not<br>exist"--> existing_license_content("Adapting content<br> by others?")
+  regulation --"Exists"--> follow_existing_norms["<em>Follow that</em>"]
+  regulation --"Does not<br>exist"--> existing_license_content("Adapting content<br> by others?")
 
   subgraph "License for individual data entries (content)"
-  existing_license_content --"No, we created the content<br>entirely by ourselves."--> metadata("Entries are facts<br>(like measurements<br>or metadata)?")
+  existing_license_content --"No, we created the content<br>entirely by ourselves."--> facts("Entries are facts<br>(like measurements<br>or metadata)?")
   existing_license_content --"Yes, it was<br>shared under a<br>free/open license."--> use_existing_license_content["<em>Use that license</em>"]
-  metadata --"Yes"--> cc0_content_metadata["CC0 1.0"]
-  metadata --"No"--> choose_license["<em>Consult flowchart for<br>software, writing,<br>image, audio, and video</em>"]
+  facts --"Yes"--> cc0_content_metadata["CC0 1.0"]
+  facts --"No"--> choose_license["<em>Consult flowchart for<br>software, writing,<br>image, audio, and video</em>"]
   end

### 2024-11-28 choose_license.qmd
Describe license situation for researchers at LMU

@@
 Copyright licenses enable authors to free up their works for reuse by others.
+
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For researchers at LMU
+
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision which license to use lies with your associated chair.
+In the (rare) case the software you developed has a technical application and constitutes an invention, you are required to officially report your invention to the university which decides on its release.
+If the university claims the invention, you may not apply a copyright license that makes a patent grant.
+More information about this process is provided in the [service portal](https://www.serviceportal.verwaltung.uni-muenchen.de/internationals/research_transfer/intellectual_property/index.html).
+More information about software patents are provided by the FSFE: [Software Patents in Europe](https://fsfe.org/activities/swpat/swpat.en.html).
+:::
 

### 2024-11-28 choose_license.qmd
Mention the other case as well

@@
 If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision which license to use lies with your associated chair.
-In the (rare) case the software you developed has a technical application and constitutes an invention, you are required to officially report your invention to the university which decides on its release.
+In the (rare) case the software you developed has a technical application and it falls under the legal definition of an invention, you are required to officially report your invention to the university which decides on its release.
 If the university claims the invention, you may not apply a copyright license that makes a patent grant.
-More information about this process is provided in the [service portal](https://www.serviceportal.verwaltung.uni-muenchen.de/internationals/research_transfer/intellectual_property/index.html).
+In all other cases you are free to license your research project as you like.
+
+More information about th invention disclosure process can be found in the [service portal](https://www.serviceportal.verwaltung.uni-muenchen.de/internationals/research_transfer/intellectual_property/index.html).
 More information about software patents are provided by the FSFE: [Software Patents in Europe](https://fsfe.org/activities/swpat/swpat.en.html).

### 2024-11-28 choose_license.qmd
Minor improvements and fixes

--- 2024-11-28 choose_license.qmd
@@
 If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos).
-For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the CC\ BY-SA\ 4.0.
-For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and CC\ BY-SA\ 4.0 are acceptable.
+For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/).
+For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
 :::
@@
 
-One should be cautious about the restrictions of licenses applied to the following types of works:
+One should take special care when applying license to the following types of works:
 

### 2024-11-28 intro.qmd
Minor improvements and fixes

@@
 
-## The importance of sharing
+## The Importance of Sharing
 
@@
 However, with only the article available, it could take months to come up with a working solution, if possible at all.
-This situation has been put aptly by @Buckheit1995, distilling an idea by Jon Claerbout:
+This situation has been put aptly by @Buckheit1995, distilling an idea by geophysicist Jon Claerbout:
 
@@
 
-## Linking results and computations
+## Linking Results and Computations
 
@@
 
-## Best practices
+## Best Practices
 
-When creating the research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
+When creating a research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 

### 2024-11-28 setup.qmd
Minor improvements and fixes

@@
 
-Then, one can open an R session, by simply typing `R` into the terminal.
+Then, one can open an R session by simply typing `R` into the terminal.
 Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
@@
 ::: {#tip-apaquarto .callout-tip}
-#### Not a psychologist?
+#### Not a Psychologist?
 
@@
 ::: {#tip-update-quarto .callout-tip}
-If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however `apaquarto` sometimes requires more recent versions.
+If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
 :::

### 2024-11-28 setup.qmd
Use appropriate callouts

@@
 
-::: {#wrn-legal-restrictions .callout-warning}
+::: {#imp-legal-restrictions .callout-important}
 ### Consider Legal Restrictions Before Sharing
@@
 
-::: {#wrn-take-copyright-seriously .callout-warning}
+::: {#imp-take-copyright-seriously .callout-important}
 ### Take Copyright Seriously

### 2024-11-28 choose_license.qmd
Use appropriate callouts
Fix diagram

@@
 lang: en
-format:
-  html:
-    mermaid:
-      theme: neutral
 engine: knitr
@@
 
-::: {#cau-exclusive-rights .callout-caution}
+::: {#wrn-exclusive-rights .callout-warning}
 ### Giving Publishers Exclusive Rights
@@
   subgraph "License for combination of data (database)"
-  choose_license --> switch_license["<em>Depending on content license</em>"]
+  choose_license --> switch_license["<em>Depending on<br>content license</em>"]
   use_existing_license_content --> switch_license

### 2024-11-28 choose_license.qmd
Explain output exception

@@
 
-In most cases, the output of software, like images or tables, does not depend on the software's license.
-Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
-However, if the output is based on data, it can be considered a derivative work and the license of the data also applies.
+In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @apsillers2019].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
+However, if the output is based on data, it can be considered a derivative work of the data and the license of the data also applies.
 For example, maps may be considered as a derivative of the geographic data they are based on.

### 2024-11-28 choose_license.qmd
Fix links

@@
 If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
-- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
-And if a document contained source code covered by the GPLv3, the same license would also apply to the document [@MadHatter2022].
+- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].
+And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
 - __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
 For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017].
-Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
+Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared.
 If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
 Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
-- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
+- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
 :::

### 2024-11-28 setup.qmd
Add citation for Fowler quote
Address citation of software and data

@@
 > 
-> --- Martin Fowler, British software engineer
+> --- @Fowler1999, p. 15
+
+## Citing Software and Data
+
+If you rely on software or data by others in your research, the question arises whether and how to cite it in your publications.
+Put simply, all data relied upon should be cited to allow for precise identification and access.
+From the "eight core principles of data citation" [@Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/)]:
+
+> **Principle 1 – Importance**: "Data should be considered legitimate, citable products of research.
+Data citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications."
+> 
+> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited."
+> 
+> **Principle 5 – Access**: "Data citations should facilitate access to the data themselves and to such associated metadata, documentation, code, and other materials, as are necessary for both humans and machines to make informed use of the referenced data."
+> 
+> **Principle 7 – Specificity and Verifiability**: "Data citations should facilitate identification of, access to, and verification of the specific data that support a claim.
+Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific time slice, version and/or granular portion of data retrieved subsequently is the same as was originally cited."
+
+When it comes to software, the answer is a little more nuanced – you can consult @fig-software-citation for advice whether to cite it.
+As with data, citations should allow for exact identification and access.
+From the "software citation principles" [@Smith2016]:
+
+> **1\.
+Importance**: Software should be considered a legitimate and citable product of research.
+Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be included in the metadata of the citing work, for example in the reference list of a journal article, and should not be omitted or separated.
+Software should be cited on the same basis as any other research product such as a paper or a book, that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
+> 
+> **5\.
+Accessibility**: Software citations should facilitate access to the software itself and to its associated metadata, documentation, data, and other materials necessary for both humans and machines to make informed use of the referenced software.
+> 
+> **6\.
+Specificity**: Software citations should facilitate identification of, and access to, the specific version of software that was used.
+Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
+
+To help with generating the citation, you can use the [CiteAs](https://citeas.org/) service.
+
+::: {#fig-software-citation}
+```{mermaid}
+flowchart TB
+  asks_for_citation("Does the software<br>ask you to cite it?")
+  critical_or_unique_contribution("Did the software<br>play a critical part<br>in or contributed<br>something unique<br>to your research?")
+  manipulate("Did the software<br>manipulate or create<br>your data, software,<br>or other outputs?")
+  relies_on_credit("Do the authors of<br>the software rely<br>on academic credit<br>for funding?")
+  cite[Cite!]
+  nocite[Don't cite!]
+
+  asks_for_citation --"Yes"--> cite
+  asks_for_citation --"No"--> critical_or_unique_contribution
+  critical_or_unique_contribution --"Yes"--> cite
+  critical_or_unique_contribution --"No"--> manipulate
+  manipulate --"Yes"--> cite
+  manipulate --"No"--> relies_on_credit
+  relies_on_credit --"Yes"--> cite
+  relies_on_credit --"No"--> nocite
+```
+
+"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Simplified from original.
+:::
 

### 2024-11-28 choose_license.qmd
Fix typo

--- 2024-11-28 choose_license.qmd
@@
 lang: en
+format:
+  html:
+    mermaid:
+      theme: neutral
 engine: knitr
@@
 
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For researchers at LMU
+
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision which license to use lies with your associated chair.
+In the (rare) case the software you developed has a technical application and it falls under the legal definition of an invention, you are required to officially report your invention to the university which decides on its release.
+If the university claims the invention, you may not apply a copyright license that makes a patent grant.
+In all other cases you are free to license your research project as you like.
+
+More information about the invention disclosure process can be found in the [service portal](https://www.serviceportal.verwaltung.uni-muenchen.de/internationals/research_transfer/intellectual_property/index.html).
+More information about software patents are provided by the FSFE: [Software Patents in Europe](https://fsfe.org/activities/swpat/swpat.en.html).
+:::
+
 A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
@@
 
-::: {#wrn-exclusive-rights .callout-warning}
+::: {#cau-exclusive-rights .callout-caution}
 ### Giving Publishers Exclusive Rights
@@
 If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos).
-For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/).
-For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
+For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the CC\ BY-SA\ 4.0.
+For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and CC\ BY-SA\ 4.0 are acceptable.
 :::
@@
 
-One should take special care when applying license to the following types of works:
+One should be cautious about the restrictions of licenses applied to the following types of works:
 
@@
   subgraph "License for combination of data (database)"
-  choose_license --> switch_license["<em>Depending on<br>content license</em>"]
+  choose_license --> switch_license["<em>Depending on content license</em>"]
   use_existing_license_content --> switch_license

### 2024-11-28 setup.qmd
Add total number of software citation principles
Mention CC BY 4.0 license for software citation principles

@@
 Put simply, all data relied upon should be cited to allow for precise identification and access.
-From the "eight core principles of data citation" [@Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/)]:
+From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
 
@@
 As with data, citations should allow for exact identification and access.
-From the "software citation principles" [@Smith2016]:
+From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):
 
@@
 
-To help with generating the citation, you can use the [CiteAs](https://citeas.org/) service.
+To help with generating the citation, you can use the [CiteAs service](https://citeas.org/).
 

### 2024-11-29 setup.qmd
Use global language and tidy _quarto.yml
Add Console notice

@@
 title: "Project Setup"
-lang: en
 engine: knitr
@@
 
-```r
+```{.r filename="Console"}
 table_info <- c(
@@
 
-```r
+```{.r filename="Console"}
 # Read data and create schema

### 2024-11-29 about.qmd
Improve license wording
Improve license statement on about page

@@
 
-If not indicated otherwise, this tutorial is licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/). None of the discussion in this tutorial constitutes legal advice.
+Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
+None of the discussion in this tutorial constitutes legal advice.
 
-Any screenshots of [RStudio](https://github.com/rstudio/rstudio) are released under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/gpl-3.0.html).
+The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are available under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/gpl-3.0.html).
+
+::: {.nte-agplv3 .callout-note collapse="true"}
+### GNU Affero General Public License v3
 
@@
 ```
+:::
 

### 2024-11-29 make_readme.qmd
Minor improvements and fixes
Use global language and tidy _quarto.yml
Elaborate on citing code and data

@@
 title: "Make a README"
-lang: en
 engine: knitr
@@
 
-## What does a README contain?
+## What Does a README Contain?
 
@@
 What software needs to be installed?
-Mention all dependencies here that are not explicitly managed by `renv`, such as the system dependencies of R packages as well as the version of Quarto.
-An R package's system dependency is any additional software that you need to install on your computer in order to use a particular R package.
-For example, the R package [`reticulate`](https://rstudio.github.io/reticulate/) allows to run Python code from within R.
-However, in order to actually use it, one has to additionally install Python itself as it does not come together with `reticulate` -- rather, it is a system dependency.
+R itself and the R packages are already documented as this project uses `renv`.
+Therefore you can focus on all other dependencies, such as the system dependencies of R packages as well as the version of Quarto.^[As of August 2024, a proposal for `renv` to record the version of Quarto has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] Also, don't forget to mention software that you have used for any manual steps.
 See @sec-dependencies for additional information.
@@
 An overview over the system dependencies of R packages can be created using the function `pak::pkg_sysreqs()`.
-In combination with `renv`, we can obtain the system dependencies of all R packages in the current project:
+In combination with `renv`, we can obtain the system dependencies of all R packages the current project directly depends on:
 
 ```{.r filename="Console"}
-pak::pkg_sysreqs(renv::dependencies()$Package)
+# First, install pak
+renv::install("pak")
+
+# Then, identify the system dependencies of your direct dependencies
+renv::dependencies()$Package |>
+  unique() |>
+  pak::pkg_sysreqs()
 ```
@@
 
-We can see that `make` and `pandoc` were identified as system dependencies. One can obtain their version by running them with the `--version` argument:
+We can see that `make` and `pandoc` were identified as system dependencies. Often, one can obtain their version by running them with the `--version` argument:
 
@@
 ```
+
+Of course, all the system dependencies identified until now may have dependencies on their own. Use your own judgement to decide when not to dig deeper.
 

### 2024-11-29 setup.qmd
Elaborate on citing code and data

@@
 
-## Citing Software and Data
+## Cite Data and Software
 
-If you rely on software or data by others in your research, the question arises whether and how to cite it in your publications.
+If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.
+
+### Data
+
 Put simply, all data relied upon should be cited to allow for precise identification and access.
@@
 
-When it comes to software, the answer is a little more nuanced – you can consult @fig-software-citation for advice whether to cite it.
+Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
+
+::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
+#### Hint for citing the data set
+
+As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:
+
+```{r, echo = -1}
+invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
+citation("palmerpenguins")
+```
+
+As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).
+
+Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`.
+Then, add a sentence to the manuscript such as follows:
+
+```md
+The analyzed data are by @horst2020.
+```
+
+Render the document to check that the citation is displayed properly.
+
+```{.bash filename="Terminal"}
+quarto render Manuscript.qmd
+```
+:::
+
+### Software
+
+When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies.
+You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not.
 As with data, citations should allow for exact identification and access.
@@
 
-To help with generating the citation, you can use the [CiteAs service](https://citeas.org/).
+In practice, the first step is to identify all pieces of software the project relies on.
+A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before.
+Then there are the individual R packages, Quarto extensions, and $\TeX$ packages.
+All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper.
+For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit.
+A system dependency is additional software that you require on your computer apart from the R package.
 
@@
 
+Now, add references for the software you would like to cite to the manuscript.
+In the following, we will demonstrate this for R and all R packages by using the R package `grateful`.
+For arbitrary software, you can use the [CiteAs service](https://citeas.org/) to create appropriate citations.
+
+Add the following code chunk to the end of the discussion in the manuscript:
+
+``````{.qmd filename="Manuscript.qmd"}
+```{{r}}
+#| echo: false
+
+grateful::cite_packages(
+    output = "paragraph",
+    out.dir = ".",
+    omit = NULL,
+    dependencies = TRUE,
+    passive.voice = TRUE,
+    bib.file = "grateful-refs"
+)
+```
+``````
+
+This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that the detection can fail ] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
+
+```{.yml filename="Manuscript.qmd"}
+bibliography:
+  - Bibliography.bib
+  - grateful-refs.bib
+```
+
+Use `renv` to view, install, and record the newly used package `grateful`:
+
+```{.r filename="Console"}
+renv::status()
+renv::install()
+renv::snapshot()
+```
+
+Finally, render the document again and commit the changes:
+
+```{.bash filename="Terminal"}
+quarto render Manuscript.qmd
+
+git status
+git add .
+git commit -m "Cite data and software"
+```
+
 ## The Last Mile
@@
 `renv` only records the versions of R packages and of R itself.
-This means that potential system dependencies of R packages and other tools utilized in the project are not documented anywhere, including Quarto.^[As of August 2024, a proposal to record the version of Quarto has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] We will manually write them down when [creating a README](make_readme.qmd).
+This means that everything we have not decided to cite in the previous step is not documented anywhere.
+We will cover system dependencies when [creating a README](make_readme.qmd).
 For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).

### 2024-11-29 _Instructions.qmd
Use "data set" rather than "dataset"

--- 2024-05-16 Instructions.qmd
@@
 title: "Workshop Instructions: Reproducible Scientific Manuscripts"
+lang: en
 ---
@@
 4. Copy a figure into the project folder and insert it into the document.
-5. Create a table, either by hand or by importing your dataset into the project folder and using a package to generate a table.
+5. Create a table, either by hand or by importing your data set into the project folder and using a package to generate a table.
 6. Add your bibliography as a .bib file.

### 2024-11-29 setup.qmd
Use "data set" rather than "dataset"

@@
 
-Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic dataset^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
+Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
 For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
@@
 table_info <- c(
-  title = "penguins dataset",
+  title = "penguins data set",
   description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"

### 2024-12-03 choose_license.qmd
Use global language and tidy _quarto.yml
Various minor improvements

--- 2024-11-28 choose_license.qmd
@@
 title: "Choose a License"
-lang: en
 engine: knitr
@@
 
-Researchers are only allowed to share material they have the necessary rights to, for example, through a free/open license.
+Researchers are only allowed to share material they have the necessary rights to.
+This includes copyrights, which are granted by free/open copyright licenses.
 These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.
@@
 
-A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
-And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
+A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss are non-exclusive licenses and do not limit the author or copyright holder in their rights.
+And while you _could_ write your own license, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
 By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
@@
 
-When you have your article published under a Creative Commons license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
-This would give them more rights than the readers of the article have through the chosen Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
+Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you.
+This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
@@
 
-First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
+First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
 The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder.
@@
   start("We want<br>to choose<br>a license.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
 
@@
 - __Attribution__ means whether recipients of your work are required to provide attribution to you.
-For software licenses, this is called a copyright notice.
+For software licenses, the attribution is part of a copyright notice.
 Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
@@
   start("We want to choose a<br>license for software,<br>writing, image, audio,<br>or video.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
 
@@
   start("We want to choose a<br>license for data.") --"We adapted a database by<br>others shared under a<br>free/open license."--> use_existing_license_db["<em>Use its license(s)<br>for content and database</em>"]
-  start --"We created a database<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
+  start --"We created a database<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
 

### 2024-12-03 make_readme.qmd
Various minor improvements

@@
 
-# Then, identify the system dependencies of your direct dependencies
+# Then, identify the system dependencies
+# of your direct dependencies.
 renv::dependencies()$Package |>
@@
 
-The manuscript file and the bibliograhpic references are made available by us under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/).
-The data set by Kristen Gorman has been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-`apaquarto` has also been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-All other files in this project are available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+The manuscript file and the bibliographic references are made available by us under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
+The data set by Kristen Gorman has been published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+`apaquarto` has also been published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+All other files in this project are available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
 ```

### 2024-12-03 setup.qmd
Various minor improvements

@@
 
-You can now download the data set we have prepared for you and put into your project folder: [`Data.csv`](_01_Example/Data.csv){download=""}
+You can now download the data set we have prepared for you and put it into your project folder: [`Data.csv`](_01_Example/Data.csv){download=""}
 
@@
 
-The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
+The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)^[Created with `palmerpenguins` v0.1.1 using `write.csv(palmerpenguins::penguins, file = "Data.csv", row.names = FALSE)`.] and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
 It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
@@
 
-When you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so.
-Only build on existing work for which you are given an express grant of relevant rights.
+If you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so.
+Only build on existing work for which you receive an express grant of relevant rights.
 How do you know you are allowed to copy, edit, and share the two files linked above?
@@
 
-- __Fail early, often and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
+- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
 
@@
 Use linting tools^[A linting tool analyzes your code without actually running it.
-This process is called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
+Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
 
@@
 
-This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that the detection can fail ] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
+This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that this automatic detection can miss packages in some circumstances, therefore always verify the rendered result.] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
 
@@
 
-Use `renv` to view, install, and record the newly used package `grateful`:
+Use `renv` to view, install, and record the newly used package:
 

### 2024-12-04 choose_license.qmd
Various minor stylistic and typographic improvements

@@
 - __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
-DRM is an abbreviation for **d**igital **r**ights **m**anagement.
+Implementing such restrictive measures is also referred to as **d**igital **r**ights **m**anagement (DRM).
 
@@
 Put differently, one cannot share a remix of two works where one is licensed under CC\ BY-SA\ 4.0 and the other only allows non-commercial use.
-In order to avoid silos, one should only choose license which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
+In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
 If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012.
@@
 
-!["Open data-capable license interoperability" by Robbie Morrison licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
-Taken without modification from @Morrison2024](images/Morrison2024.jpg){.lightbox}
+!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024] licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)](images/Morrison2024.jpg){.lightbox}
 :::
@@
 Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
+Mostly, this just means indicating which file or folder is covered by which license, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
+Notably, no sort of registration is required.
 

### 2024-12-04 intro.qmd
Use global language and tidy _quarto.yml
Various minor stylistic and typographic improvements

@@
 title: "Introduction"
-lang: en
 ---
@@
 
-Even for articles that are use computers to apply existing methods (rather than reporting on a new method), sharing the source code and being transparent about the computational environment is imperative to making research reproducible [@Ince2012].
+Even when researchers merely apply existing methods (rather than report on a new method), sharing the source code and being transparent about the computational environment is imperative to making their results reproducible [@Ince2012].
 By reproducibility, we mean "obtaining consistent results using the same input data; computational steps, methods, and code; and conditions of analysis" [@NASEM2019, p.
@@
 The practice to interleave narrative text with code has its roots in the paradigm of literate programming, where documentation and source code are treated as equals and are arranged in a way to maximize understanding [@Knuth1984].
-Alternating text and code can be also found in notebook interfaces for exploratory programming, such as provided by Wolfram Mathematica or Jupyter Notebooks [@Kluyver2016], with the added capability of executing the code and embedding its output.
+Alternating text and code can be also found in notebook interfaces for exploratory programming [e.g., Wolfram Mathematica or Jupyter Notebooks, see @Kluyver2016] that also have the capability to execute code and embed its output.
 With _Sweave_ [@Leisch2002], ideas from both worlds -- literate programming and embedding program output -- were combined into one tool for rendering dynamic documents using the R programming language.

### 2024-12-04 choose_license.qmd
Add a few resources on licensing

@@
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) -- otherwise, consult @nte-other-work-types for some caveats.
+Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.
 
@@
 As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
-There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023], but we would like to highlight one reason in particular.
+There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023; @Vezina2020; @WikimediaCommons2009], but we would like to highlight one reason in particular.
 
@@
 For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
+Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/2004/05/25/announcingandexplainingournew20licenses/#:~:text=Share%20Alike%20Across%20Borders)) and compatible licenses ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
 There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.

### 2024-12-04 choose_license.qmd
Incorporate feedback from Felix and Malika

--- 2024-12-03 choose_license.qmd
@@
 title: "Choose a License"
+lang: en
+format:
+  html:
+    mermaid:
+      theme: neutral
 engine: knitr
@@
 
-Researchers are only allowed to share material they have the necessary rights to.
-This includes copyrights, which are granted by free/open copyright licenses.
+Researchers are only allowed to share material they have the necessary rights to, for example, through a free/open license.
 These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.
@@
 
-A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss are non-exclusive licenses and do not limit the author or copyright holder in their rights.
-And while you _could_ write your own license, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For researchers at LMU
+
+If you are an employee at LMU and your funder does not mandate the use of a particular license,
+the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+In the (rare) case that the software you developed has a technical application,
+that it falls under the legal definition of an invention,
+you are required to officially report your invention to the university which decides on its release.
+However, the [IP strategy](https://www.serviceportal.verwaltung.uni-muenchen.de/services/forschungsservices/_downloads/07_ip_strategy.pdf) of the LMU highlights that
+"LMU can make the greatest contribution to the value chain by focusing on the production of outstanding results in basic research and by helping to **make the results publicly accessible**.
+It follows that the commercial exploitation of research results has an accessory role".
+Furthermore, "as a rule, the inventor is entitled to the discretionary use of invention- and research-based results".
+Taken together, LMU emphasizes the importance of openness of research results and the freedom of researchers to share their findings widely.
+
+Only if the university claims the invention after you reported it, you may not apply a copyright license that contains a patent grant.
+In all other cases you are free to license your research project as you like.
+
+More information about the invention disclosure process can be found in the [service portal](https://www.serviceportal.verwaltung.uni-muenchen.de/internationals/research_transfer/intellectual_property/index.html).
+More information about software patents are provided by the FSFE: [Software Patents in Europe](https://fsfe.org/activities/swpat/swpat.en.html).
+:::
+
+A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
+And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
 By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
@@
 
-::: {#wrn-exclusive-rights .callout-warning}
+::: {#cau-exclusive-rights .callout-caution}
 ### Giving Publishers Exclusive Rights
 
-Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you.
-This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
+When you have your article published under a Creative Commons license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
+This would give them more rights than the readers of the article have through the chosen Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
@@
 
-First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
+First, if you adapt (i.e., modify, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
 The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder.
@@
   start("We want<br>to choose<br>a license.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 
@@
 If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
-- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].
-And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
+- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
+And if a document contained source code covered by the GPLv3, the same license would also apply to the document [@MadHatter2022].
 - __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
 For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017].
-Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared.
+Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
 If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
 Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
-- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
+- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
 :::
@@
 
-In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @apsillers2019].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
-However, if the output is based on data, it can be considered a derivative work of the data and the license of the data also applies.
+In most cases, the output of software, like images or tables, does not depend on the software's license.
+Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
+However, if the output is based on data, it can be considered a derivative work and the license of the data also applies.
 For example, maps may be considered as a derivative of the geographic data they are based on.
@@
 - __Attribution__ means whether recipients of your work are required to provide attribution to you.
-For software licenses, the attribution is part of a copyright notice.
+For software licenses, this is called a copyright notice.
 Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
@@
   start("We want to choose a<br>license for software,<br>writing, image, audio,<br>or video.") --"We adapted a work by<br>others shared under a<br>free/open license."--> use_existing_license["<em>Use its license</em>"]
-  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
+  start --"We created the work<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 
@@
   start("We want to choose a<br>license for data.") --"We adapted a database by<br>others shared under a<br>free/open license."--> use_existing_license_db["<em>Use its license(s)<br>for content and database</em>"]
-  start --"We created a database<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>concerning license?")
+  start --"We created a database<br>entirely by ourselves."--> regulation("Other regulation<br>(by community or funder)<br>regarding license?")
 

### 2024-12-10 choose_license.qmd
Add newline again
Add emphasis

--- 2024-12-04 choose_license.qmd
@@
 
-  subgraph "License for individual data entries (content)"
+  subgraph content["<strong>License for content</strong>"]
   existing_license_content --"No, we created the content<br>entirely by ourselves."--> facts("Entries are facts<br>(like measurements<br>or metadata)?")
@@
 
-  subgraph "License for combination of data (database)"
+  subgraph database["<strong>License for database</strong>"]
   choose_license --> switch_license["<em>Depending on<br>content license</em>"]

### 2024-12-10 choose_license.qmd
Add references to licensing chapter

@@
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
+To facilitate self-archiving, one can also modify the contract with publishers via a rights retention statement [@UKRN2023] / author's addendum [@SPARC2006].
+If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
+
 If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
-
-If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
 :::
@@
 For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/2004/05/25/announcingandexplainingournew20licenses/#:~:text=Share%20Alike%20Across%20Borders)) and compatible licenses ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
+Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013).
 There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
+CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
 

### 2024-12-10 choose_license.qmd
Rework adding data set
- Use pointblank for the data dictionary
- Create data set automatically
- Make filename of data set all lowercase

As for the reason: I think it doesn't make
sense to use datawizard for the data
dictionary, as a proper dictionary should
contain information that cannot be just
"calculated" from the data object. For
example, the information about valid values
has to come from the user, rather than from
the data set itself.

@@
 ```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" Data.csv
+reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" data.csv
 ```
 
-You will notice that this creates another file called `Data.csv.license` containing the relevant information:
+You will notice that this creates another file called `data.csv.license` containing the relevant information:
 
-```{.yml filename="Data.csv.license"}
+```{.yml filename="data.csv.license"}
 SPDX-FileCopyrightText: 2024 Kristen Gorman

### 2024-12-10 make_readme.qmd
Rework adding data set
- Use pointblank for the data dictionary
- Create data set automatically
- Make filename of data set all lowercase

As for the reason: I think it doesn't make
sense to use datawizard for the data
dictionary, as a proper dictionary should
contain information that cannot be just
"calculated" from the data object. For
example, the information about valid values
has to come from the user, rather than from
the data set itself.

@@
 - `Bibliography.bib`: bibliographic references used in the manuscript
-- `Data.csv`: a data set containing the simplified `palmerpenguins` data created using `write.csv(palmerpenguins::penguins, "dat.csv", row.names = FALSE)`
+- `data.csv`: a data set containing the simplified `palmerpenguins` data created using `write.csv(palmerpenguins::penguins, "dat.csv", row.names = FALSE)`
 - `data_dictionary.html`: a dictionary to the data file, created using `create_data_dictionary.R`

### 2024-12-10 setup.qmd
Various minor stylistic and typographic improvements
Rework adding data set
- Use pointblank for the data dictionary
- Create data set automatically
- Make filename of data set all lowercase

As for the reason: I think it doesn't make
sense to use datawizard for the data
dictionary, as a proper dictionary should
contain information that cannot be just
"calculated" from the data object. For
example, the information about valid values
has to come from the user, rather than from
the data set itself.

@@
 
-Then, one can open an R session by simply typing `R` into the terminal.
-Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
-If not, set it using `setwd("code-publishing-exercise")`.
-Then, initialize `renv`:
+Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Then, initialize `renv`:
 
@@
 
-You can now download the data set we have prepared for you and put it into your project folder: [`Data.csv`](_01_Example/Data.csv){download=""}
+```{r}
+#| echo: false
+
+write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
+```
+
+You can now download the data set we have prepared for you and put it into your project folder: [`data.csv`](data.csv){download=""}
 
@@
 
-The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)^[Created with `palmerpenguins` v0.1.1 using `write.csv(palmerpenguins::penguins, file = "Data.csv", row.names = FALSE)`.] and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
+The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
 It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
@@
 
-When publishing a data set, it is important to document the meaning (e.g., units) and possible values of its variables.
+When distributing a data set, it is important to document the meaning (e.g., units) and valid values of its variables.
 This is typically done with a _data dictionary_ (also called a _codebook_).
-In the following, we will demonstrate how to create a simple data dictionary using the R packages [`tinylabels`](https://cran.r-project.org/package=tinylabels), [`datawizard`](https://easystats.github.io/datawizard/), and [`tinytable`](https://vincentarelbundock.github.io/tinytable/).
-You can install them now using:
+In the following, we will demonstrate how to create a simple data dictionary using the R package [`pointblank`](https://rstudio.github.io/pointblank/).
+You can install it now using:
+
+::: {.column-margin}
+![pointblank: Data Validation and Organization of Metadata for Local and Remote Tables](images/pointblank.png){width=250px}
+:::
 
 ```{.r filename="Console"}
-renv::install(c(
-  "tinylabels",
-  "datawizard",
-  "tinytable"
-))
+renv::install("pointblank")
 ```
@@
 
-Using `tinylabels` we can add labels to the variables of a `data.frame` in R:^[Note that the code provided does not alter the data file -- no labels will be added to `Data.csv`.
-The labels are only added to a (temporary) copy of the data set within R in order to create the data dictionary.]
+First, we write down everything we know about the data set. This includes:
 
-```{r, echo = -1, eval = -2, filename="create_data_dictionary.R"}
-dat <- read.csv(file.path("_01_Example", "Data.csv"))
-dat <- read.csv("Data.csv")
-
-descriptions <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-tinylabels::variable_label(dat) <- descriptions
-```
-
-Subsequently, `datawizard` can be employed to create the data dictionary containing the name and label, but also some other information about each variable:
-
-::: {.column-margin}
-![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
-:::
+- a general description of the data set
+- descriptions of all columns
+- valid values, where applicable
 
 ```{r, filename="create_data_dictionary.R"}
-(dict <- datawizard::data_codebook(dat) |>
-  subset(select = -.row_id) |>
-  tinytable::tt())
-```
-
-Finally, we can store the data dictionary inside an HTML file using the R package `tinytable` and put the HTML file into the project folder as well.
-
-::: {.column-margin}
-![tinytable: Simple and Configurable Tables](images/tinytable.svg){width=250px}
-:::
-
-```{.r filename="create_data_dictionary.R"}
-tinytable::save_tt(dict, output = "data_dictionary.html")
-```
-
-A human-readable data dictionary is necessary for making one's research reproducible and the example we provided demonstrates only the bare minimum.
-A full data documentation including measurement instruments, sampling procedures, appropriate weighting, contact information, and more information about the study can be created with the R package [`pointblank`](https://rstudio.github.io/pointblank/articles/INFO-1.html).
-And one could go even further by making the information machine-readable in a standardized way.
-We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
-
-::: {#nte-frictionless .callout-note collapse="true"}
-### Create Machine-Readable Variable Documentation
-
-This example demonstrates how the title and description of the data set, the description of the variables and their possible values are stored in a machine-readable way.
-
-```{.r filename="Console"}
 table_info <- c(
-  title = "penguins data set",
+  title = "palmerpenguins::penguins",
   description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
@@
 )
+
 vals <- list(
@@
 )
+```
+
+Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, or contact information.
+In this case, as the data have already been published, we only store a reference to its source:
+
+```{r, class.output = "r code-overflow-wrap", echo = FALSE}
+
+# We have provided the data set as CSV file to the readers.
+# Therefore, we cannot assume or require that readers have
+# the R package palmerpenguins installed. Instead, we create
+# the citation on our end and hide how we obtained it.
+
+dat_source <- citation("palmerpenguins", auto = TRUE) |>
+  format(bibtex = FALSE, style = "text")
+dat_source |>
+  gsub("\\n", " ", x = _) |>
+  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
+  paste0("dat_source <- \"", ... = _, "\"") |>
+  cat()
+```
+
+Then, we use `pointblank` to create a data dictionary with this information.
+
+```{r, filename="create_data_dictionary.R"}
+vals <- sapply(vals, \(x) {
+  paste0(
+    "(",
+    knitr::combine_words(x, and = " or ", before = "`", after = "`"),
+    ")"
+  )
+})
+
+dat <- read.csv("data.csv")
+
+dict <- pointblank::create_informant(
+  dat,
+  tbl_name = NA,
+  label = table_info[["title"]],
+  lang = "en"
+) |>
+  pointblank::info_tabular(
+    Description = table_info[["description"]],
+    Source = dat_source
+  ) |>
+  pointblank::info_columns_from_tbl(stack(descriptions)[2:1]) |>
+  pointblank::info_columns_from_tbl(stack(vals)[2:1]) |>
+  pointblank::get_informant_report(
+    title = "Data Dictionary for `data.csv`"
+  )
+
+dict
+```
+
+Finally, we can store the data dictionary inside an HTML file and put the HTML file into the project folder as well.
+
+```{.r filename="create_data_dictionary.R"}
+pointblank::export_report(dict, filename = "data_dictionary.html")
+```
+
+For a more elaborate introduction to `pointblank`, you can read their [Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html).
+
+One could go even further by making the information machine-readable in a standardized way.
+We provide an optional example of that in @nte-frictionless.
+If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
+
+::: {#nte-frictionless .callout-note collapse="true"}
+### Create Machine-Readable Variable Documentation
+
+This example demonstrates how the title and description of the data set, the description of the variables and their valid values are stored in a machine-readable way.
+As before, we also provide a reference to the source.
+
+```{r, class.output = "r code-overflow-wrap", echo = FALSE}
+
+# We have provided the data set as CSV file to the readers.
+# Therefore, we cannot assume or require that readers have
+# the R package palmerpenguins installed. Instead, we create
+# the citation on our end and hide how we obtained it.
+
+citation("palmerpenguins", auto = TRUE)$url |>
+  paste0("dat_source <- \"", ... = _, "\"") |>
+  cat()
 ```
@@
 # Read data and create schema
-dat_filename <- "Data.csv"
+dat_filename <- "data.csv"
 dat <- read.csv(dat_filename)
@@
 
-# Record possible values
+# Record valid values
 dat_schema$fields <- lapply(dat_schema$fields, \(x) {
@@
       title = "CC0 1.0 Universal"
+    )),
+    sources = list(list(
+      title = "CRAN",
+      path = dat_source
     ))
@@
 This creates the metadata file `datapackage.json` in the current directory.
-Make sure it is located in the same folder as `Data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
+Make sure it is located in the same folder as `data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
 :::

### 2024-12-13 choose_license.qmd
Don't mention SIL OFL anymore
While accepted by FSF and OSI, it would be
inconsistent to recommend a (somewhat)
non-commercial license here.

@@
 - __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
-If the intent is that only derived fonts, if published, have to be put under the same license, the [SIL Open Font License\ 1.1](https://openfontlicense.org/) is an appropriate choice.
-Note, however, that it doesn't require attribution for usage.
-If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
+As a non-copyleft license, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is a good choice for fonts.
 - __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].

### 2024-12-13 choose_license.qmd
Add SIL OFL again and remove CC0
SIL OFL is the de facto standard license
(e.g., the only accepted license by Google
Fonts and recommended by Fedora), while CC0
is not even mentioned.

@@
 - __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
-As a non-copyleft license, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is a good choice for fonts.
+The [SIL Open Font License\ 1.1](https://openfontlicense.org/) is the _de facto_ standard license for fonts, requiring that derived fonts, if published, have to be put under the same license.
+It doesn't require attribution for usage (e.g., when embedded in a document), but forbids that the font is sold by itself.
+Selling the font is allowed, however, if it is bundled with other software.
 - __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].

### 2024-12-15 choose_license.qmd
Mention full text and code disclosure requirements

@@
 Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
+- __Add license text__ means that recipients need to distribute the full text of the license along with the work (rather than only mentioning its license, for example, via a link).
 - __State changes__ means that recipients need to indicate if changes were made.
+- __Disclose code__ means that recipients need to disclose the source code when distributing modified versions of your work.
 - __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
@@
 
-  type --"Software"--> code_sa("Attribution?<br>State changes?<br>Copyleft?")
+  type --"Software"--> code_sa("Attribution?<br>Add license text?<br>State changes?<br>Disclose code?<br>Copyleft?")
   type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>State changes?<br>Anti-DRM?<br>Copyleft?")
 
-  code_sa --"Attribution &<br>State changes"--> apache["Apache 2.0"]
-  code_sa --"Attribution &<br>Weak copyleft"--> mpl["MPL 2.0"]
-  code_sa --"Attribution &<br>State changes &<br>Strong copyleft"--> agpl["AGPLv3"]
+  code_sa --"Attribution &<br>Add license text &<br>State changes"--> apache["Apache 2.0"]
+  code_sa --"Attribution &<br>Disclose code &<br>Weak copyleft"--> mpl["MPL 2.0"]
+  code_sa --"Attribution &<br>Add license text &<br>State changes &<br>Disclose code &<br>Strong copyleft"--> agpl["AGPLv3"]
 

### 2025-01-16 about.qmd
Improve mention of RStudio's license

@@
 
-The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are available under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/gpl-3.0.html).
+The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are available under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/agpl-3.0.html).
+The source code is available on [GitHub](https://github.com/rstudio/rstudio/tree/v2024.04.2%2B764).
 
@@
 
-This program is free software: you can redistribute it and/or modify it under the terms of version 3 of the GNU Affero General Public License as published by the Free Software Foundation.
+Unless you have received this program directly from Posit Software pursuant to the terms of a commercial license agreement with Posit Software, then RStudio is licensed to you under the AGPLv3, the terms of which are included below.
+Details on obtaining the source code for this distribution are included in the file SOURCE.
 
-This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
-See the GNU Affero General Public License for more details.
+RStudio includes other open source software whose license terms can be found in the file NOTICE.
+RStudio for Windows is bundled with binary copies of several utilities, whose license terms can be found in the file SOURCE.
 
-You should have received a copy of the GNU Affero General Public License along with this program.  If not, see <https://www.gnu.org/licenses/>.
+                    GNU AFFERO GENERAL PUBLIC LICENSE
+                       Version 3, 19 November 2007
+
+ Copyright (C) 2007 Free Software Foundation, Inc. <http://fsf.org/>
+ Everyone is permitted to copy and distribute verbatim copies
+ of this license document, but changing it is not allowed.
+
+                            Preamble
+
+  The GNU Affero General Public License is a free, copyleft license for
+software and other kinds of works, specifically designed to ensure
+cooperation with the community in the case of network server software.
+
+  The licenses for most software and other practical works are designed
+to take away your freedom to share and change the works.  By contrast,
+our General Public Licenses are intended to guarantee your freedom to
+share and change all versions of a program--to make sure it remains free
+software for all its users.
+
+  When we speak of free software, we are referring to freedom, not
+price.  Our General Public Licenses are designed to make sure that you
+have the freedom to distribute copies of free software (and charge for
+them if you wish), that you receive source code or can get it if you
+want it, that you can change the software or use pieces of it in new
+free programs, and that you know you can do these things.
+
+  Developers that use our General Public Licenses protect your rights
+with two steps: (1) assert copyright on the software, and (2) offer
+you this License which gives you legal permission to copy, distribute
+and/or modify the software.
+
+  A secondary benefit of defending all users' freedom is that
+improvements made in alternate versions of the program, if they
+receive widespread use, become available for other developers to
+incorporate.  Many developers of free software are heartened and
+encouraged by the resulting cooperation.  However, in the case of
+software used on network servers, this result may fail to come about.
+The GNU General Public License permits making a modified version and
+letting the public access it on a server without ever releasing its
+source code to the public.
+
+  The GNU Affero General Public License is designed specifically to
+ensure that, in such cases, the modified source code becomes available
+to the community.  It requires the operator of a network server to
+provide the source code of the modified version running there to the
+users of that server.  Therefore, public use of a modified version, on
+a publicly accessible server, gives the public access to the source
+code of the modified version.
+
+  An older license, called the Affero General Public License and
+published by Affero, was designed to accomplish similar goals.  This is
+a different license, not a version of the Affero GPL, but Affero has
+released a new version of the Affero GPL which permits relicensing under
+this license.
+
+  The precise terms and conditions for copying, distribution and
+modification follow.
+
+                       TERMS AND CONDITIONS
+
+  0. Definitions.
+
+  "This License" refers to version 3 of the GNU Affero General Public License.
+
+  "Copyright" also means copyright-like laws that apply to other kinds of
+works, such as semiconductor masks.
+
+  "The Program" refers to any copyrightable work licensed under this
+License.  Each licensee is addressed as "you".  "Licensees" and
+"recipients" may be individuals or organizations.
+
+  To "modify" a work means to copy from or adapt all or part of the work
+in a fashion requiring copyright permission, other than the making of an
+exact copy.  The resulting work is called a "modified version" of the
+earlier work or a work "based on" the earlier work.
+
+  A "covered work" means either the unmodified Program or a work based
+on the Program.
+
+  To "propagate" a work means to do anything with it that, without
+permission, would make you directly or secondarily liable for
+infringement under applicable copyright law, except executing it on a
+computer or modifying a private copy.  Propagation includes copying,
+distribution (with or without modification), making available to the
+public, and in some countries other activities as well.
+
+  To "convey" a work means any kind of propagation that enables other
+parties to make or receive copies.  Mere interaction with a user through
+a computer network, with no transfer of a copy, is not conveying.
+
+  An interactive user interface displays "Appropriate Legal Notices"
+to the extent that it includes a convenient and prominently visible
+feature that (1) displays an appropriate copyright notice, and (2)
+tells the user that there is no warranty for the work (except to the
+extent that warranties are provided), that licensees may convey the
+work under this License, and how to view a copy of this License.  If
+the interface presents a list of user commands or options, such as a
+menu, a prominent item in the list meets this criterion.
+
+  1. Source Code.
+
+  The "source code" for a work means the preferred form of the work
+for making modifications to it.  "Object code" means any non-source
+form of a work.
+
+  A "Standard Interface" means an interface that either is an official
+standard defined by a recognized standards body, or, in the case of
+interfaces specified for a particular programming language, one that
+is widely used among developers working in that language.
+
+  The "System Libraries" of an executable work include anything, other
+than the work as a whole, that (a) is included in the normal form of
+packaging a Major Component, but which is not part of that Major
+Component, and (b) serves only to enable use of the work with that
+Major Component, or to implement a Standard Interface for which an
+implementation is available to the public in source code form.  A
+"Major Component", in this context, means a major essential component
+(kernel, window system, and so on) of the specific operating system
+(if any) on which the executable work runs, or a compiler used to
+produce the work, or an object code interpreter used to run it.
+
+  The "Corresponding Source" for a work in object code form means all
+the source code needed to generate, install, and (for an executable
+work) run the object code and to modify the work, including scripts to
+control those activities.  However, it does not include the work's
+System Libraries, or general-purpose tools or generally available free
+programs which are used unmodified in performing those activities but
+which are not part of the work.  For example, Corresponding Source
+includes interface definition files associated with source files for
+the work, and the source code for shared libraries and dynamically
+linked subprograms that the work is specifically designed to require,
+such as by intimate data communication or control flow between those
+subprograms and other parts of the work.
+
+  The Corresponding Source need not include anything that users
+can regenerate automatically from other parts of the Corresponding
+Source.
+
+  The Corresponding Source for a work in source code form is that
+same work.
+
+  2. Basic Permissions.
+
+  All rights granted under this License are granted for the term of
+copyright on the Program, and are irrevocable provided the stated
+conditions are met.  This License explicitly affirms your unlimited
+permission to run the unmodified Program.  The output from running a
+covered work is covered by this License only if the output, given its
+content, constitutes a covered work.  This License acknowledges your
+rights of fair use or other equivalent, as provided by copyright law.
+
+  You may make, run and propagate covered works that you do not
+convey, without conditions so long as your license otherwise remains
+in force.  You may convey covered works to others for the sole purpose
+of having them make modifications exclusively for you, or provide you
+with facilities for running those works, provided that you comply with
+the terms of this License in conveying all material for which you do
+not control copyright.  Those thus making or running the covered works
+for you must do so exclusively on your behalf, under your direction
+and control, on terms that prohibit them from making any copies of
+your copyrighted material outside their relationship with you.
+
+  Conveying under any other circumstances is permitted solely under
+the conditions stated below.  Sublicensing is not allowed; section 10
+makes it unnecessary.
+
+  3. Protecting Users' Legal Rights From Anti-Circumvention Law.
+
+  No covered work shall be deemed part of an effective technological
+measure under any applicable law fulfilling obligations under article
+11 of the WIPO copyright treaty adopted on 20 December 1996, or
+similar laws prohibiting or restricting circumvention of such
+measures.
+
+  When you convey a covered work, you waive any legal power to forbid
+circumvention of technological measures to the extent such circumvention
+is effected by exercising rights under this License with respect to
+the covered work, and you disclaim any intention to limit operation or
+modification of the work as a means of enforcing, against the work's
+users, your or third parties' legal rights to forbid circumvention of
+technological measures.
+
+  4. Conveying Verbatim Copies.
+
+  You may convey verbatim copies of the Program's source code as you
+receive it, in any medium, provided that you conspicuously and
+appropriately publish on each copy an appropriate copyright notice;
+keep intact all notices stating that this License and any
+non-permissive terms added in accord with section 7 apply to the code;
+keep intact all notices of the absence of any warranty; and give all
+recipients a copy of this License along with the Program.
+
+  You may charge any price or no price for each copy that you convey,
+and you may offer support or warranty protection for a fee.
+
+  5. Conveying Modified Source Versions.
+
+  You may convey a work based on the Program, or the modifications to
+produce it from the Program, in the form of source code under the
+terms of section 4, provided that you also meet all of these conditions:
+
+    a) The work must carry prominent notices stating that you modified
+    it, and giving a relevant date.
+
+    b) The work must carry prominent notices stating that it is
+    released under this License and any conditions added under section
+    7.  This requirement modifies the requirement in section 4 to
+    "keep intact all notices".
+
+    c) You must license the entire work, as a whole, under this
+    License to anyone who comes into possession of a copy.  This
+    License will therefore apply, along with any applicable section 7
+    additional terms, to the whole of the work, and all its parts,
+    regardless of how they are packaged.  This License gives no
+    permission to license the work in any other way, but it does not
+    invalidate such permission if you have separately received it.
+
+    d) If the work has interactive user interfaces, each must display
+    Appropriate Legal Notices; however, if the Program has interactive
+    interfaces that do not display Appropriate Legal Notices, your
+    work need not make them do so.
+
+  A compilation of a covered work with other separate and independent
+works, which are not by their nature extensions of the covered work,
+and which are not combined with it such as to form a larger program,
+in or on a volume of a storage or distribution medium, is called an
+"aggregate" if the compilation and its resulting copyright are not
+used to limit the access or legal rights of the compilation's users
+beyond what the individual works permit.  Inclusion of a covered work
+in an aggregate does not cause this License to apply to the other
+parts of the aggregate.
+
+  6. Conveying Non-Source Forms.
+
+  You may convey a covered work in object code form under the terms
+of sections 4 and 5, provided that you also convey the
+machine-readable Corresponding Source under the terms of this License,
+in one of these ways:
+
+    a) Convey the object code in, or embodied in, a physical product
+    (including a physical distribution medium), accompanied by the
+    Corresponding Source fixed on a durable physical medium
+    customarily used for software interchange.
+
+    b) Convey the object code in, or embodied in, a physical product
+    (including a physical distribution medium), accompanied by a
+    written offer, valid for at least three years and valid for as
+    long as you offer spare parts or customer support for that product
+    model, to give anyone who possesses the object code either (1) a
+    copy of the Corresponding Source for all the software in the
+    product that is covered by this License, on a durable physical
+    medium customarily used for software interchange, for a price no
+    more than your reasonable cost of physically performing this
+    conveying of source, or (2) access to copy the
+    Corresponding Source from a network server at no charge.
+
+    c) Convey individual copies of the object code with a copy of the
+    written offer to provide the Corresponding Source.  This
+    alternative is allowed only occasionally and noncommercially, and
+    only if you received the object code with such an offer, in accord
+    with subsection 6b.
+
+    d) Convey the object code by offering access from a designated
+    place (gratis or for a charge), and offer equivalent access to the
+    Corresponding Source in the same way through the same place at no
+    further charge.  You need not require recipients to copy the
+    Corresponding Source along with the object code.  If the place to
+    copy the object code is a network server, the Corresponding Source
+    may be on a different server (operated by you or a third party)
+    that supports equivalent copying facilities, provided you maintain
+    clear directions next to the object code saying where to find the
+    Corresponding Source.  Regardless of what server hosts the
+    Corresponding Source, you remain obligated to ensure that it is
+    available for as long as needed to satisfy these requirements.
+
+    e) Convey the object code using peer-to-peer transmission, provided
+    you inform other peers where the object code and Corresponding
+    Source of the work are being offered to the general public at no
+    charge under subsection 6d.
+
+  A separable portion of the object code, whose source code is excluded
+from the Corresponding Source as a System Library, need not be
+included in conveying the object code work.
+
+  A "User Product" is either (1) a "consumer product", which means any
+tangible personal property which is normally used for personal, family,
+or household purposes, or (2) anything designed or sold for incorporation
+into a dwelling.  In determining whether a product is a consumer product,
+doubtful cases shall be resolved in favor of coverage.  For a particular
+product received by a particular user, "normally used" refers to a
+typical or common use of that class of product, regardless of the status
+of the particular user or of the way in which the particular user
+actually uses, or expects or is expected to use, the product.  A product
+is a consumer product regardless of whether the product has substantial
+commercial, industrial or non-consumer uses, unless such uses represent
+the only significant mode of use of the product.
+
+  "Installation Information" for a User Product means any methods,
+procedures, authorization keys, or other information required to install
+and execute modified versions of a covered work in that User Product from
+a modified version of its Corresponding Source.  The information must
+suffice to ensure that the continued functioning of the modified object
+code is in no case prevented or interfered with solely because
+modification has been made.
+
+  If you convey an object code work under this section in, or with, or
+specifically for use in, a User Product, and the conveying occurs as
+part of a transaction in which the right of possession and use of the
+User Product is transferred to the recipient in perpetuity or for a
+fixed term (regardless of how the transaction is characterized), the
+Corresponding Source conveyed under this section must be accompanied
+by the Installation Information.  But this requirement does not apply
+if neither you nor any third party retains the ability to install
+modified object code on the User Product (for example, the work has
+been installed in ROM).
+
+  The requirement to provide Installation Information does not include a
+requirement to continue to provide support service, warranty, or updates
+for a work that has been modified or installed by the recipient, or for
+the User Product in which it has been modified or installed.  Access to a
+network may be denied when the modification itself materially and
+adversely affects the operation of the network or violates the rules and
+protocols for communication across the network.
+
+  Corresponding Source conveyed, and Installation Information provided,
+in accord with this section must be in a format that is publicly
+documented (and with an implementation available to the public in
+source code form), and must require no special password or key for
+unpacking, reading or copying.
+
+  7. Additional Terms.
+
+  "Additional permissions" are terms that supplement the terms of this
+License by making exceptions from one or more of its conditions.
+Additional permissions that are applicable to the entire Program shall
+be treated as though they were included in this License, to the extent
+that they are valid under applicable law.  If additional permissions
+apply only to part of the Program, that part may be used separately
+under those permissions, but the entire Program remains governed by
+this License without regard to the additional permissions.
+
+  When you convey a copy of a covered work, you may at your option
+remove any additional permissions from that copy, or from any part of
+it.  (Additional permissions may be written to require their own
+removal in certain cases when you modify the work.)  You may place
+additional permissions on material, added by you to a covered work,
+for which you have or can give appropriate copyright permission.
+
+  Notwithstanding any other provision of this License, for material you
+add to a covered work, you may (if authorized by the copyright holders of
+that material) supplement the terms of this License with terms:
+
+    a) Disclaiming warranty or limiting liability differently from the
+    terms of sections 15 and 16 of this License; or
+
+    b) Requiring preservation of specified reasonable legal notices or
+    author attributions in that material or in the Appropriate Legal
+    Notices displayed by works containing it; or
+
+    c) Prohibiting misrepresentation of the origin of that material, or
+    requiring that modified versions of such material be marked in
+    reasonable ways as different from the original version; or
+
+    d) Limiting the use for publicity purposes of names of licensors or
+    authors of the material; or
+
+    e) Declining to grant rights under trademark law for use of some
+    trade names, trademarks, or service marks; or
+
+    f) Requiring indemnification of licensors and authors of that
+    material by anyone who conveys the material (or modified versions of
+    it) with contractual assumptions of liability to the recipient, for
+    any liability that these contractual assumptions directly impose on
+    those licensors and authors.
+
+  All other non-permissive additional terms are considered "further
+restrictions" within the meaning of section 10.  If the Program as you
+received it, or any part of it, contains a notice stating that it is
+governed by this License along with a term that is a further
+restriction, you may remove that term.  If a license document contains
+a further restriction but permits relicensing or conveying under this
+License, you may add to a covered work material governed by the terms
+of that license document, provided that the further restriction does
+not survive such relicensing or conveying.
+
+  If you add terms to a covered work in accord with this section, you
+must place, in the relevant source files, a statement of the
+additional terms that apply to those files, or a notice indicating
+where to find the applicable terms.
+
+  Additional terms, permissive or non-permissive, may be stated in the
+form of a separately written license, or stated as exceptions;
+the above requirements apply either way.
+
+  8. Termination.
+
+  You may not propagate or modify a covered work except as expressly
+provided under this License.  Any attempt otherwise to propagate or
+modify it is void, and will automatically terminate your rights under
+this License (including any patent licenses granted under the third
+paragraph of section 11).
+
+  However, if you cease all violation of this License, then your
+license from a particular copyright holder is reinstated (a)
+provisionally, unless and until the copyright holder explicitly and
+finally terminates your license, and (b) permanently, if the copyright
+holder fails to notify you of the violation by some reasonable means
+prior to 60 days after the cessation.
+
+  Moreover, your license from a particular copyright holder is
+reinstated permanently if the copyright holder notifies you of the
+violation by some reasonable means, this is the first time you have
+received notice of violation of this License (for any work) from that
+copyright holder, and you cure the violation prior to 30 days after
+your receipt of the notice.
+
+  Termination of your rights under this section does not terminate the
+licenses of parties who have received copies or rights from you under
+this License.  If your rights have been terminated and not permanently
+reinstated, you do not qualify to receive new licenses for the same
+material under section 10.
+
+  9. Acceptance Not Required for Having Copies.
+
+  You are not required to accept this License in order to receive or
+run a copy of the Program.  Ancillary propagation of a covered work
+occurring solely as a consequence of using peer-to-peer transmission
+to receive a copy likewise does not require acceptance.  However,
+nothing other than this License grants you permission to propagate or
+modify any covered work.  These actions infringe copyright if you do
+not accept this License.  Therefore, by modifying or propagating a
+covered work, you indicate your acceptance of this License to do so.
+
+  10. Automatic Licensing of Downstream Recipients.
+
+  Each time you convey a covered work, the recipient automatically
+receives a license from the original licensors, to run, modify and
+propagate that work, subject to this License.  You are not responsible
+for enforcing compliance by third parties with this License.
+
+  An "entity transaction" is a transaction transferring control of an
+organization, or substantially all assets of one, or subdividing an
+organization, or merging organizations.  If propagation of a covered
+work results from an entity transaction, each party to that
+transaction who receives a copy of the work also receives whatever
+licenses to the work the party's predecessor in interest had or could
+give under the previous paragraph, plus a right to possession of the
+Corresponding Source of the work from the predecessor in interest, if
+the predecessor has it or can get it with reasonable efforts.
+
+  You may not impose any further restrictions on the exercise of the
+rights granted or affirmed under this License.  For example, you may
+not impose a license fee, royalty, or other charge for exercise of
+rights granted under this License, and you may not initiate litigation
+(including a cross-claim or counterclaim in a lawsuit) alleging that
+any patent claim is infringed by making, using, selling, offering for
+sale, or importing the Program or any portion of it.
+
+  11. Patents.
+
+  A "contributor" is a copyright holder who authorizes use under this
+License of the Program or a work on which the Program is based.  The
+work thus licensed is called the contributor's "contributor version".
+
+  A contributor's "essential patent claims" are all patent claims
+owned or controlled by the contributor, whether already acquired or
+hereafter acquired, that would be infringed by some manner, permitted
+by this License, of making, using, or selling its contributor version,
+but do not include claims that would be infringed only as a
+consequence of further modification of the contributor version.  For
+purposes of this definition, "control" includes the right to grant
+patent sublicenses in a manner consistent with the requirements of
+this License.
+
+  Each contributor grants you a non-exclusive, worldwide, royalty-free
+patent license under the contributor's essential patent claims, to
+make, use, sell, offer for sale, import and otherwise run, modify and
+propagate the contents of its contributor version.
+
+  In the following three paragraphs, a "patent license" is any express
+agreement or commitment, however denominated, not to enforce a patent
+(such as an express permission to practice a patent or covenant not to
+sue for patent infringement).  To "grant" such a patent license to a
+party means to make such an agreement or commitment not to enforce a
+patent against the party.
+
+  If you convey a covered work, knowingly relying on a patent license,
+and the Corresponding Source of the work is not available for anyone
+to copy, free of charge and under the terms of this License, through a
+publicly available network server or other readily accessible means,
+then you must either (1) cause the Corresponding Source to be so
+available, or (2) arrange to deprive yourself of the benefit of the
+patent license for this particular work, or (3) arrange, in a manner
+consistent with the requirements of this License, to extend the patent
+license to downstream recipients.  "Knowingly relying" means you have
+actual knowledge that, but for the patent license, your conveying the
+covered work in a country, or your recipient's use of the covered work
+in a country, would infringe one or more identifiable patents in that
+country that you have reason to believe are valid.
+
+  If, pursuant to or in connection with a single transaction or
+arrangement, you convey, or propagate by procuring conveyance of, a
+covered work, and grant a patent license to some of the parties
+receiving the covered work authorizing them to use, propagate, modify
+or convey a specific copy of the covered work, then the patent license
+you grant is automatically extended to all recipients of the covered
+work and works based on it.
+
+  A patent license is "discriminatory" if it does not include within
+the scope of its coverage, prohibits the exercise of, or is
+conditioned on the non-exercise of one or more of the rights that are
+specifically granted under this License.  You may not convey a covered
+work if you are a party to an arrangement with a third party that is
+in the business of distributing software, under which you make payment
+to the third party based on the extent of your activity of conveying
+the work, and under which the third party grants, to any of the
+parties who would receive the covered work from you, a discriminatory
+patent license (a) in connection with copies of the covered work
+conveyed by you (or copies made from those copies), or (b) primarily
+for and in connection with specific products or compilations that
+contain the covered work, unless you entered into that arrangement,
+or that patent license was granted, prior to 28 March 2007.
+
+  Nothing in this License shall be construed as excluding or limiting
+any implied license or other defenses to infringement that may
+otherwise be available to you under applicable patent law.
+
+  12. No Surrender of Others' Freedom.
+
+  If conditions are imposed on you (whether by court order, agreement or
+otherwise) that contradict the conditions of this License, they do not
+excuse you from the conditions of this License.  If you cannot convey a
+covered work so as to satisfy simultaneously your obligations under this
+License and any other pertinent obligations, then as a consequence you may
+not convey it at all.  For example, if you agree to terms that obligate you
+to collect a royalty for further conveying from those to whom you convey
+the Program, the only way you could satisfy both those terms and this
+License would be to refrain entirely from conveying the Program.
+
+  13. Remote Network Interaction; Use with the GNU General Public License.
+
+  Notwithstanding any other provision of this License, if you modify the
+Program, your modified version must prominently offer all users
+interacting with it remotely through a computer network (if your version
+supports such interaction) an opportunity to receive the Corresponding
+Source of your version by providing access to the Corresponding Source
+from a network server at no charge, through some standard or customary
+means of facilitating copying of software.  This Corresponding Source
+shall include the Corresponding Source for any work covered by version 3
+of the GNU General Public License that is incorporated pursuant to the
+following paragraph.
+
+  Notwithstanding any other provision of this License, you have
+permission to link or combine any covered work with a work licensed
+under version 3 of the GNU General Public License into a single
+combined work, and to convey the resulting work.  The terms of this
+License will continue to apply to the part which is the covered work,
+but the work with which it is combined will remain governed by version
+3 of the GNU General Public License.
+
+  14. Revised Versions of this License.
+
+  The Free Software Foundation may publish revised and/or new versions of
+the GNU Affero General Public License from time to time.  Such new versions
+will be similar in spirit to the present version, but may differ in detail to
+address new problems or concerns.
+
+  Each version is given a distinguishing version number.  If the
+Program specifies that a certain numbered version of the GNU Affero General
+Public License "or any later version" applies to it, you have the
+option of following the terms and conditions either of that numbered
+version or of any later version published by the Free Software
+Foundation.  If the Program does not specify a version number of the
+GNU Affero General Public License, you may choose any version ever published
+by the Free Software Foundation.
+
+  If the Program specifies that a proxy can decide which future
+versions of the GNU Affero General Public License can be used, that proxy's
+public statement of acceptance of a version permanently authorizes you
+to choose that version for the Program.
+
+  Later license versions may give you additional or different
+permissions.  However, no additional obligations are imposed on any
+author or copyright holder as a result of your choosing to follow a
+later version.
+
+  15. Disclaimer of Warranty.
+
+  THERE IS NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY
+APPLICABLE LAW.  EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT
+HOLDERS AND/OR OTHER PARTIES PROVIDE THE PROGRAM "AS IS" WITHOUT WARRANTY
+OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO,
+THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM
+IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST OF
+ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+
+  16. Limitation of Liability.
+
+  IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
+WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MODIFIES AND/OR CONVEYS
+THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR DAMAGES, INCLUDING ANY
+GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL DAMAGES ARISING OUT OF THE
+USE OR INABILITY TO USE THE PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF
+DATA OR DATA BEING RENDERED INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD
+PARTIES OR A FAILURE OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS),
+EVEN IF SUCH HOLDER OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF
+SUCH DAMAGES.
+
+  17. Interpretation of Sections 15 and 16.
+
+  If the disclaimer of warranty and limitation of liability provided
+above cannot be given local legal effect according to their terms,
+reviewing courts shall apply local law that most closely approximates
+an absolute waiver of all civil liability in connection with the
+Program, unless a warranty or assumption of liability accompanies a
+copy of the Program in return for a fee.
+
+                     END OF TERMS AND CONDITIONS
+
+            How to Apply These Terms to Your New Programs
+
+  If you develop a new program, and you want it to be of the greatest
+possible use to the public, the best way to achieve this is to make it
+free software which everyone can redistribute and change under these terms.
+
+  To do so, attach the following notices to the program.  It is safest
+to attach them to the start of each source file to most effectively
+state the exclusion of warranty; and each file should have at least
+the "copyright" line and a pointer to where the full notice is found.
+
+    <one line to give the program's name and a brief idea of what it does.>
+    Copyright (C) <year>  <name of author>
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Affero General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Affero General Public License for more details.
+
+    You should have received a copy of the GNU Affero General Public License
+    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+
+Also add information on how to contact you by electronic and paper mail.
+
+  If your software can interact with users remotely through a computer
+network, you should also make sure that it provides a way for users to
+get its source.  For example, if your program is a web application, its
+interface could display a "Source" link that leads users to an archive
+of the code.  There are many ways you could offer source, and different
+solutions will be better for different programs; see section 13 for the
+specific requirements.
+
+  You should also get your employer (if you work as a programmer) or school,
+if any, to sign a "copyright disclaimer" for the program, if necessary.
+For more information on this, and how to apply and follow the GNU AGPL, see
+<http://www.gnu.org/licenses/>.
 ```

### 2025-01-17 choose_license.qmd
Shorten section on other restrictions
- Make the argument more succinct
- Remove the references as they are
  not that good

@@
 As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
-There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023; @Vezina2020; @WikimediaCommons2009], but we would like to highlight one reason in particular.
-
-With the rights provided by free/open licenses comes the possibility to build on and combine multiple works by different authors, which is essential for any cultural and scientific activity.
-However, this is only possible if the various licenses involved are compatible with each other.
-For example, one is allowed to remix two figures if the first is licensed under CC\ BY\ 4.0 and the second under CC\ BY-SA\ 4.0 because the licenses were written to be compatible with each other [compare @CC2024FAQ].
-Also note that CC\ BY-SA\ 4.0 is [one-way compatible](https://creativecommons.org/compatiblelicenses) with the GPLv3, which in turn is compatible with the AGPLv3.
-
-However, applying a restriction such as only permitting non-commercial use or not allowing derivative works creates _silos_ of works which are mutually incompatible with each other.
-Put differently, one cannot share a remix of two works where one is licensed under CC\ BY-SA\ 4.0 and the other only allows non-commercial use.
+Otherwise, creating adaptations from multiple works by different authors would either limit commercial use for everybody (including the authors) or would not be possible at all (when including materials with a copyleft license), effectively creating _silos_ of works which are mutually incompatible with each other.
+This is especially relevant for work types for which adaptations are created frequently, such as text, data, and source code.
 In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].

### 2025-01-22 choose_license.qmd
Simplify after consultation

--- 2024-12-13 choose_license.qmd
@@
 
-A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss are non-exclusive licenses and do not limit the author or copyright holder in their rights.
-And while you _could_ write your own license, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For researchers at LMU
+
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
+Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
+:::
+
+A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss do not limit the author or copyright holder in their rights.
+And while you _could_ write your own, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
 By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
@@
 
-::: {#wrn-exclusive-rights .callout-warning}
+::: {#cau-exclusive-rights .callout-caution}
 ### Giving Publishers Exclusive Rights
 
-Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you.
-This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
+When you have your article published under a Creative Commons license, some publishers still demand an exclusive publishing and distribution license or a copyright assignment.
+This would give them more rights than the readers of the article have through the chosen Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
 Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
-To facilitate self-archiving, one can also modify the contract with publishers via a rights retention statement [@UKRN2023] / author's addendum [@SPARC2006].
+If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
+
 If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
-
-If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
 :::
@@
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.
+Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) -- otherwise, consult @nte-other-work-types for some caveats.
 
@@
 
-One should take special care when applying license to the following types of works:
+One should be cautious about the restrictions of licenses applied to the following types of works:
 
 - __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
-The [SIL Open Font License\ 1.1](https://openfontlicense.org/) is the _de facto_ standard license for fonts, requiring that derived fonts, if published, have to be put under the same license.
-It doesn't require attribution for usage (e.g., when embedded in a document), but forbids that the font is sold by itself.
-Selling the font is allowed, however, if it is bundled with other software.
-- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].
-And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
+If the intent is that only derived fonts, if published, have to be put under the same license, the [SIL Open Font License\ 1.1](https://openfontlicense.org/) is an appropriate choice.
+Note, however, that it doesn't require attribution for usage.
+If no copyleft mechanism is intended, [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) also works for fonts.
+- __templates and LaTeX packages:__ If a template or, for this purpose, a LaTeX package, is licensed under a copyleft software license such as the [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/), every work that is a derivative of the template or that uses the LaTeX package has to put under the same license, if shared [@Koppor2016].
+And if a document contained source code covered by the GPLv3, the same license would also apply to the document [@MadHatter2022].
 - __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
 For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017].
-Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared.
+Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to licensed under CC\ BY-SA\ 4.0 (or a compatible license) if shared.
 If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
 Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
-- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
+- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or even another license).
 :::
@@
 - __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
-Implementing such restrictive measures is also referred to as **d**igital **r**ights **m**anagement (DRM).
+DRM is an abbreviation for **d**igital **r**ights **m**anagement.
 
@@
 As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
-There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023; @Vezina2020; @WikimediaCommons2009], but we would like to highlight one reason in particular.
+There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023], but we would like to highlight one reason in particular.
 
@@
 Put differently, one cannot share a remix of two works where one is licensed under CC\ BY-SA\ 4.0 and the other only allows non-commercial use.
-In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
+In order to avoid silos, one should only choose license which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
 If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012.
@@
 
-!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024] licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)](images/Morrison2024.jpg){.lightbox}
+!["Open data-capable license interoperability" by Robbie Morrison licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
+Taken without modification from @Morrison2024](images/Morrison2024.jpg){.lightbox}
 :::
@@
 Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which file or folder is covered by which license, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
-Notably, no sort of registration is required.
+Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
 
@@
 For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013).
+Since then, the possibility to relicense under compatible licenses has been added ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)), a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)).
 There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
-CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
 
@@
 ```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" data.csv
+reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" Data.csv
 ```
 
-You will notice that this creates another file called `data.csv.license` containing the relevant information:
+You will notice that this creates another file called `Data.csv.license` containing the relevant information:
 
-```{.yml filename="data.csv.license"}
+```{.yml filename="Data.csv.license"}
 SPDX-FileCopyrightText: 2024 Kristen Gorman
@@
 
-  subgraph content["<strong>License for content</strong>"]
+  subgraph "License for individual data entries (content)"
   existing_license_content --"No, we created the content<br>entirely by ourselves."--> facts("Entries are facts<br>(like measurements<br>or metadata)?")
@@
 
-  subgraph database["<strong>License for database</strong>"]
-  choose_license --> switch_license["<em>Depending on<br>content license</em>"]
+  subgraph "License for combination of data (database)"
+  choose_license --> switch_license["<em>Depending on content license</em>"]
   use_existing_license_content --> switch_license

### 2025-01-22 choose_license.qmd

--- 2025-01-17 choose_license.qmd
@@
 Copyright licenses enable authors to free up their works for reuse by others.
+
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For researchers at LMU
+
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
+Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
+:::
 

### 2025-01-22 index.qmd

--- 2024-11-27 index.qmd
@@
 
-Quarto is a tool to create reproducible documents, but not the only one.
-You may also use [RMarkdown](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), or one of the countless others.
+Quarto is a tool to render dynamic documents, but not the only one.
+You may also use [`rmarkdown`](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), [Org Mode](https://orgmode.org/), [StatTag](https://sites.northwestern.edu/stattag/), or one of the countless others.
 :::

### 2025-01-22 intro.qmd
Provide exact page reference

@@
 However, with only the article available, it could take months to come up with a working solution, if possible at all.
-This situation has been put aptly by @Buckheit1995, distilling an idea by geophysicist Jon Claerbout:
+This situation has been put aptly by Buckheit & Donoho [-@Buckheit1995, p.
+59], distilling an idea by the geophysicist Jon Claerbout:
 
-> "An article about computational science in a scientific publication is not the scholarship itself, it is merely advertising of the scholarship.
+> "An article about computational science in a scientific publication is __not__ the scholarship itself, it is merely __advertising__ of the scholarship.
 The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."

### 2025-01-25 about.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

@@
     orcid: "0000-0003-0374-5625"
-  - name: "Josephine Zerna"
-    orcid: "0000-0003-2892-884X"
-  - name: "Christoph Scheffel"
-    email: "christoph_scheffel@tu-dresden.de"
-    orcid: "0000-0001-5963-9229"
 bibliography: grateful-refs.bib
@@
 
-Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
 None of the discussion in this tutorial constitutes legal advice.
+
+Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code chunks are also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) or the [Apache License\ 2.0](https://www.apache.org/licenses/LICENSE-2.0).
+
+The file `Manuscript.qmd` is made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
 

### 2025-01-25 archive.qmd
Use APA citation style
Use global language and tidy _quarto.yml
Mention .gitignore files during archival
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

@@
 title: "Archive Your Project"
-lang: "en"
-bibliography: literature.bib
 ---
@@
 
-The project is now ready to be deposited for long-term archival.
+The project folder is now ready to be deposited for long-term archival.
 When considering an appropriate repository, there are a few questions to ponder [excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
@@
 
-## Using Zenodo to Archive Your Project
+## Using Zenodo to Archive Your Project Folder
 
-In the following we will demonstrate how to use Zenodo to archive one's project and obtain a DOI.
+In the following we will demonstrate how to use Zenodo to archive one's project folder and obtain a DOI.
 
@@
 - __Add/remove files__: To circumvent the file limit on Zenodo, you can bundle your project folder in one ZIP file and upload that instead.
+Be sure to exclude all files mentioned in your `.gitignore` file!
 - __Fill in metadata__
@@
   - __Resource type__: Select the most appropriate one, for example, _Publication / Preprint_.
-  - __Licenses__: If your project only contains content by you and if you previously only chose one license, you can select that by choosing _Add standard_.
+  - __Licenses__: If your project folder only contains content by you and if you previously only chose one license, you can select that by choosing _Add standard_.
 Otherwise, click on _Add custom_ and describe how each file is licensed.
@@
 
-If you previously uploaded your project to a public Git repository (which we didn't cover during this tutorial), you could also add the new DOI to your README, for example, under the section "Citation".
+If you previously uploaded your project folder to a public Git repository (which we didn't cover during this tutorial), you could also add the new DOI to your README, for example, under the section "Citation".
 

### 2025-01-25 choose_license.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

@@
 
-## TL;DR
-
-Every expression of an idea, such as a literary or an artistic works, is automatically covered by copyright.
-Put simply, this means that nobody else than the copyright holder is allowed to copy, modify, or share it.
-Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
-Certain exceptions and limitations to copyright (such as fair use or the right to quote) may obviate the need for a license.
-
-Researchers are only allowed to share material they have the necessary rights to.
-This includes copyrights, which are granted by free/open copyright licenses.
-These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.
-In addition, researchers should share any work by their own under such a license.
-
-_Choosing a license_ involves:
-
-1. Recording the copyright notices and licenses for content by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
-2.
-In the simplest case, choosing two licenses (one suitable for code, one for everything else) which to apply  simultaneously to all own content (for example, again by recording the license in the file `LICENSE.txt`).
-Recipients are then free to choose under which of the two licenses to use the material.
-
-We recommend to choose [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) for all own works.
-Take your time to read the summaries behind the links and/or the actual license text to understand its legal effect.
-
-You can adapt the following wording to your use case:
-
-> Except where noted otherwise, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
-
-## A Primer on Licenses
-
-Whenever you create a literary or artistic work (such as a text, image, video, or software), the copyright law in most countries limits other people from copying, modifying, and sharing it without your express permission.
-[If the work was created as part of your job, it might be your employer who holds the copyright to the work, depending on the country and contract.]{.aside} This even applies if you make it available to others (e.g., on your website): First and foremost they are not allowed to copy, modify, or share it.
-This legal default of "all rights reserved" was created to benefit publishers, not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
-Copyright licenses enable authors to free up their works for reuse by others.
-
-::: {#nte-situation-lmu .callout-note collapse="true"}
-### For researchers at LMU
-
-If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
-The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
-Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
-:::
-
-A license is a legal document that regulates what _others_ are allowed to do with a copyright-protected work -- the licenses we discuss are non-exclusive licenses and do not limit the author or copyright holder in their rights.
-And while you _could_ write your own license, there are already many pre-formulated licenses available to choose from and to apply to your work.^[In fact, you should not create your own license if the purpose is to share your work freely and openly with others.
-By using a boilerplate license others know what to expect, also because they have been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
-
-::: {.column-margin}
-For the purpose of this tutorial, by _license_ we mean _copyright license_.
-:::
+So far, you took care to legally include works by others in your project folder.
+Up next, you will free up your project for reuse.
+This means ensuring that every bit (co-)authored by you in the project is put under a free/open license.
+It may be that your project (i.e., its files and folders) is not one work, but rather consists of multiple works available under different licenses.
+In that case, you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
+Sometimes, even different parts of a file might be subject to different licenses.
+For the purpose of this tutorial, we will consider the manuscript which you edited and the data dictionary you created.
 
@@
 
-The licenses we discuss here mostly regard copyright.
-Therefore, recipients may lack other rights such as publicity, privacy, moral, patent,^[In fact, the software licenses we recommend on this page have been specifically selected to provide an express grant of patent rights.] or trademark rights.
-For example, sharing photos that depict people is not only a matter of copyright, but also of privacy rights.
-Conversely, the licenses do not apply if recipients are allowed to use the works for other reasons such as fair use, the right to quote, or because they made a different arrangement with the author.
+::: {#nte-license-choice-tldr .callout-note}
+## License Choice TL;DR
+
+We recommend to choose free/open licenses suitable to the type of work (i.e., code, data, and everything else).
+[CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works.
+Take your time to read the summaries behind the links and/or the actual license text to understand the legal effect.
+You can adapt the following wording to your use case:
+
+> Except where noted otherwise, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
+:::
 
@@
 Which license is appropriate depends on several factors, including existing licenses in place and the type of work, but also your personal considerations.
-We strongly recommend to apply a _free/open_ license to your work, which means that the work "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015].
-Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes.
-By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work.
-Software licenses may additionally require to make the source code available to everybody the software is shared with and often require to display the full text of the license upon usage.
+We strongly recommend to apply a _free/open_ license to your work that fits to the work type and potential use cases.
 Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.
@@
 For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/).
-For data, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
-:::
-
-::: {#nte-terminology .callout-note collapse="true"}
-### A Note on Terminology
-
-The terms _free_ and _open_, especially with regard to software, come with a history.
-The Free Software Foundation (FSF) was founded in 1985 to protect "four essential freedoms" [-@FSF2024] of a program's user.
-These are the freedoms to use, study, share, and improve a program.
-Software whose users legally and practically have these freedoms (because, among other things, they have access to its source code) is considered _free_.
-The four freedoms are seen as vital for a society as a whole in the sense that they enable sharing, cooperation and ultimately freedom in general.
-The FSF maintains [a list of software licenses](https://www.gnu.org/licenses/license-list.en.html#SoftwareLicenses) that it considers to be protecting the four freedoms.
-Sometimes the term _libre_ (Spanish and French for _free_) is used to make a distinction from _gratis_ software.
-You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).
-
-The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach.
-It is concerned with developing high-quality software, for which everyone's ability to obtain, modify and contribute back the source code is considered beneficial.
-Access to the source code is one out of multiple conditions for software to be considered _open source_ by the OSI [-@OSI2007], which equally maintains [a list of approved licenses](https://opensource.org/licenses).
-
-Access to the source code is a necessary, but not sufficient requirement both for _free software_ and for _open source software_.
-Conversely, both can (and frequently are) sold for money, as their respective criteria only apply once one has access to the software.
-Throughout this tutorial, we write "free/open license" to mean a license that is approved by either the FSF or the OSI.
-Software which is neither but makes available its source code is sometimes referred to as _source-available software_.
-
-Apart from software, the term _open access_ has often been used for works that are available at no cost.
-For example, this is the commonality of bronze, green, hybrid, gold, and diamond/platinum open access _articles_, which otherwise vary in the rights that are granted to readers.
-In 2002, the Budapest Open Access Initiative declared that _open access_ additionally includes the right to use articles for any purpose, and in 2003, the Bethesda Statement and the Berlin Declaration added the right to make derivative works.
-
-Two other notable definitions include the _Open Definition_ [@OKFN2016], which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015], for which the open editing phase began in 2006.
-They are largely viewed as compatible with one another.
-
-:::: {.column-margin}
-![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
-::::
+For databases, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
 :::
@@
 
-First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license.[You can determine whether a license is free/open by searching for its name in the [SPDX License List](https://spdx.org/licenses/) and looking for at least one `Y` in the two columns _FSF Free/Libre?_ and _OSI Approved?_]{.aside} If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
+First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license.
+If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
 The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder.
@@
 Which license to choose depends on the type of work you create.
-Software licenses, for example, may consider that the source code is the preferred form for making modifications, while licenses for data can differentiate between the database and any works produced from it.
 We have created a flowchart that covers the most likely types of works you will create as a researcher: software, writing (i.e., text), images, audio, video, and data (see @fig-flowchart-simple).
@@
 
-One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license]
+One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights [see @CC2024Software]] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license [see @Wikipedia2024FDL]]
 
@@
 
-> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
+> The Quarto documents in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
 :::
 
-Note, that for data there is, at least in principle, the possibility that a different license applies to the individual entries than to the collective database.
-For example, if you were to create a database of artworks by others, those artworks would be licensed individually as chosen by the artists, but the license for the database as a whole could be chosen by you.
-The latter includes the selection of entries, field names and any _sui generis_ database rights.
+We already mentioned that a database may be licensed differently than its content.
 However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Factual data (like measurements or metadata) should be licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.
+Factual data (like measurements or metadata) should be made available with [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.
 
@@
 Selling the font is allowed, however, if it is bundled with other software.
-- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of the template or that uses the $\TeX$ package has to put under the same license, if shared [@Koppor2016].
-And if a document contained source code covered by the (A)GPLv3, the same license would also apply to the document [@MadHatter2022].
+- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of it has to be put under the same license if shared.
+If a document contained source code covered by the (A)GPLv3, it depends on the individual case whether the same license needs to be applied to the document if shared.
 - __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
@@
 
-The consequence of applying a permissive license to your work is that others may distribute their adaptations from it under stricter terms, possibly without a free/open license.
-You may prefer that everybody who receives somebody else's adaptations of your work enjoys the same rights that you granted to them.
-This can be achieved by _copyleft_ licenses.^[So called because copyleft licenses make content permanently free, thus turning copyright around: © → 🄯.
-In the realm of Creative Commons licenses, copyleft is also called _share alike_.] For software, they come in two flavors:
-
-- __Weak copyleft licenses__ only require that modifications to the software itself are licensed under the same or a compatible license if shared.
-For example, if you create and publish software under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license.
-However, people who merely _use_ your software in their own work which they make publicly available can choose any license for it.
-- __Strong copyleft licenses,__ on the other side, insist that any larger works that use the copyleft-licensed software must also be licensed under the same or a compatible license if shared.
-For example, if your software were to be put under a strong copyleft license, everybody publishing software that uses your software would need to put it under the same license.
-Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024].
-
-Note, however, that the copyleft licenses we discuss here do not mandate sharing.
-Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015].
-This means that if you only use a work internally, you do not need to share your derivative works.
-It is also worth reiterating that these licenses do not restrict the original author(s): They are still permitted to distribute their work under a different license and without sharing the source code.
+You may prefer that adaptations of your work stay free/open and thus choose a copyleft license.
+Of course, this does not restrict you as the original author: You are still permitted to distribute the work under a different license and without sharing the source code.
 
@@
 
-In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @apsillers2019].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
+In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @FSF2024].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
 However, if the output is based on data, it can be considered a derivative work of the data and the license of the data also applies.
@@
 
-```r
+```{.r filename="Console"}
 deps <- renv::dependencies()$Package |>
@@
 
-### Attribution, State Changes, and Anti-DRM?
+### Additional considerations
 
-The advanced license flowcharts also allow you to make additional decisions:
+The advanced license flowcharts also allow you to make additional decisions, such as whether attribution should be a condition of redistribution.
+As you have learned, even if attribution is not a condition of the license, there are other reasons for attributing you as an author (such as scientific citation standards or moral rights).
+All conditions are explained on the page on [copyright](copyright.qmd).
 
-- __Attribution__ means whether recipients of your work are required to provide attribution to you.
-For software licenses, the attribution is part of a copyright notice.
-Note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
-- __Add license text__ means that recipients need to distribute the full text of the license along with the work (rather than only mentioning its license, for example, via a link).
-- __State changes__ means that recipients need to indicate if changes were made.
-- __Disclose code__ means that recipients need to disclose the source code when distributing modified versions of your work.
-- __Anti-DRM__ means that when others share your work they are not allowed to apply technological measures restricting anything that the license permits.
-Implementing such restrictive measures is also referred to as **d**igital **r**ights **m**anagement (DRM).
-
-From the Creative Commons licenses, only [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) does not require providing attribution.
-All software licenses in @fig-flowchart-non-data require providing attribution, although only [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) and [AGPLv3](https://choosealicense.com/licenses/agpl-3.0/) require others to indicate changes.
-For data, the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) allows for technological measures that restrict the work only if a version of the database is provided in parallel without such measures.
+From the Creative Commons licenses, only [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) does not require providing attribution, and all software licenses in @fig-flowchart-non-data require providing attribution.
+As an author, you are free to waive license conditions such as attribution, however.
 
@@
 
-As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
+Creative Commons also offers licenses which prevent commercial use of a work and/or sharing of derivative works.
+A free/open license, however, must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
 Otherwise, creating adaptations from multiple works by different authors would either limit commercial use for everybody (including the authors) or would not be possible at all (when including materials with a copyleft license), effectively creating _silos_ of works which are mutually incompatible with each other.
@@
 
-## Which Licenses to Choose for a Project?
-
-So far, we only discussed how to choose a license for works _of one type_.
-But what if you want to share a project with all kinds of files?
-For example, the project from this tutorial (among other things) includes a data file, a manuscript file with intermingled code and writing, and an R file.
-And what if you also want to share files by others, as is the case with `apaquarto` which you may have installed in this project?
-
-The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
-Sometimes, even different parts of a file might be subject to different licenses.
-The easiest approach is to make note of every foreign work included in your project and record its license.
-Then, multi-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.
-
 ## Applying the License
 
-Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect.
-Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which file or folder is covered by which license, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
-Notably, no sort of registration is required.
+Having selected the license(s) of your choice, we encourage you to read through the full license text (or at least a legal summary) to understand the legal effect.
+The following applies to all licenses mentioned here:
 
-::: {#cau-license-versions .callout-caution}
-### License Versions Are Important
+- Applying the license is irrevocable. You may stop distributing your work, but everybody who already received it may continue to share it.
+- As an author applying the license, you are not bound by the license terms.
+You cannot, however, give others an exclusive license to your content anymore (as some journals needlessly demand, see @wrn-exclusive-rights).
+- By applying the license you share with others your right to use your work for commercial purposes, including selling it.
+- If you apply the license to an image, the license [may also apply](https://commons.wikimedia.org/wiki/Commons:Same_work) to higher resolution version of your image.
+- If you apply a copyleft license to an image, a document in which the image is used may not constitute a derivative work and, as such, may not be affected by the copyleft condition.
+- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_] or if the work was created as part of your job, you may not be able to apply the license.
+Check with your collecting society or your employer in that case.
 
-You may have noticed that we mostly refer to licenses using a name _and_ a version number.
-This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses.
-For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013).
-There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
-CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For Researchers at LMU Munich
 
-For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
+Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
 :::
 
-For example, the `apaquarto` extension that you included in your project is a work by others.^[Unless you are the author of the `apaquarto` extension.] You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with any terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice.
-In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.
+For Creative Commons licenses, the following is additionally important to know:
+
+- By applying the license you waive any moral, publicity, privacy, or other similar personality rights to the extent necessary to use the work.
+- The attribution requirement may be implemented "in any reasonable manner", possibly not conforming to your precise expectations.
+
+Applying a license to a work mostly just means indicating which file or folder is covered by which license(s) (no sort of registration is required). There are multiple standards on how to do so:
+
+- You have already learned about the file `LICENSE.txt`.
+It usually contains detailed information, such as full license texts.
+This is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
+- A summary of the licensing situation is often given in a `README` file, whose creation will be discussed [later](make_readme.qmd).
+- If only parts of a file are covered by a certain license, then it might make sense to indicate this directly in the file.
+- The license of embedded elements (e.g., images) is either given directly beneath to where they are embedded, at the end of the respective page, or collectively on an separate page.
+- Individual programming languages also have their own way of stating which license a package is distributed under.
+For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
+- Software projects often indicate the license in every file, following the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable (explained in @nte-reuse).
+- The following two metadata standards for increasing citability also allow to indicate the license:
+  - [CodeMeta](https://codemeta.github.io/): All information are stored in a file called `codemeta.json`, which can be created [on their website](https://codemeta.github.io/codemeta-generator/).
+Alternatively, one can use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
+  - [Citation File Format](https://citation-file-format.github.io/) (CFF): Here, everything is stored in a file called `CITATION.cff`.
+It is used by GitHub to directly show a recommended citation for a repository.
+The file can be created [on their website](https://citation-file-format.github.io/cff-initializer-javascript/).
+Alternatively, the R package [`cffr`](https://cran.r-project.org/package=cffr) may be used for that purpose.
 
@@
 
-While it is common to state the chosen license(s) in the README, usually one of the following actions is taken in addition:
-
-In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim.
-This is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
-However, if the project is not completely covered by one single license, this practice may become unwieldy.
-For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.
-
-Individual programming languages also have their own way of stating which license a package is distributed under.
-For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
-
-Finally, @nte-reuse explains how to use the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable. This is the approach we recommend taking.
-
 ::: {#nte-reuse .callout-note collapse="true"}
@@
 Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/) which allows communicating the license choice unequivocally.
-We will be using that to indicate the license for every file in your project, along with the year of publication and the copyright holder.
+We will be using that to indicate the license for every file in your project folder, along with the year of publication and the copyright holder.
 To do this, we add a comment to the beginning of every file and include the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`.
@@
 
-Regardless of how exactly the licenses are added to the project, this is a good opportunity to verify one last time that all third party content is provided to you under a free/open license and that you comply with it.
-Please add a license to your project now, either creating a file `LICENSE.txt` or following the REUSE standard.
+The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) or (at your option) under the [Apache License\ 2.0](https://www.apache.org/licenses/LICENSE-2.0), as is all code in this tutorial (e.g., for creating the data dictionary).
+Therefore, you can apply any license to your modified version (as long as you follow the terms of the existing license).
+Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).
 
-## Wrap-up
+::: {#tip-manuscript-licensetxt .callout-tip collapse="true"}
+### Addition to `LICENSE.txt` (Solution)
 
-If you would like to learn more about copyright and licenses you might find the following resources interesting:
+In the following example, the files from this tutorial are used under the terms of CC0\ 1.0 and therefore, the requirements of the Apache license\ 2.0 need not be followed.
 
-- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2014
-- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
-- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
+The edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.
+
+```{.txt .code-overflow-wrap filename="LICENSE.txt"}
+The Quarto documents stored in "Manuscript.qmd", "Manuscript.tex", and "Manuscript.pdf" by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under AGPLv3 (or later) <https://www.gnu.org/licenses/agpl-3.0.html> or (at your option) under CC BY-SA 4.0 <https://creativecommons.org/licenses/by-sa/4.0/>.
+
+The data dictionary (stored in "data_dictionary.html") and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
+```
+
+As the AGPLv3 requires reusers adding the full license text, we recommend to also copy it to `LICENSE.txt` (if not already in there).
+
+<details>
+<summary>Full text of AGPLv3</summary>
+
+```{.txt .code-overflow-wrap filename="LICENSE.txt"}
+                    GNU AFFERO GENERAL PUBLIC LICENSE
+                       Version 3, 19 November 2007
+
+ Copyright (C) 2007 Free Software Foundation, Inc. <https://fsf.org/>
+ Everyone is permitted to copy and distribute verbatim copies
+ of this license document, but changing it is not allowed.
+
+                            Preamble
+
+  The GNU Affero General Public License is a free, copyleft license for
+software and other kinds of works, specifically designed to ensure
+cooperation with the community in the case of network server software.
+
+  The licenses for most software and other practical works are designed
+to take away your freedom to share and change the works.  By contrast,
+our General Public Licenses are intended to guarantee your freedom to
+share and change all versions of a program--to make sure it remains free
+software for all its users.
+
+  When we speak of free software, we are referring to freedom, not
+price.  Our General Public Licenses are designed to make sure that you
+have the freedom to distribute copies of free software (and charge for
+them if you wish), that you receive source code or can get it if you
+want it, that you can change the software or use pieces of it in new
+free programs, and that you know you can do these things.
+
+  Developers that use our General Public Licenses protect your rights
+with two steps: (1) assert copyright on the software, and (2) offer
+you this License which gives you legal permission to copy, distribute
+and/or modify the software.
+
+  A secondary benefit of defending all users' freedom is that
+improvements made in alternate versions of the program, if they
+receive widespread use, become available for other developers to
+incorporate.  Many developers of free software are heartened and
+encouraged by the resulting cooperation.  However, in the case of
+software used on network servers, this result may fail to come about.
+The GNU General Public License permits making a modified version and
+letting the public access it on a server without ever releasing its
+source code to the public.
+
+  The GNU Affero General Public License is designed specifically to
+ensure that, in such cases, the modified source code becomes available
+to the community.  It requires the operator of a network server to
+provide the source code of the modified version running there to the
+users of that server.  Therefore, public use of a modified version, on
+a publicly accessible server, gives the public access to the source
+code of the modified version.
+
+  An older license, called the Affero General Public License and
+published by Affero, was designed to accomplish similar goals.  This is
+a different license, not a version of the Affero GPL, but Affero has
+released a new version of the Affero GPL which permits relicensing under
+this license.
+
+  The precise terms and conditions for copying, distribution and
+modification follow.
+
+                       TERMS AND CONDITIONS
+
+  0. Definitions.
+
+  "This License" refers to version 3 of the GNU Affero General Public License.
+
+  "Copyright" also means copyright-like laws that apply to other kinds of
+works, such as semiconductor masks.
+
+  "The Program" refers to any copyrightable work licensed under this
+License.  Each licensee is addressed as "you".  "Licensees" and
+"recipients" may be individuals or organizations.
+
+  To "modify" a work means to copy from or adapt all or part of the work
+in a fashion requiring copyright permission, other than the making of an
+exact copy.  The resulting work is called a "modified version" of the
+earlier work or a work "based on" the earlier work.
+
+  A "covered work" means either the unmodified Program or a work based
+on the Program.
+
+  To "propagate" a work means to do anything with it that, without
+permission, would make you directly or secondarily liable for
+infringement under applicable copyright law, except executing it on a
+computer or modifying a private copy.  Propagation includes copying,
+distribution (with or without modification), making available to the
+public, and in some countries other activities as well.
+
+  To "convey" a work means any kind of propagation that enables other
+parties to make or receive copies.  Mere interaction with a user through
+a computer network, with no transfer of a copy, is not conveying.
+
+  An interactive user interface displays "Appropriate Legal Notices"
+to the extent that it includes a convenient and prominently visible
+feature that (1) displays an appropriate copyright notice, and (2)
+tells the user that there is no warranty for the work (except to the
+extent that warranties are provided), that licensees may convey the
+work under this License, and how to view a copy of this License.  If
+the interface presents a list of user commands or options, such as a
+menu, a prominent item in the list meets this criterion.
+
+  1. Source Code.
+
+  The "source code" for a work means the preferred form of the work
+for making modifications to it.  "Object code" means any non-source
+form of a work.
+
+  A "Standard Interface" means an interface that either is an official
+standard defined by a recognized standards body, or, in the case of
+interfaces specified for a particular programming language, one that
+is widely used among developers working in that language.
+
+  The "System Libraries" of an executable work include anything, other
+than the work as a whole, that (a) is included in the normal form of
+packaging a Major Component, but which is not part of that Major
+Component, and (b) serves only to enable use of the work with that
+Major Component, or to implement a Standard Interface for which an
+implementation is available to the public in source code form.  A
+"Major Component", in this context, means a major essential component
+(kernel, window system, and so on) of the specific operating system
+(if any) on which the executable work runs, or a compiler used to
+produce the work, or an object code interpreter used to run it.
+
+  The "Corresponding Source" for a work in object code form means all
+the source code needed to generate, install, and (for an executable
+work) run the object code and to modify the work, including scripts to
+control those activities.  However, it does not include the work's
+System Libraries, or general-purpose tools or generally available free
+programs which are used unmodified in performing those activities but
+which are not part of the work.  For example, Corresponding Source
+includes interface definition files associated with source files for
+the work, and the source code for shared libraries and dynamically
+linked subprograms that the work is specifically designed to require,
+such as by intimate data communication or control flow between those
+subprograms and other parts of the work.
+
+  The Corresponding Source need not include anything that users
+can regenerate automatically from other parts of the Corresponding
+Source.
+
+  The Corresponding Source for a work in source code form is that
+same work.
+
+  2. Basic Permissions.
+
+  All rights granted under this License are granted for the term of
+copyright on the Program, and are irrevocable provided the stated
+conditions are met.  This License explicitly affirms your unlimited
+permission to run the unmodified Program.  The output from running a
+covered work is covered by this License only if the output, given its
+content, constitutes a covered work.  This License acknowledges your
+rights of fair use or other equivalent, as provided by copyright law.
+
+  You may make, run and propagate covered works that you do not
+convey, without conditions so long as your license otherwise remains
+in force.  You may convey covered works to others for the sole purpose
+of having them make modifications exclusively for you, or provide you
+with facilities for running those works, provided that you comply with
+the terms of this License in conveying all material for which you do
+not control copyright.  Those thus making or running the covered works
+for you must do so exclusively on your behalf, under your direction
+and control, on terms that prohibit them from making any copies of
+your copyrighted material outside their relationship with you.
+
+  Conveying under any other circumstances is permitted solely under
+the conditions stated below.  Sublicensing is not allowed; section 10
+makes it unnecessary.
+
+  3. Protecting Users' Legal Rights From Anti-Circumvention Law.
+
+  No covered work shall be deemed part of an effective technological
+measure under any applicable law fulfilling obligations under article
+11 of the WIPO copyright treaty adopted on 20 December 1996, or
+similar laws prohibiting or restricting circumvention of such
+measures.
+
+  When you convey a covered work, you waive any legal power to forbid
+circumvention of technological measures to the extent such circumvention
+is effected by exercising rights under this License with respect to
+the covered work, and you disclaim any intention to limit operation or
+modification of the work as a means of enforcing, against the work's
+users, your or third parties' legal rights to forbid circumvention of
+technological measures.
+
+  4. Conveying Verbatim Copies.
+
+  You may convey verbatim copies of the Program's source code as you
+receive it, in any medium, provided that you conspicuously and
+appropriately publish on each copy an appropriate copyright notice;
+keep intact all notices stating that this License and any
+non-permissive terms added in accord with section 7 apply to the code;
+keep intact all notices of the absence of any warranty; and give all
+recipients a copy of this License along with the Program.
+
+  You may charge any price or no price for each copy that you convey,
+and you may offer support or warranty protection for a fee.
+
+  5. Conveying Modified Source Versions.
+
+  You may convey a work based on the Program, or the modifications to
+produce it from the Program, in the form of source code under the
+terms of section 4, provided that you also meet all of these conditions:
+
+    a) The work must carry prominent notices stating that you modified
+    it, and giving a relevant date.
+
+    b) The work must carry prominent notices stating that it is
+    released under this License and any conditions added under section
+    7.  This requirement modifies the requirement in section 4 to
+    "keep intact all notices".
+
+    c) You must license the entire work, as a whole, under this
+    License to anyone who comes into possession of a copy.  This
+    License will therefore apply, along with any applicable section 7
+    additional terms, to the whole of the work, and all its parts,
+    regardless of how they are packaged.  This License gives no
+    permission to license the work in any other way, but it does not
+    invalidate such permission if you have separately received it.
+
+    d) If the work has interactive user interfaces, each must display
+    Appropriate Legal Notices; however, if the Program has interactive
+    interfaces that do not display Appropriate Legal Notices, your
+    work need not make them do so.
+
+  A compilation of a covered work with other separate and independent
+works, which are not by their nature extensions of the covered work,
+and which are not combined with it such as to form a larger program,
+in or on a volume of a storage or distribution medium, is called an
+"aggregate" if the compilation and its resulting copyright are not
+used to limit the access or legal rights of the compilation's users
+beyond what the individual works permit.  Inclusion of a covered work
+in an aggregate does not cause this License to apply to the other
+parts of the aggregate.
+
+  6. Conveying Non-Source Forms.
+
+  You may convey a covered work in object code form under the terms
+of sections 4 and 5, provided that you also convey the
+machine-readable Corresponding Source under the terms of this License,
+in one of these ways:
+
+    a) Convey the object code in, or embodied in, a physical product
+    (including a physical distribution medium), accompanied by the
+    Corresponding Source fixed on a durable physical medium
+    customarily used for software interchange.
+
+    b) Convey the object code in, or embodied in, a physical product
+    (including a physical distribution medium), accompanied by a
+    written offer, valid for at least three years and valid for as
+    long as you offer spare parts or customer support for that product
+    model, to give anyone who possesses the object code either (1) a
+    copy of the Corresponding Source for all the software in the
+    product that is covered by this License, on a durable physical
+    medium customarily used for software interchange, for a price no
+    more than your reasonable cost of physically performing this
+    conveying of source, or (2) access to copy the
+    Corresponding Source from a network server at no charge.
+
+    c) Convey individual copies of the object code with a copy of the
+    written offer to provide the Corresponding Source.  This
+    alternative is allowed only occasionally and noncommercially, and
+    only if you received the object code with such an offer, in accord
+    with subsection 6b.
+
+    d) Convey the object code by offering access from a designated
+    place (gratis or for a charge), and offer equivalent access to the
+    Corresponding Source in the same way through the same place at no
+    further charge.  You need not require recipients to copy the
+    Corresponding Source along with the object code.  If the place to
+    copy the object code is a network server, the Corresponding Source
+    may be on a different server (operated by you or a third party)
+    that supports equivalent copying facilities, provided you maintain
+    clear directions next to the object code saying where to find the
+    Corresponding Source.  Regardless of what server hosts the
+    Corresponding Source, you remain obligated to ensure that it is
+    available for as long as needed to satisfy these requirements.
+
+    e) Convey the object code using peer-to-peer transmission, provided
+    you inform other peers where the object code and Corresponding
+    Source of the work are being offered to the general public at no
+    charge under subsection 6d.
+
+  A separable portion of the object code, whose source code is excluded
+from the Corresponding Source as a System Library, need not be
+included in conveying the object code work.
+
+  A "User Product" is either (1) a "consumer product", which means any
+tangible personal property which is normally used for personal, family,
+or household purposes, or (2) anything designed or sold for incorporation
+into a dwelling.  In determining whether a product is a consumer product,
+doubtful cases shall be resolved in favor of coverage.  For a particular
+product received by a particular user, "normally used" refers to a
+typical or common use of that class of product, regardless of the status
+of the particular user or of the way in which the particular user
+actually uses, or expects or is expected to use, the product.  A product
+is a consumer product regardless of whether the product has substantial
+commercial, industrial or non-consumer uses, unless such uses represent
+the only significant mode of use of the product.
+
+  "Installation Information" for a User Product means any methods,
+procedures, authorization keys, or other information required to install
+and execute modified versions of a covered work in that User Product from
+a modified version of its Corresponding Source.  The information must
+suffice to ensure that the continued functioning of the modified object
+code is in no case prevented or interfered with solely because
+modification has been made.
+
+  If you convey an object code work under this section in, or with, or
+specifically for use in, a User Product, and the conveying occurs as
+part of a transaction in which the right of possession and use of the
+User Product is transferred to the recipient in perpetuity or for a
+fixed term (regardless of how the transaction is characterized), the
+Corresponding Source conveyed under this section must be accompanied
+by the Installation Information.  But this requirement does not apply
+if neither you nor any third party retains the ability to install
+modified object code on the User Product (for example, the work has
+been installed in ROM).
+
+  The requirement to provide Installation Information does not include a
+requirement to continue to provide support service, warranty, or updates
+for a work that has been modified or installed by the recipient, or for
+the User Product in which it has been modified or installed.  Access to a
+network may be denied when the modification itself materially and
+adversely affects the operation of the network or violates the rules and
+protocols for communication across the network.
+
+  Corresponding Source conveyed, and Installation Information provided,
+in accord with this section must be in a format that is publicly
+documented (and with an implementation available to the public in
+source code form), and must require no special password or key for
+unpacking, reading or copying.
+
+  7. Additional Terms.
+
+  "Additional permissions" are terms that supplement the terms of this
+License by making exceptions from one or more of its conditions.
+Additional permissions that are applicable to the entire Program shall
+be treated as though they were included in this License, to the extent
+that they are valid under applicable law.  If additional permissions
+apply only to part of the Program, that part may be used separately
+under those permissions, but the entire Program remains governed by
+this License without regard to the additional permissions.
+
+  When you convey a copy of a covered work, you may at your option
+remove any additional permissions from that copy, or from any part of
+it.  (Additional permissions may be written to require their own
+removal in certain cases when you modify the work.)  You may place
+additional permissions on material, added by you to a covered work,
+for which you have or can give appropriate copyright permission.
+
+  Notwithstanding any other provision of this License, for material you
+add to a covered work, you may (if authorized by the copyright holders of
+that material) supplement the terms of this License with terms:
+
+    a) Disclaiming warranty or limiting liability differently from the
+    terms of sections 15 and 16 of this License; or
+
+    b) Requiring preservation of specified reasonable legal notices or
+    author attributions in that material or in the Appropriate Legal
+    Notices displayed by works containing it; or
+
+    c) Prohibiting misrepresentation of the origin of that material, or
+    requiring that modified versions of such material be marked in
+    reasonable ways as different from the original version; or
+
+    d) Limiting the use for publicity purposes of names of licensors or
+    authors of the material; or
+
+    e) Declining to grant rights under trademark law for use of some
+    trade names, trademarks, or service marks; or
+
+    f) Requiring indemnification of licensors and authors of that
+    material by anyone who conveys the material (or modified versions of
+    it) with contractual assumptions of liability to the recipient, for
+    any liability that these contractual assumptions directly impose on
+    those licensors and authors.
+
+  All other non-permissive additional terms are considered "further
+restrictions" within the meaning of section 10.  If the Program as you
+received it, or any part of it, contains a notice stating that it is
+governed by this License along with a term that is a further
+restriction, you may remove that term.  If a license document contains
+a further restriction but permits relicensing or conveying under this
+License, you may add to a covered work material governed by the terms
+of that license document, provided that the further restriction does
+not survive such relicensing or conveying.
+
+  If you add terms to a covered work in accord with this section, you
+must place, in the relevant source files, a statement of the
+additional terms that apply to those files, or a notice indicating
+where to find the applicable terms.
+
+  Additional terms, permissive or non-permissive, may be stated in the
+form of a separately written license, or stated as exceptions;
+the above requirements apply either way.
+
+  8. Termination.
+
+  You may not propagate or modify a covered work except as expressly
+provided under this License.  Any attempt otherwise to propagate or
+modify it is void, and will automatically terminate your rights under
+this License (including any patent licenses granted under the third
+paragraph of section 11).
+
+  However, if you cease all violation of this License, then your
+license from a particular copyright holder is reinstated (a)
+provisionally, unless and until the copyright holder explicitly and
+finally terminates your license, and (b) permanently, if the copyright
+holder fails to notify you of the violation by some reasonable means
+prior to 60 days after the cessation.
+
+  Moreover, your license from a particular copyright holder is
+reinstated permanently if the copyright holder notifies you of the
+violation by some reasonable means, this is the first time you have
+received notice of violation of this License (for any work) from that
+copyright holder, and you cure the violation prior to 30 days after
+your receipt of the notice.
+
+  Termination of your rights under this section does not terminate the
+licenses of parties who have received copies or rights from you under
+this License.  If your rights have been terminated and not permanently
+reinstated, you do not qualify to receive new licenses for the same
+material under section 10.
+
+  9. Acceptance Not Required for Having Copies.
+
+  You are not required to accept this License in order to receive or
+run a copy of the Program.  Ancillary propagation of a covered work
+occurring solely as a consequence of using peer-to-peer transmission
+to receive a copy likewise does not require acceptance.  However,
+nothing other than this License grants you permission to propagate or
+modify any covered work.  These actions infringe copyright if you do
+not accept this License.  Therefore, by modifying or propagating a
+covered work, you indicate your acceptance of this License to do so.
+
+  10. Automatic Licensing of Downstream Recipients.
+
+  Each time you convey a covered work, the recipient automatically
+receives a license from the original licensors, to run, modify and
+propagate that work, subject to this License.  You are not responsible
+for enforcing compliance by third parties with this License.
+
+  An "entity transaction" is a transaction transferring control of an
+organization, or substantially all assets of one, or subdividing an
+organization, or merging organizations.  If propagation of a covered
+work results from an entity transaction, each party to that
+transaction who receives a copy of the work also receives whatever
+licenses to the work the party's predecessor in interest had or could
+give under the previous paragraph, plus a right to possession of the
+Corresponding Source of the work from the predecessor in interest, if
+the predecessor has it or can get it with reasonable efforts.
+
+  You may not impose any further restrictions on the exercise of the
+rights granted or affirmed under this License.  For example, you may
+not impose a license fee, royalty, or other charge for exercise of
+rights granted under this License, and you may not initiate litigation
+(including a cross-claim or counterclaim in a lawsuit) alleging that
+any patent claim is infringed by making, using, selling, offering for
+sale, or importing the Program or any portion of it.
+
+  11. Patents.
+
+  A "contributor" is a copyright holder who authorizes use under this
+License of the Program or a work on which the Program is based.  The
+work thus licensed is called the contributor's "contributor version".
+
+  A contributor's "essential patent claims" are all patent claims
+owned or controlled by the contributor, whether already acquired or
+hereafter acquired, that would be infringed by some manner, permitted
+by this License, of making, using, or selling its contributor version,
+but do not include claims that would be infringed only as a
+consequence of further modification of the contributor version.  For
+purposes of this definition, "control" includes the right to grant
+patent sublicenses in a manner consistent with the requirements of
+this License.
+
+  Each contributor grants you a non-exclusive, worldwide, royalty-free
+patent license under the contributor's essential patent claims, to
+make, use, sell, offer for sale, import and otherwise run, modify and
+propagate the contents of its contributor version.
+
+  In the following three paragraphs, a "patent license" is any express
+agreement or commitment, however denominated, not to enforce a patent
+(such as an express permission to practice a patent or covenant not to
+sue for patent infringement).  To "grant" such a patent license to a
+party means to make such an agreement or commitment not to enforce a
+patent against the party.
+
+  If you convey a covered work, knowingly relying on a patent license,
+and the Corresponding Source of the work is not available for anyone
+to copy, free of charge and under the terms of this License, through a
+publicly available network server or other readily accessible means,
+then you must either (1) cause the Corresponding Source to be so
+available, or (2) arrange to deprive yourself of the benefit of the
+patent license for this particular work, or (3) arrange, in a manner
+consistent with the requirements of this License, to extend the patent
+license to downstream recipients.  "Knowingly relying" means you have
+actual knowledge that, but for the patent license, your conveying the
+covered work in a country, or your recipient's use of the covered work
+in a country, would infringe one or more identifiable patents in that
+country that you have reason to believe are valid.
+
+  If, pursuant to or in connection with a single transaction or
+arrangement, you convey, or propagate by procuring conveyance of, a
+covered work, and grant a patent license to some of the parties
+receiving the covered work authorizing them to use, propagate, modify
+or convey a specific copy of the covered work, then the patent license
+you grant is automatically extended to all recipients of the covered
+work and works based on it.
+
+  A patent license is "discriminatory" if it does not include within
+the scope of its coverage, prohibits the exercise of, or is
+conditioned on the non-exercise of one or more of the rights that are
+specifically granted under this License.  You may not convey a covered
+work if you are a party to an arrangement with a third party that is
+in the business of distributing software, under which you make payment
+to the third party based on the extent of your activity of conveying
+the work, and under which the third party grants, to any of the
+parties who would receive the covered work from you, a discriminatory
+patent license (a) in connection with copies of the covered work
+conveyed by you (or copies made from those copies), or (b) primarily
+for and in connection with specific products or compilations that
+contain the covered work, unless you entered into that arrangement,
+or that patent license was granted, prior to 28 March 2007.
+
+  Nothing in this License shall be construed as excluding or limiting
+any implied license or other defenses to infringement that may
+otherwise be available to you under applicable patent law.
+
+  12. No Surrender of Others' Freedom.
+
+  If conditions are imposed on you (whether by court order, agreement or
+otherwise) that contradict the conditions of this License, they do not
+excuse you from the conditions of this License.  If you cannot convey a
+covered work so as to satisfy simultaneously your obligations under this
+License and any other pertinent obligations, then as a consequence you may
+not convey it at all.  For example, if you agree to terms that obligate you
+to collect a royalty for further conveying from those to whom you convey
+the Program, the only way you could satisfy both those terms and this
+License would be to refrain entirely from conveying the Program.
+
+  13. Remote Network Interaction; Use with the GNU General Public License.
+
+  Notwithstanding any other provision of this License, if you modify the
+Program, your modified version must prominently offer all users
+interacting with it remotely through a computer network (if your version
+supports such interaction) an opportunity to receive the Corresponding
+Source of your version by providing access to the Corresponding Source
+from a network server at no charge, through some standard or customary
+means of facilitating copying of software.  This Corresponding Source
+shall include the Corresponding Source for any work covered by version 3
+of the GNU General Public License that is incorporated pursuant to the
+following paragraph.
+
+  Notwithstanding any other provision of this License, you have
+permission to link or combine any covered work with a work licensed
+under version 3 of the GNU General Public License into a single
+combined work, and to convey the resulting work.  The terms of this
+License will continue to apply to the part which is the covered work,
+but the work with which it is combined will remain governed by version
+3 of the GNU General Public License.
+
+  14. Revised Versions of this License.
+
+  The Free Software Foundation may publish revised and/or new versions of
+the GNU Affero General Public License from time to time.  Such new versions
+will be similar in spirit to the present version, but may differ in detail to
+address new problems or concerns.
+
+  Each version is given a distinguishing version number.  If the
+Program specifies that a certain numbered version of the GNU Affero General
+Public License "or any later version" applies to it, you have the
+option of following the terms and conditions either of that numbered
+version or of any later version published by the Free Software
+Foundation.  If the Program does not specify a version number of the
+GNU Affero General Public License, you may choose any version ever published
+by the Free Software Foundation.
+
+  If the Program specifies that a proxy can decide which future
+versions of the GNU Affero General Public License can be used, that proxy's
+public statement of acceptance of a version permanently authorizes you
+to choose that version for the Program.
+
+  Later license versions may give you additional or different
+permissions.  However, no additional obligations are imposed on any
+author or copyright holder as a result of your choosing to follow a
+later version.
+
+  15. Disclaimer of Warranty.
+
+  THERE IS NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY
+APPLICABLE LAW.  EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT
+HOLDERS AND/OR OTHER PARTIES PROVIDE THE PROGRAM "AS IS" WITHOUT WARRANTY
+OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO,
+THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM
+IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST OF
+ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+
+  16. Limitation of Liability.
+
+  IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
+WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MODIFIES AND/OR CONVEYS
+THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR DAMAGES, INCLUDING ANY
+GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL DAMAGES ARISING OUT OF THE
+USE OR INABILITY TO USE THE PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF
+DATA OR DATA BEING RENDERED INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD
+PARTIES OR A FAILURE OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS),
+EVEN IF SUCH HOLDER OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF
+SUCH DAMAGES.
+
+  17. Interpretation of Sections 15 and 16.
+
+  If the disclaimer of warranty and limitation of liability provided
+above cannot be given local legal effect according to their terms,
+reviewing courts shall apply local law that most closely approximates
+an absolute waiver of all civil liability in connection with the
+Program, unless a warranty or assumption of liability accompanies a
+copy of the Program in return for a fee.
+
+                     END OF TERMS AND CONDITIONS
+
+            How to Apply These Terms to Your New Programs
+
+  If you develop a new program, and you want it to be of the greatest
+possible use to the public, the best way to achieve this is to make it
+free software which everyone can redistribute and change under these terms.
+
+  To do so, attach the following notices to the program.  It is safest
+to attach them to the start of each source file to most effectively
+state the exclusion of warranty; and each file should have at least
+the "copyright" line and a pointer to where the full notice is found.
+
+    <one line to give the program's name and a brief idea of what it does.>
+    Copyright (C) <year>  <name of author>
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Affero General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Affero General Public License for more details.
+
+    You should have received a copy of the GNU Affero General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+
+Also add information on how to contact you by electronic and paper mail.
+
+  If your software can interact with users remotely through a computer
+network, you should also make sure that it provides a way for users to
+get its source.  For example, if your program is a web application, its
+interface could display a "Source" link that leads users to an archive
+of the code.  There are many ways you could offer source, and different
+solutions will be better for different programs; see section 13 for the
+specific requirements.
+
+  You should also get your employer (if you work as a programmer) or school,
+if any, to sign a "copyright disclaimer" for the program, if necessary.
+For more information on this, and how to apply and follow the GNU AGPL, see
+<https://www.gnu.org/licenses/>.
+
+```
+
+</details>
+:::
 
@@
 
-  type --"Software"--> code_sa("Attribution?<br>Add license text?<br>State changes?<br>Disclose code?<br>Copyleft?")
-  type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>State changes?<br>Anti-DRM?<br>Copyleft?")
+  type --"Software"--> code_sa("Attribution?<br>Indicate license?<br>Add license text?<br>State changes?<br>Disclose code?<br>Copyleft?")
+  type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>Indicate license?<br>State changes?<br>Anti-DRM?<br>Copyleft?")
 
   code_sa --"Attribution &<br>Add license text &<br>State changes"--> apache["Apache 2.0"]
-  code_sa --"Attribution &<br>Disclose code &<br>Weak copyleft"--> mpl["MPL 2.0"]
+  code_sa --"Attribution &<br>Indicate license &<br>Disclose code &<br>Weak copyleft"--> mpl["MPL 2.0"]
   code_sa --"Attribution &<br>Add license text &<br>State changes &<br>Disclose code &<br>Strong copyleft"--> agpl["AGPLv3"]
@@
   nocode_cc --"Neither"--> cc0["CC0 1.0"]
-  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM"--> cc_by["CC BY 4.0"]
-  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
+  nocode_cc --"Attribution &<br>Indicate license &<br>State changes &<br>Anti-DRM"--> cc_by["CC BY 4.0"]
+  nocode_cc --"Attribution &<br>Indicate license &<br>State changes &<br>Anti-DRM &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]
 
@@
   cc0_content_metadata --> cc0_db["CC0 1.0"]
-  switch_license --"CC0 or<br>non-CC license"--> sa("Attribution?<br>Anti-DRM?<br>Copyleft?")
+  switch_license --"CC0 or<br>non-CC license"--> sa("Attribution?<br>Indicate license?<br>Anti-DRM?<br>Copyleft?")
   switch_license --"CC BY or<br>CC BY-SA"--> same["<em>Same license for DB</em>"]
@@
   sa --"Neither"--> cc0_db
-  sa --"Attribution &<br>Anti-DRM &<br>Copyleft"--> odbl["ODbL 1.0"]
+  sa --"Attribution &<br>Indicate license &<br>Anti-DRM &<br>Copyleft"--> odbl["ODbL 1.0"]
 

### 2025-01-25 code.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

--- 2024-12-10 setup.qmd
@@
 ---
-title: "Project Setup"
-engine: knitr
+title: "Add Code"
 ---
 
-We will start by setting up a simple example of a reproducible report.
+## Add Data Analysis
 
-## Create Quarto Project
+With the data added to the project folder, you can now analyze them and describe the results.
+For the purpose of this tutorial, we suggest to conduct a simple analysis that explores whether the bill lengths of penguins vary between sex.
+You can write your own or use the following example:
 
-First, we will need to create a new Quarto project.
+::: {#tip-t-test .callout-tip collapse="true"}
+### Data Analysis Example
 
-If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
+You can copy the following code into the _results_ section of the manuscript:
 
-![](images/setup_01.png){width=500px fig-align="center"}
+``````{.qmd .code-overflow-wrap filename="Manuscript.qmd"}
+Descriptive statistics of the data set are given in @tbl-descriptive-statistics and individual bill lengths are displayed in @fig-bill-length-comparison.
 
-Here, select _New Directory_
+```{{r}}
+#| label: "t-test"
+#| echo: false
+#| output: "asis"
 
-![](images/setup_02.png){width=600px fig-align="center"}
+# Load penguins data set
+dat <- read.csv("data.csv")
 
-And choose the project type _Quarto Project_.
+# Remove rows with NA
+dat_clean <- dat[complete.cases(dat), ]
 
-![](images/setup_03.png){width=600px fig-align="center"}
+# Treat year as categorical variable
+dat_clean$year <- as.factor(dat_clean$year)
 
-Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.
+# Perform t-test
+t_test_result <- t.test(
+  bill_length_mm ~ sex,
+  data = dat_clean,
+  var.equal = FALSE,
+  conf.level = 0.95
+)
 
-As we will use Git to track the version history of files, be sure to check _Create a git repository_.
-If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".
+# Describe results
+report::report(
+  x = t_test_result,
+  data = dat_clean,
+  rules = "gignac2016"
+) |>
+  report::as.report_text(summary = TRUE)
+```
 
-::: {.column-margin}
-![renv: A dependency management toolkit for R](images/renv.svg){width=250px}
-:::
+```{{r}}
+#| label: "tbl-descriptive-statistics"
+#| echo: false
+#| tbl-cap: "Descriptive Statistics"
+#| apa-twocolumn: true
 
-Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on.
-Using it makes it easier for others to view and obtain them at a later point in time.
-Therefore make sure that the box _Use renv with this project_ is checked.
-Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".
+report::report_sample(dat_clean, by = "sex", total = FALSE) |>
+  knitr::kable()
 
-If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.
+```
 
-![](images/setup_04.png){width=600px fig-align="center"}
+```{{r}}
+#| label: "fig-bill-length-comparison"
+#| echo: false
+#| fig-cap: "Scatter Plot of Bill Lengths by Sex With Violin Plot Showing Quartiles"
 
-Click on _Create Project_. Your RStudio window should now look similar to this:
+# Create ggplot object
+ggplot2::ggplot(dat_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
+  ggplot2::geom_violin(
+    draw_quantiles = c(0.25, 0.5, 0.75),
+    show.legend = FALSE
+  ) +
+  ggplot2::geom_jitter(width = 0.15, show.legend = FALSE) +
+  ggplot2::labs(x = "Sex", y = "Bill length in mm") +
+  ggplot2::theme_minimal()
+```
+``````
 
-![](images/setup_05.png){.lightbox fig-align="center" alt="The project `code-publishing-exercise` opened in RStudio.
-The source pane to the top left has a Quarto file open called \"code-publishing-exercise.qmd\".
-The console pane to the bottom left indicates by its output that renv is active in the current project.
-The environment pane to the top right indicates that the environment is currently empty.
-The output pane to the bottom right shows the files in the current project."}
-
-If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.
-
-Make sure that your project is in a consistent state according to `renv` by running:
+This code uses some new packages, one can use `renv` to view them:
 
@@
 
-If it reports packages that are not used, synchronize the lock file using:
+Then, install and record them with:
 
 ```{.r filename="Console"}
+renv::install()
 renv::snapshot()
-```
-
-::: {#nte-terminal .callout-note collapse="true"}
-### Without RStudio
-
-Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:
-
-```{.bash filename="Terminal"}
-quarto create project default code-publishing-exercise
-cd code-publishing-exercise/
-rm code-publishing-exercise.qmd
-git init
-git checkout -b main
-```
-
-Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Then, initialize `renv`:
-
-```{.r filename="Console"}
-renv::init()
 ```
@@
 
-You are now ready to stage and commit your files.
-You can either stage files separately or the whole project folder at once.
-If you do the latter, we recommend you to inspect the untracked changes before staging all of them:
-
-::: {.column-margin}
-In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
-:::
-
-```{.bash filename="Terminal"}
-git status
-```
-
-Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file.
-If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:
-
-```{.bash filename="Terminal"}
-git add .
-git commit -m "Initial commit"
-```
-
-If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file.
-If you don't check your changes before committing, you might accidentally commit something you'd rather not.
-
-::: {#tip-author-identity-unknown .callout-tip}
-If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:
-
-```{.bash filename="Terminal"}
-git config user.name "YOUR NAME"
-git config user.email "YOUR EMAIL ADDRESS"
-```
-
-Then, commit again.
-:::
-
-## Decide on Structure
-
-Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it.
-In general, the directory structure should facilitate understanding a project by breaking it into logical chunks.
-There is no single best solution, as a good structure depends on where a project's complexity lies.
-However, it is usually helpful if the different files and folders reflect the execution order.
-For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders.
-Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.
-
-Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021].
-We recommend you to follow the standards of your field.
-
-For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.
-
-## Add Data
-
-```{r}
-#| echo: false
-
-write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
-```
-
-You can now download the data set we have prepared for you and put it into your project folder: [`data.csv`](data.csv){download=""}
-
-::: {.column-margin}
-![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
-:::
-
-The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
-It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-
-::: {#imp-legal-restrictions .callout-important}
-### Consider Legal Restrictions Before Sharing
-
-Everything you put into the project folder will be shared publicly.
-For reasons of reproducibility, this should include the data you analyze.
-Of course, you should only share them to the extent you are allowed to, taking into account:
-
-1. applicable privacy laws (e.g., the GDPR for European citizens),
-2. contractual obligations (e.g., with your data provider),
-3. copyright of the data and their particular structure, and
-4. any _sui generis_ database right.
-
-Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
-For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
-
-Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are.
-Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.
-As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
-:::
-
-When distributing a data set, it is important to document the meaning (e.g., units) and valid values of its variables.
-This is typically done with a _data dictionary_ (also called a _codebook_).
-In the following, we will demonstrate how to create a simple data dictionary using the R package [`pointblank`](https://rstudio.github.io/pointblank/).
-You can install it now using:
-
-::: {.column-margin}
-![pointblank: Data Validation and Organization of Metadata for Local and Remote Tables](images/pointblank.png){width=250px}
-:::
-
-```{.r filename="Console"}
-renv::install("pointblank")
-```
-
-You can put the code that follows for creating the data dictionary into a new file called `create_data_dictionary.R`.
-
-First, we write down everything we know about the data set. This includes:
-
-- a general description of the data set
-- descriptions of all columns
-- valid values, where applicable
-
-```{r, filename="create_data_dictionary.R"}
-table_info <- c(
-  title = "palmerpenguins::penguins",
-  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
-)
-descriptions <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-
-vals <- list(
-  species = c("Adelie", "Gentoo", "Chinstrap"),
-  island = c("Torgersen", "Biscoe", "Dream"),
-  sex = c("male", "female"),
-  year = c(2007, 2008, 2009)
-)
-```
-
-Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, or contact information.
-In this case, as the data have already been published, we only store a reference to its source:
-
-```{r, class.output = "r code-overflow-wrap", echo = FALSE}
-
-# We have provided the data set as CSV file to the readers.
-# Therefore, we cannot assume or require that readers have
-# the R package palmerpenguins installed. Instead, we create
-# the citation on our end and hide how we obtained it.
-
-dat_source <- citation("palmerpenguins", auto = TRUE) |>
-  format(bibtex = FALSE, style = "text")
-dat_source |>
-  gsub("\\n", " ", x = _) |>
-  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
-  paste0("dat_source <- \"", ... = _, "\"") |>
-  cat()
-```
-
-Then, we use `pointblank` to create a data dictionary with this information.
-
-```{r, filename="create_data_dictionary.R"}
-vals <- sapply(vals, \(x) {
-  paste0(
-    "(",
-    knitr::combine_words(x, and = " or ", before = "`", after = "`"),
-    ")"
-  )
-})
-
-dat <- read.csv("data.csv")
-
-dict <- pointblank::create_informant(
-  dat,
-  tbl_name = NA,
-  label = table_info[["title"]],
-  lang = "en"
-) |>
-  pointblank::info_tabular(
-    Description = table_info[["description"]],
-    Source = dat_source
-  ) |>
-  pointblank::info_columns_from_tbl(stack(descriptions)[2:1]) |>
-  pointblank::info_columns_from_tbl(stack(vals)[2:1]) |>
-  pointblank::get_informant_report(
-    title = "Data Dictionary for `data.csv`"
-  )
-
-dict
-```
-
-Finally, we can store the data dictionary inside an HTML file and put the HTML file into the project folder as well.
-
-```{.r filename="create_data_dictionary.R"}
-pointblank::export_report(dict, filename = "data_dictionary.html")
-```
-
-For a more elaborate introduction to `pointblank`, you can read their [Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html).
-
-One could go even further by making the information machine-readable in a standardized way.
-We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
-
-::: {#nte-frictionless .callout-note collapse="true"}
-### Create Machine-Readable Variable Documentation
-
-This example demonstrates how the title and description of the data set, the description of the variables and their valid values are stored in a machine-readable way.
-As before, we also provide a reference to the source.
-
-```{r, class.output = "r code-overflow-wrap", echo = FALSE}
-
-# We have provided the data set as CSV file to the readers.
-# Therefore, we cannot assume or require that readers have
-# the R package palmerpenguins installed. Instead, we create
-# the citation on our end and hide how we obtained it.
-
-citation("palmerpenguins", auto = TRUE)$url |>
-  paste0("dat_source <- \"", ... = _, "\"") |>
-  cat()
-```
-
-Generally, metadata are either stored embedded into the data or externally, for example, in a separate file.
-We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately.
-Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).
-
-Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released.
-As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information.
-We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-Finally, the package is used to create a metadata file that contains the schema.
-
-```{.r filename="Console"}
-# Read data and create schema
-dat_filename <- "data.csv"
-dat <- read.csv(dat_filename)
-dat_schema <- frictionless::create_schema(dat)
-
-# Add descriptions to the fields
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  c(x, description = descriptions[[x$name]])
-})
-
-# Record valid values
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  if (x$name %in% names(vals)) {
-    modifyList(x, list(constraints = list(enum = vals[[x$name]])))
-  } else {
-    x
-  }
-})
-
-# Define missing values
-dat_schema$missingValues <- c("", "NA")
-
-# Create package with license info and write it
-dat_package <- frictionless::create_package() |>
-  frictionless::add_resource(
-    resource_name = "penguins",
-    data = dat_filename,
-    schema = dat_schema,
-    title = table_info[["title"]],
-    description = table_info[["description"]],
-    licenses = list(list(
-      name = "CC0-1.0",
-      path = "https://creativecommons.org/publicdomain/zero/1.0/",
-      title = "CC0 1.0 Universal"
-    )),
-    sources = list(list(
-      title = "CRAN",
-      path = dat_source
-    ))
-  )
-frictionless::write_package(dat_package, directory = ".")
-```
-
-This creates the metadata file `datapackage.json` in the current directory.
-Make sure it is located in the same folder as `data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
-:::
-
-Having added the data and its documentation, one can view and record the utilized packages with `renv`...
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-...and go through the commit routine:
-
-```{.bash filename="Terminal"}
-git status
-git add .
-git commit -m "Add data"
-```
-
-## Add Code
-
-In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file.
-The manuscript contains code together with a written narrative.
-Download the two files to your computer and put them into your project folder.
-
-- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
-- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}
-
-The manuscript explores differences in bill length between male and female penguins, feel free to read through it.
-
-::: {#imp-take-copyright-seriously .callout-important}
-### Take Copyright Seriously
-
-If you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so.
-Only build on existing work for which you receive an express grant of relevant rights.
-How do you know you are allowed to copy, edit, and share the two files linked above?
-:::
-
-:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
-### Hint for "Take Copyright Seriously"
-
-Have a look at the [about page](about.qmd).
-::::
-
-As the manuscript uses some new packages, install them with:
-
-```{.r filename="Console"}
-renv::install()
-```
-
-The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
-It can be installed in the project using the following command:
-
-```{.bash filename="Terminal"}
-quarto add --no-prompt wjschne/apaquarto
-```
-
-::: {#tip-apaquarto .callout-tip}
-#### Not a Psychologist?
-
-If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.
-
-Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:
-
-```{.yml filename="Manuscript.qmd"}
-format:
-  pdf:
-    pdf-engine: lualatex
-    documentclass: scrartcl
-    papersize: a4
-```
-:::
-
-Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents.
-A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
-
-```{.bash filename="Terminal"}
-quarto install tinytex
-```
-
-You should now be able to render the document using Quarto:
+After adding the data analysis, render the manuscript again:
 
@@
 
-This should create a PDF file called `Manuscript.pdf` in your project folder.
-
-::: {#tip-update-quarto .callout-tip}
-If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
-:::
-
-With the code being added, one can use `renv` again to view and record the new packages:
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-::: {#tip-renv-status .callout-tip}
-Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
-:::
-
-Finally, make your changes known to Git:
+Then, make your changes known to Git:
 
@@
 git add .
-git commit -m "Add manuscript"
+git commit -m "Add data analysis"
 ```
@@
 
-- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
+- Data may be collected from social network sites using their APIs^[An **a**pplication **p**rogramming **i**nterface provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
 - an analysis may be conducted with the help of AI providers.
@@
 
-When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:
+When you start a new session from the project folder, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:
 
@@
 
-### Coding Best Practices
+## Style Manuscript
 
-Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.
+To format manuscripts according to the requirements of a particular journal, [multiple Quarto extensions are available](https://quarto.org/docs/extensions/listing-journals.html).
+In the following, we will demonstrate the usage of [`apaquarto`](https://wjschne.github.io/apaquarto/), which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
+Of course, you may also use an extension for a different journal.
 
-- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root.
-For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
+`apaquarto` can be installed in the project folder using the following command:
 
-- __Keep it simple.__ Add complexity only when you must.
-Whenever there's a boring way to do something and a clever way, go for the boring way.
-If the code grows increasingly complex, refactor it into separate functions and files.
-
-- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.
-
-- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.
-
-- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).
-
-- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
-Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
-We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.
-
-- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
-
-- __Test your code.__ Test your code with scenarios where you know what the result should be.
-Turn bugs you discovered into test cases.
-Use linting tools^[A linting tool analyzes your code without actually running it.
-Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
-
-- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality.
-R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/).
-Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013.
-And even if you don't follow a style guide, be consistent.
-
-This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:
-
-- "Tidy design principles" [@Wickham2023Design]
-- "The Good Research Code Handbook" [@Mineault2021]
-- "Quality assurance of code for analysis and research" [@UKGovernment2020]
-- "The Art of UNIX Programming" [@Raymond2003]
-
-> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
-> 
-> --- @Fowler1999, p. 15
-
-## Cite Data and Software
-
-If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.
-
-### Data
-
-Put simply, all data relied upon should be cited to allow for precise identification and access.
-From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
-
-> **Principle 1 – Importance**: "Data should be considered legitimate, citable products of research.
-Data citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications."
-> 
-> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited."
-> 
-> **Principle 5 – Access**: "Data citations should facilitate access to the data themselves and to such associated metadata, documentation, code, and other materials, as are necessary for both humans and machines to make informed use of the referenced data."
-> 
-> **Principle 7 – Specificity and Verifiability**: "Data citations should facilitate identification of, access to, and verification of the specific data that support a claim.
-Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific time slice, version and/or granular portion of data retrieved subsequently is the same as was originally cited."
-
-Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
-
-::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
-#### Hint for citing the data set
-
-As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:
-
-```{r, echo = -1}
-invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
-citation("palmerpenguins")
+```{.bash filename="Terminal"}
+quarto add --no-prompt wjschne/apaquarto
 ```
 
-As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).
+This downloads the extension into the folder `_extensions` of your project.
 
-Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`.
-Then, add a sentence to the manuscript such as follows:
+Then, change the `format` in the YAML header of your manuscript as follows:
 
-```md
-The analyzed data are by @horst2020.
+```{.yml filename="Manuscript.qmd"}
+format:
+  apaquarto-pdf:
+    documentmode: man
+    keep-tex: true
+    pdf-engine: lualatex
+    a4paper: true
 ```
 
-Render the document to check that the citation is displayed properly.
+You can now render the manuscript with the new format:
 
@@
 ```
+
+::: {#tip-update-quarto .callout-tip}
+If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
 :::
 
-### Software
+Again, make your changes known to Git:
 
-When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies.
+```{.bash filename="Terminal"}
+git status
+git add .
+git commit -m "Apply custom manuscript style"
+```
+
+## Add Code Citation and Attribution
+
+As with data, there are multiple reasons to indicate when using code by others.
+First, there is software which is actually copied into the own project folder.
+For example, this is the case for the Quarto extension that you downloaded earlier.
+These cases matter from a copyright perspective and you need to make sure to have permission for redistribution.
+
+Under which license is `apaquarto` made available?
+
+::: {#tip-apaquarto-license .callout-tip collapse="true"}
+### `apaquarto` License (Solution)
+
+According to its GitHub page, `apaquarto` is licensed under CC0\ 1.0. This means that it can be used and distributed even without attribution.
+:::
+
+We recommend adding a short paragraph to `LICENSE.txt` to describe its license.
+
+::: {#tip-apaquarto-licensetxt .callout-tip collapse="true"}
+### Addition to `LICENSE.txt` (Solution)
+
+```{.txt .code-overflow-wrap filename="LICENSE.txt"}
+"apaquarto" stored in "_extensions/wjschne/apaquarto" by W.
+Joel Schneider available from <https://github.com/wjschne/apaquarto> is licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>
+```
+
+Again, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).
+:::
+
+Then, there is software that is used (e.g., during the data analysis), but not copied into the project folder. Here, questions of copyright due to redistribution do not apply.
+
+Of course, being transparent about used software for reasons of reproducibility and academic integrity matters in any case.
 You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not.
@@
 
-In practice, the first step is to identify all pieces of software the project relies on.
+In practice, you would identify all pieces of software the project relies on.
 A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before.
-Then there are the individual R packages, Quarto extensions, and $\TeX$ packages.
+Then there are the individual R packages^[By default, `renv` avoids copying them into the project folder.
+And even that happens (e.g., by using `renv::isolate()`) they are excluded from Git by default.], Quarto extensions, and $\TeX$ packages.
 All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper.
@@
 git add .
-git commit -m "Cite data and software"
+git commit -m "Add software citation"
 ```
+
+## Coding Best Practices
+
+Although we provide you with example code in this tutorial, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.
+
+- __Use project-relative paths.__ When you refer to a file within your project folder, write paths relative to the root (the uppermost folder in your project).
+For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
+
+- __Keep it simple.__ Add complexity only when you must.
+Whenever there's a boring way to do something and a clever way, go for the boring way.
+If the code grows increasingly complex, refactor it into separate functions and files.
+
+- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.
+
+- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.
+
+- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).
+
+- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
+Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
+We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.
+
+- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
+
+- __Test your code.__ Test your code with scenarios where you know what the result should be.
+Turn bugs you discovered into test cases.
+Use linting tools^[A linting tool analyzes your code without actually running it.
+Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
+
+- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality.
+R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/).
+Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013.
+And even if you don't follow a style guide, be consistent.
+
+This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:
+
+- "Tidy design principles" [@Wickham2023Design]
+- "The Good Research Code Handbook" [@Mineault2021]
+- "Quality assurance of code for analysis and research" [@UKGovernment2020]
+- "The Art of UNIX Programming" [@Raymond2003]
+
+> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
+> 
+> --- @Fowler1999, p. 15
 
@@
 `renv` only records the versions of R packages and of R itself.
-This means that everything we have not decided to cite in the previous step is not documented anywhere.
-We will cover system dependencies when [creating a README](make_readme.qmd).
+This means that everything we have not decided to cite or attribute is not documented anywhere.
+We will cover system dependencies when [creating a `README`](make_readme.qmd).
 For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).
@@
 ```
-
-You are now all set up to prepare your project for sharing!

### 2025-01-25 copyright.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

---
title: "Copyright"
format:
  html:
    include-in-header:
      - text: |
          <style>
          .unstyled-list > ul {
              list-style-type: none;
              padding-left: 1rem;
          }
          .unstyled-list summary {
              cursor: pointer;
          }
          </style>
---

By default, everything you put into the project folder will be shared publicly. In many instances, this will also include works by others than yourself or your co-authors, because you may use existing data, images or software. In these cases, you should make sure to only share them to the extent you are allowed to. In the following, we will be focusing on copyright, but of course, there can also be other restrictions.

::: {#nte-copyright-tldr .callout-note}
## Copyright TL;DR

Every expression of an idea, such as a literary or an artistic works, is automatically covered by copyright. Put simply, this means that nobody else than the copyright holder is allowed to copy, modify, or share it. Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights. Certain exceptions and limitations to copyright (such as fair use or the right to quote) may obviate the need for a license.

With regard to copyright, free/open licenses can give permission to share a work by others. These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.

```{=html}
<div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
  <iframe src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
</div>
```
:::

## Introduction to Copyright and Licenses

Whenever authors create a literary or artistic work (such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_] in most countries limits other people from copying, modifying, and sharing it without the author's express permission. This even applies if the author makes it available to others (e.g., on their website): First and foremost others are not allowed to copy, modify, or share it. Importantly, this legal default of "all rights reserved" was created to benefit publishers, not authors [@Fogel2006], and runs counter to many cultural and scientific processes. Copyright licenses enable authors to free up their works for reuse by others.

Note that copyright law [varies between countries](https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory) and the rule of territoriality (_lex loci protectionis_) applies: Applicable are the laws of the country where a work is used. Also, if the work was created as part of the author's job, it might be their employer who holds the copyright to the work, depending on the country and contract.

A license is a legal document that regulates what other people than the author can and cannot do with a copyright-protected work, and under which conditions. The license may be exclusive, such that it is only granted to one recipient, effectively limiting the author in dealing with their own work. It can also be non-exclusive, thus giving rights to others without the author losing them. To avoid that every author who would like to share their work writes their own license, pre-formulated "boilerplate" licenses have been created. Recognizing them facilitates both sharing (on the side of the author) and re-using (on the side of the recipient).^[In fact, authors should not create their own license texts if the purpose is to share their work freely and openly with others. By using a boilerplate license others know what to expect, also because it may have already been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]

::: {.column-margin}
For the purpose of this tutorial, by _license_ we mean _copyright license_.
:::

An important class of boilerplate licenses are _free/open_ licenses. Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015]. Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes. By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work. Software licenses may additionally require to make the source code available to everybody the software is shared with and often require to display the full text of the license upon distribution.

::: {#nte-terminology .callout-note collapse="true"}
### A Note on Terminology

The terms _free_ and _open_, especially with regard to software, come with a history. The Free Software Foundation (FSF) was founded in 1985 to protect "four essential freedoms" [-@FSF2024] of a program's user. These are the freedoms to use, study, share, and improve a program. Software whose users legally and practically have these freedoms (because, among other things, they have access to its source code) is considered _free_. The four freedoms are seen as vital for a society as a whole in the sense that they enable sharing, cooperation and ultimately freedom in general. The FSF maintains [a list of software licenses](https://www.gnu.org/licenses/license-list.en.html#SoftwareLicenses) that it considers to be protecting the four freedoms. Sometimes the term _libre_ (Spanish and French for _free_) is used to make a distinction from _gratis_ software. You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).

The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach. It is concerned with developing high-quality software, for which everyone's ability to obtain, modify and contribute back the source code is considered beneficial. Access to the source code is one out of multiple conditions for software to be considered _open source_ by the OSI [-@OSI2007], which equally maintains [a list of approved licenses](https://opensource.org/licenses).

Access to the source code is a necessary, but not sufficient requirement both for _free software_ and for _open source software_. Conversely, both can (and frequently are) sold for money, as their respective criteria only apply once one has access to the software. Throughout this tutorial, we write "free/open license" to mean a license that is approved by either the FSF or the OSI. Software which is neither but makes available its source code is sometimes referred to as _source-available software_.

Apart from software, the term _open access_ has often been used for works that are available at no cost. For example, this is the commonality of bronze, green, hybrid, gold, and diamond/platinum open access _articles_, which otherwise vary in the rights that are granted to readers. In 2002, the Budapest Open Access Initiative declared that _open access_ additionally includes the right to use articles for any purpose, and in 2003, the Bethesda Statement and the Berlin Declaration added the right to make derivative works.

Two other notable definitions include the _Open Definition_ [@OKFN2016], which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015], for which the open editing phase began in 2006. They are largely viewed as compatible with one another.

:::: {.column-margin}
![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
::::
:::

### Can I Share That Work?

To determine whether a particular work can be included in the project folder and shared publicly, the first step is to search for its license. For example, this tutorial has an [about page](about.qmd) detailing its licensing, whereas other websites sometimes state the license in the footer. Other places to look include a project's `README` file or a dedicated `LICENSE` or `COPYING` file. If a project has been stored in a research repository, yet another means of communicating its copyright status might exist. For some repositories, external attribution generators exist:

- Wikimedia Commons: [Attribution Generator](https://lizenzhinweisgenerator.de/?lang=en)
- Various software repositories: [CiteAs](https://citeas.org/)

::: {#tip-private-use .callout-tip}
#### Consider Complicated Licensing Situations

For works by one or only few authors, it is possible that one license consistently applies to the whole content. More often than not, however, the situation is a little more complicated:

- Different parts of a work may be available under different licenses: A document may incorporate direct quotes or images by others. In this case, the license of the document may not cover those. In the best case, the authors indicate which parts of the work are covered by which license.
- A work may be available under different licenses at the choice of the (re-)user. This is also called multi-licensing. For example, authors might allow to user a work under any version of a license.
- In contrast, it may also be the case that one has to comply with the conditions of multiple licenses at the same time. This is the case, for example, if a derivative work is published under a different license than the original work.
:::

If you cannot find any license, the work might be in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law. For example, the work may not cross the [threshold of originality](https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality), it may not [qualify for copyright protection](https://commons.wikimedia.org/wiki/Commons:Unprotected_works) -- such as official works by the government --, or the copyright protection might have expired. According to the [Berne convention](https://en.wikipedia.org/wiki/Berne_Convention), copyright expires at earliest 50 years after the author's death, but in [many countries](https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country) it expires even later -- you can use a [public domain calculator](https://lawflow.org/public-domain-calculators/) to check. If no license notice is present and the work is not in the public domain, it is _in copyright_ and you are generally not allowed to share it with others. You can, however, ask the author to put the corresponding work under a free/open license -- Wikimedia Commons provides an [email template](https://commons.wikimedia.org/wiki/Commons:Email_templates) to obtain such a permission. Without permission you can check you national laws for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_] Examples include the right to quote, [trivial use](https://commons.wikimedia.org/wiki/Commons:De_minimis), and the US-American [fair use](https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?).

Having found a license, you need to determine (1) whether it is a standard (boilerplate) license, (2) whether it is a free/open, and (3) whether it is appropriate for the particular type of work and use case.

1. __Standard licenses__ are usually stated by their name (or an abbreviation), which may include a version number. Most standard licenses were assigned an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification, though the identifier is often not indicated for a particular work (which would be best practice). Also, not all standard licenses have SPDX identifiers. You can view a [list of all licenses with SPDX identifiers here](https://spdx.org/licenses/).
2. __Free/open__: We have compiled lists of the most common free/open licenses below, categorized by the type of work they apply to. Be sure to look for an exact match, as there are often many licenses with very similar names. If a particular license is not listed below, we also provide links to other resources which you can use to determine whether a license is free/open.
3. __Type of work and use case__: Many licenses were written with a particular type of work in mind, along with a certain use case. Software licenses, for example, may consider that the source code is the preferred form for making modifications and that software can be encumbered by patents, while licenses for data can differentiate between the database and any works produced from it. Take caution if a license applies to a type work for which it was not originally created and conflicts with your use case. Sometimes, the license name hints at the intended type of work, but when in doubt always read the full license text (and/or consult with a lawyer) to determine whether the conditions are acceptable to you.

::: panel-tabset
### Software

For a software license to be considered free/open, it must be approved by either the [Open Source Initiative](https://opensource.org/licenses) (OSI) or the [Free Software Foundation](https://www.gnu.org/licenses/license-list.html#SoftwareLicenses) (FSF). Some of the most common free/open software licenses are listed below:

:::: unstyled-list
- <details>
  <summary>GNU Affero General Public License v3.0 (`AGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/agpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU General Public License v3.0 (`GPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/gpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU Lesser General Public License v3.0 (`LGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.gnu.org/licenses/lgpl-3.0.html#license-text)
  </details>
- <details>
  <summary>Apache License 2.0 (`Apache-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes

  [Link to full text](https://www.apache.org/licenses/LICENSE-2.0)
  </details>
- <details>
  <summary>Mozilla Public License 2.0 (`MPL-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.mozilla.org/en-US/MPL/2.0/)
  </details>
- <details>
  <summary>MIT License (`MIT`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/mit/#license-text)
  </details>
- <details>
  <summary>BSD 3-Clause "New" or "Revised" License (`BSD-3-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-3-clause/#license-text)
  </details>
- <details>
  <summary>BSD 2-Clause "Simplified" License (`BSD-2-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-2-clause/#license-text)
  </details>
::::

If you did not find a particular software license in this list and consult the resources by the OSI and/or FSF linked above, be cautious because they also approve special-purpose licenses that are unfit for software in the general case. For example, the license CC0\ 1.0 expressively does not grant patent rights and one should only use software under that license if it is most likely not covered by patents.

### Writing, Images, Audio, Video

Many authors of text and media put their work under a license created by Creative Commons. Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use or the distribution of modified works (such as sharing a translated version). Therefore, we recommend to only use works made available to you under CC0 or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`), but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`). In their most recent version this includes the following licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text](https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title)
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text](https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title)
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text](https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title)
  </details>
::::

Of course, if you encounter a Creative Commons license with a previous version that satisfies the description above, it can be also considered free/open. If you encounter a license not from Creative Commons and want to determine whether it is free/open, you can consult the [lists of licenses accepted at Wikimedia Commons](https://commons.wikimedia.org/wiki/Commons:Copyright_tags). Note, however, that some of the licenses there might be inappropriate for text and media. For example, licenses for software or for documentation commonly require to distribute a work together with the full license text, which is impractical for usage in print media.

### Data

Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are. Additionally, in some jurisdictions, collections of data (i.e., databases) can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.

As a consequence, a database may be licensed differently than its content. For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists, but the license for the database as a whole might have been chosen by the curator, including the selection of entries, field names and any _sui generis_ database rights.  Watch out for any signs that the license for the database and its content are not the same. In the following, we list some popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text](https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title)
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text](https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title)
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text](https://creativecommons.org/licenses/by-sa/4.0/legalcode)
  </details>
::::

Note that Creative Commons licenses prior to version 4.0 are not suitable for databases. You might also encounter the following, less popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>Open Data Commons Public Domain Dedication & License 1.0 (`PDDL-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text](https://opendatacommons.org/licenses/pddl/1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 2.0 (`CDLA-Permissive-2.0`)</summary>
  __Major conditions when sharing works__:

  - Add license text

  [Link to full text](https://cdla.dev/permissive-2-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 1.0 (`CDLA-Permissive-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://cdla.dev/permissive-1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Sharing 1.0 (`CDLA-Sharing-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Copyleft

  [Link to full text](https://cdla.dev/sharing-1-0/)
  </details>
::::

Finally, you should know about the following two free/open licenses because they only cover the database, but not its individual contents or the works produced from it. If any of these licenses is used, the license for the content must be stated separately.

:::: unstyled-list
- <details>
  <summary>Open Data Commons Open Database License v1.0 (`ODbL-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Anti-DRM
  - Copyleft

  [Link to full text](https://opendatacommons.org/licenses/odbl/1-0/)
  </details>
- <details>
  <summary>Open Data Commons Attribution License v1.0 (`ODC-By-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://opendatacommons.org/licenses/by/1-0/)
  </details>
::::

If you encounter a license not listed here and want to determine whether it is free/open, you can consult the [lists of licenses accepted at Wikimedia Commons](https://commons.wikimedia.org/wiki/Commons:Copyright_tags). Note, however, that some of the licenses there might be inappropriate for data, because they do not mention _sui generis_ database rights.
:::

::: {#cau-license-versions .callout-caution}
### License Versions Are Important

You may have noticed that in many cases a version number is part of the license name. This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses. For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002](https://creativecommons.org/weblog/entry/3476). Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights have been covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013). There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically (i.e., you have to follow the terms of that exact version). CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).

For the GNU licenses it is even recommended to state whether a work is licensed _only_ under the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
:::

### Under Which Conditions?

Finally, if you verified that the work has been made available to you under a free/open license, you need to understand the conditions under which you are allowed to include the work in your project folder and share it with others -- regardless of whether you modified it. In the following, we will explain the most important conditions of the licenses listed above. For other licenses we recommend reading their text in full and/or consulting with a lawyer. The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com](https://choosealicense.com/appendix/), which you can use to get a general idea of a license (but which does not obviate reading the license text in full). Specifically for software licenses the _Open Source Automation Development Lab_ provides [useful checklists detailing the respective obligations](https://www.osadl.org/Checklists/).^[Accessing the checklists requires a (free) account on their website.]

::: {#nte-private-use .callout-note}
#### Private Use

In the following, we only discuss the conditions under which works may be _shared_. This is because free/open licenses put no conditions on private use, where the work is not made available to others.
:::

#### Attribution

Licenses may require that you provide attribution to the author(s)^[or others whose names are provided with the original work] and to retain any copyright notices,^[Such as `Copyright (c) 2025 Jane Doe`] although the specifics vary between licenses. For example, Creative Commons licenses other than CC0 require that the name of the author, the title of the work, and a link to the work are provided.

Projects with many authors often explicitly allow for an attribution to the collective as a whole rather than being mentioned individually. For example, the flexible attribution requirements of Creative Commons licenses allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009]. And software projects may use copyright notices that only contain the project name, such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].

Where to put the attribution depends on the type of the work, but in general it is best to keep it close to where the work is used and to provide all required information. Only linking to a page that provides the attribution (as with Wikipedia articles) most likely will not suffice in general.

#### State Changes

If a work has been modified, licenses may require that this is indicated. For example, if an image under a Creative Commons license other than CC0 is reused in a cropped version, reusers could add "adapted from original" to indicate that changes have occurred.

#### Indicate License / Add License Text

Licenses may either require that one mentions the applicable license by name and/or link ("Indicate license") or to even copy the full license text along with the used work ("Add license text"). Of course, the latter may not be feasible if the work is printed and smaller than the license text. For example, if the license of an image requires to print the full license text, the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].

#### Disclose Code

There are various situations in which one interacts with software without having access to the underlying source code. For example, software can be distributed as a binary ("executable") file or it can be used over the internet. In both cases, licenses may require that the source code is provided to its users.

#### Anti-DRM

You may have encountered images or videos which you could not download due to a copy protection. Such a protection is also called **d**igital **r**ights **m**anagement (DRM). For works published under a license with an anti-DRM clause, reusers may not apply DRM to the work or must provide a DRM-free version in parallel (depending on the exact license).

#### Copyleft

If you modified the work and put a sufficient amount of creativity into it, the result is called a _derivative work_. It is (again) protected by copyright and even if you share it, nobody except you is allowed to copy, modify, or distribute it further. You can put it under a free/open license but don't have to. The only exception is if the original work has been made available under a "copyleft" license. In that case, if you share the work with others you must put it under the same license as the original work^[or sometimes also under a later version or under a compatible license] -- or you commit a copyright violation.^[Note, however, that [copyleft licenses are not automatic](https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic): A derivative work of a copylefted work is not "implicitly" under the same license -- others would commit a copyright violation if they just took it. Only if the derivative author explicitly makes it available under the same license, others may use the derivative work under the same terms.] This explains the origin of the term _copyleft_, as it "turns copyright around" to make works permanently free: ©\ →\ 🄯. In the realm of Creative Commons licenses, copyleft is also called _share alike_. The opposite of a copyleft license is sometimes also called a _permissive_ license.

What is considered creation of a derivative work varies between jurisdictions, but the following are typical examples: translating a text into a different language, taking a picture of an artwork, adding a song to a video, adapting a computer program to fit own use cases, or taking a [screenshot of a computer program](https://commons.wikimedia.org/wiki/Commons:Screenshots). In all of these cases, if the original work is under a copyleft license, then sharing of the resulting derivative work is only allowed if it is put under the same license.

The question is then: Which uses of a work do not create a derivative work? In many cases, if an unaltered excerpt "is used to illuminate an idea or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made. The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p. 44] and not be "merged with other material into a new and larger work" [@Kreutzer2024, p. 45]. For example, embedding images in a document and clearly separating them from the rest or assembling works as part of a collection typically would not create a derivative work. However, due to a lack of case law, there is only little guidance on what constitutes a derivative work in general.

Also note that the copyleft licenses we discuss here do not mandate sharing. Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015]. This means that if you only use a work internally, you do not need to share your derivative works.

::: {#nte-weak-strong-copyleft .callout-note collapse="true"}
##### Weak and Strong Copyleft

For software, copyleft licenses come in two flavors:

- __Weak copyleft licenses__ only require that modifications to the software itself are licensed under the same or a compatible license if shared. For example, if you create and publish software under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license. However, people who merely _use_ your software in their own work which they make publicly available can choose any license for it.
- __Strong copyleft licenses,__ on the other side, insist that any larger works that use the copyleft-licensed software must also be licensed under the same or a compatible license if shared. For example, if your software were to be put under a strong copyleft license, everybody publishing software that uses your software would need to put it under the same license. Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024GPL].
:::

### Further Remarks

The description of the conditions above is a simplified summary -- read the full text of the license before sharing the respective work with others. For example, Creative Commons licenses other than CC0 also require that

- you retain a notice about the disclaimer of warranties,
- you retain an indication of any previous modifications,
- the attribution does not imply endorsement by the original author, and
- you remove the attribution again upon request by the original author.

::: {#tip-cc-licenses-checklist .callout-tip}
#### Required Information under Creative Commons Licenses

For most Creative Commons licenses, providing the following information is required (if available):

- [ ] Author name
- [ ] Copyright notice
- [ ] Title of the work
  - [ ] Link to the work
- [ ] License name
  - [ ] Link to the license
- [ ] Link to warranty disclaimer
- [ ] Indicate modification
:::

Obviously, if you are in doubt whether a condition of the license applies in your use case, it is better to either follow it or to ask the author for clarification or explicit permission. This is especially important given the existence of [_copyleft trolls_](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling): People who distribute works under a free/open license but send out statutory damages claims after minor violations without providing reusers an opportunity to fix their error. In addition, in many European countries attribution may be necessary regardless due to the author's moral rights. For these reasons, we recommend that you provide attribution for all works under a free/open license -- even if the license does not strictly require it (e.g., CC0). For this reason, you need to make note of every foreign work included in your project folder and record its license.

::: {#imp-non-copyright-restrictions .callout-important}
#### Non-copyright Restrictions

The free/open licenses we discuss here mostly grant copyrights (with some of them explicitly granting patent rights and _sui generis_ database rights). Therefore, you may lack other rights such as personality, privacy, moral, or trademark rights. For example, [sharing photos that depict people](https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people) is not only a matter of copyright, but also of privacy rights. Also, their commercial use may require the consent of the depicted person. Finally, note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
:::

## Adding an Image

Let's practice what you learned by adding an image to the manuscript. We'll use [this picture of a penguin from Flickr](https://flic.kr/p/2pEKnUr). First, try to answer the following questions:

1. Under which license is the picture available?
2. Is it free/open?
3. Is it appropriate for the type of work and use case?
4. Under which conditions may the work be shared?

::: {#tip-example-license .callout-tip collapse="true"}
### License of Penguin Picture (Solution)

1. The picture has been put under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/)
2. The license is free/open
3. The license is appropriate for images
4. The license contains no (major) conditions
:::

You can now download the image and move it to your project folder (maybe call it `penguin.jpg`). In your manuscript file, find the line that says "add image here", remove it, and add the image instead, including proper attribution.

::: {#tip-example-license .callout-tip collapse="true"}
### Attribution for Penguin Picture (Solution)

Although the license CC0 doesn't require it, we still recommend providing proper attribution.

- Author name: `steve b`
- Copyright notice: (not available)
- Title of the work: `Chinstrap Penguin`
  - Link to the work: <https://www.flickr.com/photos/192320315@N02/53603388557/> or <https://flic.kr/p/2pEKnUr>
- License name: `CC0 1.0`
  - Link to the license: <https://creativecommons.org/publicdomain/zero/1.0/>
- Link to warranty disclaimer: (not available)
- Indicate modification: (no existing modification notices)

```{.md .code-overflow-wrap}
!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" licensed under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/)](penguin.jpg)
```

If it is likely that the article gets printed, you might want to ensure that the link to the work is retained in that case:

```{.md .code-overflow-wrap}
!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/) from [`flic.kr/p/2pEKnUr`](https://flic.kr/p/2pEKnUr)](penguin.jpg)
```
:::

Let's render the document and see if the changes were successful (open `Manuscript.pdf` afterwards):

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```

In addition, we recommend creating a file called `LICENSE.txt` to describe the license of the image:

```{.txt .code-overflow-wrap filename="LICENSE.txt"}
"Chinstrap Penguin" stored in "penguin.jpg" by "steve b" available from <https://flic.kr/p/2pEKnUr> is licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>
```

If the license required adding the full license text, you would also need to copy it to `LICENSE.txt`.

Finally, you can store the changes in the version control system:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add image"
```

## Wrap-up

If you would like to learn more about copyright and licenses you might find the following resources interesting:

- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2024
- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
- "Rechtsfragen bei Open Science: Ein Leitfaden" by @Kreutzer2021

### 2025-01-25 data.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

--- 2024-12-10 setup.qmd
@@
 ---
-title: "Project Setup"
+title: "Add Data"
 engine: knitr
@@
 
-We will start by setting up a simple example of a reproducible report.
-
-## Create Quarto Project
-
-First, we will need to create a new Quarto project.
-
-If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
-
-![](images/setup_01.png){width=500px fig-align="center"}
-
-Here, select _New Directory_
-
-![](images/setup_02.png){width=600px fig-align="center"}
-
-And choose the project type _Quarto Project_.
-
-![](images/setup_03.png){width=600px fig-align="center"}
-
-Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.
-
-As we will use Git to track the version history of files, be sure to check _Create a git repository_.
-If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".
-
-::: {.column-margin}
-![renv: A dependency management toolkit for R](images/renv.svg){width=250px}
-:::
-
-Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on.
-Using it makes it easier for others to view and obtain them at a later point in time.
-Therefore make sure that the box _Use renv with this project_ is checked.
-Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".
-
-If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.
-
-![](images/setup_04.png){width=600px fig-align="center"}
-
-Click on _Create Project_. Your RStudio window should now look similar to this:
-
-![](images/setup_05.png){.lightbox fig-align="center" alt="The project `code-publishing-exercise` opened in RStudio.
-The source pane to the top left has a Quarto file open called \"code-publishing-exercise.qmd\".
-The console pane to the bottom left indicates by its output that renv is active in the current project.
-The environment pane to the top right indicates that the environment is currently empty.
-The output pane to the bottom right shows the files in the current project."}
-
-If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.
-
-Make sure that your project is in a consistent state according to `renv` by running:
-
-```{.r filename="Console"}
-renv::status()
-```
-
-If it reports packages that are not used, synchronize the lock file using:
-
-```{.r filename="Console"}
-renv::snapshot()
-```
-
-::: {#nte-terminal .callout-note collapse="true"}
-### Without RStudio
-
-Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:
-
-```{.bash filename="Terminal"}
-quarto create project default code-publishing-exercise
-cd code-publishing-exercise/
-rm code-publishing-exercise.qmd
-git init
-git checkout -b main
-```
-
-Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Then, initialize `renv`:
-
-```{.r filename="Console"}
-renv::init()
-```
-:::
-
-You are now ready to stage and commit your files.
-You can either stage files separately or the whole project folder at once.
-If you do the latter, we recommend you to inspect the untracked changes before staging all of them:
-
-::: {.column-margin}
-In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
-:::
-
-```{.bash filename="Terminal"}
-git status
-```
-
-Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file.
-If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:
-
-```{.bash filename="Terminal"}
-git add .
-git commit -m "Initial commit"
-```
-
-If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file.
-If you don't check your changes before committing, you might accidentally commit something you'd rather not.
-
-::: {#tip-author-identity-unknown .callout-tip}
-If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:
-
-```{.bash filename="Terminal"}
-git config user.name "YOUR NAME"
-git config user.email "YOUR EMAIL ADDRESS"
-```
-
-Then, commit again.
-:::
-
-## Decide on Structure
-
-Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it.
-In general, the directory structure should facilitate understanding a project by breaking it into logical chunks.
-There is no single best solution, as a good structure depends on where a project's complexity lies.
-However, it is usually helpful if the different files and folders reflect the execution order.
-For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders.
-Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.
-
-Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021].
-We recommend you to follow the standards of your field.
-
-For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.
-
-## Add Data
+## Add Data File
 
 ```{r}
-#| echo: false
+#| include: false
 
+invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
+stopifnot(packageDescription("palmerpenguins", fields = "License") == "CC0")
 write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
@@
 
-You can now download the data set we have prepared for you and put it into your project folder: [`data.csv`](data.csv){download=""}
+You can now download the data set we have prepared for you and put it into your project folder:
+
+[data.csv](data.csv){.btn .btn-lg .btn-info download=""}
 
@@
 The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
-It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+It was made available by Allison Horst, Alison Hill, and Kristen Gorman under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
 
@@
 For reasons of reproducibility, this should include the data you analyze.
-Of course, you should only share them to the extent you are allowed to, taking into account:
+Of course, you should only share them to the extent you are allowed to.
+Besides copyright and similar rights, you need to take into account applicable privacy laws (e.g., the GDPR for European citizens) and contractual obligations (e.g., with your data provider).
 
-1. applicable privacy laws (e.g., the GDPR for European citizens),
-2. contractual obligations (e.g., with your data provider),
-3. copyright of the data and their particular structure, and
-4. any _sui generis_ database right.
+Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
+For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
+Another resource to look at is the presentation "Creating reproducible packages when data are confidential" by @Vilhuber2024.
+:::
 
-Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
-For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
-
-Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are.
-Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.
-As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
-:::
+## Add Data Dictionary
 
@@
 
-Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, or contact information.
-In this case, as the data have already been published, we only store a reference to its source:
+Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, applied preprocessing steps, or contact information.
+In our case, as the data have already been published, we only store a reference to its source.
 
-```{r, class.output = "r code-overflow-wrap", echo = FALSE}
+As the data set is from the R package `palmerpenguins`, if you have it installed you can use the function `citation()` to create such a reference:
+
+```{r, filename="create_data_dictionary.R"}
+dat_source <- citation("palmerpenguins", auto = TRUE) |>
+  format(bibtex = FALSE, style = "text")
+```
+
+Without the package `palmerpenguins` installed, you can find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation) and run the following:
+
+```{r}
+#| echo: false
+#| class-output: "r code-overflow-wrap"
+#| attr-output: 'filename="create_data_dictionary.R"'
 
@@
 # the citation on our end and hide how we obtained it.
-
-dat_source <- citation("palmerpenguins", auto = TRUE) |>
-  format(bibtex = FALSE, style = "text")
 dat_source |>
@@
 
-For a more elaborate introduction to `pointblank`, you can read their [Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html).
+For a more elaborate introduction to `pointblank`, you can read its "[Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html)".
 
@@
 
-```{r, class.output = "r code-overflow-wrap", echo = FALSE}
+```{r}
+#| echo: false
+#| class-output: "r code-overflow-wrap"
+#| attr-output: 'filename="Console"'
 
@@
 As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information.
-We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under CC0\ 1.0.
 Finally, the package is used to create a metadata file that contains the schema.
@@
 
-Having added the data and its documentation, one can view and record the utilized packages with `renv`...
+Having added the data and its documentation, one can view and record the utilized packages with `renv`:
 
@@
 
-...and go through the commit routine:
+## Add Data Citation and Attribution
 
-```{.bash filename="Terminal"}
-git status
-git add .
-git commit -m "Add data"
-```
-
-## Add Code
-
-In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file.
-The manuscript contains code together with a written narrative.
-Download the two files to your computer and put them into your project folder.
-
-- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
-- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}
-
-The manuscript explores differences in bill length between male and female penguins, feel free to read through it.
-
-::: {#imp-take-copyright-seriously .callout-important}
-### Take Copyright Seriously
-
-If you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so.
-Only build on existing work for which you receive an express grant of relevant rights.
-How do you know you are allowed to copy, edit, and share the two files linked above?
-:::
-
-:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
-### Hint for "Take Copyright Seriously"
-
-Have a look at the [about page](about.qmd).
-::::
-
-As the manuscript uses some new packages, install them with:
-
-```{.r filename="Console"}
-renv::install()
-```
-
-The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
-It can be installed in the project using the following command:
-
-```{.bash filename="Terminal"}
-quarto add --no-prompt wjschne/apaquarto
-```
-
-::: {#tip-apaquarto .callout-tip}
-#### Not a Psychologist?
-
-If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.
-
-Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:
-
-```{.yml filename="Manuscript.qmd"}
-format:
-  pdf:
-    pdf-engine: lualatex
-    documentclass: scrartcl
-    papersize: a4
-```
-:::
-
-Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents.
-A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
-
-```{.bash filename="Terminal"}
-quarto install tinytex
-```
-
-You should now be able to render the document using Quarto:
-
-```{.bash filename="Terminal"}
-quarto render Manuscript.qmd
-```
-
-This should create a PDF file called `Manuscript.pdf` in your project folder.
-
-::: {#tip-update-quarto .callout-tip}
-If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
-:::
-
-With the code being added, one can use `renv` again to view and record the new packages:
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-::: {#tip-renv-status .callout-tip}
-Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
-:::
-
-Finally, make your changes known to Git:
-
-```{.bash filename="Terminal"}
-git status
-git add .
-git commit -m "Add manuscript"
-```
-
-::: {#wrn-credentials .callout-warning}
-### Beware of Credentials
-
-Sometimes, a data analysis requires the interaction with online services:
-
-- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
-- an analysis may be conducted with the help of AI providers.
-
-In these cases, make sure that the code you check in to Git does not contain any credentials that are required for accessing these services.
-Instead, make use of environment variables which are defined in a location that is excluded from version control.
-When programming with R, you can define them in a file called `.Renviron` in the root of your project folder:
-
-```{.ini filename=".Renviron"}
-MY_FIRST_KEY="your_api_key_here"
-MY_SECOND_KEY="your_api_key_here"
-```
-
-When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:
-
-```r
-query_api(..., api_key = Sys.getenv("MY_FIRST_KEY"))
-```
-
-Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that contains credentials, you can follow @Chacon2024.
-:::
-
-### Coding Best Practices
-
-Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.
-
-- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root.
-For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
-
-- __Keep it simple.__ Add complexity only when you must.
-Whenever there's a boring way to do something and a clever way, go for the boring way.
-If the code grows increasingly complex, refactor it into separate functions and files.
-
-- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.
-
-- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.
-
-- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).
-
-- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
-Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
-We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.
-
-- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
-
-- __Test your code.__ Test your code with scenarios where you know what the result should be.
-Turn bugs you discovered into test cases.
-Use linting tools^[A linting tool analyzes your code without actually running it.
-Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
-
-- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality.
-R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/).
-Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013.
-And even if you don't follow a style guide, be consistent.
-
-This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:
-
-- "Tidy design principles" [@Wickham2023Design]
-- "The Good Research Code Handbook" [@Mineault2021]
-- "Quality assurance of code for analysis and research" [@UKGovernment2020]
-- "The Art of UNIX Programming" [@Raymond2003]
-
-> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
-> 
-> --- @Fowler1999, p. 15
-
-## Cite Data and Software
-
-If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.
-
-### Data
-
-Put simply, all data relied upon should be cited to allow for precise identification and access.
+All data relied upon should be cited in the manuscript to allow for precise identification and access.
 From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
@@
 
-Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
+Now, it's your turn to add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
 
-::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
-#### Hint for citing the data set
+::: {#tip-cite-palmerpenguins .callout-tip collapse="true"}
+#### Citing the Data Set (Solution)
 
-As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:
+You can find an appropriate BibTeX entry on [the package website](https://allisonhorst.github.io/palmerpenguins/#citation) or with the function `citation()`:^[Note that this function requires to have the respective package installed.]
 
-```{r, echo = -1}
-invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
-citation("palmerpenguins")
+```{r}
+#| class.output: "bib code-overflow-scroll"
+#| attr-output: 'filename="Bibliography.bib"'
+
+citation("palmerpenguins", auto = TRUE) |>
+  transform(key = "horst2020") |>
+  toBibtex()
 ```
 
-As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).
+Copy the BibTeX entry to the file `Bibliography.bib`. Then, find the line in the manuscript that says "cite data here" and replace it with a sentence such as the following:
 
-Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`.
-Then, add a sentence to the manuscript such as follows:
-
-```md
+```{.md filename="Manuscript.qmd"}
 The analyzed data are by @horst2020.
@@
 
-### Software
+While citation happens in the manuscript for reasons of academic integrity and reproducibility, to comply with any licenses you also may need to provide attribution within your project folder.
+Even though the data file we use here does not require attribution, we recommend adding a short paragraph to `LICENSE.txt`:
 
-When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies.
-You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not.
-As with data, citations should allow for exact identification and access.
-From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):
+```{r}
+#| echo: false
+#| class-output: "txt code-overflow-wrap"
+#| attr-output: 'filename="LICENSE.txt"'
 
-> **1\.
-Importance**: Software should be considered a legitimate and citable product of research.
-Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be included in the metadata of the citing work, for example in the reference list of a journal article, and should not be omitted or separated.
-Software should be cited on the same basis as any other research product such as a paper or a book, that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
-> 
-> **5\.
-Accessibility**: Software citations should facilitate access to the software itself and to its associated metadata, documentation, data, and other materials necessary for both humans and machines to make informed use of the referenced software.
-> 
-> **6\.
-Specificity**: Software citations should facilitate identification of, and access to, the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-
-In practice, the first step is to identify all pieces of software the project relies on.
-A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before.
-Then there are the individual R packages, Quarto extensions, and $\TeX$ packages.
-All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper.
-For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit.
-A system dependency is additional software that you require on your computer apart from the R package.
-
-::: {#fig-software-citation}
-```{mermaid}
-flowchart TB
-  asks_for_citation("Does the software<br>ask you to cite it?")
-  critical_or_unique_contribution("Did the software<br>play a critical part<br>in or contributed<br>something unique<br>to your research?")
-  manipulate("Did the software<br>manipulate or create<br>your data, software,<br>or other outputs?")
-  relies_on_credit("Do the authors of<br>the software rely<br>on academic credit<br>for funding?")
-  cite[Cite!]
-  nocite[Don't cite!]
-
-  asks_for_citation --"Yes"--> cite
-  asks_for_citation --"No"--> critical_or_unique_contribution
-  critical_or_unique_contribution --"Yes"--> cite
-  critical_or_unique_contribution --"No"--> manipulate
-  manipulate --"Yes"--> cite
-  manipulate --"No"--> relies_on_credit
-  relies_on_credit --"Yes"--> cite
-  relies_on_credit --"No"--> nocite
+cat(paste0(
+  "The penguins data stored in \"data.csv\" by Allison Horst, Alison Hill, and Kristen Gorman available from <",
+  citation("palmerpenguins", auto = TRUE)$url,
+  "> are licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>"
+))
 ```
 
-"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Simplified from original.
-:::
+As before, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).
 
-Now, add references for the software you would like to cite to the manuscript.
-In the following, we will demonstrate this for R and all R packages by using the R package `grateful`.
-For arbitrary software, you can use the [CiteAs service](https://citeas.org/) to create appropriate citations.
-
-Add the following code chunk to the end of the discussion in the manuscript:
-
-``````{.qmd filename="Manuscript.qmd"}
-```{{r}}
-#| echo: false
-
-grateful::cite_packages(
-    output = "paragraph",
-    out.dir = ".",
-    omit = NULL,
-    dependencies = TRUE,
-    passive.voice = TRUE,
-    bib.file = "grateful-refs"
-)
-```
-``````
-
-This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that this automatic detection can miss packages in some circumstances, therefore always verify the rendered result.] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
-
-```{.yml filename="Manuscript.qmd"}
-bibliography:
-  - Bibliography.bib
-  - grateful-refs.bib
-```
-
-Use `renv` to view, install, and record the newly used package:
-
-```{.r filename="Console"}
-renv::status()
-renv::install()
-renv::snapshot()
-```
-
-Finally, render the document again and commit the changes:
-
-```{.bash filename="Terminal"}
-quarto render Manuscript.qmd
-
-git status
-git add .
-git commit -m "Cite data and software"
-```
-
-## The Last Mile
-
-`renv` only records the versions of R packages and of R itself.
-This means that everything we have not decided to cite in the previous step is not documented anywhere.
-We will cover system dependencies when [creating a README](make_readme.qmd).
-For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).
-Do run the following:
-
-```{.bash filename="Terminal"}
-quarto use binder
-```
-
-This will create a few additional files which facilitate reconstructing the computational environment in the future.^[Either using [repo2docker](https://repo2docker.readthedocs.io/) or the public [binder](https://mybinder.org/) service.] As always, commit your changes:
+Finally, you can go through the commit routine:
 
@@
 git add .
-git commit -m "Add repo2docker config"
+git commit -m "Add data"
 ```
-
-You are now all set up to prepare your project for sharing!

### 2025-01-25 intro.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

@@
 
-In the following, we will provide a brief introduction to the concept of _research compendia_.
+In the following, we will set the stage by highlighting the importance of sharing all materials, documenting their usage, and linking results with their computations, leading to the creation _research compendia_.
 
@@
 
+## Document Materials
+
+However, there's more to code publishing than sharing.
+Researchers should document (i.e., track) which data, software, images, texts, and other materials contributed to their work.
+In more elaborate terms, researchers should preserve the _provenance_ of shared materials, for at least three tangible reasons:
+
+- **Academic integrity**: Providing citations for works that are not one's own avoids plagiarism.
+- **Complying with the law**: Providing attribution and license texts (among other things) may be a legal obligation when redistributing materials.
+- **Facilitating reproductions**: Stating utilized software and data with their exact versions helps reproducers.
+
+Of course, when and how to reserve provenance varies for these reasons.
+Rephrasing someone's text avoids issues of copyright, but may be plagiarism.
+And bibliographies of various styles may comply with scientific citation standards, but if they omit the first names of authors or the version number of computer programs, they might not help with matters of copyright or reproducibility.
+
+In this tutorial, we'll consider all of these reasons important.
+And there are even more: Documenting the provenance puts research into context and allows others to understand how it came about.
+Also, cited authors benefit from the citation as they gain potential readers.
+
 ## Linking Results and Computations
 
-This tutorial not only covers sharing the source code, but also connecting it to the results through the creation of dynamic documents.
+For the purpose of this tutorial, tracking the provenance of results deserves particular attention.
+This means connecting them to the source code through the creation of dynamic documents.
 Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article.
-Dynamic documents bundled together with any necessary data and auxiliary software are called a [research compendium](https://research-compendium.science/) [@Gentleman2007].
+Dynamic documents bundled together with any necessary data and auxiliary software are called [research compendia](https://research-compendium.science/) [@Gentleman2007].
 
@@
 
-When creating a research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
+To recap, in this tutorial you will share your materials, document their usage, and connect results to the underlying source code through the creation of a research compendium.
+Make sure to consider the following things along the way [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 

### 2025-01-25 make_readme.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

@@
 
-Once you have prepared your project and settled on a license, it is time to add a final touch.
+Having settled on a license, it is time to add a final touch.
 Imagine returning to your project in five years, having forgotten most of the details of what you did exactly.
@@
 See @sec-dependencies for additional information.
+It may make sense to provide these information in the README even if you already cite all programs and their version numbers in your manuscript.
 
@@
 
-:   Which files does the project contain?
+:   Which files does the project folder contain?
 How are they organized?
@@
 
-:   Is there a recommended way to cite this project?
-Is there a published article associated with it that you would like to have cited?
-We recommend to increase citability by using the [CodeMeta generator](https://codemeta.github.io/codemeta-generator/) to create a file called `codemeta.json`.
-Alternatively, one can also use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
-Then, put it into the project folder.
-Another common standard is the [Citation File Format](https://citation-file-format.github.io/) (CFF), which is an equally valid option.
+:   Is there a recommended way to cite this project? Is there a published article associated with it that you would like to have cited?
 
@@
 
-:   Under which licenses are the works in this project available?
+:   Under which licenses are the works in this project folder available?
 
@@
 ```txt
-── Install scripts ─────────────────────── 
-dnf install -y make pandoc
+Finding R package dependencies ... Done!
+── Install scripts ──────────────────────── Fedora 40 ── 
+dnf install -y make pandoc libcurl-devel \
+  openssl-devel libicu-devel v8-devel libxml2-devel \
+  git
 
-── Packages and their system dependencies 
+── Packages and their system dependencies ──────────────
+blastula  – pandoc
+curl      – libcurl-devel, openssl-devel
 fs        – make
 knitr     – pandoc
+openssl   – openssl-devel
+remotes   – git
 rmarkdown – pandoc
 sass      – make
+stringi   – libicu-devel
+V8        – v8-devel
+xml2      – libxml2-devel
 ```
 
-We can see that `make` and `pandoc` were identified as system dependencies. Often, one can obtain their version by running them with the `--version` argument:
+We can see that the programs `make`, `pandoc`, and `git` were identified as system dependencies. Often, one can obtain their version by running them with the `--version` argument:
 
@@
 pandoc --version
+git --version
 ```
+
+The other system dependencies are libraries, and the method to identify the installed version varies between operating systems. For example, on Fedora Linux, one can run the following:
+
+```{.bash .code-overflow-wrap filename="Terminal"}
+rpm -q libcurl-devel openssl-devel libicu-devel v8-devel libxml2-devel
+```
+
+However, they are only relevant if the R packages were installed from source, rather than as binary package.
 
@@
 
-Next, one can open a new R session in the root directory of this project and run the following, which should install all required R packages at their recorded versions:
+Next, one can open a new R session in the root folder of this project and run the following, which should install all required R packages at their recorded versions:
 
@@
 
-The most important file in this project is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:
+The most important file in this project folder is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:
 
 - `Bibliography.bib`: bibliographic references used in the manuscript
-- `data.csv`: a data set containing the simplified `palmerpenguins` data created using `write.csv(palmerpenguins::penguins, "dat.csv", row.names = FALSE)`
+- `data.csv`: a data set containing the simplified `palmerpenguins` data
 - `data_dictionary.html`: a dictionary to the data file, created using `create_data_dictionary.R`
@@
 
+You need to add your name in the following example:
+
 ```{.md .code-overflow-wrap filename="README.md"}
@@
 
-> Zerna & Scheffel (2024): "A Study on Penguins: A Minimal Reproducible Example". Unpublished manuscript.
+> Zerna, Scheffel, & <YOUR NAME> (2024): "A Study on Penguins: A Minimal Reproducible Example". Unpublished manuscript.
 ```
@@
 
+Of course, you would use the same license for the manuscript that you chose in the previous step. You need to add your name in the following example:
+
 ```{.md .code-overflow-wrap filename="README.md"}
@@
 
-The manuscript file and the bibliographic references are made available by us under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/).
-The data set by Kristen Gorman has been published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-`apaquarto` has also been published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-All other files in this project are available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) or (at your option) under the [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later).
+For further copyright information, see `LICENSE.txt`.
 ```

### 2025-01-25 setup.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

--- 2024-11-29 setup.qmd
@@
 
+Throughout this tutorial, you will need to run both R code and system commands (primarily `git` and `quarto`).
+Within RStudio, R code can be run by going to the tab _Console_, while system commands are executed in the tab _Terminal_.
+We also indicate where to run your code directly above each code snippet.
+
 Make sure that your project is in a consistent state according to `renv` by running:
@@
 ```
+
+::: {#tip-renv-status .callout-tip}
+Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
+:::
 
@@
 
-Then, one can open an R session by simply typing `R` into the terminal.
-Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`.
-If not, set it using `setwd("code-publishing-exercise")`.
-Then, initialize `renv`:
+Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Now, initialize `renv`:
 
@@
 
-Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it.
-In general, the directory structure should facilitate understanding a project by breaking it into logical chunks.
+Before adding your project files, it is helpful to decide on a folder structure, that is, how to call each file and where to put it.
+In general, the folder structure should facilitate understanding a project by breaking it into logical chunks.
 There is no single best solution, as a good structure depends on where a project's complexity lies.
@@
 
-Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021].
+Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021; @ArayaSalas2024] or journal [@Vilhuber2021].
 We recommend you to follow the standards of your field.
 
-For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.
+In this simple project, we will explore differences in bill length between male and female penguins. You can put all your files in the root folder of your project.
 
-## Add Data
+![Artwork by \@allison_horst, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), adapted from [original](https://allisonhorst.github.io/palmerpenguins/articles/art.html)](images/culmen_depth.png)
 
-You can now download the data set we have prepared for you and put into your project folder: [`Data.csv`](_01_Example/Data.csv){download=""}
+## Add Manuscript
 
-::: {.column-margin}
-![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
-:::
+In order to get you started, we have prepared a manuscript for you, alongside a bibliography file. Download the two files to your computer and put them into your project folder.
 
-The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
-It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+[Manuscript.qmd](_01_Example/Manuscript.qmd){.btn .btn-lg .btn-info download=""} [Bibliography.bib](_01_Example/Bibliography.bib){.btn .btn-lg .btn-info download=""}
 
-::: {#imp-legal-restrictions .callout-important}
-### Consider Legal Restrictions Before Sharing
+The manuscript already contains the narrative, but the results are still missing.
+It will be your job to conduct the data analysis.
+For now, add yourself as a co-author in the YAML header of the manuscript.
 
-Everything you put into the project folder will be shared publicly.
-For reasons of reproducibility, this should include the data you analyze.
-Of course, you should only share them to the extent you are allowed to, taking into account:
-
-1. applicable privacy laws (e.g., the GDPR for European citizens),
-2. contractual obligations (e.g., with your data provider),
-3. copyright of the data and their particular structure, and
-4. any _sui generis_ database right.
-
-Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
-For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
-
-Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are.
-Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.
-As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
-:::
-
-When publishing a data set, it is important to document the meaning (e.g., units) and possible values of its variables.
-This is typically done with a _data dictionary_ (also called a _codebook_).
-In the following, we will demonstrate how to create a simple data dictionary using the R packages [`tinylabels`](https://cran.r-project.org/package=tinylabels), [`datawizard`](https://easystats.github.io/datawizard/), and [`tinytable`](https://vincentarelbundock.github.io/tinytable/).
-You can install them now using:
-
-```{.r filename="Console"}
-renv::install(c(
-  "tinylabels",
-  "datawizard",
-  "tinytable"
-))
-```
-
-You can put the code that follows for creating the data dictionary into a new file called `create_data_dictionary.R`.
-
-Using `tinylabels` we can add labels to the variables of a `data.frame` in R:^[Note that the code provided does not alter the data file -- no labels will be added to `Data.csv`.
-The labels are only added to a (temporary) copy of the data set within R in order to create the data dictionary.]
-
-```{r, echo = -1, eval = -2, filename="create_data_dictionary.R"}
-dat <- read.csv(file.path("_01_Example", "Data.csv"))
-dat <- read.csv("Data.csv")
-
-descriptions <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-tinylabels::variable_label(dat) <- descriptions
-```
-
-Subsequently, `datawizard` can be employed to create the data dictionary containing the name and label, but also some other information about each variable:
-
-::: {.column-margin}
-![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
-:::
-
-```{r, filename="create_data_dictionary.R"}
-(dict <- datawizard::data_codebook(dat) |>
-  subset(select = -.row_id) |>
-  tinytable::tt())
-```
-
-Finally, we can store the data dictionary inside an HTML file using the R package `tinytable` and put the HTML file into the project folder as well.
-
-::: {.column-margin}
-![tinytable: Simple and Configurable Tables](images/tinytable.svg){width=250px}
-:::
-
-```{.r filename="create_data_dictionary.R"}
-tinytable::save_tt(dict, output = "data_dictionary.html")
-```
-
-A human-readable data dictionary is necessary for making one's research reproducible and the example we provided demonstrates only the bare minimum.
-A full data documentation including measurement instruments, sampling procedures, appropriate weighting, contact information, and more information about the study can be created with the R package [`pointblank`](https://rstudio.github.io/pointblank/articles/INFO-1.html).
-And one could go even further by making the information machine-readable in a standardized way.
-We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
-
-::: {#nte-frictionless .callout-note collapse="true"}
-### Create Machine-Readable Variable Documentation
-
-This example demonstrates how the title and description of the data set, the description of the variables and their possible values are stored in a machine-readable way.
-
-```{.r filename="Console"}
-table_info <- c(
-  title = "penguins data set",
-  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
-)
-descriptions <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-vals <- list(
-  species = c("Adelie", "Gentoo", "Chinstrap"),
-  island = c("Torgersen", "Biscoe", "Dream"),
-  sex = c("male", "female"),
-  year = c(2007, 2008, 2009)
-)
-```
-
-Generally, metadata are either stored embedded into the data or externally, for example, in a separate file.
-We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately.
-Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).
-
-Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released.
-As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information.
-We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
-Finally, the package is used to create a metadata file that contains the schema.
-
-```{.r filename="Console"}
-# Read data and create schema
-dat_filename <- "Data.csv"
-dat <- read.csv(dat_filename)
-dat_schema <- frictionless::create_schema(dat)
-
-# Add descriptions to the fields
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  c(x, description = descriptions[[x$name]])
-})
-
-# Record possible values
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  if (x$name %in% names(vals)) {
-    modifyList(x, list(constraints = list(enum = vals[[x$name]])))
-  } else {
-    x
-  }
-})
-
-# Define missing values
-dat_schema$missingValues <- c("", "NA")
-
-# Create package with license info and write it
-dat_package <- frictionless::create_package() |>
-  frictionless::add_resource(
-    resource_name = "penguins",
-    data = dat_filename,
-    schema = dat_schema,
-    title = table_info[["title"]],
-    description = table_info[["description"]],
-    licenses = list(list(
-      name = "CC0-1.0",
-      path = "https://creativecommons.org/publicdomain/zero/1.0/",
-      title = "CC0 1.0 Universal"
-    ))
-  )
-frictionless::write_package(dat_package, directory = ".")
-```
-
-This creates the metadata file `datapackage.json` in the current directory.
-Make sure it is located in the same folder as `Data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
-:::
-
-Having added the data and its documentation, one can view and record the utilized packages with `renv`...
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-...and go through the commit routine:
-
-```{.bash filename="Terminal"}
-git status
-git add .
-git commit -m "Add data"
-```
-
-## Add Code
-
-In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file.
-The manuscript contains code together with a written narrative.
-Download the two files to your computer and put them into your project folder.
-
-- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
-- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}
-
-The manuscript explores differences in bill length between male and female penguins, feel free to read through it.
-
-::: {#imp-take-copyright-seriously .callout-important}
-### Take Copyright Seriously
-
-When you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so.
-Only build on existing work for which you are given an express grant of relevant rights.
-How do you know you are allowed to copy, edit, and share the two files linked above?
-:::
-
-:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
-### Hint for "Take Copyright Seriously"
-
-Have a look at the [about page](about.qmd).
-::::
-
-As the manuscript uses some new packages, install them with:
-
-```{.r filename="Console"}
-renv::install()
-```
-
-The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
-It can be installed in the project using the following command:
-
-```{.bash filename="Terminal"}
-quarto add --no-prompt wjschne/apaquarto
-```
-
-::: {#tip-apaquarto .callout-tip}
-#### Not a Psychologist?
-
-If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.
-
-Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:
-
-```{.yml filename="Manuscript.qmd"}
-format:
-  pdf:
-    pdf-engine: lualatex
-    documentclass: scrartcl
-    papersize: a4
-```
-:::
-
-Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents.
+To create a PDF from the manuscript, you need to have a $\TeX$ distribution installed on your computer.
 A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
@@
 
-::: {#tip-update-quarto .callout-tip}
-If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
-:::
-
-With the code being added, one can use `renv` again to view and record the new packages:
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-::: {#tip-renv-status .callout-tip}
-Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
-:::
-
-Finally, make your changes known to Git:
+You can now make your changes known to Git:
 
@@
 
-::: {#wrn-credentials .callout-warning}
-### Beware of Credentials
-
-Sometimes, a data analysis requires the interaction with online services:
-
-- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
-- an analysis may be conducted with the help of AI providers.
-
-In these cases, make sure that the code you check in to Git does not contain any credentials that are required for accessing these services.
-Instead, make use of environment variables which are defined in a location that is excluded from version control.
-When programming with R, you can define them in a file called `.Renviron` in the root of your project folder:
-
-```{.ini filename=".Renviron"}
-MY_FIRST_KEY="your_api_key_here"
-MY_SECOND_KEY="your_api_key_here"
-```
-
-When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:
-
-```r
-query_api(..., api_key = Sys.getenv("MY_FIRST_KEY"))
-```
-
-Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that contains credentials, you can follow @Chacon2024.
-:::
-
-### Coding Best Practices
-
-Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.
-
-- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root.
-For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
-
-- __Keep it simple.__ Add complexity only when you must.
-Whenever there's a boring way to do something and a clever way, go for the boring way.
-If the code grows increasingly complex, refactor it into separate functions and files.
-
-- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.
-
-- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.
-
-- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).
-
-- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
-Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
-We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.
-
-- __Fail early, often and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
-
-- __Test your code.__ Test your code with scenarios where you know what the result should be.
-Turn bugs you discovered into test cases.
-Use linting tools^[A linting tool analyzes your code without actually running it.
-This process is called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
-
-- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality.
-R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/).
-Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013.
-And even if you don't follow a style guide, be consistent.
-
-This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:
-
-- "Tidy design principles" [@Wickham2023Design]
-- "The Good Research Code Handbook" [@Mineault2021]
-- "Quality assurance of code for analysis and research" [@UKGovernment2020]
-- "The Art of UNIX Programming" [@Raymond2003]
-
-> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
-> 
-> --- @Fowler1999, p. 15
-
-## Cite Data and Software
-
-If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.
-
-### Data
-
-Put simply, all data relied upon should be cited to allow for precise identification and access.
-From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
-
-> **Principle 1 – Importance**: "Data should be considered legitimate, citable products of research.
-Data citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications."
-> 
-> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited."
-> 
-> **Principle 5 – Access**: "Data citations should facilitate access to the data themselves and to such associated metadata, documentation, code, and other materials, as are necessary for both humans and machines to make informed use of the referenced data."
-> 
-> **Principle 7 – Specificity and Verifiability**: "Data citations should facilitate identification of, access to, and verification of the specific data that support a claim.
-Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific time slice, version and/or granular portion of data retrieved subsequently is the same as was originally cited."
-
-Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
-
-::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
-#### Hint for citing the data set
-
-As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:
-
-```{r, echo = -1}
-invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
-citation("palmerpenguins")
-```
-
-As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).
-
-Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`.
-Then, add a sentence to the manuscript such as follows:
-
-```md
-The analyzed data are by @horst2020.
-```
-
-Render the document to check that the citation is displayed properly.
-
-```{.bash filename="Terminal"}
-quarto render Manuscript.qmd
-```
-:::
-
-### Software
-
-When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies.
-You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not.
-As with data, citations should allow for exact identification and access.
-From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):
-
-> **1\.
-Importance**: Software should be considered a legitimate and citable product of research.
-Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be included in the metadata of the citing work, for example in the reference list of a journal article, and should not be omitted or separated.
-Software should be cited on the same basis as any other research product such as a paper or a book, that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
-> 
-> **5\.
-Accessibility**: Software citations should facilitate access to the software itself and to its associated metadata, documentation, data, and other materials necessary for both humans and machines to make informed use of the referenced software.
-> 
-> **6\.
-Specificity**: Software citations should facilitate identification of, and access to, the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-
-In practice, the first step is to identify all pieces of software the project relies on.
-A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before.
-Then there are the individual R packages, Quarto extensions, and $\TeX$ packages.
-All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper.
-For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit.
-A system dependency is additional software that you require on your computer apart from the R package.
-
-::: {#fig-software-citation}
-```{mermaid}
-flowchart TB
-  asks_for_citation("Does the software<br>ask you to cite it?")
-  critical_or_unique_contribution("Did the software<br>play a critical part<br>in or contributed<br>something unique<br>to your research?")
-  manipulate("Did the software<br>manipulate or create<br>your data, software,<br>or other outputs?")
-  relies_on_credit("Do the authors of<br>the software rely<br>on academic credit<br>for funding?")
-  cite[Cite!]
-  nocite[Don't cite!]
-
-  asks_for_citation --"Yes"--> cite
-  asks_for_citation --"No"--> critical_or_unique_contribution
-  critical_or_unique_contribution --"Yes"--> cite
-  critical_or_unique_contribution --"No"--> manipulate
-  manipulate --"Yes"--> cite
-  manipulate --"No"--> relies_on_credit
-  relies_on_credit --"Yes"--> cite
-  relies_on_credit --"No"--> nocite
-```
-
-"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Simplified from original.
-:::
-
-Now, add references for the software you would like to cite to the manuscript.
-In the following, we will demonstrate this for R and all R packages by using the R package `grateful`.
-For arbitrary software, you can use the [CiteAs service](https://citeas.org/) to create appropriate citations.
-
-Add the following code chunk to the end of the discussion in the manuscript:
-
-``````{.qmd filename="Manuscript.qmd"}
-```{{r}}
-#| echo: false
-
-grateful::cite_packages(
-    output = "paragraph",
-    out.dir = ".",
-    omit = NULL,
-    dependencies = TRUE,
-    passive.voice = TRUE,
-    bib.file = "grateful-refs"
-)
-```
-``````
-
-This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that the detection can fail ] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
-
-```{.yml filename="Manuscript.qmd"}
-bibliography:
-  - Bibliography.bib
-  - grateful-refs.bib
-```
-
-Use `renv` to view, install, and record the newly used package `grateful`:
-
-```{.r filename="Console"}
-renv::status()
-renv::install()
-renv::snapshot()
-```
-
-Finally, render the document again and commit the changes:
-
-```{.bash filename="Terminal"}
-quarto render Manuscript.qmd
-
-git status
-git add .
-git commit -m "Cite data and software"
-```
-
-## The Last Mile
-
-`renv` only records the versions of R packages and of R itself.
-This means that everything we have not decided to cite in the previous step is not documented anywhere.
-We will cover system dependencies when [creating a README](make_readme.qmd).
-For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).
-Do run the following:
-
-```{.bash filename="Terminal"}
-quarto use binder
-```
-
-This will create a few additional files which facilitate reconstructing the computational environment in the future.^[Either using [repo2docker](https://repo2docker.readthedocs.io/) or the public [binder](https://mybinder.org/) service.] As always, commit your changes:
-
-```{.bash filename="Terminal"}
-git status
-git add .
-git commit -m "Add repo2docker config"
-```
-
-You are now all set up to prepare your project for sharing!
+You are now all set up to prepare your project folder for sharing!

### 2025-01-25 about.qmd
Only license code as CC0
This makes it less complicated to describe,
although we deviate now from our
recommendations (regarding CC0's lack of patent
grant).

@@
 
-Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code chunks are also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) or the [Apache License\ 2.0](https://www.apache.org/licenses/LICENSE-2.0).
+Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code chunks are also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
 

### 2025-01-25 choose_license.qmd
Only license code as CC0
This makes it less complicated to describe,
although we deviate now from our
recommendations (regarding CC0's lack of patent
grant).

@@
 
-The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) or (at your option) under the [Apache License\ 2.0](https://www.apache.org/licenses/LICENSE-2.0), as is all code in this tutorial (e.g., for creating the data dictionary).
-Therefore, you can apply any license to your modified version (as long as you follow the terms of the existing license).
+The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary).
+Therefore, you can apply any license to your modified version.
 Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).
@@
 
-In the following example, the files from this tutorial are used under the terms of CC0\ 1.0 and therefore, the requirements of the Apache license\ 2.0 need not be followed.
-
-The edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.
+In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.
 

### 2025-01-25 about.qmd
Only license code as CC0
This makes it less complicated to describe,
although we deviate now from our
recommendations (regarding CC0's lack of patent
grant).

@@
 
-Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code chunks are also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+Except where noted otherwise, the text in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code without any text is also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
 

### 2025-01-25 about.qmd
Add full license texts

--- 2025-01-25 about.qmd
@@
 
-Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/); the code chunks are also (at your option) available under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+Except where noted otherwise, the narrative text in this tutorial is licensed under [CC\ BY-SA\ 4.0](LICENSE-CC-BY-SA.txt); the code without any narrative text is also (at your option) available under [CC0\ 1.0](LICENSE-CC0.txt).
 
-The file `Manuscript.qmd` is made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+The file `Manuscript.qmd` is made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).
 
-The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are available under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/agpl-3.0.html).
+The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are Copyright (C) 2024 Posit PBC and are available under the [GNU Affero General Public License v3](LICENSE-AGPL.txt).
 The source code is available on [GitHub](https://github.com/rstudio/rstudio/tree/v2024.04.2%2B764).
-
-::: {.nte-agplv3 .callout-note collapse="true"}
-### GNU Affero General Public License v3
-
-```{.txt .code-overflow-wrap}
-Copyright (C) 2024 Posit Software, PBC
-
-Unless you have received this program directly from Posit Software pursuant to the terms of a commercial license agreement with Posit Software, then RStudio is licensed to you under the AGPLv3, the terms of which are included below.
-Details on obtaining the source code for this distribution are included in the file SOURCE.
-
-RStudio includes other open source software whose license terms can be found in the file NOTICE.
-RStudio for Windows is bundled with binary copies of several utilities, whose license terms can be found in the file SOURCE.
-
-                    GNU AFFERO GENERAL PUBLIC LICENSE
-                       Version 3, 19 November 2007
-
- Copyright (C) 2007 Free Software Foundation, Inc. <http://fsf.org/>
- Everyone is permitted to copy and distribute verbatim copies
- of this license document, but changing it is not allowed.
-
-                            Preamble
-
-  The GNU Affero General Public License is a free, copyleft license for
-software and other kinds of works, specifically designed to ensure
-cooperation with the community in the case of network server software.
-
-  The licenses for most software and other practical works are designed
-to take away your freedom to share and change the works.  By contrast,
-our General Public Licenses are intended to guarantee your freedom to
-share and change all versions of a program--to make sure it remains free
-software for all its users.
-
-  When we speak of free software, we are referring to freedom, not
-price.  Our General Public Licenses are designed to make sure that you
-have the freedom to distribute copies of free software (and charge for
-them if you wish), that you receive source code or can get it if you
-want it, that you can change the software or use pieces of it in new
-free programs, and that you know you can do these things.
-
-  Developers that use our General Public Licenses protect your rights
-with two steps: (1) assert copyright on the software, and (2) offer
-you this License which gives you legal permission to copy, distribute
-and/or modify the software.
-
-  A secondary benefit of defending all users' freedom is that
-improvements made in alternate versions of the program, if they
-receive widespread use, become available for other developers to
-incorporate.  Many developers of free software are heartened and
-encouraged by the resulting cooperation.  However, in the case of
-software used on network servers, this result may fail to come about.
-The GNU General Public License permits making a modified version and
-letting the public access it on a server without ever releasing its
-source code to the public.
-
-  The GNU Affero General Public License is designed specifically to
-ensure that, in such cases, the modified source code becomes available
-to the community.  It requires the operator of a network server to
-provide the source code of the modified version running there to the
-users of that server.  Therefore, public use of a modified version, on
-a publicly accessible server, gives the public access to the source
-code of the modified version.
-
-  An older license, called the Affero General Public License and
-published by Affero, was designed to accomplish similar goals.  This is
-a different license, not a version of the Affero GPL, but Affero has
-released a new version of the Affero GPL which permits relicensing under
-this license.
-
-  The precise terms and conditions for copying, distribution and
-modification follow.
-
-                       TERMS AND CONDITIONS
-
-  0. Definitions.
-
-  "This License" refers to version 3 of the GNU Affero General Public License.
-
-  "Copyright" also means copyright-like laws that apply to other kinds of
-works, such as semiconductor masks.
-
-  "The Program" refers to any copyrightable work licensed under this
-License.  Each licensee is addressed as "you".  "Licensees" and
-"recipients" may be individuals or organizations.
-
-  To "modify" a work means to copy from or adapt all or part of the work
-in a fashion requiring copyright permission, other than the making of an
-exact copy.  The resulting work is called a "modified version" of the
-earlier work or a work "based on" the earlier work.
-
-  A "covered work" means either the unmodified Program or a work based
-on the Program.
-
-  To "propagate" a work means to do anything with it that, without
-permission, would make you directly or secondarily liable for
-infringement under applicable copyright law, except executing it on a
-computer or modifying a private copy.  Propagation includes copying,
-distribution (with or without modification), making available to the
-public, and in some countries other activities as well.
-
-  To "convey" a work means any kind of propagation that enables other
-parties to make or receive copies.  Mere interaction with a user through
-a computer network, with no transfer of a copy, is not conveying.
-
-  An interactive user interface displays "Appropriate Legal Notices"
-to the extent that it includes a convenient and prominently visible
-feature that (1) displays an appropriate copyright notice, and (2)
-tells the user that there is no warranty for the work (except to the
-extent that warranties are provided), that licensees may convey the
-work under this License, and how to view a copy of this License.  If
-the interface presents a list of user commands or options, such as a
-menu, a prominent item in the list meets this criterion.
-
-  1. Source Code.
-
-  The "source code" for a work means the preferred form of the work
-for making modifications to it.  "Object code" means any non-source
-form of a work.
-
-  A "Standard Interface" means an interface that either is an official
-standard defined by a recognized standards body, or, in the case of
-interfaces specified for a particular programming language, one that
-is widely used among developers working in that language.
-
-  The "System Libraries" of an executable work include anything, other
-than the work as a whole, that (a) is included in the normal form of
-packaging a Major Component, but which is not part of that Major
-Component, and (b) serves only to enable use of the work with that
-Major Component, or to implement a Standard Interface for which an
-implementation is available to the public in source code form.  A
-"Major Component", in this context, means a major essential component
-(kernel, window system, and so on) of the specific operating system
-(if any) on which the executable work runs, or a compiler used to
-produce the work, or an object code interpreter used to run it.
-
-  The "Corresponding Source" for a work in object code form means all
-the source code needed to generate, install, and (for an executable
-work) run the object code and to modify the work, including scripts to
-control those activities.  However, it does not include the work's
-System Libraries, or general-purpose tools or generally available free
-programs which are used unmodified in performing those activities but
-which are not part of the work.  For example, Corresponding Source
-includes interface definition files associated with source files for
-the work, and the source code for shared libraries and dynamically
-linked subprograms that the work is specifically designed to require,
-such as by intimate data communication or control flow between those
-subprograms and other parts of the work.
-
-  The Corresponding Source need not include anything that users
-can regenerate automatically from other parts of the Corresponding
-Source.
-
-  The Corresponding Source for a work in source code form is that
-same work.
-
-  2. Basic Permissions.
-
-  All rights granted under this License are granted for the term of
-copyright on the Program, and are irrevocable provided the stated
-conditions are met.  This License explicitly affirms your unlimited
-permission to run the unmodified Program.  The output from running a
-covered work is covered by this License only if the output, given its
-content, constitutes a covered work.  This License acknowledges your
-rights of fair use or other equivalent, as provided by copyright law.
-
-  You may make, run and propagate covered works that you do not
-convey, without conditions so long as your license otherwise remains
-in force.  You may convey covered works to others for the sole purpose
-of having them make modifications exclusively for you, or provide you
-with facilities for running those works, provided that you comply with
-the terms of this License in conveying all material for which you do
-not control copyright.  Those thus making or running the covered works
-for you must do so exclusively on your behalf, under your direction
-and control, on terms that prohibit them from making any copies of
-your copyrighted material outside their relationship with you.
-
-  Conveying under any other circumstances is permitted solely under
-the conditions stated below.  Sublicensing is not allowed; section 10
-makes it unnecessary.
-
-  3. Protecting Users' Legal Rights From Anti-Circumvention Law.
-
-  No covered work shall be deemed part of an effective technological
-measure under any applicable law fulfilling obligations under article
-11 of the WIPO copyright treaty adopted on 20 December 1996, or
-similar laws prohibiting or restricting circumvention of such
-measures.
-
-  When you convey a covered work, you waive any legal power to forbid
-circumvention of technological measures to the extent such circumvention
-is effected by exercising rights under this License with respect to
-the covered work, and you disclaim any intention to limit operation or
-modification of the work as a means of enforcing, against the work's
-users, your or third parties' legal rights to forbid circumvention of
-technological measures.
-
-  4. Conveying Verbatim Copies.
-
-  You may convey verbatim copies of the Program's source code as you
-receive it, in any medium, provided that you conspicuously and
-appropriately publish on each copy an appropriate copyright notice;
-keep intact all notices stating that this License and any
-non-permissive terms added in accord with section 7 apply to the code;
-keep intact all notices of the absence of any warranty; and give all
-recipients a copy of this License along with the Program.
-
-  You may charge any price or no price for each copy that you convey,
-and you may offer support or warranty protection for a fee.
-
-  5. Conveying Modified Source Versions.
-
-  You may convey a work based on the Program, or the modifications to
-produce it from the Program, in the form of source code under the
-terms of section 4, provided that you also meet all of these conditions:
-
-    a) The work must carry prominent notices stating that you modified
-    it, and giving a relevant date.
-
-    b) The work must carry prominent notices stating that it is
-    released under this License and any conditions added under section
-    7.  This requirement modifies the requirement in section 4 to
-    "keep intact all notices".
-
-    c) You must license the entire work, as a whole, under this
-    License to anyone who comes into possession of a copy.  This
-    License will therefore apply, along with any applicable section 7
-    additional terms, to the whole of the work, and all its parts,
-    regardless of how they are packaged.  This License gives no
-    permission to license the work in any other way, but it does not
-    invalidate such permission if you have separately received it.
-
-    d) If the work has interactive user interfaces, each must display
-    Appropriate Legal Notices; however, if the Program has interactive
-    interfaces that do not display Appropriate Legal Notices, your
-    work need not make them do so.
-
-  A compilation of a covered work with other separate and independent
-works, which are not by their nature extensions of the covered work,
-and which are not combined with it such as to form a larger program,
-in or on a volume of a storage or distribution medium, is called an
-"aggregate" if the compilation and its resulting copyright are not
-used to limit the access or legal rights of the compilation's users
-beyond what the individual works permit.  Inclusion of a covered work
-in an aggregate does not cause this License to apply to the other
-parts of the aggregate.
-
-  6. Conveying Non-Source Forms.
-
-  You may convey a covered work in object code form under the terms
-of sections 4 and 5, provided that you also convey the
-machine-readable Corresponding Source under the terms of this License,
-in one of these ways:
-
-    a) Convey the object code in, or embodied in, a physical product
-    (including a physical distribution medium), accompanied by the
-    Corresponding Source fixed on a durable physical medium
-    customarily used for software interchange.
-
-    b) Convey the object code in, or embodied in, a physical product
-    (including a physical distribution medium), accompanied by a
-    written offer, valid for at least three years and valid for as
-    long as you offer spare parts or customer support for that product
-    model, to give anyone who possesses the object code either (1) a
-    copy of the Corresponding Source for all the software in the
-    product that is covered by this License, on a durable physical
-    medium customarily used for software interchange, for a price no
-    more than your reasonable cost of physically performing this
-    conveying of source, or (2) access to copy the
-    Corresponding Source from a network server at no charge.
-
-    c) Convey individual copies of the object code with a copy of the
-    written offer to provide the Corresponding Source.  This
-    alternative is allowed only occasionally and noncommercially, and
-    only if you received the object code with such an offer, in accord
-    with subsection 6b.
-
-    d) Convey the object code by offering access from a designated
-    place (gratis or for a charge), and offer equivalent access to the
-    Corresponding Source in the same way through the same place at no
-    further charge.  You need not require recipients to copy the
-    Corresponding Source along with the object code.  If the place to
-    copy the object code is a network server, the Corresponding Source
-    may be on a different server (operated by you or a third party)
-    that supports equivalent copying facilities, provided you maintain
-    clear directions next to the object code saying where to find the
-    Corresponding Source.  Regardless of what server hosts the
-    Corresponding Source, you remain obligated to ensure that it is
-    available for as long as needed to satisfy these requirements.
-
-    e) Convey the object code using peer-to-peer transmission, provided
-    you inform other peers where the object code and Corresponding
-    Source of the work are being offered to the general public at no
-    charge under subsection 6d.
-
-  A separable portion of the object code, whose source code is excluded
-from the Corresponding Source as a System Library, need not be
-included in conveying the object code work.
-
-  A "User Product" is either (1) a "consumer product", which means any
-tangible personal property which is normally used for personal, family,
-or household purposes, or (2) anything designed or sold for incorporation
-into a dwelling.  In determining whether a product is a consumer product,
-doubtful cases shall be resolved in favor of coverage.  For a particular
-product received by a particular user, "normally used" refers to a
-typical or common use of that class of product, regardless of the status
-of the particular user or of the way in which the particular user
-actually uses, or expects or is expected to use, the product.  A product
-is a consumer product regardless of whether the product has substantial
-commercial, industrial or non-consumer uses, unless such uses represent
-the only significant mode of use of the product.
-
-  "Installation Information" for a User Product means any methods,
-procedures, authorization keys, or other information required to install
-and execute modified versions of a covered work in that User Product from
-a modified version of its Corresponding Source.  The information must
-suffice to ensure that the continued functioning of the modified object
-code is in no case prevented or interfered with solely because
-modification has been made.
-
-  If you convey an object code work under this section in, or with, or
-specifically for use in, a User Product, and the conveying occurs as
-part of a transaction in which the right of possession and use of the
-User Product is transferred to the recipient in perpetuity or for a
-fixed term (regardless of how the transaction is characterized), the
-Corresponding Source conveyed under this section must be accompanied
-by the Installation Information.  But this requirement does not apply
-if neither you nor any third party retains the ability to install
-modified object code on the User Product (for example, the work has
-been installed in ROM).
-
-  The requirement to provide Installation Information does not include a
-requirement to continue to provide support service, warranty, or updates
-for a work that has been modified or installed by the recipient, or for
-the User Product in which it has been modified or installed.  Access to a
-network may be denied when the modification itself materially and
-adversely affects the operation of the network or violates the rules and
-protocols for communication across the network.
-
-  Corresponding Source conveyed, and Installation Information provided,
-in accord with this section must be in a format that is publicly
-documented (and with an implementation available to the public in
-source code form), and must require no special password or key for
-unpacking, reading or copying.
-
-  7. Additional Terms.
-
-  "Additional permissions" are terms that supplement the terms of this
-License by making exceptions from one or more of its conditions.
-Additional permissions that are applicable to the entire Program shall
-be treated as though they were included in this License, to the extent
-that they are valid under applicable law.  If additional permissions
-apply only to part of the Program, that part may be used separately
-under those permissions, but the entire Program remains governed by
-this License without regard to the additional permissions.
-
-  When you convey a copy of a covered work, you may at your option
-remove any additional permissions from that copy, or from any part of
-it.  (Additional permissions may be written to require their own
-removal in certain cases when you modify the work.)  You may place
-additional permissions on material, added by you to a covered work,
-for which you have or can give appropriate copyright permission.
-
-  Notwithstanding any other provision of this License, for material you
-add to a covered work, you may (if authorized by the copyright holders of
-that material) supplement the terms of this License with terms:
-
-    a) Disclaiming warranty or limiting liability differently from the
-    terms of sections 15 and 16 of this License; or
-
-    b) Requiring preservation of specified reasonable legal notices or
-    author attributions in that material or in the Appropriate Legal
-    Notices displayed by works containing it; or
-
-    c) Prohibiting misrepresentation of the origin of that material, or
-    requiring that modified versions of such material be marked in
-    reasonable ways as different from the original version; or
-
-    d) Limiting the use for publicity purposes of names of licensors or
-    authors of the material; or
-
-    e) Declining to grant rights under trademark law for use of some
-    trade names, trademarks, or service marks; or
-
-    f) Requiring indemnification of licensors and authors of that
-    material by anyone who conveys the material (or modified versions of
-    it) with contractual assumptions of liability to the recipient, for
-    any liability that these contractual assumptions directly impose on
-    those licensors and authors.
-
-  All other non-permissive additional terms are considered "further
-restrictions" within the meaning of section 10.  If the Program as you
-received it, or any part of it, contains a notice stating that it is
-governed by this License along with a term that is a further
-restriction, you may remove that term.  If a license document contains
-a further restriction but permits relicensing or conveying under this
-License, you may add to a covered work material governed by the terms
-of that license document, provided that the further restriction does
-not survive such relicensing or conveying.
-
-  If you add terms to a covered work in accord with this section, you
-must place, in the relevant source files, a statement of the
-additional terms that apply to those files, or a notice indicating
-where to find the applicable terms.
-
-  Additional terms, permissive or non-permissive, may be stated in the
-form of a separately written license, or stated as exceptions;
-the above requirements apply either way.
-
-  8. Termination.
-
-  You may not propagate or modify a covered work except as expressly
-provided under this License.  Any attempt otherwise to propagate or
-modify it is void, and will automatically terminate your rights under
-this License (including any patent licenses granted under the third
-paragraph of section 11).
-
-  However, if you cease all violation of this License, then your
-license from a particular copyright holder is reinstated (a)
-provisionally, unless and until the copyright holder explicitly and
-finally terminates your license, and (b) permanently, if the copyright
-holder fails to notify you of the violation by some reasonable means
-prior to 60 days after the cessation.
-
-  Moreover, your license from a particular copyright holder is
-reinstated permanently if the copyright holder notifies you of the
-violation by some reasonable means, this is the first time you have
-received notice of violation of this License (for any work) from that
-copyright holder, and you cure the violation prior to 30 days after
-your receipt of the notice.
-
-  Termination of your rights under this section does not terminate the
-licenses of parties who have received copies or rights from you under
-this License.  If your rights have been terminated and not permanently
-reinstated, you do not qualify to receive new licenses for the same
-material under section 10.
-
-  9. Acceptance Not Required for Having Copies.
-
-  You are not required to accept this License in order to receive or
-run a copy of the Program.  Ancillary propagation of a covered work
-occurring solely as a consequence of using peer-to-peer transmission
-to receive a copy likewise does not require acceptance.  However,
-nothing other than this License grants you permission to propagate or
-modify any covered work.  These actions infringe copyright if you do
-not accept this License.  Therefore, by modifying or propagating a
-covered work, you indicate your acceptance of this License to do so.
-
-  10. Automatic Licensing of Downstream Recipients.
-
-  Each time you convey a covered work, the recipient automatically
-receives a license from the original licensors, to run, modify and
-propagate that work, subject to this License.  You are not responsible
-for enforcing compliance by third parties with this License.
-
-  An "entity transaction" is a transaction transferring control of an
-organization, or substantially all assets of one, or subdividing an
-organization, or merging organizations.  If propagation of a covered
-work results from an entity transaction, each party to that
-transaction who receives a copy of the work also receives whatever
-licenses to the work the party's predecessor in interest had or could
-give under the previous paragraph, plus a right to possession of the
-Corresponding Source of the work from the predecessor in interest, if
-the predecessor has it or can get it with reasonable efforts.
-
-  You may not impose any further restrictions on the exercise of the
-rights granted or affirmed under this License.  For example, you may
-not impose a license fee, royalty, or other charge for exercise of
-rights granted under this License, and you may not initiate litigation
-(including a cross-claim or counterclaim in a lawsuit) alleging that
-any patent claim is infringed by making, using, selling, offering for
-sale, or importing the Program or any portion of it.
-
-  11. Patents.
-
-  A "contributor" is a copyright holder who authorizes use under this
-License of the Program or a work on which the Program is based.  The
-work thus licensed is called the contributor's "contributor version".
-
-  A contributor's "essential patent claims" are all patent claims
-owned or controlled by the contributor, whether already acquired or
-hereafter acquired, that would be infringed by some manner, permitted
-by this License, of making, using, or selling its contributor version,
-but do not include claims that would be infringed only as a
-consequence of further modification of the contributor version.  For
-purposes of this definition, "control" includes the right to grant
-patent sublicenses in a manner consistent with the requirements of
-this License.
-
-  Each contributor grants you a non-exclusive, worldwide, royalty-free
-patent license under the contributor's essential patent claims, to
-make, use, sell, offer for sale, import and otherwise run, modify and
-propagate the contents of its contributor version.
-
-  In the following three paragraphs, a "patent license" is any express
-agreement or commitment, however denominated, not to enforce a patent
-(such as an express permission to practice a patent or covenant not to
-sue for patent infringement).  To "grant" such a patent license to a
-party means to make such an agreement or commitment not to enforce a
-patent against the party.
-
-  If you convey a covered work, knowingly relying on a patent license,
-and the Corresponding Source of the work is not available for anyone
-to copy, free of charge and under the terms of this License, through a
-publicly available network server or other readily accessible means,
-then you must either (1) cause the Corresponding Source to be so
-available, or (2) arrange to deprive yourself of the benefit of the
-patent license for this particular work, or (3) arrange, in a manner
-consistent with the requirements of this License, to extend the patent
-license to downstream recipients.  "Knowingly relying" means you have
-actual knowledge that, but for the patent license, your conveying the
-covered work in a country, or your recipient's use of the covered work
-in a country, would infringe one or more identifiable patents in that
-country that you have reason to believe are valid.
-
-  If, pursuant to or in connection with a single transaction or
-arrangement, you convey, or propagate by procuring conveyance of, a
-covered work, and grant a patent license to some of the parties
-receiving the covered work authorizing them to use, propagate, modify
-or convey a specific copy of the covered work, then the patent license
-you grant is automatically extended to all recipients of the covered
-work and works based on it.
-
-  A patent license is "discriminatory" if it does not include within
-the scope of its coverage, prohibits the exercise of, or is
-conditioned on the non-exercise of one or more of the rights that are
-specifically granted under this License.  You may not convey a covered
-work if you are a party to an arrangement with a third party that is
-in the business of distributing software, under which you make payment
-to the third party based on the extent of your activity of conveying
-the work, and under which the third party grants, to any of the
-parties who would receive the covered work from you, a discriminatory
-patent license (a) in connection with copies of the covered work
-conveyed by you (or copies made from those copies), or (b) primarily
-for and in connection with specific products or compilations that
-contain the covered work, unless you entered into that arrangement,
-or that patent license was granted, prior to 28 March 2007.
-
-  Nothing in this License shall be construed as excluding or limiting
-any implied license or other defenses to infringement that may
-otherwise be available to you under applicable patent law.
-
-  12. No Surrender of Others' Freedom.
-
-  If conditions are imposed on you (whether by court order, agreement or
-otherwise) that contradict the conditions of this License, they do not
-excuse you from the conditions of this License.  If you cannot convey a
-covered work so as to satisfy simultaneously your obligations under this
-License and any other pertinent obligations, then as a consequence you may
-not convey it at all.  For example, if you agree to terms that obligate you
-to collect a royalty for further conveying from those to whom you convey
-the Program, the only way you could satisfy both those terms and this
-License would be to refrain entirely from conveying the Program.
-
-  13. Remote Network Interaction; Use with the GNU General Public License.
-
-  Notwithstanding any other provision of this License, if you modify the
-Program, your modified version must prominently offer all users
-interacting with it remotely through a computer network (if your version
-supports such interaction) an opportunity to receive the Corresponding
-Source of your version by providing access to the Corresponding Source
-from a network server at no charge, through some standard or customary
-means of facilitating copying of software.  This Corresponding Source
-shall include the Corresponding Source for any work covered by version 3
-of the GNU General Public License that is incorporated pursuant to the
-following paragraph.
-
-  Notwithstanding any other provision of this License, you have
-permission to link or combine any covered work with a work licensed
-under version 3 of the GNU General Public License into a single
-combined work, and to convey the resulting work.  The terms of this
-License will continue to apply to the part which is the covered work,
-but the work with which it is combined will remain governed by version
-3 of the GNU General Public License.
-
-  14. Revised Versions of this License.
-
-  The Free Software Foundation may publish revised and/or new versions of
-the GNU Affero General Public License from time to time.  Such new versions
-will be similar in spirit to the present version, but may differ in detail to
-address new problems or concerns.
-
-  Each version is given a distinguishing version number.  If the
-Program specifies that a certain numbered version of the GNU Affero General
-Public License "or any later version" applies to it, you have the
-option of following the terms and conditions either of that numbered
-version or of any later version published by the Free Software
-Foundation.  If the Program does not specify a version number of the
-GNU Affero General Public License, you may choose any version ever published
-by the Free Software Foundation.
-
-  If the Program specifies that a proxy can decide which future
-versions of the GNU Affero General Public License can be used, that proxy's
-public statement of acceptance of a version permanently authorizes you
-to choose that version for the Program.
-
-  Later license versions may give you additional or different
-permissions.  However, no additional obligations are imposed on any
-author or copyright holder as a result of your choosing to follow a
-later version.
-
-  15. Disclaimer of Warranty.
-
-  THERE IS NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY
-APPLICABLE LAW.  EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT
-HOLDERS AND/OR OTHER PARTIES PROVIDE THE PROGRAM "AS IS" WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO,
-THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM
-IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST OF
-ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-  16. Limitation of Liability.
-
-  IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MODIFIES AND/OR CONVEYS
-THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR DAMAGES, INCLUDING ANY
-GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL DAMAGES ARISING OUT OF THE
-USE OR INABILITY TO USE THE PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF
-DATA OR DATA BEING RENDERED INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD
-PARTIES OR A FAILURE OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS),
-EVEN IF SUCH HOLDER OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF
-SUCH DAMAGES.
-
-  17. Interpretation of Sections 15 and 16.
-
-  If the disclaimer of warranty and limitation of liability provided
-above cannot be given local legal effect according to their terms,
-reviewing courts shall apply local law that most closely approximates
-an absolute waiver of all civil liability in connection with the
-Program, unless a warranty or assumption of liability accompanies a
-copy of the Program in return for a fee.
-
-                     END OF TERMS AND CONDITIONS
-
-            How to Apply These Terms to Your New Programs
-
-  If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make it
-free software which everyone can redistribute and change under these terms.
-
-  To do so, attach the following notices to the program.  It is safest
-to attach them to the start of each source file to most effectively
-state the exclusion of warranty; and each file should have at least
-the "copyright" line and a pointer to where the full notice is found.
-
-    <one line to give the program's name and a brief idea of what it does.>
-    Copyright (C) <year>  <name of author>
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU Affero General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU Affero General Public License for more details.
-
-    You should have received a copy of the GNU Affero General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>.
-
-Also add information on how to contact you by electronic and paper mail.
-
-  If your software can interact with users remotely through a computer
-network, you should also make sure that it provides a way for users to
-get its source.  For example, if your program is a web application, its
-interface could display a "Source" link that leads users to an archive
-of the code.  There are many ways you could offer source, and different
-solutions will be better for different programs; see section 13 for the
-specific requirements.
-
-  You should also get your employer (if you work as a programmer) or school,
-if any, to sign a "copyright disclaimer" for the program, if necessary.
-For more information on this, and how to apply and follow the GNU AGPL, see
-<http://www.gnu.org/licenses/>.
-```
-:::
 

### 2025-01-25 choose_license.qmd
Add full license texts

--- 2025-01-22 choose_license.qmd
@@
 
-## Which Licenses to Choose for a Project?
-
-So far, we only discussed how to choose a license for works _of one type_.
-But what if you want to share a project with all kinds of files?
-For example, the project from this tutorial (among other things) includes a data file, a manuscript file with intermingled code and writing, and an R file.
-And what if you also want to share files by others, as is the case with `apaquarto` which you may have installed in this project?
-
-The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
-Sometimes, even different parts of a file might be subject to different licenses.
-The easiest approach is to make note of every foreign work included in your project and record its license.
-Then, multi-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.
-
 ## Applying the License
 
-Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect.
-Then, you can record the license of existing content and apply the licenses of your own contributions.
-Mostly, this just means indicating which file or folder is covered by which license, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).
-Notably, no sort of registration is required.
+Having selected the license(s) of your choice, we encourage you to read through the full license text (or at least a legal summary) to understand the legal effect.
+The following applies to all licenses mentioned here:
 
-::: {#cau-license-versions .callout-caution}
-### License Versions Are Important
+- Applying the license is irrevocable. You may stop distributing your work, but everybody who already received it may continue to share it.
+- As an author applying the license, you are not bound by the license terms.
+You cannot, however, give others an exclusive license to your content anymore (as some journals needlessly demand, see @wrn-exclusive-rights).
+- By applying the license you share with others your right to use your work for commercial purposes, including selling it.
+- If you apply the license to an image, the license [may also apply](https://commons.wikimedia.org/wiki/Commons:Same_work) to higher resolution version of your image.
+- If you apply a copyleft license to an image, a document in which the image is used may not constitute a derivative work and, as such, may not be affected by the copyleft condition.
+- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_] or if the work was created as part of your job, you may not be able to apply the license.
+Check with your collecting society or your employer in that case.
 
-You may have noticed that we mostly refer to licenses using a name _and_ a version number.
-This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses.
-For example, the Creative Commons licenses (that start with `CC`) were first published in 2002.
-Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013).
-There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.
-CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
+::: {#nte-situation-lmu .callout-note collapse="true"}
+### For Researchers at LMU Munich
 
-For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
+If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
+Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
 :::
 
-For example, the `apaquarto` extension that you included in your project is a work by others.^[Unless you are the author of the `apaquarto` extension.] You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with any terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice.
-In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.
+For Creative Commons licenses, the following is additionally important to know:
+
+- By applying the license you waive any moral, publicity, privacy, or other similar personality rights to the extent necessary to use the work.
+- The attribution requirement may be implemented "in any reasonable manner", possibly not conforming to your precise expectations.
+
+Applying a license to a work mostly just means indicating which file or folder is covered by which license(s) (no sort of registration is required). There are multiple standards on how to do so:
+
+- You have already learned about the file `LICENSE.txt`.
+Either it contains the full text of one license that applies to the whole project folder -- this is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
+Alternatively, it details which license applies to what and the invidual licenses are stored in the same or in different files.
+- A summary of the licensing situation is often given in a `README` file, whose creation will be discussed [later](make_readme.qmd).
+- If only parts of a file are covered by a certain license, then it might make sense to indicate this directly in the file.
+- The license of embedded elements (e.g., images) is either given directly beneath to where they are embedded, at the end of the respective page, or collectively on an separate page.
+- Individual programming languages also have their own way of stating which license a package is distributed under.
+For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
+- Software projects often indicate the license in every file, following the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable (explained in @nte-reuse).
+- The following two metadata standards for increasing citability also allow to indicate the license:
+  - [CodeMeta](https://codemeta.github.io/): All information are stored in a file called `codemeta.json`, which can be created [on their website](https://codemeta.github.io/codemeta-generator/).
+Alternatively, one can use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
+  - [Citation File Format](https://citation-file-format.github.io/) (CFF): Here, everything is stored in a file called `CITATION.cff`.
+It is used by GitHub to directly show a recommended citation for a repository.
+The file can be created [on their website](https://citation-file-format.github.io/cff-initializer-javascript/).
+Alternatively, the R package [`cffr`](https://cran.r-project.org/package=cffr) may be used for that purpose.
 
@@
 
-Regardless of how exactly the licenses are added to the project, this is a good opportunity to verify one last time that all third party content is provided to you under a free/open license and that you comply with it.
-Please add a license to your project now, either creating a file `LICENSE.txt` or following the REUSE standard.
+The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary).
+Therefore, you can apply any license to your modified version.
+Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).
 
-## Wrap-up
+::: {#tip-manuscript-licensetxt .callout-tip collapse="true"}
+### Addition to `LICENSE.txt` (Solution)
 
-If you would like to learn more about copyright and licenses you might find the following resources interesting:
+In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.
 
-- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2014
-- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
-- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
+```{.txt .code-overflow-wrap filename="LICENSE.txt"}
+The Quarto documents stored in "Manuscript.qmd", "Manuscript.tex", and "Manuscript.pdf" by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under AGPLv3 (or later) <https://www.gnu.org/licenses/agpl-3.0.html> or (at your option) under CC BY-SA 4.0 <https://creativecommons.org/licenses/by-sa/4.0/>.
+
+The data dictionary (stored in "data_dictionary.html") and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
+```
+
+As the AGPLv3 requires reusers adding the [full license text](https://www.gnu.org/licenses/agpl-3.0.txt), we recommend to also copy it to the project folder (if not already in there).
+:::
 

### 2025-01-25 index.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.
Various minor improvements

@@
 
-In particular, we will assume you have installed [Git](https://git-scm.com/downloads), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/).
+In particular, we will assume you have installed [Git](https://git-scm.com/downloads), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/#:~:text=2%3A%20Install%20RStudio), and [Quarto](https://quarto.org/docs/download/).
 Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).
@@
 
-[Get started](intro.qmd){.btn-action-primary .btn-action .btn .btn-success .btn-lg}
+[Get started](intro.qmd){.btn .btn-lg .btn-success}

### 2025-01-26 about.qmd
Various minor improvements
Also mention bibliography file

@@
 
-cat(c("Quarto:", as.character(quarto::quarto_version())))
+cat(paste0("Quarto ", as.character(quarto::quarto_version()), " was utilized to render the files."))
 ```
@@
 ```{r}
-#| class.output: "txt code-overflow-scroll"
+#| class-output: "txt code-overflow-scroll"
 
@@
 
-The file `Manuscript.qmd` is made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).
+The files `Manuscript.qmd` and `Bibliography.bib` are made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).
 

### 2025-01-26 choose_license.qmd
Also mention bibliography file

@@
 
-The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary).
+The manuscript and the bibliography file you downloaded for this tutorial are made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary).
 Therefore, you can apply any license to your modified version.
@@
 
-In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.
+In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.
 
@@
 
-The data dictionary (stored in "data_dictionary.html") and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
+The bibliography (stored in "Bibliography.bib"), data dictionary (stored in "data_dictionary.html"), and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
 ```

### 2025-01-26 choose_license.qmd
Fix REUSE note

@@
 
-For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name -- of course, you can also choose a different license:
+For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<YOUR NAME>` with the current year and your name -- of course, you can also choose a different license:
 
 ```{.r filename="create_data_dictionary.R"}
-# SPDX-FileCopyrightText: <YEAR> <NAME>
+# SPDX-FileCopyrightText: <YEAR> <YOUR NAME>
 #
@@
 ```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="<NAME>" --license="CC0-1.0" create_data_dictionary.R
+reuse annotate --copyright="<YOUR NAME>" --license="CC0-1.0" create_data_dictionary.R
 ```
@@
 ```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="Josephine Zerna <josephine.zerna@tu-dresden.de>" --copyright="Christoph Scheffel <christoph_scheffel@tu-dresden.de>" --copyright="Florian Kohrt" --license="CC-BY-4.0" --style=html Manuscript.qmd
+reuse annotate --copyright="Josephine Zerna" --copyright="Christoph Scheffel" --license="CC0-1.0" --style=html Manuscript.qmd
 ```
@@
 <!--
-SPDX-FileCopyrightText: 2024 Christoph Scheffel <christoph_scheffel@tu-dresden.de>
-SPDX-FileCopyrightText: 2024 Florian Kohrt
-SPDX-FileCopyrightText: 2024 Josephine Zerna <josephine.zerna@tu-dresden.de>
+SPDX-FileCopyrightText: 2024 Christoph Scheffel
+SPDX-FileCopyrightText: 2024 Josephine Zerna
 
-SPDX-License-Identifier: CC-BY-4.0
+SPDX-License-Identifier: CC0-1.0
 -->
@@
 
-Note that `Manuscript.qmd` was provided to you under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), which is what you indicate with the previous comment.
-If you edited the file, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them, although you do not need to provide your real name.]
+Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which is what you indicate with the previous comment.
+As you edited the file throughout this tutorial, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them.
+Also, you do not need to provide your real name.]
 
@@
 ```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="Kristen Gorman" --license="CC0-1.0" data.csv
+reuse annotate --copyright="Allison Horst" --copyright="Alison Hill" --copyright="Kristen Gorman" --license="CC0-1.0" data.csv
 ```
@@
 ```{.yml filename="data.csv.license"}
+SPDX-FileCopyrightText: 2024 Allison Horst
+SPDX-FileCopyrightText: 2024 Alison Hill
 SPDX-FileCopyrightText: 2024 Kristen Gorman

### 2025-01-26 choose_license.qmd
Various minor improvements

@@
 Sometimes, even different parts of a file might be subject to different licenses.
-For the purpose of this tutorial, we will consider the manuscript which you edited and the data dictionary you created.
+For the purpose of this tutorial, we will consider the manuscript and the bibliography which you edited as well as the data dictionary you created.
 
@@
 
-We recommend to choose free/open licenses suitable to the type of work (i.e., code, data, and everything else).
+We recommend to choose free/open licenses suitable to the type of work (e.g., code, data, image etc.).
 [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works.
@@
 Finally, there may be some minor files which are build artifacts.
-You can either add them to your `.gitignore` file or use the CC0\ 1.0 license/waiver with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder.
+You can either add them to your `.gitignore` file or use CC0\ 1.0 with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder.
 For more information, also discussing other corner cases, you can read their [Frequently Asked Questions](https://reuse.software/faq/).

### 2025-01-26 copyright.qmd
Add full license texts
Various minor improvements

@@
 
-Licenses may either require that one mentions the applicable license by name and/or link ("Indicate license") or to even copy the full license text along with the used work ("Add license text").
+Licenses may either require that one mentions the applicable license by name and possibly link ("Indicate license") or they even require to copy the full license text along with the used work ("Add license text").
 Of course, the latter may not be feasible if the work is printed and smaller than the license text.
@@
 It is (again) protected by copyright and even if you share it, nobody except you is allowed to copy, modify, or distribute it further.
-You can put it under a free/open license but don't have to.
-The only exception is if the original work has been made available under a "copyleft" license.
+You can put it under a free/open license but don't have to.^[Of course, you still need to abide by the license conditions of the original work, such as attribution.] The only exception is if the original work has been made available under a "copyleft" license.
 In that case, if you share the work with others you must put it under the same license as the original work^[or sometimes also under a later version or under a compatible license] -- or you commit a copyright violation.^[Note, however, that [copyleft licenses are not automatic](https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic): A derivative work of a copylefted work is not "implicitly" under the same license -- others would commit a copyright violation if they just took it.
@@
 
-If the license required adding the full license text, you would also need to copy it to `LICENSE.txt`.
+If the license required adding the full license text, you would also need to copy it to the project folder.
 

### 2025-01-27 make_readme.qmd
Track all (recursive) deps

@@
 
-# Then, identify the system dependencies
-# of your direct dependencies.
-renv::dependencies()$Package |>
+# Find all dependencies
+deps <- renv::dependencies()$Package |>
    unique() |>
-  pak::pkg_sysreqs()
+   pak::pkg_deps(dependencies = NA) |>
+   getElement("package")
+
+# Identify the system dependencies
+pak::pkg_sysreqs(deps)
 ```

### 2025-01-27 data.qmd
Add full license texts
Clarify usage of renv

@@
 
-```{r, filename="create_data_dictionary.R"}
+```{r}
 dat_source <- citation("palmerpenguins", auto = TRUE) |>
@@
 
-As before, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).
+As before, if the license required adding the full license text, you would also need to copy it to the project folder (if not already in there).
 

### 2025-01-27 setup.qmd
Clarify usage of renv
Fix citations
Somehow the order of citations changes, thus
moving the "e.g." around. Now we cite them
directly in the order they appear later on.

@@
 Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on.
-Using it makes it easier for others to view and obtain them at a later point in time.
+Using it makes it easier for others to view and obtain them at the exact same version at a later point in time.
 Therefore make sure that the box _Use renv with this project_ is checked.
@@
 We also indicate where to run your code directly above each code snippet.
+If no indication is given, the code is only for demonstration purposes and does not need to be run.
 
@@
 
-Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021; @ArayaSalas2024] or journal [@Vilhuber2021].
+Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @TIER4; @Wilson2017] as well as specific to a particular programming language [e.g., @ArayaSalas2024; @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021].
 We recommend you to follow the standards of your field.

### 2025-01-27 copyright.qmd
Only load video on demand

@@
 
+The following video "The correct indication of a license" embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy)) explains the correct attribution of Creative Commons licenses.
+
 ```{=html}
+<details>
+  <summary>Load video</summary>
   <div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
-  <iframe src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
+    <iframe loading="lazy" src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
   </div>
+</details>
 ```

### 2025-01-27 code.qmd
Add full license texts
Add DOI Citation Formatter
...and move CiteAs as it is not about
attributions, but citations.

@@
 
-Again, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).
+Again, if the license required adding the full license text, you would also need to copy it to the project folder (if not already in there).
 :::
@@
 In the following, we will demonstrate this for R and all R packages by using the R package `grateful`.
-For arbitrary software, you can use the [CiteAs service](https://citeas.org/) to create appropriate citations.
+For arbitrary software, you can use [CiteAs](https://citeas.org/) or the [DOI Citation Formatter](https://citation.doi.org/) to create appropriate citations.
 

### 2025-01-27 copyright.qmd
Add DOI Citation Formatter
...and move CiteAs as it is not about
attributions, but citations.

@@
 If a project has been stored in a research repository, yet another means of communicating its copyright status might exist.
-For some repositories, external attribution generators exist:
-
-- Wikimedia Commons: [Attribution Generator](https://lizenzhinweisgenerator.de/?lang=en)
-- Various software repositories: [CiteAs](https://citeas.org/)
+For media from Wikimedia Commons, an [Attribution Generator](https://lizenzhinweisgenerator.de/?lang=en) exists.
 

### 2025-01-27 intro.qmd
Various minor improvements
Minor improvements

@@
 You got interested and would like to apply this method to your own data.
-However, with only the article available, it could take months to come up with a working solution, if possible at all.
+However, with only the article available, it could take you months to come up with a working solution, if possible at all.
 This situation has been put aptly by Buckheit & Donoho [-@Buckheit1995, p.
-59], distilling an idea by the geophysicist Jon Claerbout:
+59, emphasis in original], distilling an idea by the geophysicist Jon Claerbout:
 
@@
 However, there's more to code publishing than sharing.
-Researchers should document (i.e., track) which data, software, images, texts, and other materials contributed to their work.
+Researchers should document (i.e., track) which foreign data, software, images, texts, and other materials contributed to their work.
 In more elaborate terms, researchers should preserve the _provenance_ of shared materials, for at least three tangible reasons:

### 2025-02-08 archive.qmd
Recommend against OSF for archival
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 The project folder is now ready to be deposited for long-term archival.
-When considering an appropriate repository, there are a few questions to ponder [excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
+When considering an appropriate repository,
+there are a few questions to ponder
+[excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0][cc-by]]:
 
-> __Does the digital repository give you a unique persistent digital identifier for your deposit?__
+[cc-by]: https://creativecommons.org/licenses/by/4.0/
+
+> __Does the digital repository give you a unique persistent digital identifier
+> for your deposit?__
 > 
-> Persistent digital identifiers are identifiers that are provided under the assumption that they, and the digital objects they refer to, will remain available long term.
-Examples of unique persistent digital identifiers include DOIs and ARKs.
-The use of persistent digital identifiers for research outputs, including papers, data and software, and the use of these for citation purposes, is becoming prevalent in research.
+> Persistent digital identifiers are identifiers that are provided under the
+> assumption that they, and the digital objects they refer to, will remain
+> available long term. Examples of unique persistent digital identifiers include
+> DOIs and ARKs. The use of persistent digital identifiers for research outputs,
+> including papers, data and software, and the use of these for citation
+> purposes, is becoming prevalent in research.
 > 
@@
 > Digital repositories may have constraints on the size of deposits they accept.
-For example, Zenodo has a limit of 50 GB and the University of Edinburgh DataShare has a limit of 20 GB for deposits done via a browser and 100 GB otherwise.
+> For example, Zenodo has a limit of 50 GB and the University of Edinburgh
+> DataShare has a limit of 20 GB for deposits done via a browser and 100 GB
+> otherwise.
 > 
-> __Has your funder or publisher recommended or mandated a digital repository to use?__
+> __Has your funder or publisher recommended or mandated a digital repository
+> to use?__
 > 
-> If both a funder and publisher mandate two different digital repositories you could deposit within them both, or discuss with each to see if they will agree for you to use one only, even if it is not the one they mandate.
+> If both a funder and publisher mandate two different digital repositories you
+> could deposit within them both, or discuss with each to see if they will agree
+> for you to use one only, even if it is not the one they mandate.
 > 
@@
 > 
-> These can include terms and conditions of use; availability of the service (e.g.
-does the service regularly go down for updates or does it run 24/7); how resilient the service is to breakdowns; where deposits are stored; how they are stored; what are the plans for stored deposits if the repository goes out of business; how secure the storage is; how regularly it is backed up; how long backups are kept for; and what disaster recovery plans are in place.
-Knowing the quality of service can help you judge whether your deposit will be available to others, and yourself, even in the face of disaster.
-It also allows you, and your institution, to develop contingency plans should such drastic circumstances arise.
+> These can include terms and conditions of use; availability of the service
+> (e.g. does the service regularly go down for updates or does it run 24/7);
+> how resilient the service is to breakdowns; where deposits are stored;
+> how they are stored; what are the plans for stored deposits if the repository
+> goes out of business; how secure the storage is; how regularly it is backed up;
+> how long backups are kept for; and what disaster recovery plans are in place.
+> Knowing the quality of service can help you judge whether your deposit will be
+> available to others, and yourself, even in the face of disaster. It also
+> allows you, and your institution, to develop contingency plans should such
+> drastic circumstances arise.
 > 
@@
 > 
-> You want to know that the digital repository will host your deposit for as long as you need it too, not just 6 months or so.
-The policies or terms and conditions of the digital repository should state this.
-For example, Zenodo states that their lifetime is currently 20 years and figshare guarantees 10 years of persistent availability.
+> You want to know that the digital repository will host your deposit for as
+> long as you need it too, not just 6 months or so. The policies or terms and
+> conditions of the digital repository should state this. For example, Zenodo
+> states that their lifetime is currently 20 years and figshare guarantees
+> 10 years of persistent availability.
 > 
@@
 > 
-> If a digital repository expects a one-off payment, to subsidise its hosting, you may be able to pay for this out of your project budget or your institution may pay this for you.
-It is recommended that you avoid any digital repository that expects a regular fee for the duration for which they hold the deposit, since it would most likely be unsustainable for your institution to commit to ever-increasing fees over a long term.
-There are many free digital repositories available.
+> If a digital repository expects a one-off payment, to subsidise its hosting,
+> you may be able to pay for this out of your project budget or your institution
+> may pay this for you. It is recommended that you avoid any digital repository
+> that expects a regular fee for the duration for which they hold the deposit,
+> since it would most likely be unsustainable for your institution to commit to
+> ever-increasing fees over a long term. There are many free digital
+> repositories available.
 > 
@@
 > 
-> The digital repository may be accredited or certified by an independent agency that reviews digital repositories according to various criteria related to its policies and processes.
-The presence of an accreditation or certification may provide you with an additional degree of confidence that the digital repository meets your requirements.
-For example, the Data Seal of Approval can be awarded to digital repositories who satisfy a set of quality guidelines relating to how they manage, preserve and make available the deposits they hold.
-The presence of a Data Seal of Approval could give you additional confidence that a digital repository is suitable for holding your deposit.
+> The digital repository may be accredited or certified by an independent agency
+> that reviews digital repositories according to various criteria related to its
+> policies and processes. The presence of an accreditation or certification may
+> provide you with an additional degree of confidence that the digital
+> repository meets your requirements. For example, the Data Seal of Approval can
+> be awarded to digital repositories who satisfy a set of quality guidelines
+> relating to how they manage, preserve and make available the deposits they hold.
+> The presence of a Data Seal of Approval could give you additional confidence
+> that a digital repository is suitable for holding your deposit.
 
 Additional guidance is provided by @Daquin2023, @Garijo2022, and @ScienceEurope2021.
-You can use the registry of research data repositories [re3data.org](https://www.re3data.org/) to find one that matches your criteria.
-We think that the [Software Heritage code archive](https://www.softwareheritage.org/) and [Zenodo](https://zenodo.org/) are good choices, but don't forget to also check your local institutions for recommendations.
+You can use the registry of research data repositories [re3data.org][re3data]
+to find one that matches your criteria.
+We think that the [Software Heritage code archive][sw-heritage]
+and [Zenodo][zenodo] are good choices,
+but don't forget to also check your local institutions for recommendations.
+Note that the Open Science Framework (OSF) is not recommended
+because it allows the deletion of files and projects after their publication.
+
+[re2data]: https://www.re3data.org/
+[sw-heritage]: https://www.softwareheritage.org/
+[zenodo]: https://zenodo.org/
 
@@
 
-In the following we will demonstrate how to use Zenodo to archive one's project folder and obtain a DOI.
+In the following we will demonstrate how to use Zenodo
+to archive one's project folder and obtain a DOI.
 
@@
 
-Note that during this tutorial, we will be using Zenodo's sandbox instance located at [`sandbox.zenodo.org`](https://sandbox.zenodo.org/) in order to learn how to use it while not wasting Zenodo's resources.
-For real projects, use the production instance at [`zenodo.org`](https://zenodo.org/), which requires a separate account.
+Note that during this tutorial, we will be using Zenodo's sandbox instance
+located at [`sandbox.zenodo.org`][zenodo-sandbox] in order to learn
+how to use it while not wasting Zenodo's resources.
+For real projects, use the production instance at [`zenodo.org`][zenodo],
+which requires a separate account.
 :::
 
-First, create an account on the Zenodo Sandbox by clicking on _Sign up_ in the upper right corner and submitting the provided form.
-Then, confirm your email address (also check your spam folder if you cannot find the confirmation email).
+[zenodo-sandbox]: https://sandbox.zenodo.org/
 
-Now you can follow the guide "[Create new upload](https://help.zenodo.org/docs/deposit/create-new-upload/)", taking into account the following notes specific to each step:
+First, create an account on the Zenodo Sandbox
+by clicking on _Sign up_ in the upper right corner and submitting the provided form.
+Then, confirm your email address
+(also check your spam folder if you cannot find the confirmation email).
 
-- __Add/remove files__: To circumvent the file limit on Zenodo, you can bundle your project folder in one ZIP file and upload that instead.
+Now you can follow the guide "[Create new upload][zenodo-upload]",
+taking into account the following notes specific to each step:
+
+[zenodo-upload]: https://help.zenodo.org/docs/deposit/create-new-upload/
+
+- __Add/remove files__:
+  To circumvent the file limit on Zenodo, you can bundle your project folder
+  in one ZIP file and upload that instead.
   Be sure to exclude all files mentioned in your `.gitignore` file!
 - __Fill in metadata__
-  - __Digital Object Identifier__ (DOI): Usually, you will want a DOI, therefore you need to choose _No_ for the question whether you already have a DOI for this upload, and click on _Get a DOI now!_.
-  - __Resource type__: Select the most appropriate one, for example, _Publication / Preprint_.
-  - __Licenses__: If your project folder only contains content by you and if you previously only chose one license, you can select that by choosing _Add standard_.
+  - __Digital Object Identifier__ (DOI):
+    Usually, you will want a DOI,
+    therefore you need to choose _No_ for the question
+    whether you already have a DOI for this upload, and click on _Get a DOI now!_.
+  - __Resource type__:
+    Select the most appropriate one, for example, _Publication / Preprint_.
+  - __Licenses__:
+    If your project folder only contains content by you
+    and if you previously only chose one license,
+    you can select that by choosing _Add standard_.
     Otherwise, click on _Add custom_ and describe how each file is licensed.
-If you have been using REUSE to record license information, you can run enter `SPDX Document` as title and run the reuse tool with `reuse spdx -o reuse.spdx` to create the file `reuse.spdx`.
-Then you can paste the content of `reuse.spdx` into the description field of the custom license as it contains information about the license of each file.
+    If you have been using REUSE to record license information,
+    you can run enter `SPDX Document` as title
+    and run the reuse tool with `reuse spdx -o reuse.spdx`
+    to create the file `reuse.spdx`.
+    Then you can paste the content of `reuse.spdx`
+    into the description field of the custom license
+    as it contains information about the license of each file.
 
-If you previously uploaded your project folder to a public Git repository (which we didn't cover during this tutorial), you could also add the new DOI to your README, for example, under the section "Citation".
+If you previously uploaded your project folder to a public Git repository
+(which we didn't cover during this tutorial),
+you could also add the new DOI to your README,
+for example, under the section "Citation".
 
@@
 There are situations in which you are not allowed to share your data publicly.
-In this case, consider using a repository specifically for research data, which later grants access for valid uses individually.
+In this case, consider using a repository specifically for research data,
+which later grants access for valid uses individually.
 
-For further information, you may be interested in [Software Deposit Guidance for Researchers](https://softwaresaved.github.io/software-deposit-guidance/).
+For further information, you may be interested in
+[Software Deposit Guidance for Researchers](https://softwaresaved.github.io/software-deposit-guidance/).

### 2025-02-08 choose_license.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 Up next, you will free up your project for reuse.
-This means ensuring that every bit (co-)authored by you in the project is put under a free/open license.
-It may be that your project (i.e., its files and folders) is not one work, but rather consists of multiple works available under different licenses.
-In that case, you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project).
+This means ensuring that every bit (co-)authored by you in the project
+is put under a free/open license.
+It may be that your project (i.e., its files and folders) is not one work,
+but rather consists of multiple works available under different licenses.
+In that case, you need to indicate the license on a per-file or per-folder basis
+(rather than choosing one for the whole project).
 Sometimes, even different parts of a file might be subject to different licenses.
-For the purpose of this tutorial, we will consider the manuscript and the bibliography which you edited as well as the data dictionary you created.
+For the purpose of this tutorial,
+we will consider the manuscript and the bibliography which you edited
+as well as the data dictionary you created.
 
@@
 
-Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you.
-This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022].
-Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has.
-To facilitate self-archiving, one can also modify the contract with publishers via a rights retention statement [@UKRN2023] / author's addendum [@SPARC2006].
-If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
+Some journals offer to publish your article under a Creative Commons license,
+but still demand an exclusive publishing and distribution license
+or a copyright assignment from you.
+This would give them more rights than the readers of the article have
+through the respective Creative Commons license [@Rumsey2022]
+and exceeds by far what is necessary to make publication possible [@Suber2022].
+Consequently, authors should oppose this practice and grant publishers the same rights
+that every other reader of the article has.
+To facilitate self-archiving, one can also modify the contract with publishers
+via a rights retention statement [@UKRN2023] / author's addendum [@SPARC2006].
+If you have published a closed-access paper before,
+you can consult [ShareYourPaper](https://shareyourpaper.org/)
+for legal options to still make it available free of charge to readers.
 
-If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
+If your chosen publisher insists on an exclusive license,
+you may at least retain the copyright for your figures
+-- follow the guide "Retaining copyright for figures in academic publications
+to allow easy citation and reuse" by @Elson2016 to learn how to do that.
 :::
@@
 
-We recommend to choose free/open licenses suitable to the type of work (e.g., code, data, image etc.).
-[CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works.
-Take your time to read the summaries behind the links and/or the actual license text to understand the legal effect.
+We recommend to choose free/open licenses suitable to the type of work
+(e.g., code, data, image etc.).
+[CC0\ 1.0][cc0] and [Apache\ 2.0][apache2] are good choices
+to apply simultaneously to all own works.
+Take your time to read the summaries behind the links
+and/or the actual license text to understand the legal effect.
 You can adapt the following wording to your use case:
 
-> Except where noted otherwise, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
+[cc0]: https://creativecommons.org/publicdomain/zero/1.0/
+[apache2]: https://choosealicense.com/licenses/apache-2.0/
+
+> Except where noted otherwise, all files in this project
+> are made available under CC0\ 1.0
+> or (at your option) under the terms of the Apache License\ 2.0.
 :::
@@
 Many boilerplate licenses are available to apply to your work.
-Which license is appropriate depends on several factors, including existing licenses in place and the type of work, but also your personal considerations.
-We strongly recommend to apply a _free/open_ license to your work that fits to the work type and potential use cases.
-Because there are many free/open licenses available, the licenses discussed here only represent a recommended subset.
+Which license is appropriate depends on several factors,
+including existing licenses in place and the type of work,
+but also your personal considerations.
+We strongly recommend to apply a _free/open_ license to your work
+that fits to the work type and potential use cases.
+Because there are many free/open licenses available,
+the licenses discussed here only represent a recommended subset.
 
 ::: {.column-margin}
-If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos).
-For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/).
-For databases, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
+If you would like to choose a license not listed here,
+it should be appropriate for the type of work in question
+and be compatible with the dominant copyleft license in the respective community
+(see also @nte-silos).
+For software that's almost universally the (A)GPLv3
+and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0][cc-by-sa].
+For databases, no dominant copyleft license has emerged yet,
+so any of [ODbL\ 1.0][odbl], [CDLA\ Sharing\ 1.0][cdla-sharing],
+and [CC\ BY-SA\ 4.0][cc-by-sa] are acceptable.
 :::
+
+[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/
+[odbl]: https://opendatacommons.org/licenses/odbl/summary/
+[cdla-sharing]: https://cdla.dev/sharing-1-0/
 
@@
 
-First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license.
-If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications.
-The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder.
-Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.
+First, if you adapt (i.e., modify, remix, build on) a work by others
+you need to determine if it is provided to you under a free/open license.
+If yes, we recommend you to make your contribution available
+under the same license.[^required-with-copyleft]
+For example, if you adapt code published in another paper,
+choose the same license for your modifications.
+The same applies if there are strong community norms
+to use a particular free/open license[^heuristic]
+or if a particular license choice is mandated by your funder.
+Importantly, as discussed before, you are generally not allowed
+to adapt a work _not_ published under a free/open license.
+
+[^required-with-copyleft]: Copyleft licenses even require you to choose the same
+or a compatible license.
+
+[^heuristic]: Of course, this is only a heuristic
+and there might be good reasons to deviate from community norms.
 
@@
 
-If you create a new work and no strong community norms suggest a particular license, you need to choose the license yourself.
+If you create a new work and no strong community norms suggest a particular license,
+you need to choose the license yourself.
 Which license to choose depends on the type of work you create.
-We have created a flowchart that covers the most likely types of works you will create as a researcher: software, writing (i.e., text), images, audio, video, and data (see @fig-flowchart-simple).
-This flowchart always recommends the most permissive license possible to maximize reuse -- though we provide two additional flowcharts below that allow for more choices.
+We have created a flowchart that covers
+the most likely types of works you will create as a researcher:
+software, writing (i.e., text), images, audio, video, and data
+(see @fig-flowchart-simple).
+This flowchart always recommends the most permissive license possible to maximize reuse
+-- though we provide two additional flowcharts below that allow for more choices.
 Click on the name of a license to learn more about it.
@@
 
-One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights [see @CC2024Software]] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license [see @Wikipedia2024FDL]]
+One may wonder which license to apply in this case,
+because Creative Commons licenses are not recommended for source code[^disclaim-patents]
+and applying software licenses to PDFs or images
+can lead to confusion or nuisance.[^require-full-text]
 
-One solution is to make such a work simultaneously available under two (or more) licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video].
-This is called multi-licensing and makes it easier to reuse both the rendered document as well as the code.
+[^disclaim-patents]: because (among other reasons) they explicitly disclaim
+any conveyance of patent rights [see @CC2024Software]
+
+[^require-full-text]: because they often require
+to display the full text of the license [see @Wikipedia2024FDL]
+
+One solution is to make such a work simultaneously available
+under two (or more) licenses, at the choice of the recipient:
+Either under a specified software license,
+or under a Creative Commons license[^media].
+This is called multi-licensing and makes it easier
+to reuse both the rendered document as well as the code.
 For example, one could write:
 
-> The Quarto documents in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
+[^media]: that is, a license for writing, image, audio, and video
+
+> The Quarto documents in this project are made available under CC0\ 1.0
+> or (at your option) under the terms of the Apache License\ 2.0.
 :::
@@
 We already mentioned that a database may be licensed differently than its content.
-However, if the content was created by you, we recommend you to choose the same license for both content and database.
-Factual data (like measurements or metadata) should be made available with [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.
+However, if the content was created by you,
+we recommend you to choose the same license for both content and database.
+Factual data (like measurements or metadata) should be made available
+with [CC0\ 1.0][cc0] [@Villa2016Copyleft; @Villa2016CC0]
+-- otherwise, consult @nte-other-work-types for some caveats.
 
@@
 
-- __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared.
-The [SIL Open Font License\ 1.1](https://openfontlicense.org/) is the _de facto_ standard license for fonts, requiring that derived fonts, if published, have to be put under the same license.
-It doesn't require attribution for usage (e.g., when embedded in a document), but forbids that the font is sold by itself.
+- __fonts:__ Copyleft licenses applied to fonts can be a special case:
+  If a font is put under the license [CC\ BY-SA\ 4.0][cc-by-sa],
+  any documents containing texts using that font will probably be derivative works
+  and have to be put under the same license if shared.
+  The [SIL Open Font License\ 1.1][ofl] is the _de facto_ standard license for fonts,
+  requiring that derived fonts, if published, have to be put under the same license.
+  It doesn't require attribution for usage (e.g., when embedded in a document),
+  but forbids that the font is sold by itself.
   Selling the font is allowed, however, if it is bundled with other software.
-- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of it has to be put under the same license if shared.
-If a document contained source code covered by the (A)GPLv3, it depends on the individual case whether the same license needs to be applied to the document if shared.
-- __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license.
-For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017].
-Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared.
-If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database.
-Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
-- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
+- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package,
+  is licensed under a copyleft license such as the (A)GPLv3,
+  every work that is a derivative of it has to be put under the same license if shared.
+  If a document contained source code covered by the (A)GPLv3,
+  it depends on the individual case
+  whether the same license needs to be applied to the document if shared.
+- __database content:__ If the work produced from a database (the "output")
+  is a derivative of the content in the database,
+  the output is subject to the restrictions laid out in the license.
+  For example, if geospatial data were to be licensed under [CC\ BY\ 4.0][cc-by],
+  all maps produced from the data would likely need to fulfill
+  this license's obligation for unrestricted access (its Anti-DRM provision)
+  if shared [see @Poole2017].
+  Similarly, following an example from @Matt2009,
+  if one were to choose the copyleft license [CC\ BY-SA\ 4.0][cc-by-sa]
+  for this purpose, any map that is a derivative of the data
+  would also need to be licensed under [CC\ BY-SA\ 4.0][cc-by-sa]
+  (or a compatible license) if shared.
+  If the intention is to only have [derivative databases][osm-produced-work]
+  under the same license, one might want to choose the [ODbL\ 1.0][odbl] for the database,
+  as it was specifically designed not to apply to works produced from the data in the database.
+  Otherwise [CC0\ 1.0][cc0] is an excellent choice for data.
+- __works in the public domain__: If a work is already in the public domain,
+  it should be marked using the [PDM 1.0][pdm],
+  rather than applying a waiver such as the [CC0\ 1.0][cc0] (or another license).
 :::
+https://creativecommons.org/licenses/by/4.0/
+[ofl]: https://openfontlicense.org/
+[cc-by]: https://creativecommons.org/licenses/by/4.0/
+[pdm]: https://creativecommons.org/publicdomain/mark/1.0/
+[osm-produced-work]: https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline
 
@@
 
-You may prefer that adaptations of your work stay free/open and thus choose a copyleft license.
-Of course, this does not restrict you as the original author: You are still permitted to distribute the work under a different license and without sharing the source code.
+You may prefer that adaptations of your work stay free/open
+and thus choose a copyleft license.
+Of course, this does not restrict you as the original author:
+You are still permitted to distribute the work under a different license
+and without sharing the source code.
 
@@
 
-In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @FSF2024].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner.
-However, if the output is based on data, it can be considered a derivative work of the data and the license of the data also applies.
+In most cases, the output of software, like images or tables,
+does not depend on the software's license.[^output-exception]
+Therefore, if you use an R package under a copyleft license to create a figure,
+you are likely the copyright owner.
+However, if the output is based on data,
+it can be considered a derivative work of the data and the license of the data also applies.
 For example, maps may be considered as a derivative of the geographic data they are based on.
 
-It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license.
-Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense].
+[^output-exception]: The exception is when the software is part of the output [see @FSF2024].
 
-You can learn which license an installed package uses via `packageDescription("<PACKAGE_NAME>", fields = "License")`.
-And to identify which licenses are being used by the R packages your project depends on, you can use the following code:
+It is disputed whether software that uses an R package
+under AGPLv3 or GPLv3[^gplv3-explanation]
+can only be published under a GPL-compatible license
+-- or even has to be published under the same license.
+Posit, the company behind RStudio,
+does not believe that to be the case [see also @Wickham2023PackagesLicense].
+
+[^gplv3-explanation]: a variant of the AGPLv3 that does not cover software running as a service
+
+You can learn which license an installed package uses via
+`packageDescription("<PACKAGE_NAME>", fields = "License")`.
+And to identify which licenses are being used by the R packages your project depends on,
+you can use the following code:
 
@@
 
-We have prepared two advanced license flowcharts, one for software, writing, images, audio, and video in @fig-flowchart-non-data and one for data in @fig-flowchart-data where you can make additional choices.
-Note, however, that especially the advanced flowchart for licensing data is quite complex and we recommend you to seek legal counsel if you want to be sure.
+We have prepared two advanced license flowcharts,
+one for software, writing, images, audio, and video in @fig-flowchart-non-data
+and one for data in @fig-flowchart-data where you can make additional choices.
+Note, however, that especially the advanced flowchart for licensing data is quite complex
+and we recommend you to seek legal counsel if you want to be sure.
 
@@
 
-The advanced license flowcharts also allow you to make additional decisions, such as whether attribution should be a condition of redistribution.
-As you have learned, even if attribution is not a condition of the license, there are other reasons for attributing you as an author (such as scientific citation standards or moral rights).
+The advanced license flowcharts also allow you to make additional decisions,
+such as whether attribution should be a condition of redistribution.
+As you have learned, even if attribution is not a condition of the license,
+there are other reasons for attributing you as an author
+(such as scientific citation standards or moral rights).
 All conditions are explained on the page on [copyright](copyright.qmd).
 
-From the Creative Commons licenses, only [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) does not require providing attribution, and all software licenses in @fig-flowchart-non-data require providing attribution.
+From the Creative Commons licenses,
+only [CC0\ 1.0][cc0] does not require providing attribution,
+and all software licenses in @fig-flowchart-non-data require providing attribution.
 As an author, you are free to waive license conditions such as attribution, however.
@@
 
-Creative Commons also offers licenses which prevent commercial use of a work and/or sharing of derivative works.
-A free/open license, however, must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use.
-Otherwise, creating adaptations from multiple works by different authors would either limit commercial use for everybody (including the authors) or would not be possible at all (when including materials with a copyleft license), effectively creating _silos_ of works which are mutually incompatible with each other.
-This is especially relevant for work types for which adaptations are created frequently, such as text, data, and source code.
-In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
-If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012.
+Creative Commons also offers licenses which prevent commercial use of a work
+and/or sharing of derivative works.
+A free/open license, however, must allow creating derivative works
+and must allow exercising the rights granted by it _for any purpose_, including commercial use.
+Otherwise, creating adaptations from multiple works by different authors would either
+limit commercial use for everybody (including the authors)
+or would not be possible at all (when including materials with a copyleft license),
+effectively creating _silos_ of works which are mutually incompatible with each other.
+This is especially relevant for work types for which adaptations are created frequently,
+such as text, data, and source code.
+In order to avoid silos, one should only choose licenses which are compatible
+with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014].
+If you would like to learn more about the different types of compatibility,
+we recommend you to read the article
+"A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012.
 The following diagram provides an overview of the compatibility of various licenses:
 
-!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024] licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)](images/Morrison2024.jpg){.lightbox}
+!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024]
+licensed under [CC\ BY\ 4.0][cc-by]](images/Morrison2024.jpg){.lightbox}
 :::
@@
 
-Having selected the license(s) of your choice, we encourage you to read through the full license text (or at least a legal summary) to understand the legal effect.
+Having selected the license(s) of your choice,
+we encourage you to read through the full license text (or at least a legal summary)
+to understand the legal effect.
 The following applies to all licenses mentioned here:
 
-- Applying the license is irrevocable. You may stop distributing your work, but everybody who already received it may continue to share it.
+- Applying the license is irrevocable.
+  You may stop distributing your work,
+  but everybody who already received it may continue to share it.
 - As an author applying the license, you are not bound by the license terms.
-You cannot, however, give others an exclusive license to your content anymore (as some journals needlessly demand, see @wrn-exclusive-rights).
-- By applying the license you share with others your right to use your work for commercial purposes, including selling it.
-- If you apply the license to an image, the license [may also apply](https://commons.wikimedia.org/wiki/Commons:Same_work) to higher resolution version of your image.
-- If you apply a copyleft license to an image, a document in which the image is used may not constitute a derivative work and, as such, may not be affected by the copyleft condition.
-- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_] or if the work was created as part of your job, you may not be able to apply the license.
+  You cannot, however, give others an exclusive license to your content anymore
+  (as some journals needlessly demand, see @wrn-exclusive-rights).
+- By applying the license you share with others your right
+  to use your work for commercial purposes, including selling it.
+- If you apply the license to an image,
+  the license [may also apply][commons-same-work] to a higher resolution version of your image.
+- If you apply a copyleft license to an image, a document
+  in which the image is used may not constitute a derivative work
+  and, as such, may not be affected by the copyleft condition.
+- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_]
+  or if the work was created as part of your job,
+  you may not be able to apply the license.
   Check with your collecting society or your employer in that case.
+
+[commons-same-work]: https://commons.wikimedia.org/wiki/Commons:Same_work
 
@@
 
-If you are an employee at LMU and your funder does not mandate the use of a particular license, the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
-The only exception is the case that you developed an invention you would like to exploit commercially with a patent.
-Then, you are required to officially report your invention to the IP management team of the LMU, which will work out the next steps with you.
+If you are an employee at LMU and your funder does not mandate the use of a particular license,
+the decision of which license to use lies with the holder of your chair (_Lehrstuhlinhaber\*in_).
+The only exception is the case that you developed an invention
+you would like to exploit commercially with a patent.
+Then, you are required to officially report your invention to the IP management team of the LMU,
+which will work out the next steps with you.
 :::
@@
 
-- By applying the license you waive any moral, publicity, privacy, or other similar personality rights to the extent necessary to use the work.
-- The attribution requirement may be implemented "in any reasonable manner", possibly not conforming to your precise expectations.
+- By applying the license you waive any moral, publicity, privacy,
+  or other similar personality rights to the extent necessary to use the work.
+- The attribution requirement may be implemented "in any reasonable manner",
+  possibly not conforming to your precise expectations.
 
-Applying a license to a work mostly just means indicating which file or folder is covered by which license(s) (no sort of registration is required). There are multiple standards on how to do so:
+Applying a license to a work mostly just means indicating
+which file or folder is covered by which license(s)
+(no sort of registration is required).
+There are multiple standards on how to do so:
 
 - You have already learned about the file `LICENSE.txt`.
-Either it contains the full text of one license that applies to the whole project folder -- this is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/).
-Alternatively, it details which license applies to what and the invidual licenses are stored in the same or in different files.
-- A summary of the licensing situation is often given in a `README` file, whose creation will be discussed [later](make_readme.qmd).
-- If only parts of a file are covered by a certain license, then it might make sense to indicate this directly in the file.
-- The license of embedded elements (e.g., images) is either given directly beneath to where they are embedded, at the end of the respective page, or collectively on an separate page.
-- Individual programming languages also have their own way of stating which license a package is distributed under.
-For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
-- Software projects often indicate the license in every file, following the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable (explained in @nte-reuse).
-- The following two metadata standards for increasing citability also allow to indicate the license:
-  - [CodeMeta](https://codemeta.github.io/): All information are stored in a file called `codemeta.json`, which can be created [on their website](https://codemeta.github.io/codemeta-generator/).
-Alternatively, one can use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
-  - [Citation File Format](https://citation-file-format.github.io/) (CFF): Here, everything is stored in a file called `CITATION.cff`.
+  Either it contains the full text of one license that applies to the whole project folder
+  -- this is a practice propagated by GitHub,
+  which provides instructions for and comparisons of many licenses
+  via [ChooseALicense.com](https://choosealicense.com/licenses/).
+  Alternatively, it details which license applies to what
+  and the individual licenses are stored in the same or in different files.
+
+- A summary of the licensing situation is often given in a `README` file,
+  whose creation will be discussed [later](make_readme.qmd).
+
+- If only parts of a file are covered by a certain license,
+  then it might make sense to indicate this directly in the file.
+
+- The license of embedded elements (e.g., images) is either given
+  directly beneath to where they are embedded,
+  at the end of the respective page,
+  or collectively on an separate page.
+
+- Individual programming languages also have their own way
+  of stating which license a package is distributed under.
+  For R packages, this is usually set by the field `License`
+  in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
+
+- Software projects often indicate the license in every file,
+  following the [REUSE](https://reuse.software/) specification
+  to make the choice of license machine-readable (explained in @nte-reuse).
+
+- The following two metadata standards for increasing citability
+  also allow to indicate the license:
+
+  - [CodeMeta](https://codemeta.github.io/):
+    All information are stored in a file called `codemeta.json`,
+    which can be created [on their website][codemeta-gen].
+    Alternatively, one can use the R package [`codemetar`][codemetar]
+    to create this file.
+
+  - [Citation File Format](https://citation-file-format.github.io/) (CFF):
+    Here, everything is stored in a file called `CITATION.cff`.
     It is used by GitHub to directly show a recommended citation for a repository.
-The file can be created [on their website](https://citation-file-format.github.io/cff-initializer-javascript/).
-Alternatively, the R package [`cffr`](https://cran.r-project.org/package=cffr) may be used for that purpose.
+    The file can be created [on their website][cff-gen].
+    Alternatively, the R package [`cffr`][cffr] may be used for that purpose.
+
+[codemeta-gen]: https://codemeta.github.io/codemeta-generator/
+[codemetar]: https://docs.ropensci.org/codemetar/
+[cff-gen]: https://citation-file-format.github.io/cff-initializer-javascript/
+[cffr]: https://cran.r-project.org/package=cffr
 
@@
 
-For all the licenses recommended in this tutorial, the organizations that created these licenses provide more information on how to apply them to your work:
+For all the licenses recommended in this tutorial,
+the organizations that created these licenses provide more information
+on how to apply them to your work:
 
@@
 
-Creative Commons even provides a range of considerations for licensors and licensees [@CC2013] and an [interactive chooser](https://creativecommons.org/choose/) which you can use to create text snippet that you can copy and paste to the desired location.
+Creative Commons even provides a range of considerations for licensors and licensees [@CC2013]
+and an [interactive chooser](https://creativecommons.org/choose/)
+which you can use to create text snippet that you can copy and paste to the desired location.
 :::
@@
 
-Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/) which allows communicating the license choice unequivocally.
-We will be using that to indicate the license for every file in your project folder, along with the year of publication and the copyright holder.
-To do this, we add a comment to the beginning of every file and include the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`.
+Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/)
+which allows communicating the license choice unequivocally.
+We will be using that to indicate the license for every file in your project folder,
+along with the year of publication and the copyright holder.
+To do this, we add a comment to the beginning of every file
+and include the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`.
 How this works depends on the file type, as the syntax for a comment varies.
 
-For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<YOUR NAME>` with the current year and your name -- of course, you can also choose a different license:
+For example, if you previously created the file `create_data_dictionary.R`,
+you can now add the following comment to the beginning of the file,
+replacing `<YEAR>` and `<YOUR NAME>` with the current year and your name
+-- of course, you can also choose a different license:
 
@@
 
-You need to use `#` to start the comment because this is the symbol that starts comment lines in R scripts.
-Alternatively, you can use the [reuse tool](https://github.com/fsfe/reuse-tool) to add these information for you.
+You need to use `#` to start the comment
+because this is the symbol that starts comment lines in R scripts.
+Alternatively, you can use the [reuse tool](https://github.com/fsfe/reuse-tool)
+to add these information for you.
 After installing it with...
@@
 
-...you can add the copyright information using the following command -- the current year will be added automatically:
+...you can add the copyright information using the following command
+-- the current year will be added automatically:
 
@@
 In many cases, the reuse tool will figure out the appropriate comment style for you.
-If this is not the case, as currently with Quarto files, you can tell it directly which comment style to use (`html` in this case):
+If this is not the case, as currently with Quarto files,
+you can tell it directly which comment style to use (`html` in this case):
 
@@
 
-Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which is what you indicate with the previous comment.
-As you edited the file throughout this tutorial, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them.
+Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0][cc0],
+which is what you indicate with the previous comment.
+As you edited the file throughout this tutorial,
+you may also add yourself.^[For licenses that require that modifications are indicated,
+this is an easy way to comply with them.
 Also, you do not need to provide your real name.]
 
-Sometimes, there are file types which do not allow for adding the license information inside them, such as PDF and CSV files.
+Sometimes, there are file types which do not allow
+for adding the license information inside them, such as PDF and CSV files.
 For these, a corresponding `.license` file can be created.
-Try the following command which indicates that the data were published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
+Try the following command which indicates that the data were published
+under [CC0\ 1.0][cc0]:
 
@@
 
-You will notice that this creates another file called `data.csv.license` containing the relevant information:
+You will notice that this creates another file called `data.csv.license`
+containing the relevant information:
 
@@
 
-If you want to indicate the license for all files in a particular folder, you can create a file called `REUSE.toml` and add an `[[annotations]]` table for them:
+If you want to indicate the license for all files in a particular folder,
+you can create a file called `REUSE.toml` and add an `[[annotations]]` table for them:
 
@@
 Finally, there may be some minor files which are build artifacts.
-You can either add them to your `.gitignore` file or use CC0\ 1.0 with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder.
-For more information, also discussing other corner cases, you can read their [Frequently Asked Questions](https://reuse.software/faq/).
+You can either add them to your `.gitignore` file or use CC0\ 1.0
+with a copyright tag such as `SPDX-FileCopyrightText: NONE`
+to assert that there is no copyright holder.
+For more information, also discussing other corner cases,
+you can read their [Frequently Asked Questions](https://reuse.software/faq/).
 
@@
 
-The manuscript and the bibliography file you downloaded for this tutorial are made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary).
+The manuscript and the bibliography file you downloaded for this tutorial
+are made available to you under [CC0\ 1.0][cc0]
+as is all code in this tutorial (e.g., for creating the data dictionary).
 Therefore, you can apply any license to your modified version.
-Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).
+Please apply the license of your choice now,
+using any of the methods mentioned above (except the `README`, which we will cover later).
 
@@
 
-In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.
+In the following example, the edited manuscript files have been multi-licensed
+under [AGPLv3][agpl] (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0][cc-by-sa],
+while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.
+
+[agpl]: https://www.gnu.org/licenses/agpl-3.0.html
 
@@
 
-As the AGPLv3 requires reusers adding the [full license text](https://www.gnu.org/licenses/agpl-3.0.txt), we recommend to also copy it to the project folder (if not already in there).
+As the AGPLv3 requires reusers adding the [full license text](https://www.gnu.org/licenses/agpl-3.0.txt),
+we recommend to also copy it to the project folder (if not already in there).
 :::

### 2025-02-08 code.qmd
Minor improvements
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 With the data added to the project folder, you can now analyze them and describe the results.
-For the purpose of this tutorial, we suggest to conduct a simple analysis that explores whether the bill lengths of penguins vary between sex.
+For the purpose of this tutorial, we suggest to conduct a simple analysis
+that explores whether the bill lengths of penguins vary between sex.
 You can write your own or use the following example:
@@
 
-- Data may be collected from social network sites using their APIs^[An **a**pplication **p**rogramming **i**nterface provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
+- Data may be collected from social network sites using their APIs[^api-explanation]
+  or downloaded from a data repository, or
 - an analysis may be conducted with the help of AI providers.
 
-In these cases, make sure that the code you check in to Git does not contain any credentials that are required for accessing these services.
-Instead, make use of environment variables which are defined in a location that is excluded from version control.
-When programming with R, you can define them in a file called `.Renviron` in the root of your project folder:
+[^api-explanation]: An **a**pplication **p**rogramming **i**nterface
+provides the capability to interact with other software using a programming language.
+
+In these cases, make sure that the code you check in to Git
+does not contain any credentials that are required for accessing these services.
+Instead, make use of environment variables which are defined
+in a location that is excluded from version control.
+When programming with R, you can define them
+in a file called `.Renviron` in the root of your project folder:
 
@@
 
-When you start a new session from the project folder, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:
+When you start a new session from the project folder,
+the file is automatically read by R
+and the environment variables can be accessed using `Sys.getenv()`:
 
@@
 
-Make sure that `.Renviron` is added to your `.gitignore` file in order to exclude it from the Git repository. If you already committed a file that contains credentials, you can follow @Chacon2024.
+Make sure that `.Renviron` is added to your `.gitignore` file
+in order to exclude it from the Git repository.
+If you already committed a file that contains credentials, you can follow @Chacon2024.
 :::
@@
 
-To format manuscripts according to the requirements of a particular journal, [multiple Quarto extensions are available](https://quarto.org/docs/extensions/listing-journals.html).
-In the following, we will demonstrate the usage of [`apaquarto`](https://wjschne.github.io/apaquarto/), which typesets documents according to the requirements of the American Psychological Association [-@APA2020].
+To format manuscripts according to the requirements of a particular journal,
+[multiple Quarto extensions are available][apaquarto-extensions].
+In the following, we will demonstrate the usage of [`apaquarto`][apaquarto],
+which typesets documents according to the requirements
+of the American Psychological Association [-@APA2020].
 Of course, you may also use an extension for a different journal.
+
+[apaquarto]: https://wjschne.github.io/apaquarto/
+[apaquarto-extensions]: https://quarto.org/docs/extensions/listing-journals.html
 
@@
 ::: {#tip-update-quarto .callout-tip}
-If the PDF file cannot be created, try updating Quarto. It comes bundled with RStudio, however, `apaquarto` sometimes requires more recent versions.
+If the PDF file cannot be created, try updating Quarto.
+It comes bundled with RStudio,
+however, `apaquarto` sometimes requires more recent versions.
 :::
@@
 For example, this is the case for the Quarto extension that you downloaded earlier.
-These cases matter from a copyright perspective and you need to make sure to have permission for redistribution.
+These cases matter from a copyright perspective
+and you need to make sure to have permission for redistribution.
 
@@
 
-According to its GitHub page, `apaquarto` is licensed under CC0\ 1.0. This means that it can be used and distributed even without attribution.
+According to its GitHub page, `apaquarto` is licensed under CC0\ 1.0.
+This means that it can be used and distributed even without attribution.
 :::
@@
 
+It is best to provide attribution even if the license does not require it.
+
 ```{.txt .code-overflow-wrap filename="LICENSE.txt"}
@@
 
-Again, if the license required adding the full license text, you would also need to copy it to the project folder (if not already in there).
+Again, if the license required adding the full license text,
+you would also need to copy it to the project folder (if not already in there).
 :::
 
-Then, there is software that is used (e.g., during the data analysis), but not copied into the project folder. Here, questions of copyright due to redistribution do not apply.
+Then, there is software that is used (e.g., during the data analysis),
+but not copied into the project folder.
+Here, questions of copyright due to redistribution do not apply.
 
-Of course, being transparent about used software for reasons of reproducibility and academic integrity matters in any case.
-You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not.
+Of course, being transparent about used software
+for reasons of reproducibility and academic integrity matters in any case.
+You can consult @fig-software-citation for general advice
+whether to cite a particular piece of software or not.
 As with data, citations should allow for exact identification and access.
-From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):
+From the six "software citation principles" by @Smith2016,
+licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):
 
-> **1\.
-Importance**: Software should be considered a legitimate and citable product of research.
-Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be included in the metadata of the citing work, for example in the reference list of a journal article, and should not be omitted or separated.
-Software should be cited on the same basis as any other research product such as a paper or a book, that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
+> **1\. Importance**: Software should be considered a legitimate and citable
+> product of research. Software citations should be accorded the same importance
+> in the scholarly record as citations of other research products, such as
+> publications and data; they should be included in the metadata of the citing
+> work, for example in the reference list of a journal article, and should not
+> be omitted or separated. Software should be cited on the same basis as any
+> other research product such as a paper or a book, that is, authors should cite
+> the appropriate set of software products just as they cite the appropriate set
+> of papers.
 > 
-> **5\.
-Accessibility**: Software citations should facilitate access to the software itself and to its associated metadata, documentation, data, and other materials necessary for both humans and machines to make informed use of the referenced software.
+> **5\. Accessibility**: Software citations should facilitate access to the
+> software itself and to its associated metadata, documentation, data, and other
+> materials necessary for both humans and machines to make informed use of the
+> referenced software.
 > 
-> **6\.
-Specificity**: Software citations should facilitate identification of, and access to, the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
+> **6\. Specificity**: Software citations should facilitate identification of,
+> and access to, the specific version of software that was used. Software
+> identification should be as specific as necessary, such as using version
+> numbers, revision numbers, or variants such as platforms.
 
 In practice, you would identify all pieces of software the project relies on.
-A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before.
-Then there are the individual R packages^[By default, `renv` avoids copying them into the project folder.
-And even that happens (e.g., by using `renv::isolate()`) they are excluded from Git by default.], Quarto extensions, and $\TeX$ packages.
-All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper.
-For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit.
-A system dependency is additional software that you require on your computer apart from the R package.
+A few of them are obvious, such as R itself, Quarto,
+and the $\TeX$ distribution we installed before.
+Then there are the individual R packages[^renv-copying],
+Quarto extensions, and $\TeX$ packages.
+All of them, in turn, may have dependencies
+and it is up to you decide when not to dig deeper.
+For example, some R packages are only thin wrappers around other R packages
+or around system dependencies which also might deserve credit.
+A system dependency is additional software
+that you require on your computer apart from the R package.
+
+[^renv-copying]: By default, `renv` avoids copying them into the project folder.
+And even that happens (e.g., by using `renv::isolate()`)
+they are excluded from Git by default.
 
@@
 
-"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Simplified from original.
+"Should I cite the software?" by @Brown2016 licensed under [CC\ BY-SA\ 4.0][cc-by-sa].
+Simplified from original.
 :::
 
+[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/
+
 Now, add references for the software you would like to cite to the manuscript.
-In the following, we will demonstrate this for R and all R packages by using the R package `grateful`.
-For arbitrary software, you can use [CiteAs](https://citeas.org/) or the [DOI Citation Formatter](https://citation.doi.org/) to create appropriate citations.
+In the following, we will demonstrate this for R and all R packages
+by using the R package `grateful`.
+For arbitrary software,
+you can use [CiteAs][citeas] or the [DOI Citation Formatter][doi-citation]
+to create appropriate citations.
+
+[citeas]: https://citeas.org/
+[doi-citation]: https://citation.doi.org/
 
@@
 
-This will automatically create a paragraph citing all used packages and generate the bibliography file `grateful-refs.bib`.^[Note that this automatic detection can miss packages in some circumstances, therefore always verify the rendered result.] Then, in the YAML header, add `grateful-refs.bib` by setting the `bibliography` as follows:
+This will automatically create a paragraph citing all used packages
+and generate the bibliography file `grateful-refs.bib`.[^automatic.note]
+Then, in the YAML header,
+add `grateful-refs.bib` by setting the `bibliography` as follows:
+
+[^automatic.note]: Note that this automatic detection
+can miss packages in some circumstances,
+therefore always verify the rendered result.
 
@@
 
-Although we provide you with example code in this tutorial, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.
+Although we provide you with example code in this tutorial,
+a few things remain to be said about best practices
+when it comes to writing code that is readable and maintainable.
 
-- __Use project-relative paths.__ When you refer to a file within your project folder, write paths relative to the root (the uppermost folder in your project).
-For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.
+- __Use project-relative paths.__
+  When you refer to a file within your project folder,
+  write paths relative to the root (the uppermost folder in your project).
+  For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`,
+  instead write `images/result.png`.
 
-- __Keep it simple.__ Add complexity only when you must.
-Whenever there's a boring way to do something and a clever way, go for the boring way.
-If the code grows increasingly complex, refactor it into separate functions and files.
+- __Keep it simple.__
+  Add complexity only when you must.
+  Whenever there's a boring way to do something and a clever way,
+  go for the boring way.
+  If the code grows increasingly complex,
+  refactor it into separate functions and files.
 
-- __Don't repeat yourself.__ Use variables and functions before you start to write (or copy-paste) the same thing twice.
+- __Don't repeat yourself.__
+  Use variables and functions before you start
+  to write (or copy-paste) the same thing twice.
 
-- __Use comments to explain why you do things.__ The code already shows what you do. Use comments to summarize it and explain why you do it.
+- __Use comments to explain why you do things.__
+  The code already shows what you do.
+  Use comments to summarize it and explain why you do it.
 
-- __Don't reinvent the wheel.__ With R, chances are that what you need to do is greatly facilitated by a package from one of many high-quality collections such as [rOpenSci](https://ropensci.org/packages/), [r-lib](https://github.com/r-lib), [Tidyverse](https://www.tidyverse.org/packages/), or [fastverse](https://fastverse.github.io/fastverse/).
+- __Don't reinvent the wheel.__
+  With R, chances are that what you need to do is greatly facilitated
+  by a package from one of many high-quality collections
+  such as [rOpenSci][ropensci], [r-lib][r-lib], [Tidyverse][tidyverse],
+  or [fastverse][fastverse].
 
-- __Think twice about your dependencies.__ Every dependency increases the risk of irreproducibility in the future.
-Prefer packages that are well-maintained and light on dependencies^[You can use the function `pak::pkg_deps()` to count the total number of package dependencies in R.].
-We also recommend you to read "When should you take a dependency?" by @Wickham2023PackagesDependency.
+- __Think twice about your dependencies.__
+  Every dependency increases the risk of irreproducibility in the future.
+  Prefer packages that are well-maintained and light on dependencies[^pkg-deps].
+  We also recommend you to read "When should you take a dependency?"
+  by @Wickham2023PackagesDependency.
 
-- __Fail early, often, and noisily.__ Whenever you expect a certain state, use assertions to be sure. In R, you can use `stopifnot()` to make sure that a condition is actually true.
+- __Fail early, often, and noisily.__
+  Whenever you expect a certain state, use assertions to be sure.
+  In R, you can use `stopifnot()` to make sure that a condition is actually true.
 
-- __Test your code.__ Test your code with scenarios where you know what the result should be.
-Turn bugs you discovered into test cases.
-Use linting tools^[A linting tool analyzes your code without actually running it.
-Therefore, this process is also called static code analysis.] to identify common mistakes in your code, for example, the R package [`lintr`](https://lintr.r-lib.org/).
+- __Test your code.__
+  Test your code with scenarios where you know what the result should be.
+  Turn bugs you discovered into test cases. Use linting tools[^linting-explanation]
+  to identify common mistakes in your code,
+  for example, the R package [`lintr`](https://lintr.r-lib.org/).
 
-- __Read through a style guide and follow it.__ A style guide is a set of stylistic conventions that improve the code quality.
-R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide" and use the R package [`styler`](https://styler.r-lib.org/).
+- __Read through a style guide and follow it.__
+  A style guide is a set of stylistic conventions that improve the code quality.
+  R users are recommended to read Wickham's [-@Wickham2022Style] "Tidyverse style guide"
+  and use the R package [`styler`](https://styler.r-lib.org/).
   Python users may benefit from reading the "Style Guide for Python Code" by @VanRossum2013.
@@
 
-This is only a brief summary and there is much more to be learned about coding practices. If you want to dive deeper we recommend the following resources:
+[ropensci]: https://ropensci.org/packages/
+[r-lib]: https://github.com/r-lib
+[tidyverse]: https://www.tidyverse.org/packages/
+[fastverse]: https://fastverse.github.io/fastverse/
+
+[^pkg-deps]: You can use the function `pak::pkg_deps()`
+to count the total number of package dependencies in R.
+
+[^linting-explanation]: A linting tool analyzes your code without actually running it. Therefore, this process is also called static code analysis.
+
+This is only a brief summary and there is much more to be learned about coding practices.
+If you want to dive deeper we recommend the following resources:
 
@@
 
-> "Any fool can write code that a computer can understand. Good programmers write code that humans can understand."
+> "Any fool can write code that a computer can understand.
+> Good programmers write code that humans can understand."
 > 
@@
 `renv` only records the versions of R packages and of R itself.
-This means that everything we have not decided to cite or attribute is not documented anywhere.
+This means that everything we have not decided to cite or attribute
+is not documented anywhere.
 We will cover system dependencies when [creating a `README`](make_readme.qmd).
-For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies).
+For now, however, there is one simple step you can take
+to record the version of Quarto (and a few other dependencies).
 Do run the following:
@@
 
-This will create a few additional files which facilitate reconstructing the computational environment in the future.^[Either using [repo2docker](https://repo2docker.readthedocs.io/) or the public [binder](https://mybinder.org/) service.] As always, commit your changes:
+This will create a few additional files which facilitate
+reconstructing the computational environment in the future.[^binder-note]
+As always, commit your changes:
+
+[^binder-note]: Either using [repo2docker](https://repo2docker.readthedocs.io/)
+or the public [binder](https://mybinder.org/) service.
 

### 2025-02-08 copyright.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 By default, everything you put into the project folder will be shared publicly.
-In many instances, this will also include works by others than yourself or your co-authors, because you may use existing data, images or software.
+In many instances, this will also include works by others than yourself or your co-authors,
+because you may use existing data, images or software.
 In these cases, you should make sure to only share them to the extent you are allowed to.
-In the following, we will be focusing on copyright, but of course, there can also be other restrictions.
+In the following, we will be focusing on copyright,
+but of course, there can also be other restrictions.
 
@@
 
-Every expression of an idea, such as a literary or an artistic works, is automatically covered by copyright.
-Put simply, this means that nobody else than the copyright holder is allowed to copy, modify, or share it.
-Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
-Certain exceptions and limitations to copyright (such as fair use or the right to quote) may obviate the need for a license.
+Every expression of an idea, such as a literary or an artistic works,
+is automatically covered by copyright.
+Put simply, this means that nobody else than the copyright holder
+is allowed to copy, modify, or share it.
+Copyright licenses can grant others some of the necessary rights,
+but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
+Certain exceptions and limitations to copyright (such as fair use or the right to quote)
+may obviate the need for a license.
 
 With regard to copyright, free/open licenses can give permission to share a work by others.
-These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things.
+These licenses often require to attribute the authors,
+indicate whether changes were made,
+and provide the text of the license, among other things.
 
-The following video "The correct indication of a license" embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy)) explains the correct attribution of Creative Commons licenses.
+The following video "The correct indication of a license"
+embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy))
+explains the correct attribution of works under Creative Commons licenses.
 
@@
 
-Whenever authors create a literary or artistic work (such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_] in most countries limits other people from copying, modifying, and sharing it without the author's express permission.
-This even applies if the author makes it available to others (e.g., on their website): First and foremost others are not allowed to copy, modify, or share it.
-Importantly, this legal default of "all rights reserved" was created to benefit publishers, not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
+Whenever authors create a literary or artistic work
+(such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_]
+in most countries limits other people from copying, modifying, and sharing it
+without the author's express permission.
+This even applies if the author makes it available to others (e.g., on their website):
+First and foremost others are not allowed to copy, modify, or share it.
+Importantly, this legal default of "all rights reserved" was created to benefit publishers,
+not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
 Copyright licenses enable authors to free up their works for reuse by others.
 
-Note that copyright law [varies between countries](https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory) and the rule of territoriality (_lex loci protectionis_) applies: Applicable are the laws of the country where a work is used.
-Also, if the work was created as part of the author's job, it might be their employer who holds the copyright to the work, depending on the country and contract.
+Note that copyright law [varies between countries][commons-rules]
+and the rule of territoriality (_lex loci protectionis_) applies:
+Applicable are the laws of the country where a work is used.
+Also, if the work was created as part of the author's job,
+it might be their employer who holds the copyright to the work,
+depending on the country and contract.
 
-A license is a legal document that regulates what other people than the author can and cannot do with a copyright-protected work, and under which conditions.
-The license may be exclusive, such that it is only granted to one recipient, effectively limiting the author in dealing with their own work.
+[commons-rules]: https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory
+
+A license is a legal document that regulates what other people than the author
+can and cannot do with a copyright-protected work, and under which conditions.
+The license may be exclusive, such that it is only granted to one recipient,
+effectively limiting the author in dealing with their own work.
 It can also be non-exclusive, thus giving rights to others without the author losing them.
-To avoid that every author who would like to share their work writes their own license, pre-formulated "boilerplate" licenses have been created.
-Recognizing them facilitates both sharing (on the side of the author) and re-using (on the side of the recipient).^[In fact, authors should not create their own license texts if the purpose is to share their work freely and openly with others.
-By using a boilerplate license others know what to expect, also because it may have already been [tested](https://sfconservancy.org/copyleft-compliance/) in [courts](https://legaldb.creativecommons.org/).]
+To avoid that every author who would like to share their work writes their own license,
+pre-formulated "boilerplate" licenses have been created.
+Recognizing them facilitates both sharing (on the side of the author)
+and re-using (on the side of the recipient).^[In fact, authors should not
+create their own license texts if the purpose is to share their work freely
+and openly with others.
+By using a boilerplate license others know what to expect,
+also because it may have already been [tested][sfconservancy] in [courts][cc-db].]
+
+[sfconservancy]: https://sfconservancy.org/copyleft-compliance/
+[cc-db]: https://legaldb.creativecommons.org/
 
@@
 An important class of boilerplate licenses are _free/open_ licenses.
-Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified, by anyone, for any purpose" [@Moller2015].
-Importantly, this also means that others do not need to ask or notify the author and that they can use it for commercial purposes.
-By the choice of license, authors can, however, demand that they are credited, that the original license is indicated, that modifications are indicated, that derivative works are only shared under the same license, and that no further restrictions are imposed on the work.
-Software licenses may additionally require to make the source code available to everybody the software is shared with and often require to display the full text of the license upon distribution.
+Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified,
+by anyone, for any purpose" [@Moller2015].
+Importantly, this also means that others do not need to ask or notify the author
+and that they can use it for commercial purposes.
+By the choice of license, authors can, however, demand that they are credited,
+that the original license is indicated, that modifications are indicated,
+that derivative works are only shared under the same license,
+and that no further restrictions are imposed on the work.
+Software licenses may additionally require to make the source code available
+to everybody the software is shared with
+and often require to display the full text of the license upon distribution.
 
@@
 The terms _free_ and _open_, especially with regard to software, come with a history.
-The Free Software Foundation (FSF) was founded in 1985 to protect "four essential freedoms" [-@FSF2024] of a program's user.
+The Free Software Foundation (FSF) was founded in 1985
+to protect "four essential freedoms" [-@FSF2024] of a program's user.
 These are the freedoms to use, study, share, and improve a program.
-Software whose users legally and practically have these freedoms (because, among other things, they have access to its source code) is considered _free_.
-The four freedoms are seen as vital for a society as a whole in the sense that they enable sharing, cooperation and ultimately freedom in general.
-The FSF maintains [a list of software licenses](https://www.gnu.org/licenses/license-list.en.html#SoftwareLicenses) that it considers to be protecting the four freedoms.
-Sometimes the term _libre_ (Spanish and French for _free_) is used to make a distinction from _gratis_ software.
+Software whose users legally and practically have these freedoms
+(because, among other things, they have access to its source code) is considered _free_.
+The four freedoms are seen as vital for a society as a whole
+in the sense that they enable sharing, cooperation and ultimately freedom in general.
+The FSF maintains [a list of software licenses][fsf-list]
+that it considers to be protecting the four freedoms.
+Sometimes the term _libre_ (Spanish and French for _free_)
+is used to make a distinction from _gratis_ software.
 You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).
 
+[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses
+
 The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach.
-It is concerned with developing high-quality software, for which everyone's ability to obtain, modify and contribute back the source code is considered beneficial.
-Access to the source code is one out of multiple conditions for software to be considered _open source_ by the OSI [-@OSI2007], which equally maintains [a list of approved licenses](https://opensource.org/licenses).
+It is concerned with developing high-quality software,
+for which everyone's ability to obtain, modify and contribute back the source code
+is considered beneficial.
+Access to the source code is one out of multiple conditions for software
+to be considered _open source_ by the OSI [-@OSI2007],
+which equally maintains [a list of approved licenses][osi-list].
 
-Access to the source code is a necessary, but not sufficient requirement both for _free software_ and for _open source software_.
-Conversely, both can (and frequently are) sold for money, as their respective criteria only apply once one has access to the software.
-Throughout this tutorial, we write "free/open license" to mean a license that is approved by either the FSF or the OSI.
-Software which is neither but makes available its source code is sometimes referred to as _source-available software_.
+[osi-list]: https://opensource.org/licenses
 
-Apart from software, the term _open access_ has often been used for works that are available at no cost.
-For example, this is the commonality of bronze, green, hybrid, gold, and diamond/platinum open access _articles_, which otherwise vary in the rights that are granted to readers.
-In 2002, the Budapest Open Access Initiative declared that _open access_ additionally includes the right to use articles for any purpose, and in 2003, the Bethesda Statement and the Berlin Declaration added the right to make derivative works.
+Access to the source code is a necessary, but not sufficient requirement
+both for _free software_ and for _open source software_.
+Conversely, both can (and frequently are) sold for money,
+as their respective criteria only apply once one has access to the software.
+Throughout this tutorial, we write "free/open license"
+to mean a license that is approved by either the FSF or the OSI.
+Software which is neither but makes available its source code
+is sometimes referred to as _source-available software_.
 
-Two other notable definitions include the _Open Definition_ [@OKFN2016], which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015], for which the open editing phase began in 2006.
+Apart from software, the term _open access_ has often been used for works
+that are available at no cost.
+For example, this is the commonality of bronze, green, hybrid, gold,
+and diamond/platinum open access _articles_,
+which otherwise vary in the rights that are granted to readers.
+In 2002, the Budapest Open Access Initiative declared
+that _open access_ additionally includes the right to use articles for any purpose,
+and in 2003, the Bethesda Statement and the Berlin Declaration added
+the right to make derivative works.
+
+Two other notable definitions include the _Open Definition_ [@OKFN2016],
+which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015],
+for which the open editing phase began in 2006.
 They are largely viewed as compatible with one another.
@@
 
-To determine whether a particular work can be included in the project folder and shared publicly, the first step is to search for its license.
-For example, this tutorial has an [about page](about.qmd) detailing its licensing, whereas other websites sometimes state the license in the footer.
+To determine whether a particular work can be included in the project folder
+and shared publicly, the first step is to search for its license.
+For example, this tutorial has an [about page](about.qmd) detailing its licensing,
+whereas other websites sometimes state the license in the footer.
 Other places to look include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
-If a project has been stored in a research repository, yet another means of communicating its copyright status might exist.
-For media from Wikimedia Commons, an [Attribution Generator](https://lizenzhinweisgenerator.de/?lang=en) exists.
+If a project has been stored in a research repository,
+yet another means of communicating its copyright status might exist.
+For media from Wikimedia Commons, an [Attribution Generator][commons-gen] exists.
+
+[commons-gen]: https://lizenzhinweisgenerator.de/?lang=en
 
@@
 
-For works by one or only few authors, it is possible that one license consistently applies to the whole content. More often than not, however, the situation is a little more complicated:
+For works by one or only few authors, it is possible
+that one license consistently applies to the whole content.
+More often than not, however, the situation is a little more complicated:
 
-- Different parts of a work may be available under different licenses: A document may incorporate direct quotes or images by others.
+- Different parts of a work may be available under different licenses:
+  A document may incorporate direct quotes or images by others.
   In this case, the license of the document may not cover those.
   In the best case, the authors indicate which parts of the work are covered by which license.
-- A work may be available under different licenses at the choice of the (re-)user. This is also called multi-licensing. For example, authors might allow to user a work under any version of a license.
-- In contrast, it may also be the case that one has to comply with the conditions of multiple licenses at the same time.
-This is the case, for example, if a derivative work is published under a different license than the original work.
+- A work may be available under different licenses at the choice of the (re-)user.
+  This is also called multi-licensing.
+  For example, authors might allow to user a work under any version of a license.
+- In contrast, it may also be the case that one has to comply
+  with the conditions of multiple licenses at the same time.
+  This is the case, for example, if a derivative work is published
+  under a different license than the original work.
 :::
 
-If you cannot find any license, the work might be in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law.
-For example, the work may not cross the [threshold of originality](https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality), it may not [qualify for copyright protection](https://commons.wikimedia.org/wiki/Commons:Unprotected_works) -- such as official works by the government --, or the copyright protection might have expired.
-According to the [Berne convention](https://en.wikipedia.org/wiki/Berne_Convention), copyright expires at earliest 50 years after the author's death, but in [many countries](https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country) it expires even later -- you can use a [public domain calculator](https://lawflow.org/public-domain-calculators/) to check.
-If no license notice is present and the work is not in the public domain, it is _in copyright_ and you are generally not allowed to share it with others.
-You can, however, ask the author to put the corresponding work under a free/open license -- Wikimedia Commons provides an [email template](https://commons.wikimedia.org/wiki/Commons:Email_templates) to obtain such a permission.
-Without permission you can check you national laws for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_] Examples include the right to quote, [trivial use](https://commons.wikimedia.org/wiki/Commons:De_minimis), and the US-American [fair use](https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?).
+If you cannot find any license, the work might be
+in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law.
+For example, the work may not cross the [threshold of originality][commons-threshold],
+it may not [qualify for copyright protection][commons-unprotected]
+-- such as official works by the government --,
+or the copyright protection might have expired.
+According to the [Berne convention][berne-convention],
+copyright expires at earliest 50 years after the author's death,
+but in [many countries][commons-pd] it expires even later
+-- you can use a [public domain calculator][pd-calc] to check.
+If no license notice is present and the work is not in the public domain,
+it is _in copyright_ and you are generally not allowed to share it with others.
+You can, however, ask the author to put the corresponding work under a free/open license
+-- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.
+Without permission you can check you national laws
+for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_]
+Examples include the right to quote, [trivial use][commons-minimis],
+and the US-American [fair use][wiki-fair-use].
 
-Having found a license, you need to determine (1) whether it is a standard (boilerplate) license, (2) whether it is a free/open, and (3) whether it is appropriate for the particular type of work and use case.
+[commons-threshold]: https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality
+[commons-unprotected]: https://commons.wikimedia.org/wiki/Commons:Unprotected_works
+[berne-convention]: https://en.wikipedia.org/wiki/Berne_Convention
+[commons-pd]: https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country
+[pd-calc]: https://lawflow.org/public-domain-calculators/
+[commons-templates]: https://commons.wikimedia.org/wiki/Commons:Email_templates
+[commons-minimis]: https://commons.wikimedia.org/wiki/Commons:De_minimis
+[wiki-fair-use]: https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?
 
-1.
-__Standard licenses__ are usually stated by their name (or an abbreviation), which may include a version number.
-Most standard licenses were assigned an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification, though the identifier is often not indicated for a particular work (which would be best practice).
+Having found a license, you need to determine
+(1) whether it is a standard (boilerplate) license,
+(2) whether it is a free/open,
+and (3) whether it is appropriate for the particular type of work and use case.
+
+1. __Standard licenses__ are usually stated by their name (or an abbreviation),
+   which may include a version number.
+   Most standard licenses were assigned
+   an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification,
+   though the identifier is often not indicated for a particular work
+   (which would be best practice).
    Also, not all standard licenses have SPDX identifiers.
    You can view a [list of all licenses with SPDX identifiers here](https://spdx.org/licenses/).
-2.
-__Free/open__: We have compiled lists of the most common free/open licenses below, categorized by the type of work they apply to.
-Be sure to look for an exact match, as there are often many licenses with very similar names.
-If a particular license is not listed below, we also provide links to other resources which you can use to determine whether a license is free/open.
-3.
-__Type of work and use case__: Many licenses were written with a particular type of work in mind, along with a certain use case.
-Software licenses, for example, may consider that the source code is the preferred form for making modifications and that software can be encumbered by patents, while licenses for data can differentiate between the database and any works produced from it.
-Take caution if a license applies to a type work for which it was not originally created and conflicts with your use case.
-Sometimes, the license name hints at the intended type of work, but when in doubt always read the full license text (and/or consult with a lawyer) to determine whether the conditions are acceptable to you.
+2. __Free/open__: We have compiled lists of the most common free/open licenses below,
+   categorized by the type of work they apply to.
+   Be sure to look for an exact match,
+   as there are often many licenses with very similar names.
+   If a particular license is not listed below, we also provide links to other resources
+   which you can use to determine whether a license is free/open.
+3. __Type of work and use case__:
+   Many licenses were written with a particular type of work in mind,
+   along with a certain use case.
+   Software licenses, for example, may consider that the source code
+   is the preferred form for making modifications
+   and that software can be encumbered by patents,
+   while licenses for data can differentiate between the database
+   and any works produced from it.
+   Take caution if a license applies to a type work for which it was not originally created
+   and conflicts with your use case.
+   Sometimes, the license name hints at the intended type of work,
+   but when in doubt always read the full license text (and/or consult with a lawyer)
+   to determine whether the conditions are acceptable to you.
 
@@
 
-For a software license to be considered free/open, it must be approved by either the [Open Source Initiative](https://opensource.org/licenses) (OSI) or the [Free Software Foundation](https://www.gnu.org/licenses/license-list.html#SoftwareLicenses) (FSF).
+For a software license to be considered free/open,
+it must be approved by either the [Open Source Initiative][osi-list] (OSI)
+or the [Free Software Foundation][fsf-list] (FSF).
 Some of the most common free/open software licenses are listed below:
@@
 
-If you did not find a particular software license in this list and consult the resources by the OSI and/or FSF linked above, be cautious because they also approve special-purpose licenses that are unfit for software in the general case.
-For example, the license CC0\ 1.0 expressively does not grant patent rights and one should only use software under that license if it is most likely not covered by patents.
+If you did not find a particular software license in this list
+and consult the resources by the OSI and/or FSF linked above,
+be cautious because they also approve special-purpose licenses
+that are unfit for software in the general case.
+For example, the license CC0\ 1.0 expressively does not grant patent rights
+and one should only use software under that license if it is most likely not covered by patents.
 
@@
 Many authors of text and media put their work under a license created by Creative Commons.
-Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use or the distribution of modified works (such as sharing a translated version).
-Therefore, we recommend to only use works made available to you under CC0 or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`), but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`).
+Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use
+or the distribution of modified works (such as sharing a translated version).
+Therefore, we recommend to only use works made available to you under CC0
+or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`),
+but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`).
 In their most recent version this includes the following licenses:
@@
 
-  [Link to full text](https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title)
+  [Link to full text][cc0-text]
   </details>
@@
 
-  [Link to full text](https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title)
+  [Link to full text][cc-by-text]
   </details>
@@
 
-  [Link to full text](https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title)
+  [Link to full text][cc-by-sa-text]
   </details>
@@
 
-Of course, if you encounter a Creative Commons license with a previous version that satisfies the description above, it can be also considered free/open.
-If you encounter a license not from Creative Commons and want to determine whether it is free/open, you can consult the [lists of licenses accepted at Wikimedia Commons](https://commons.wikimedia.org/wiki/Commons:Copyright_tags).
+[cc0-text]: https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title
+[cc-by-text]: https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title
+[cc-by-sa-text]: https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title
+
+Of course, if you encounter a Creative Commons license with a previous version
+that satisfies the description above, it can be also considered free/open.
+If you encounter a license not from Creative Commons and want to determine whether it is free/open,
+you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
 Note, however, that some of the licenses there might be inappropriate for text and media.
-For example, licenses for software or for documentation commonly require to distribute a work together with the full license text, which is impractical for usage in print media.
+For example, licenses for software or for documentation commonly require
+to distribute a work together with the full license text,
+which is impractical for usage in print media.
+
+[commons-tags]: https://commons.wikimedia.org/wiki/Commons:Copyright_tags
 
@@
 
-Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are.
-Additionally, in some jurisdictions, collections of data (i.e., databases) can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database.
+Purely factual data such as measurements are usually not copyrightable,
+but literary or artistic works that cross the threshold of originality are.
+Additionally, in some jurisdictions, collections of data (i.e., databases)
+can be subject to _sui generis_ database rights
+which prevent extracting substantial parts of a database.
 
 As a consequence, a database may be licensed differently than its content.
-For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists, but the license for the database as a whole might have been chosen by the curator, including the selection of entries, field names and any _sui generis_ database rights.
+For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists,
+but the license for the database as a whole might have been chosen by the curator,
+including the selection of entries, field names and any _sui generis_ database rights.
 Watch out for any signs that the license for the database and its content are not the same.
@@
 
-  [Link to full text](https://creativecommons.org/licenses/by-sa/4.0/legalcode)
+  [Link to full text][cc-by-sa-text]
   </details>
@@
 
-Note that Creative Commons licenses prior to version 4.0 are not suitable for databases. You might also encounter the following, less popular free/open data licenses:
+Note that Creative Commons licenses prior to version 4.0 are not suitable for databases.
+You might also encounter the following, less popular free/open data licenses:
 
@@
 
-Finally, you should know about the following two free/open licenses because they only cover the database, but not its individual contents or the works produced from it.
+Finally, you should know about the following two free/open licenses because they only cover the database,
+but not its individual contents or the works produced from it.
 If any of these licenses is used, the license for the content must be stated separately.
@@
 
-If you encounter a license not listed here and want to determine whether it is free/open, you can consult the [lists of licenses accepted at Wikimedia Commons](https://commons.wikimedia.org/wiki/Commons:Copyright_tags).
-Note, however, that some of the licenses there might be inappropriate for data, because they do not mention _sui generis_ database rights.
+If you encounter a license not listed here and want to determine whether it is free/open,
+you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
+Note, however, that some of the licenses there might be inappropriate for data,
+because they do not mention _sui generis_ database rights.
 :::
@@
 You may have noticed that in many cases a version number is part of the license name.
-This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses.
-For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002](https://creativecommons.org/weblog/entry/3476).
-Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights have been covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013).
-There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically (i.e., you have to follow the terms of that exact version).
-CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
+This is because the organizations that created the licenses sometimes publish updated versions
+to accommodate for developments in copyright law and the communities that use the licenses.
+For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002][cc-v1].
+Since then, the possibility to relicense under later ([v2.0][cc-v2], 2004)
+and compatible licenses ([v3.0][cc-v3], 2007) has been added,
+a 30-day window to correct license violations has been established to combat [copyleft trolls][commons-trolls],
+and _sui generis_ database rights have been covered explicitly ([v4.0][cc-v4], 2013).
+There are many more [subtle differences between license versions][cc-versions],
+therefore it is important to indicate which license version exactly one is referring to,
+as the license of a work does not "update" automatically
+(i.e., you have to follow the terms of that exact version).
+CC0, officially introduced in 2009, is still at version [1.0][cc0-v1].
 
-For the GNU licenses it is even recommended to state whether a work is licensed _only_ under the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
+[cc-v1]: https://creativecommons.org/weblog/entry/3476
+[cc-v2]: https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders
+[cc-v3]: https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included
+[cc-v4]: https://creativecommons.org/weblog/entry/40768
+[cc0-v1]: http://www.creativecommons.org/weblog/entry/13304
+[cc-versions]: https://wiki.creativecommons.org/wiki/License_Versions
+[commons-trolls]: https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling
+
+For the GNU licenses it is even recommended to state whether a work is licensed
+_only_ under the indicated version of the license or, alternatively,
+also under newer versions of the license [@Stallman2022Version].
 :::
@@
 
-Finally, if you verified that the work has been made available to you under a free/open license, you need to understand the conditions under which you are allowed to include the work in your project folder and share it with others -- regardless of whether you modified it.
+Finally, if you verified that the work has been made available to you under a free/open license,
+you need to understand the conditions under which you are allowed to include the work in your project folder
+and share it with others -- regardless of whether you modified it.
 In the following, we will explain the most important conditions of the licenses listed above.
 For other licenses we recommend reading their text in full and/or consulting with a lawyer.
-The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com](https://choosealicense.com/appendix/), which you can use to get a general idea of a license (but which does not obviate reading the license text in full).
-Specifically for software licenses the _Open Source Automation Development Lab_ provides [useful checklists detailing the respective obligations](https://www.osadl.org/Checklists/).^[Accessing the checklists requires a (free) account on their website.]
+The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com][cal-appendix],
+which you can use to get a general idea of a license
+(but which does not obviate reading the license text in full).
+Specifically for software licenses the _Open Source Automation Development Lab_
+provides [useful checklists detailing the respective obligations][osadl].[^osadl-note]
+
+[cal-appendix]: https://choosealicense.com/appendix/
+[osadl]: https://www.osadl.org/Checklists/
+[^osadl-note]: Accessing the checklists requires a (free) account on their website.
 
@@
 
-In the following, we only discuss the conditions under which works may be _shared_. This is because free/open licenses put no conditions on private use, where the work is not made available to others.
+In the following, we only discuss the conditions under which works may be _shared_.
+This is because free/open licenses put no conditions on private use,
+where the work is not made available to others.
 :::
@@
 
-Licenses may require that you provide attribution to the author(s)^[or others whose names are provided with the original work] and to retain any copyright notices,^[Such as `Copyright (c) 2025 Jane Doe`] although the specifics vary between licenses.
-For example, Creative Commons licenses other than CC0 require that the name of the author, the title of the work, and a link to the work are provided.
+Licenses may require that you provide attribution to the author(s)[^others]
+and to retain any copyright notices,[^example-notice]
+although the specifics vary between licenses.
+For example, Creative Commons licenses other than CC0 require
+that the name of the author, the title of the work, and a link to the work are provided.
 
-Projects with many authors often explicitly allow for an attribution to the collective as a whole rather than being mentioned individually.
-For example, the flexible attribution requirements of Creative Commons licenses allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009].
-And software projects may use copyright notices that only contain the project name, such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].
+[^others]: or others whose names are provided with the original work
+[^example-notice]: Such as `Copyright (c) 2025 Jane Doe`
 
-Where to put the attribution depends on the type of the work, but in general it is best to keep it close to where the work is used and to provide all required information.
-Only linking to a page that provides the attribution (as with Wikipedia articles) most likely will not suffice in general.
+Projects with many authors often explicitly allow for an attribution
+to the collective as a whole rather than being mentioned individually.
+For example, the flexible attribution requirements of Creative Commons licenses
+allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009].
+And software projects may use copyright notices that only contain the project name,
+such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].
+
+Where to put the attribution depends on the type of the work,
+but in general it is best to keep it close to where the work is used
+and to provide all required information.
+Only linking to a page that provides the attribution (as with Wikipedia articles)
+most likely will not suffice in general.
 
@@
 If a work has been modified, licenses may require that this is indicated.
-For example, if an image under a Creative Commons license other than CC0 is reused in a cropped version, reusers could add "adapted from original" to indicate that changes have occurred.
+For example, if an image under a Creative Commons license other than CC0
+is reused in a cropped version, reusers could add "adapted from original"
+to indicate that changes have occurred.
 
@@
 
-Licenses may either require that one mentions the applicable license by name and possibly link ("Indicate license") or they even require to copy the full license text along with the used work ("Add license text").
-Of course, the latter may not be feasible if the work is printed and smaller than the license text.
-For example, if the license of an image requires to print the full license text, the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].
+Licenses may either require that one mentions the applicable license
+by name and possibly link ("Indicate license")
+or they even require to copy the full license text along with the used work ("Add license text").
+Of course, the latter may not be feasible
+if the work is printed and smaller than the license text.
+For example, if the license of an image requires to print the full license text,
+the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].
 
@@
 
-There are various situations in which one interacts with software without having access to the underlying source code.
-For example, software can be distributed as a binary ("executable") file or it can be used over the internet.
+There are various situations in which one interacts with software
+without having access to the underlying source code.
+For example, software can be distributed as a binary ("executable") file
+or it can be used over the internet.
 In both cases, licenses may require that the source code is provided to its users.
@@
 Such a protection is also called **d**igital **r**ights **m**anagement (DRM).
-For works published under a license with an anti-DRM clause, reusers may not apply DRM to the work or must provide a DRM-free version in parallel (depending on the exact license).
+For works published under a license with an anti-DRM clause,
+reusers may not apply DRM to the work or must provide a DRM-free version in parallel
+(depending on the exact license).
 
@@
 
-If you modified the work and put a sufficient amount of creativity into it, the result is called a _derivative work_.
-It is (again) protected by copyright and even if you share it, nobody except you is allowed to copy, modify, or distribute it further.
-You can put it under a free/open license but don't have to.^[Of course, you still need to abide by the license conditions of the original work, such as attribution.] The only exception is if the original work has been made available under a "copyleft" license.
-In that case, if you share the work with others you must put it under the same license as the original work^[or sometimes also under a later version or under a compatible license] -- or you commit a copyright violation.^[Note, however, that [copyleft licenses are not automatic](https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic): A derivative work of a copylefted work is not "implicitly" under the same license -- others would commit a copyright violation if they just took it.
-Only if the derivative author explicitly makes it available under the same license, others may use the derivative work under the same terms.] This explains the origin of the term _copyleft_, as it "turns copyright around" to make works permanently free: ©\ →\ 🄯.
+If you modified the work and put a sufficient amount of creativity into it,
+the result is called a _derivative work_.
+It is (again) protected by copyright and even if you share it,
+nobody except you is allowed to copy, modify, or distribute it further.
+You can put it under a free/open license but don't have to.[^still-abide]
+The only exception is if the original work has been made available under a "copyleft" license.
+In that case, if you share the work with others
+you must put it under the same license as the original work[^or-compatible]
+-- or you commit a copyright violation.[^not-automatic]
+This explains the origin of the term _copyleft_,
+as it "turns copyright around" to make works permanently free: ©\ →\ 🄯.
 In the realm of Creative Commons licenses, copyleft is also called _share alike_.
@@
 
-What is considered creation of a derivative work varies between jurisdictions, but the following are typical examples: translating a text into a different language, taking a picture of an artwork, adding a song to a video, adapting a computer program to fit own use cases, or taking a [screenshot of a computer program](https://commons.wikimedia.org/wiki/Commons:Screenshots).
-In all of these cases, if the original work is under a copyleft license, then sharing of the resulting derivative work is only allowed if it is put under the same license.
+[^still-abide]: Of course, you still need to abide
+by the license conditions of the original work, such as attribution.
+
+[^or-compatible]: or sometimes also under a later version or under a compatible license
+
+[^not-automatic]: Note, however, that [copyleft licenses are not automatic][commons-automatic]:
+A derivative work of a copylefted work is not "implicitly" under the same license
+-- others would commit a copyright violation if they just took it.
+Only if the derivative author explicitly makes it available under the same license,
+others may use the derivative work under the same terms.
+
+[commons-automatic]: https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic
+
+What is considered creation of a derivative work varies between jurisdictions,
+but the following are typical examples:
+translating a text into a different language,
+taking a picture of an artwork,
+adding a song to a video, adapting a computer program to fit own use cases,
+or taking a [screenshot of a computer program][commons-screenshots].
+In all of these cases, if the original work is under a copyleft license,
+then sharing of the resulting derivative work is only allowed
+if it is put under the same license.
+
+[commons-screenshots]: https://commons.wikimedia.org/wiki/Commons:Screenshots
 
 The question is then: Which uses of a work do not create a derivative work?
-In many cases, if an unaltered excerpt "is used to illuminate an idea or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made.
-The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p.
-44] and not be "merged with other material into a new and larger work" [@Kreutzer2024, p.
-45].
-For example, embedding images in a document and clearly separating them from the rest or assembling works as part of a collection typically would not create a derivative work.
-However, due to a lack of case law, there is only little guidance on what constitutes a derivative work in general.
+In many cases, if an unaltered excerpt "is used to illuminate an idea
+or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made.
+The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p. 44]
+and not be "merged with other material into a new and larger work" [@Kreutzer2024, p. 45].
+For example, embedding images in a document and clearly separating them from the rest
+or assembling works as part of a collection typically would not create a derivative work.
+However, due to a lack of case law, there is only little guidance
+on what constitutes a derivative work in general.
 
@@
 
-- __Weak copyleft licenses__ only require that modifications to the software itself are licensed under the same or a compatible license if shared.
-For example, if you create and publish software under a weak copyleft license, others who modify it and put their version on the internet have to apply the same license.
-However, people who merely _use_ your software in their own work which they make publicly available can choose any license for it.
-- __Strong copyleft licenses,__ on the other side, insist that any larger works that use the copyleft-licensed software must also be licensed under the same or a compatible license if shared.
-For example, if your software were to be put under a strong copyleft license, everybody publishing software that uses your software would need to put it under the same license.
+- __Weak copyleft licenses__ only require that modifications to the software itself
+  are licensed under the same or a compatible license if shared.
+  For example, if you create and publish software under a weak copyleft license,
+  others who modify it and put their version on the internet have to apply the same license.
+  However, people who merely _use_ your software in their own work which they make publicly available
+  can choose any license for it.
+- __Strong copyleft licenses,__ on the other side, insist that any larger works
+  that use the copyleft-licensed software must also be licensed
+  under the same or a compatible license if shared.
+  For example, if your software were to be put under a strong copyleft license,
+  everybody publishing software that uses your software would need to put it under the same license.
   Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024GPL].
@@
 
-The description of the conditions above is a simplified summary -- read the full text of the license before sharing the respective work with others.
+The description of the conditions above is a simplified summary
+-- read the full text of the license before sharing the respective work with others.
 For example, Creative Commons licenses other than CC0 also require that
@@
 
-For most Creative Commons licenses, providing the following information is required (if available):
+For most Creative Commons licenses,
+providing the following information is required (if available):
 
@@
 
-Obviously, if you are in doubt whether a condition of the license applies in your use case, it is better to either follow it or to ask the author for clarification or explicit permission.
-This is especially important given the existence of [_copyleft trolls_](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling): People who distribute works under a free/open license but send out statutory damages claims after minor violations without providing reusers an opportunity to fix their error.
-In addition, in many European countries attribution may be necessary regardless due to the author's moral rights.
-For these reasons, we recommend that you provide attribution for all works under a free/open license -- even if the license does not strictly require it (e.g., CC0).
-For this reason, you need to make note of every foreign work included in your project folder and record its license.
+Obviously, if you are in doubt whether a condition of the license applies in your use case,
+it is better to either follow it or to ask the author for clarification or explicit permission.
+This is especially important given the existence of [_copyleft trolls_][commons-trolls]:
+People who distribute works under a free/open license
+but send out statutory damages claims after minor violations
+without providing reusers an opportunity to fix their error.
+In addition, in many European countries attribution may be necessary regardless
+due to the author's moral rights.
+For these reasons, we recommend that you provide attribution
+for all works under a free/open license
+-- even if the license does not strictly require it (e.g., CC0).
+For this reason, you need to make note of every foreign work
+included in your project folder and record its license.
 
@@
 
-The free/open licenses we discuss here mostly grant copyrights (with some of them explicitly granting patent rights and _sui generis_ database rights).
+The free/open licenses we discuss here mostly grant copyrights
+(with some of them explicitly granting patent rights and _sui generis_ database rights).
 Therefore, you may lack other rights such as personality, privacy, moral, or trademark rights.
-For example, [sharing photos that depict people](https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people) is not only a matter of copyright, but also of privacy rights.
+For example, [sharing photos that depict people][commons-photographs]
+is not only a matter of copyright, but also of privacy rights.
 Also, their commercial use may require the consent of the depicted person.
-Finally, note that even if attribution is not a requirement of the license, good scientific practice demands that appropriate citations are made.
+Finally, note that even if attribution is not a requirement of the license,
+good scientific practice demands that appropriate citations are made.
 :::
+
+[commons-photographs]: https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people
 
@@
 
-Let's practice what you learned by adding an image to the manuscript. We'll use [this picture of a penguin from Flickr](https://flic.kr/p/2pEKnUr). First, try to answer the following questions:
+Let's practice what you learned by adding an image to the manuscript.
+We'll use [this picture of a penguin from Flickr](https://flic.kr/p/2pEKnUr).
+First, try to answer the following questions:
 
@@
 
-1. The picture has been put under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/)
+1. The picture has been put under the license [CC0\ 1.0][cc0]
 2. The license is free/open
@@
 
+[cc0]: https://creativecommons.org/publicdomain/zero/1.0/
+
 You can now download the image and move it to your project folder (maybe call it `penguin.jpg`).
-In your manuscript file, find the line that says "add image here", remove it, and add the image instead, including proper attribution.
+In your manuscript file, find the line that says "add image here", remove it,
+and add the image instead, including proper attribution.
 
@@
 
-Although the license CC0 doesn't require it, we still recommend providing proper attribution.
+Although the license CC0 doesn't require it,
+we still recommend providing proper attribution.
 
@@
 
-If it is likely that the article gets printed, you might want to ensure that the link to the work is retained in that case:
+If it is likely that the article gets printed,
+you might want to ensure that the link to the work is retained in that case:
 
@@
 
-Let's render the document and see if the changes were successful (open `Manuscript.pdf` afterwards):
+Let's render the document and see if the changes were successful
+(open `Manuscript.pdf` afterwards):
 
@@
 
-In addition, we recommend creating a file called `LICENSE.txt` to describe the license of the image:
+In addition, we recommend creating a file called `LICENSE.txt`
+to describe the license of the image:
 
@@
 
-If the license required adding the full license text, you would also need to copy it to the project folder.
+If the license required adding the full license text,
+you would also need to copy it to the project folder.
 
@@
 
-If you would like to learn more about copyright and licenses you might find the following resources interesting:
+If you would like to learn more about copyright and licenses
+you might find the following resources interesting:
 
-- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2024
-- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
+- "Open Content -- A Practical Guide to Using Creative Commons Licences"
+  by @Kreutzer2024
+- "Creative Commons Certificate for Educators, Academic Librarians,
+  and Open Culture" by @CC2024Certificate
 - "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004

### 2025-02-08 data.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 
-You can now download the data set we have prepared for you and put it into your project folder:
+You can now download the data set we have prepared for you
+and put it into your project folder:
 
@@
 
-The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica.
-It was made available by Allison Horst, Alison Hill, and Kristen Gorman under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
+The data set is from the package [`palmerpenguins`][palmerpenguins]
+(v`r packageVersion("palmerpenguins")`)
+and contains the recorded bill lengths and sex of penguins
+living on three islands in the Palmer Archipelago, Antarctica.
+It was made available by Allison Horst, Alison Hill, and Kristen Gorman
+under the license [CC0\ 1.0][cc0].
+
+[palmerpenguins]: https://allisonhorst.github.io/palmerpenguins/
+[cc0]: https://creativecommons.org/publicdomain/zero/1.0/
 
@@
 Of course, you should only share them to the extent you are allowed to.
-Besides copyright and similar rights, you need to take into account applicable privacy laws (e.g., the GDPR for European citizens) and contractual obligations (e.g., with your data provider).
+Besides copyright and similar rights, you need to take into account
+applicable privacy laws (e.g., the GDPR for European citizens)
+and contractual obligations (e.g., with your data provider).
 
-Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from.
-For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).
-Another resource to look at is the presentation "Creating reproducible packages when data are confidential" by @Vilhuber2024.
+Privacy laws and contractual obligations may require you
+to create a completely anonymized or synthetic data set[^anon-synth-tools] (if possible),
+or prohibit any sharing of data, in which case you should provide a reference
+to a data repository where they can be obtained from.
+For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)"
+by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023
+and have a look at [the accompanying slides](https://osf.io/z6gcu).
+Another resource to look at is the presentation
+"Creating reproducible packages when data are confidential" by @Vilhuber2024.
 :::
+
+[^anon-synth-tools]: For example, using [Amnesia](https://amnesia.openaire.eu/),
+[ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools),
+or [Synthpop](https://www.synthpop.org.uk/).
 
@@
 
-When distributing a data set, it is important to document the meaning (e.g., units) and valid values of its variables.
+When distributing a data set, it is important to document the meaning (e.g., units)
+and valid values of its variables.
 This is typically done with a _data dictionary_ (also called a _codebook_).
-In the following, we will demonstrate how to create a simple data dictionary using the R package [`pointblank`](https://rstudio.github.io/pointblank/).
+In the following, we will demonstrate how to create a simple data dictionary
+using the R package [`pointblank`][pointblank].
 You can install it now using:
+
+[pointblank]: https://rstudio.github.io/pointblank/
 
@@
 
-You can put the code that follows for creating the data dictionary into a new file called `create_data_dictionary.R`.
+You can put the code that follows for creating the data dictionary
+into a new file called `create_data_dictionary.R`.
 
-First, we write down everything we know about the data set. This includes:
+First, we write down everything we know about the data set.
+This includes:
 
@@
 
-Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, applied preprocessing steps, or contact information.
-In our case, as the data have already been published, we only store a reference to its source.
+Depending on the type of data, it may also be necessary
+to describe measurement instruments, sampling procedures, appropriate weighting,
+applied preprocessing steps, or contact information.
+In our case, as the data have already been published,
+we only store a reference to its source.
 
-As the data set is from the R package `palmerpenguins`, if you have it installed you can use the function `citation()` to create such a reference:
+As the data set is from the R package `palmerpenguins`,
+if you have it installed you can use the function `citation()`
+to create such a reference:
 
@@
 
-Without the package `palmerpenguins` installed, you can find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation) and run the following:
+Without the package `palmerpenguins` installed,
+you can find a [suggested citation on its website][palmerpenguins-citation]
+and run the following:
+
+[palmerpenguins-citation]: https://allisonhorst.github.io/palmerpenguins/#citation
 
@@
 
-Finally, we can store the data dictionary inside an HTML file and put the HTML file into the project folder as well.
+Finally, we can store the data dictionary inside an HTML file
+and put the HTML file into the project folder as well.
 
@@
 
-For a more elaborate introduction to `pointblank`, you can read its "[Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html)".
+For a more elaborate introduction to `pointblank`,
+you can read its "[Intro to Information Management][pointblank-intro]".
+
+[pointblank-intro]: https://rstudio.github.io/pointblank/articles/INFO-1.html
 
@@
 We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".
+If you want to learn more about the sharing of research data,
+have a look at the tutorial "[FAIR research data management][fair-tutorial]".
+
+[fair-tutorial]: https://lmu-osc.github.io/FAIR-Data-Management/
 
@@
 
-This example demonstrates how the title and description of the data set, the description of the variables and their valid values are stored in a machine-readable way.
+This example demonstrates how the title and description of the data set,
+the description of the variables and their valid values are stored in a machine-readable way.
 As before, we also provide a reference to the source.
@@
 
-Generally, metadata are either stored embedded into the data or externally, for example, in a separate file.
-We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately.
+Generally, metadata are either stored embedded into the data or externally,
+for example, in a separate file.
+We will use the "[frictionless data](https://frictionlessdata.io/)" standard,
+where metadata are stored separately.
 Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).
 
-Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released.
-As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information.
-We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under CC0\ 1.0.
+Specifically, one can use the R package [`frictionless`][frictionless]
+to create a _schema_ which describes the structure of the data.[^frictionless-note]
+For the purpose of the following code,
+it is just a nested list that we edit to include our own information.
+We also explicitly record in the schema
+that missing values are stored in the data file as `NA`
+and that the data are licensed under CC0\ 1.0.
 Finally, the package is used to create a metadata file that contains the schema.
+
+[frictionless]: https://docs.ropensci.org/frictionless/
+
+[^frictionless-note]: In June 2024, [version 2](https://datapackage.org/)
+of the frictionless data standard has been released.
+As of November 2024, the R package `frictionless` only supports the first version,
+though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).
 
@@
 This creates the metadata file `datapackage.json` in the current directory.
-Make sure it is located in the same folder as `data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
+Make sure it is located in the same folder as `data.csv`,
+as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
 :::
 
-Having added the data and its documentation, one can view and record the utilized packages with `renv`:
+Having added the data and its documentation,
+one can view and record the utilized packages with `renv`:
 
@@
 
-All data relied upon should be cited in the manuscript to allow for precise identification and access.
-From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):
+All data relied upon should be cited in the manuscript
+to allow for precise identification and access.
+From the "eight core principles of data citation" by @Starr2015,
+licensed under [CC0\ 1.0][cc0]:
 
-> **Principle 1 – Importance**: "Data should be considered legitimate, citable products of research.
-Data citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications."
+> **Principle 1 – Importance**: "Data should be considered legitimate,
+> citable products of research. Data citations should be accorded the same
+> importance in the scholarly record as citations of other research objects,
+> such as publications."
 > 
-> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited."
+> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a
+> claim relies upon data, the corresponding data should be cited."
 > 
-> **Principle 5 – Access**: "Data citations should facilitate access to the data themselves and to such associated metadata, documentation, code, and other materials, as are necessary for both humans and machines to make informed use of the referenced data."
+> **Principle 5 – Access**: "Data citations should facilitate access to the data
+> themselves and to such associated metadata, documentation, code, and other
+> materials, as are necessary for both humans and machines to make informed use
+> of the referenced data."
 > 
-> **Principle 7 – Specificity and Verifiability**: "Data citations should facilitate identification of, access to, and verification of the specific data that support a claim.
-Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific time slice, version and/or granular portion of data retrieved subsequently is the same as was originally cited."
+> **Principle 7 – Specificity and Verifiability**: "Data citations should
+> facilitate identification of, access to, and verification of the specific data
+> that support a claim. Citations or citation metadata should include
+> information about provenance and fixity sufficient to facilitate verifying
+> that the specific time slice, version and/or granular portion of data
+> retrieved subsequently is the same as was originally cited."
 
-Now, it's your turn to add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?
+Now, it's your turn to add an appropriate citation for the data set to the manuscript.
+Does your citation adhere to the principles above?
 
@@
 
-You can find an appropriate BibTeX entry on [the package website](https://allisonhorst.github.io/palmerpenguins/#citation) or with the function `citation()`:^[Note that this function requires to have the respective package installed.]
+You can find an appropriate BibTeX entry on [the package website][palmerpenguins-citation]
+or with the function `citation()`:^[Note that this function requires
+to have the respective package installed.]
 
@@
 
-Copy the BibTeX entry to the file `Bibliography.bib`. Then, find the line in the manuscript that says "cite data here" and replace it with a sentence such as the following:
+Copy the BibTeX entry to the file `Bibliography.bib`.
+Then, find the line in the manuscript that says "cite data here"
+and replace it with a sentence such as the following:
 
@@
 
-While citation happens in the manuscript for reasons of academic integrity and reproducibility, to comply with any licenses you also may need to provide attribution within your project folder.
-Even though the data file we use here does not require attribution, we recommend adding a short paragraph to `LICENSE.txt`:
+While citation happens in the manuscript for reasons of academic integrity and reproducibility,
+to comply with any licenses you also may need to provide attribution within your project folder.
+Even though the data file we use here does not require attribution,
+we recommend adding a short paragraph to `LICENSE.txt`:
 
@@
 
-As before, if the license required adding the full license text, you would also need to copy it to the project folder (if not already in there).
+As before, if the license required adding the full license text,
+you would also need to copy it to the project folder (if not already in there).
 

### 2025-02-08 intro.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

--- 2025-01-25 intro.qmd
@@
 
-In the following, we will set the stage by highlighting the importance of sharing all materials, documenting their usage, and linking results with their computations, leading to the creation _research compendia_.
+In the following, we will set the stage by
+highlighting the importance of sharing all materials,
+documenting their usage,
+and linking results with their computations,
+leading to the creation _research compendia_.
 
@@
 
-Suppose you are reading an article about a new imaging method to turn seismological data into subsurface images.
-The article describes the ideas that went into developing this method and presents a few examples to illustrate its superiority over previous approaches.
+Suppose you are reading an article about a new imaging method
+to turn seismological data into subsurface images.
+The article describes the ideas that went into developing this method
+and presents a few examples to illustrate its superiority over previous approaches.
 You got interested and would like to apply this method to your own data.
-However, with only the article available, it could take months to come up with a working solution, if possible at all.
-This situation has been put aptly by Buckheit & Donoho [-@Buckheit1995, p.
-59], distilling an idea by the geophysicist Jon Claerbout:
+However, with only the article available, it could take months
+to come up with a working solution, if possible at all.
+This situation has been put aptly by
+Buckheit & Donoho [-@Buckheit1995, p. 59, emphasis in original],
+distilling an idea by the geophysicist Jon Claerbout:
 
-> "An article about computational science in a scientific publication is __not__ the scholarship itself, it is merely __advertising__ of the scholarship.
-The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."
+> "An article about computational science in a scientific publication
+> is __not__ the scholarship itself, it is merely __advertising__ of the scholarship.
+The actual scholarship is the complete software development environment
+and the complete set of instructions which generated the figures."
 
-Even when researchers merely apply existing methods (rather than report on a new method), sharing the source code and being transparent about the computational environment is imperative to making their results reproducible [@Ince2012].
-By reproducibility, we mean "obtaining consistent results using the same input data; computational steps, methods, and code; and conditions of analysis" [@NASEM2019, p.
-46].
+Even when researchers merely apply existing methods (rather than report on a new method),
+sharing the source code and being transparent about the computational environment
+is imperative to making their results reproducible [@Ince2012].
+By reproducibility, we mean "obtaining consistent results using the same
+input data; computational steps, methods, and code;
+and conditions of analysis" [@NASEM2019, p. 46].
 
@@
 However, there's more to code publishing than sharing.
-Researchers should document (i.e., track) which data, software, images, texts, and other materials contributed to their work.
-In more elaborate terms, researchers should preserve the _provenance_ of shared materials, for at least three tangible reasons:
+Researchers should document (i.e., track) which data, software, images, texts,
+and other materials contributed to their work.
+In more elaborate terms, researchers should preserve the _provenance_ of shared materials,
+for at least three tangible reasons:
 
 - **Academic integrity**: Providing citations for works that are not one's own avoids plagiarism.
-- **Complying with the law**: Providing attribution and license texts (among other things) may be a legal obligation when redistributing materials.
-- **Facilitating reproductions**: Stating utilized software and data with their exact versions helps reproducers.
+- **Complying with the law**: Providing attribution and license texts
+  (among other things) may be a legal obligation when redistributing materials.
+- **Facilitating reproductions**: Stating utilized software and data
+  with their exact versions helps reproducers.
 
@@
 Rephrasing someone's text avoids issues of copyright, but may be plagiarism.
-And bibliographies of various styles may comply with scientific citation standards, but if they omit the first names of authors or the version number of computer programs, they might not help with matters of copyright or reproducibility.
+And bibliographies of various styles may comply with scientific citation standards,
+but if they omit the first names of authors or the version number of computer programs,
+they might not help with matters of copyright or reproducibility.
 
 In this tutorial, we'll consider all of these reasons important.
-And there are even more: Documenting the provenance puts research into context and allows others to understand how it came about.
+And there are even more: Documenting the provenance puts research into context
+and allows others to understand how it came about.
 Also, cited authors benefit from the citation as they gain potential readers.
@@
 
-For the purpose of this tutorial, tracking the provenance of results deserves particular attention.
+For the purpose of this tutorial,
+tracking the provenance of results deserves particular attention.
 This means connecting them to the source code through the creation of dynamic documents.
-Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article.
-Dynamic documents bundled together with any necessary data and auxiliary software are called [research compendia](https://research-compendium.science/) [@Gentleman2007].
+Rather than manually copying numerical results, figures, or tables,
+they are inserted automatically upon rendering of the article.
+Dynamic documents bundled together with any necessary data and auxiliary software
+are called [research compendia](https://research-compendium.science/) [@Gentleman2007].
 
-The practice to interleave narrative text with code has its roots in the paradigm of literate programming, where documentation and source code are treated as equals and are arranged in a way to maximize understanding [@Knuth1984].
-Alternating text and code can be also found in notebook interfaces for exploratory programming [e.g., Wolfram Mathematica or Jupyter Notebooks, see @Kluyver2016] that also have the capability to execute code and embed its output.
-With _Sweave_ [@Leisch2002], ideas from both worlds -- literate programming and embedding program output -- were combined into one tool for rendering dynamic documents using the R programming language.
-It is the predecessor of the R package `knitr` [@Xie2015] which is being used under the hood in this tutorial.^[Specifically, Quarto employs `knitr` to execute chunks of R code.]
+The practice to interleave narrative text with code has its roots
+in the paradigm of literate programming, where documentation and source code
+are treated as equals and are arranged in a way to maximize understanding [@Knuth1984].
+Alternating text and code can be also found in notebook interfaces for exploratory programming
+[e.g., Wolfram Mathematica or Jupyter Notebooks, see @Kluyver2016]
+that also have the capability to execute code and embed its output.
+With _Sweave_ [@Leisch2002], ideas from both world
+-- literate programming and embedding program output --
+were combined into one tool for rendering dynamic documents using the R programming language.
+It is the predecessor of the R package `knitr` [@Xie2015]
+which is being used under the hood in this tutorial.[^knitr]
+
+[^knitr]: Specifically, Quarto employs `knitr` to execute chunks of R code.
 
 Linking results with their computations has benefits for authors and readers.
-For the author, articles always contain the most recent version of figures, as they are updated automatically when the computation changes.
-For the readers, it enables understanding exactly how a particular result was obtained if they get access to the underlying research compendium.
+For the author, articles always contain the most recent version of figures,
+as they are updated automatically when the computation changes.
+For the readers, it enables understanding exactly how a particular result was obtained
+if they get access to the underlying research compendium.
 
@@
 
-To recap, in this tutorial you will share your materials, document their usage, and connect results to the underlying source code through the creation of a research compendium.
-Make sure to consider the following things along the way [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
+To recap, in this tutorial you will share your materials, document their usage,
+and connect results to the underlying source code through the creation of a research compendium.
+Make sure to consider the following things along the way
+[by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:
 
-> __Does the research compendium contain everything needed to reproduce a predefined outcome in an organized and parsimonious way?__
+> __Does the research compendium contain everything needed to reproduce a
+> predefined outcome in an organized and parsimonious way?__
 > 
-> 1. __Completeness:__ The research compendium contains all of the objects needed to reproduce a predefined outcome.
-> 2. __Organization:__ It is easy to understand and keep track of the various objects in the research compendium and their relationship over time.
-> 3. __Economy:__ Fewer extraneous objects in the compendium mean fewer things that can break and require less maintenance over time.
+> 1. __Completeness:__ The research compendium contains all of the objects
+>    needed to reproduce a predefined outcome.
+> 2. __Organization:__ It is easy to understand and keep track of the various
+>    objects in the research compendium and their relationship over time.
+> 3. __Economy:__ Fewer extraneous objects in the compendium mean fewer things
+>    that can break and require less maintenance over time.
 > 
-> __Is descriptive information about the research compendium and its components available and easy to understand?__
+> __Is descriptive information about the research compendium and its components
+> available and easy to understand?__
 > 
-> 4. __Transparency:__ The research compendium provides full disclosure of the research process that produced the scientific claim.
-> 5. __Documentation:__ Information describing compendium objects is provided in enough detail to enable independent understanding and use of the compendium.
+> 4. __Transparency:__ The research compendium provides full disclosure of
+>    the research process that produced the scientific claim.
+> 5. __Documentation:__ Information describing compendium objects is provided
+>    in enough detail to enable independent understanding and use of the compendium.
 > 
-> __Is information about how the research compendium and its components can be used available and easy to understand?__
+> __Is information about how the research compendium and its components
+> can be used available and easy to understand?__
 > 
-> 6. __Access:__ It is clear who can use what, how, and under what conditions, with open access preferred.
-> 7. __Provenance:__ The origin of the components of the research compendium and how each has changed over time is evident.
+> 6. __Access:__ It is clear who can use what, how, and under what conditions,
+>    with open access preferred.
+> 7. __Provenance:__ The origin of the components of the research compendium
+>    and how each has changed over time is evident.
 > 
@@
 > 
-> 8. __Metadata:__ Information about the research compendium and its components is embedded in a standardized, machine-readable code.
-> 9. __Automation:__ As much as possible, the computational workflow is script- or workflow-based so that the workflow can be re-executed using minimal actions.
+> 8. __Metadata:__ Information about the research compendium and its component
+>    is embedded in a standardized, machine-readable code.
+> 9. __Automation:__ As much as possible, the computational workflow is script-
+>    or workflow-based so that the workflow can be re-executed using minimal actions.
 > 
-> __Is there a plan for reviewing the research compendium for FAIR and computational reproducibility standards over time?__
+> __Is there a plan for reviewing the research compendium for FAIR
+> and computational reproducibility standards over time?__
 > 
-> 10. __Review:__ A series of managed activities are needed to ensure continued access to and functionality of the research compendium and its components for as long as necessary.
+> 10. __Review:__ A series of managed activities are needed to ensure
+>    continued access to and functionality of the research compendium and its components for as long as necessary.
 
-Although this tutorial guides you through the creation of a research compendium, you are invited to revisit these questions after completion and check whether and how each point was addressed (or not).
+Although this tutorial guides you through the creation of a research compendium,
+you are invited to revisit these questions after completion and check
+whether and how each point was addressed (or not).
 Further, you can consult them as a checklist for future projects.

### 2025-02-08 make_readme.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 Having settled on a license, it is time to add a final touch.
-Imagine returning to your project in five years, having forgotten most of the details of what you did exactly.
-What would be useful to know in order to quickly understand what is going on in the project?
+Imagine returning to your project in five years,
+having forgotten most of the details of what you did exactly.
+What would be useful to know in order to quickly understand
+what is going on in the project?
 This is what needs to be described in the README.
-While you could just start writing along, most READMEs have some sections in common which we describe below.
+While you could just start writing along,
+most READMEs have some sections in common which we describe below.
 
@@
 
-:   Badges typically report quick facts, like [whether the project is actively maintained](https://www.repostatus.org/), [how many dependencies it has](https://tinyverse.netlify.app/), or [whether it has been published in a package repository](https://www.r-pkg.org/services#badges).
-Sometimes, the license, any associated DOIs, or quality metrics like code coverage by unit tests are communicated via badges as well.
+:   Badges typically report quick facts, like
+    [whether the project is actively maintained][repostatus],
+    [how many dependencies it has][tinyverse],
+    or [whether it has been published in a package repository][metacran].
+    Sometimes, the license, any associated DOIs,
+    or quality metrics like code coverage by unit tests
+    are communicated via badges as well.
+
+[repostatus]: https://www.repostatus.org/
+[tinyverse]: https://tinyverse.netlify.app/
+[metacran]: https://www.r-pkg.org/services#badges
 
@@
 
-:   Is there anything you can show that demonstrates how the project can be used? Screenshots or other visuals can make the README more appealing.
+:   Is there anything you can show that demonstrates how the project can be used?
+    Screenshots or other visuals can make the README more appealing.
 
@@
 
-:   What steps need to be taken to run the project?
-What software needs to be installed?
+:   What steps need to be taken to run the project? What software needs to be installed?
     R itself and the R packages are already documented as this project uses `renv`.
-Therefore you can focus on all other dependencies, such as the system dependencies of R packages as well as the version of Quarto.^[As of August 2024, a proposal for `renv` to record the version of Quarto has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).] Also, don't forget to mention software that you have used for any manual steps.
+    Therefore you can focus on all other dependencies,
+    such as the system dependencies of R packages
+    as well as the version of Quarto.[^renv-quarto]
+    Also, don't forget to mention software that you have used for any manual steps.
     See @sec-dependencies for additional information.
-It may make sense to provide these information in the README even if you already cite all programs and their version numbers in your manuscript.
+    It may make sense to provide these information in the README
+    even if you already cite all programs and their version numbers in your manuscript.
+
+[^renv-quarto]: As of August 2024, a proposal for `renv` to record the version of Quarto
+has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).
 
@@
 
-:   Which files does the project folder contain?
-How are they organized?
-How can one run the project -- is there a master script or a particular order in which any scripts need to be executed?
+:   Which files does the project folder contain? How are they organized?
+    How can one run the project -- is there a master script
+    or a particular order in which any scripts need to be executed?
     How long does it take to run all scripts?
@@
 
-:   Do you offer support or help, for example, via GitHub discussions or a mailing list?
+:   Do you offer support or help, for example, via GitHub discussions
+    or a mailing list?
 
@@
 
-:   If the project is active: Can other people contribute? How? Do you accept contributions? Do you review issues? This section is sometimes outsourced into a file called `CONTRIBUTING.md`.
+:   If the project is active: Can other people contribute? How?
+    Do you accept contributions? Do you review issues?
+    This section is sometimes outsourced into a file called `CONTRIBUTING.md`.
 
@@
 
-:   Who was involved in creating this project? This involves you, your co-authors, and anybody you accepted contributions from.
+:   Who was involved in creating this project? This involves you,
+    your co-authors, and anybody you accepted contributions from.
 
@@
 
-:   Is there a recommended way to cite this project? Is there a published article associated with it that you would like to have cited?
+:   Is there a recommended way to cite this project?
+    Is there a published article associated with it
+    that you would like to have cited?
 
@@
 
-An overview over the system dependencies of R packages can be created using the function `pak::pkg_sysreqs()`.
-In combination with `renv`, we can obtain the system dependencies of all R packages the current project directly depends on:
+An overview over the system dependencies of R packages can be created
+using the function `pak::pkg_sysreqs()`.
+In combination with `renv`, we can obtain the system dependencies
+of all R packages the current project directly depends on:
 
@@
 
-We can see that the programs `make`, `pandoc`, and `git` were identified as system dependencies. Often, one can obtain their version by running them with the `--version` argument:
+We can see that the programs `make`, `pandoc`,
+and `git` were identified as system dependencies.
+Often, one can obtain their version by running them with the `--version` argument:
 
@@
 
-The other system dependencies are libraries, and the method to identify the installed version varies between operating systems. For example, on Fedora Linux, one can run the following:
+The other system dependencies are libraries,
+and the method to identify the installed version varies between operating systems.
+For example, on Fedora Linux, one can run the following:
 
@@
 
-However, they are only relevant if the R packages were installed from source, rather than as binary package.
+However, they are only relevant if the R packages were installed from source,
+rather than as binary package.
 
-We also know that we need Quarto to create the PDF, so let's find out its version as well:
+We also know that we need Quarto to create the PDF,
+so let's find out its version as well:
 
@@
 
-If you installed `apaquarto` or any other Quarto extension, one can query their versions as follows:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]
+If you installed `apaquarto` or any other Quarto extension,
+one can query their versions as follows:[^already-included]
+
+[^already-included]: Luckily, the extensions are included in the project folder,
+so technically their version is already recorded in the project's files.
 
@@
 
-Finally, we know that we installed a $\TeX$ distribution to create the PDF, so let's find out its version by running:
+Finally, we know that we installed a $\TeX$ distribution to create the PDF,
+so let's find out its version by running:
 
@@
 
-The output is quite long and it might look slightly different for you, but the relevant sections are the following:
+The output is quite long and it might look slightly different for you,
+but the relevant sections are the following:
 
@@
 
-Of course, all the system dependencies identified until now may have dependencies on their own. Use your own judgement to decide when not to dig deeper.
+Of course, all the system dependencies identified until now
+may have dependencies on their own. Use your own judgement to decide when not to dig deeper.
 
@@
 
-Create your README now as the file `README.md`. If you feel stuck, you can have a look at the following examples:
+Create your README now as the file `README.md`. If you feel stuck,
+you can have a look at the following examples:
 
@@
 - `data.csv`: a data set containing the simplified `palmerpenguins` data
-- `data_dictionary.html`: a dictionary to the data file, created using `create_data_dictionary.R`
+- `data_dictionary.html`: a dictionary to the data file,
+  created using `create_data_dictionary.R`
 

### 2025-02-08 setup.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

@@
 
-If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
+If you haven't already, open RStudio
+-- see @nte-terminal for how to use the terminal instead.
+Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.
 
@@
 
-Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.
+Finally, enter the name of the directory where our report will be created in,
+for example `code-publishing-exercise`.
 
-As we will use Git to track the version history of files, be sure to check _Create a git repository_.
-If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".
+As we will use Git to track the version history of files,
+be sure to check _Create a git repository_.
+If you don't know what Git is, have a look at the tutorial
+"[Introduction to version control with git and GitHub within RStudio][git-tutorial]".
+
+[git-tutorial]: https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/
 
@@
 
-Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on.
-Using it makes it easier for others to view and obtain them at the exact same version at a later point in time.
+Also, we will utilize the package [`renv`][renv]
+to track the R packages our project depends on.
+Using it makes it easier for others to view and obtain them
+at the exact same version at a later point in time.
 Therefore make sure that the box _Use renv with this project_ is checked.
-Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".
+Again, if this is the first time you are hearing about `renv`,
+have a look at the tutorial "[Introduction to {renv}][renv-tutorial]".
 
-If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.
+[renv]: https://rstudio.github.io/renv/
+[renv-tutorial]: https://lmu-osc.github.io/introduction-to-renv/
+
+If you are already familiar with Markdown and Quarto,
+you can uncheck the box _Use visual markdown editor_.
 
@@
 
-Click on _Create Project_. Your RStudio window should now look similar to this:
+Click on _Create Project_.
+Your RStudio window should now look similar to this:
 
@@
 
-If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.
+If, like in the image, a Quarto file with some demo content was opened automatically,
+you can close and delete it, for example, using RStudio's file manager.
 
-Throughout this tutorial, you will need to run both R code and system commands (primarily `git` and `quarto`).
-Within RStudio, R code can be run by going to the tab _Console_, while system commands are executed in the tab _Terminal_.
+Throughout this tutorial, you will need to run both R code and system commands
+(primarily `git` and `quarto`).
+Within RStudio, R code can be run by going to the tab _Console_,
+while system commands are executed in the tab _Terminal_.
 We also indicate where to run your code directly above each code snippet.
-If no indication is given, the code is only for demonstration purposes and does not need to be run.
+If no indication is given,
+the code is only for demonstration purposes and does not need to be run.
 
@@
 ::: {#tip-renv-status .callout-tip}
-Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
+Always run `renv::status()` and resolve any inconsistencies
+before you commit code to your project.
+This way, every commit represents a working state of your project.
 :::
@@
 
-Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:
+Without RStudio, one can create a Quarto project
+with version control and `renv` enabled by typing the following into a terminal:
 
@@
 
-Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Now, initialize `renv`:
+Then, one can open an R session by simply typing `R` into the terminal.
+Next, make sure that `getwd()` indicates that
+the working directory is `code-publishing-exercise`.
+Now, initialize `renv`:
 
@@
 You can either stage files separately or the whole project folder at once.
-If you do the latter, we recommend you to inspect the untracked changes before staging all of them:
+If you do the latter,
+we recommend you to inspect the untracked changes before staging all of them:
 
 ::: {.column-margin}
-In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
+In file paths, a period (`.`) means "the current directory",
+while two periods (`..`) mean "the parent directory".
+Therefore `git add .` means "stage the current directory for committing".
 :::
@@
 
-Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file.
-If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:
+Since no commits have been made so far,
+this should include every file that is not covered by the `.gitignore` file.
+If everything can be staged for committing
+-- as is the case in this tutorial -- you can follow up with:
 
@@
 
-If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file.
-If you don't check your changes before committing, you might accidentally commit something you'd rather not.
+If you see a file you'd rather not commit,
+delete it or add its name to the `.gitignore` file.
+If you don't check your changes before committing,
+you might accidentally commit something you'd rather not.
 
 ::: {#tip-author-identity-unknown .callout-tip}
-If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:
+If `git commit` fails with the message `Author identity unknown`,
+you need to tell Git who you are.
+Run the following commands to set your name and email address:
 
@@
 
-Before adding your project files, it is helpful to decide on a folder structure, that is, how to call each file and where to put it.
-In general, the folder structure should facilitate understanding a project by breaking it into logical chunks.
+Before adding your project files, it is helpful to decide on a folder structure,
+that is, how to call each file and where to put it.
+In general, the folder structure should facilitate understanding a project
+by breaking it into logical chunks.
 There is no single best solution, as a good structure depends on where a project's complexity lies.
 However, it is usually helpful if the different files and folders reflect the execution order.
-For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders.
-Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.
+For example, if there are multiple data processing stages,
+one can possibly differentiate input (raw data), intermediate (processed data),
+and output files (e.g., figures) and put them into separate folders.
+Similarly, the corresponding code files (e.g., preparation, modeling, visualization)
+can be prefixed with increasing numbers.
 
-Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @TIER4; @Wilson2017] as well as specific to a particular programming language [e.g., @ArayaSalas2024; @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021].
+Luckily, there are already numerous proposals for how to organize one's project files,
+both general [e.g., @TIER4; @Wilson2017] as well as specific
+to a particular programming language [e.g., @ArayaSalas2024; @Marwick2018; @Vuorre2021]
+or journal [@Vilhuber2021].
 We recommend you to follow the standards of your field.
 
-In this simple project, we will explore differences in bill length between male and female penguins. You can put all your files in the root folder of your project.
+In this simple project, we will explore differences in bill length
+between male and female penguins.
+You can put all your files in the root folder of your project.
 
@@
 
-In order to get you started, we have prepared a manuscript for you, alongside a bibliography file. Download the two files to your computer and put them into your project folder.
+In order to get you started, we have prepared a manuscript for you,
+alongside a bibliography file.
+Download the two files to your computer and put them into your project folder.
 
@@
 
-To create a PDF from the manuscript, you need to have a $\TeX$ distribution installed on your computer.
-A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:
+To create a PDF from the manuscript,
+you need to have a $\TeX$ distribution installed on your computer.
+A lightweight choice is [TinyTeX](https://yihui.org/tinytex/),
+which can be installed with Quarto as follows:
 

### 2025-02-08 choose_license.qmd
Fix typos
Introduced in 9e90838

@@
 :::
-https://creativecommons.org/licenses/by/4.0/
+
 [ofl]: https://openfontlicense.org/
@@
 
-!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024]
-licensed under [CC\ BY\ 4.0][cc-by]](images/Morrison2024.jpg){.lightbox}
+!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024] licensed under [CC\ BY\ 4.0][cc-by]](images/Morrison2024.jpg){.lightbox}
 :::

### 2025-02-09 index.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.
Address Quarto vs. rmarkdown differences

@@
 
-This tutorial covers how to make your code available to others. It wraps up a series of tutorials which we recommend to go through in advance:
+This tutorial covers how to make your code available to others.
+It wraps up a series of tutorials which we recommend to go through in advance:
 
-- [Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/): This tutorial explains how to set up a simple project with version control.
-- [Collaborative coding with GitHub and RStudio](https://lmu-osc.github.io/Collaborative-RStudio-GitHub/): Here you can learn how to collaborate with others on code projects using GitHub.
-- [Introduction to R](https://lmu-osc.github.io/introduction-to-R/): This is an introduction to the R programming language.
-- [Introduction to `renv`](https://lmu-osc.github.io/introduction-to-renv/): Here you can learn how to manage the dependencies of an R project.
-- [Introduction to Quarto](https://lmu-osc.github.io/introduction-to-Quarto/): Learn how to combine prose and code to make documents more reproducible.
-- [Introduction to Zotero](https://lmu-osc.github.io/introduction-to-zotero/): Get introduced to the reference management software Zotero.
-- [FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/): Learn how to make your data findable, accessible, interoperable, and reusable.
+- [Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/):
+  This tutorial explains how to set up a simple project with version control.
+- [Collaborative coding with GitHub and RStudio](https://lmu-osc.github.io/Collaborative-RStudio-GitHub/):
+  Here you can learn how to collaborate with others on code projects using GitHub.
+- [Introduction to R](https://lmu-osc.github.io/introduction-to-R/):
+  This is an introduction to the R programming language.
+- [Introduction to `renv`](https://lmu-osc.github.io/introduction-to-renv/):
+  Here you can learn how to manage the dependencies of an R project.
+- [Introduction to Quarto](https://lmu-osc.github.io/introduction-to-Quarto/):
+  Learn how to combine prose and code to make documents more reproducible.
+- [Introduction to Zotero](https://lmu-osc.github.io/introduction-to-zotero/):
+  Get introduced to the reference management software Zotero.
+- [FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/):
+  Learn how to make your data findable, accessible, interoperable, and reusable.
 
-While the general principles taught in this tutorial apply to most researchers publishing code, we will make very specific suggestions what tools to use.
+While the general principles taught in this tutorial apply to most researchers
+publishing code, we will make very specific suggestions what tools to use.
 
-In particular, we will assume you have installed [Git](https://git-scm.com/downloads), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/#:~:text=2%3A%20Install%20RStudio), and [Quarto](https://quarto.org/docs/download/).
-Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).
+In particular, we will assume you have installed
+[Git](https://git-scm.com/downloads),
+[R](https://cloud.r-project.org/),
+[RStudio](https://posit.co/download/rstudio-desktop/#:~:text=2%3A%20Install%20RStudio),
+and [Quarto](https://quarto.org/docs/download/).
+Windows users may additionally need
+[RTools](https://cloud.r-project.org/bin/windows/Rtools/).
 
@@
 Quarto is a tool to render dynamic documents, but not the only one.
-You may also use [`rmarkdown`](https://rmarkdown.rstudio.com/), [Jupyter Book](https://jupyterbook.org/), [Stencila](https://stencila.io/), [Curvenote](https://curvenote.dev/), [showyourwork!](https://show-your.work/), [Org Mode](https://orgmode.org/), [StatTag](https://sites.northwestern.edu/stattag/), or one of the countless others.
+
+It is the successor of the R package [`rmarkdown`](https://rmarkdown.rstudio.com/)
+and mostly backwards-compatible.
+`rmarkdown` continues to be supported,
+but the development of new features happens in Quarto.
+The main difference is that Quarto has first-class support for Python.
+In other words, Python users of Quarto do not need to have R installed.
+
+While this tutorial focuses on Quarto,
+you can also create reproducible research projects with the following tools:
+
+- [Jupyter Book](https://jupyterbook.org/)
+- [Stencila](https://stencila.io/)
+- [Curvenote](https://curvenote.dev/)
+- [showyourwork!](https://show-your.work/)
+- [Org Mode](https://orgmode.org/)
+- [StatTag](https://sites.northwestern.edu/stattag/)
 :::
 
-We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:
+We will start with creating a simple report which incorporates R code.
+Then, we will focus on what is required to publish your work:
 

### 2025-02-09 choose_license.qmd
Rework adding data set (again)
This reverses commit 6d3ca2b.
Instead of pointblank, which
has many dependencies, datawizard
and tinytable are used (again).
tinylabels is not necessary,
however, as adding labels can be
already achieved with datawizard
alone.

Main motivation was that basic
summary statistics should be
part of every data dictionary,
and althugh this can be achieved
with pointblank, it requires more
manual work. Also, value labels
are catered for automatically,
while with pointblank one would
first have to think where exactly
to indicate them (probably as
subsections to the "colu

@@
 
-For example, if you previously created the file `create_data_dictionary.R`,
-you can now add the following comment to the beginning of the file,
-replacing `<YEAR>` and `<YOUR NAME>` with the current year and your name
--- of course, you can also choose a different license:
-
-```{.r filename="create_data_dictionary.R"}
-# SPDX-FileCopyrightText: <YEAR> <YOUR NAME>
-#
-# SPDX-License-Identifier: CC0-1.0
-```
-
-You need to use `#` to start the comment
-because this is the symbol that starts comment lines in R scripts.
-Alternatively, you can use the [reuse tool](https://github.com/fsfe/reuse-tool)
-to add these information for you.
-After installing it with...
-
-```{.bash filename="Terminal"}
-pipx install reuse
-```
-
-...you can add the copyright information using the following command
--- the current year will be added automatically:
-
-```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="<YOUR NAME>" --license="CC0-1.0" create_data_dictionary.R
-```
-
-In many cases, the reuse tool will figure out the appropriate comment style for you.
-If this is not the case, as currently with Quarto files,
-you can tell it directly which comment style to use (`html` in this case):
-
-```{.bash .code-overflow-wrap filename="Terminal"}
-reuse annotate --copyright="Josephine Zerna" --copyright="Christoph Scheffel" --license="CC0-1.0" --style=html Manuscript.qmd
-```
-
-This adds the following header to `Manuscript.qmd`:
+For example, the file `Manuscript.qmd` was provided to you under [CC0\ 1.0][cc0].
+You can indicate that by adding the following comment to the beginning of the file:
 
@@
 
-Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0][cc0],
-which is what you indicate with the previous comment.
+You need to use `<!--` and `-->` to start and end the comment
+because this is the convention that starts comment lines in Quarto files.
+Alternatively, you can use the [reuse tool](https://github.com/fsfe/reuse-tool)
+to add these information for you.
+After installing it with...
+
+```{.bash filename="Terminal"}
+pipx install reuse
+```
+
+...you can add the copyright information using the following command
+-- the current year will be added automatically.
+In many cases, the reuse tool will figure out the appropriate comment style for you.
+If this is not the case, as currently with Quarto files,
+you can tell it directly which comment style to use (`html` in this case):
+
+```{.bash .code-overflow-wrap filename="Terminal"}
+reuse annotate --copyright="Josephine Zerna" --copyright="Christoph Scheffel" --license="CC0-1.0" --style=html Manuscript.qmd
+```
+
 As you edited the file throughout this tutorial,
@@
 Also, you do not need to provide your real name.]
+
+```{.bash .code-overflow-wrap filename="Terminal"}
+reuse annotate --copyright="<YOUR NAME>" --license="CC0-1.0" --style=html Manuscript.qmd
+```
 
@@
 are made available to you under [CC0\ 1.0][cc0]
-as is all code in this tutorial (e.g., for creating the data dictionary).
+as is all code in this tutorial (e.g., for the data analysis).
 Therefore, you can apply any license to your modified version.
@@
 
-The bibliography (stored in "Bibliography.bib"), data dictionary (stored in "data_dictionary.html"), and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
+The bibliography (stored in "Bibliography.bib") by Florian Kohrt and <YOUR NAME> is available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
 ```

### 2025-02-09 data.qmd
Rework adding data set (again)
This reverses commit 6d3ca2b.
Instead of pointblank, which
has many dependencies, datawizard
and tinytable are used (again).
tinylabels is not necessary,
however, as adding labels can be
already achieved with datawizard
alone.

Main motivation was that basic
summary statistics should be
part of every data dictionary,
and althugh this can be achieved
with pointblank, it requires more
manual work. Also, value labels
are catered for automatically,
while with pointblank one would
first have to think where exactly
to indicate them (probably as
subsections to the "colu

@@
 
-When distributing a data set, it is important to document the meaning (e.g., units)
-and valid values of its variables.
+Whether or not distributing the data set,
+it is important to document the meaning (e.g., units)
+and values of its variables.
 This is typically done with a _data dictionary_ (also called a _codebook_).
 In the following, we will demonstrate how to create a simple data dictionary
-using the R package [`pointblank`][pointblank].
-You can install it now using:
+using the R packages [`datawizard`][datawizard] and [`tinytable`][tinytable].
+You can install them now using:
 
-[pointblank]: https://rstudio.github.io/pointblank/
-
-::: {.column-margin}
-![pointblank: Data Validation and Organization of Metadata for Local and Remote Tables](images/pointblank.png){width=250px}
-:::
+[datawizard]: https://easystats.github.io/datawizard/
+[tinytable]: https://vincentarelbundock.github.io/tinytable/
 
 ```{.r filename="Console"}
-renv::install("pointblank")
+renv::install(c(
+  "datawizard",
+  "tinytable"
+))
 ```
 
-You can put the code that follows for creating the data dictionary
-into a new file called `create_data_dictionary.R`.
+You can put the code that follows in a separate document.
+Create it by clicking on _File_ > _New File_  > _Quarto Document..._.
+Choose a title such as `Data Dictionary`,
+select _HTML_ as format,
+uncheck the use of the visual markdown editor, and click on _Create_.
+Remove everything except the YAML header (between the `---`).
+To make the HTML file self-contained,
+also set `embed-resources: true` such that the YAML header looks as follows:
 
-First, we write down everything we know about the data set.
-This includes:
+```{.yml filename="data_dictionary.qmd"}
+---
+title: "Data Dictionary"
+format:
+  html:
+    embed-resources: true
+---
+```
 
-- a general description of the data set
-- descriptions of all columns
-- valid values, where applicable
+Then, save it as `data_dictionary.qmd` by clicking on _File_ > _Save_.
 
-```{r, filename="create_data_dictionary.R"}
-table_info <- c(
-  title = "palmerpenguins::penguins",
-  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
-)
-descriptions <- c(
+To create the actual data dictionary, first write a description for all columns
+so others can understand what the variable names mean.
+Where necessary, also document their value
+-- this is especially important if their meaning is non-obvious.
+In the following, we demonstrate this by storing the penguins' binomial name
+along with the English name.
+
+``````{cat}
+#| engine.opts = list(file = "_data_dictionary.qmd"),
+#| class.source = "md",
+#| filename = "data_dictionary.qmd"
+
+```{r}
+#| echo: false
+
+vars <- c(
   species = "a character string denoting penguin species",
@@
 vals <- list(
-  species = c("Adelie", "Gentoo", "Chinstrap"),
-  island = c("Torgersen", "Biscoe", "Dream"),
-  sex = c("male", "female"),
-  year = c(2007, 2008, 2009)
+  species = c(
+    Adelie = "Pygoscelis adeliae",
+    Gentoo = "Pygoscelis papua",
+    Chinstrap = "Pygoscelis antarcticus"
   )
+)
+```
+``````
+
+Then, load the data and use `datawizard`
+to add the descriptions to the `data.frame`:[^not-permanent]
+
+[^not-permanent]: Note that the code provided does not alter the data file
+-- no description will be added to `data.csv`.
+The descriptions are only added to a (temporary) copy of the data set within R
+to create the data dictionary.
+
+::: {.column-margin}
+![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
+:::
+
+``````{cat}
+#| engine.opts = list(file = "_data_dictionary.qmd", append = TRUE),
+#| class.source = "md",
+#| filename = "data_dictionary.qmd"
+
+```{r}
+#| echo: false
+
+dat <- read.csv("data.csv")
+
+for (x in names(vars)) {
+  if (x %in% names(vals)) {
+    dat <- datawizard::assign_labels(
+      dat,
+      select = I(x),
+      variable = vars[[x]],
+      values = vals[[x]]
+    )
+  } else {
+    dat <- datawizard::assign_labels(
+      dat,
+      select = I(x),
+      variable = vars[[x]]
+    )
+  }
+}
+```
+
+``````
+
+Next, you can use `datawizard` again to create
+the data dictionary containing the descriptions,
+but also some other information about each variable
+(e.g., the number of missing values).
+Finally, you can use `tinytable` to print the table.
+
+::: {.column-margin}
+![tinytable: Simple and Configurable Tables](images/tinytable.svg){width=250px}
+:::
+
+``````{cat}
+#| engine.opts = list(file = "_data_dictionary.qmd", append = TRUE),
+#| class.source = "md",
+#| filename = "data_dictionary.qmd"
+
+```{r}
+#| echo: false
+#| column: "body-outset"
+
+(dict <- datawizard::data_codebook(dat) |>
+  subset(select = -c(ID, .row_id)) |>
+  tinytable::tt() |>
+  tinytable::format_tt(replace = "")) # hide NAs
+```
+
+``````
+
+```{r}
+#| child: "_data_dictionary.qmd"
+
 ```
@@
 
-As the data set is from the R package `palmerpenguins`,
-if you have it installed you can use the function `citation()`
-to create such a reference:
+The data set is from the R package `palmerpenguins`.
+If you had it installed
+you could use the function `citation()` to create such a reference:
 
 ```{r}
-dat_source <- citation("palmerpenguins", auto = TRUE) |>
+#| label: "data-citation"
+#| output: false
+
+citation("palmerpenguins", auto = TRUE) |>
   format(bibtex = FALSE, style = "text")
@@
 you can find a [suggested citation on its website][palmerpenguins-citation]
-and run the following:
+and add that to your data dictionary:
 
@@
 ```{r}
-#| echo: false
-#| class-output: "r code-overflow-wrap"
-#| attr-output: 'filename="create_data_dictionary.R"'
+#| ref.label = "data-citation",
+#| render = function(x, options) gsub("\\n", " ", x = x),
+#| echo = FALSE,
+#| class.output = "md code-overflow-wrap",
+#| attr.output = 'filename="data_dictionary.qmd"'
 
-# We have provided the data set as CSV file to the readers.
-# Therefore, we cannot assume or require that readers have
-# the R package palmerpenguins installed. Instead, we create
-# the citation on our end and hide how we obtained it.
-dat_source |>
-  gsub("\\n", " ", x = _) |>
-  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
-  paste0("dat_source <- \"", ... = _, "\"") |>
-  cat()
+# This chunk takes the output from the chunk "data-citation"
+# and renders it with all newlines replaced by whitespaces.
 ```
 
-Then, we use `pointblank` to create a data dictionary with this information.
+Finally, you can render the data dictionary by running the following:
 
-```{r, filename="create_data_dictionary.R"}
-vals <- sapply(vals, \(x) {
-  paste0(
-    "(",
-    knitr::combine_words(x, and = " or ", before = "`", after = "`"),
-    ")"
-  )
-})
-
-dat <- read.csv("data.csv")
-
-dict <- pointblank::create_informant(
-  dat,
-  tbl_name = NA,
-  label = table_info[["title"]],
-  lang = "en"
-) |>
-  pointblank::info_tabular(
-    Description = table_info[["description"]],
-    Source = dat_source
-  ) |>
-  pointblank::info_columns_from_tbl(stack(descriptions)[2:1]) |>
-  pointblank::info_columns_from_tbl(stack(vals)[2:1]) |>
-  pointblank::get_informant_report(
-    title = "Data Dictionary for `data.csv`"
-  )
-
-dict
+```{.bash filename="Terminal"}
+quarto render data_dictionary.qmd
 ```
 
-Finally, we can store the data dictionary inside an HTML file
-and put the HTML file into the project folder as well.
-
-```{.r filename="create_data_dictionary.R"}
-pointblank::export_report(dict, filename = "data_dictionary.html")
-```
-
-For a more elaborate introduction to `pointblank`,
-you can read its "[Intro to Information Management][pointblank-intro]".
-
-[pointblank-intro]: https://rstudio.github.io/pointblank/articles/INFO-1.html
+This should create the file `data_dictionary.html`
+which you open and view in your web browser.
 
@@
 the description of the variables and their valid values are stored in a machine-readable way.
-As before, we also provide a reference to the source.
+We'll reuse the descriptions we already created[^value-labels] and add a few others.
+
+[^value-labels]: Unfortunately, the descriptions of values are not reused in this example,
+as they are [not supported][enum-labels] by the specification we are using.
+
+[enum-labels]: https://specs.frictionlessdata.io/patterns/#table-schema-enum-labels-and-ordering
+
+First, store the title and description of the data set as a whole:
+
+```{.r filename="Console"}
+table_info <- c(
+  title = "penguins data set",
+  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
+)
+```
+
+As before, also provide a reference to the source.
 
@@
   cat()
+```
+
+Next, create a list of the categorical variables' valid values:
+
+```{.r filename="Console"}
+valid_vals <- list(
+  species = c("Adelie", "Gentoo", "Chinstrap"),
+  island = c("Torgersen", "Biscoe", "Dream"),
+  sex = c("male", "female"),
+  year = c(2007, 2008, 2009)
+)
+```
+
+Finally, store the descriptions of the variables we already created earlier:
+
+```{.r filename="Console"}
+vars <- c(
+  species = "a character string denoting penguin species",
+  island = "a character string denoting island in Palmer Archipelago, Antarctica",
+  bill_length_mm = "a number denoting bill length (millimeters)",
+  bill_depth_mm = "a number denoting bill depth (millimeters)",
+  flipper_length_mm = "an integer denoting flipper length (millimeters)",
+  body_mass_g = "an integer denoting body mass (grams)",
+  sex = "a character string denoting penguin sex",
+  year = "an integer denoting the study year"
+)
 ```
@@
 ```{.r filename="Console"}
+# Install {frictionless} and the required dependency {stringi}
+renv::install(c(
+  "frictionless",
+  "stringi"
+))
+
 # Read data and create schema
@@
 dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  c(x, description = descriptions[[x$name]])
+  c(x, description = vars[[x$name]])
 })
@@
 dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  if (x$name %in% names(vals)) {
-    modifyList(x, list(constraints = list(enum = vals[[x$name]])))
+  if (x[["name"]] %in% names(valid_vals)) {
+    modifyList(x, list(constraints = list(enum = valid_vals[[x$name]])))
   } else {

### 2025-02-09 make_readme.qmd
Rework adding data set (again)
This reverses commit 6d3ca2b.
Instead of pointblank, which
has many dependencies, datawizard
and tinytable are used (again).
tinylabels is not necessary,
however, as adding labels can be
already achieved with datawizard
alone.

Main motivation was that basic
summary statistics should be
part of every data dictionary,
and althugh this can be achieved
with pointblank, it requires more
manual work. Also, value labels
are catered for automatically,
while with pointblank one would
first have to think where exactly
to indicate them (probably as
subsections to the "colu

@@
 ```txt
-Finding R package dependencies ... Done!
-── Install scripts ──────────────────────── Fedora 40 ── 
-dnf install -y make pandoc libcurl-devel \
-  openssl-devel libicu-devel v8-devel libxml2-devel \
-  git
+── Install scripts ────────────────── Fedora 40 ── 
+dnf install -y make pandoc git
 
-── Packages and their system dependencies ──────────────
-blastula  – pandoc
-curl      – libcurl-devel, openssl-devel
+── Packages and their system dependencies ────────
 fs        – make
 knitr     – pandoc
-openssl   – openssl-devel
 remotes   – git
@@
 sass      – make
-stringi   – libicu-devel
-V8        – v8-devel
-xml2      – libxml2-devel
 ```
@@
 
-The other system dependencies are libraries,
-and the method to identify the installed version varies between operating systems.
-For example, on Fedora Linux, one can run the following:
-
-```{.bash .code-overflow-wrap filename="Terminal"}
-rpm -q libcurl-devel openssl-devel libicu-devel v8-devel libxml2-devel
-```
-
-However, they are only relevant if the R packages were installed from source,
-rather than as binary package.
+However, this does not work for all system dependencies.
+Specifically, it does not work for libraries -- software that is not supposed to be run on its own.
+Identifying their version is beyond the scope of this tutorial.
 
@@
 - `data_dictionary.html`: a dictionary to the data file,
-  created using `create_data_dictionary.R`
+  created using `data_dictionary.qmd`
 

### 2025-02-10 copyright.qmd
Minor improvements
- Decrease size of penguins image
- Detailed instructions for creating a text file

@@
 ```{.md .code-overflow-wrap}
-!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" licensed under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/)](penguin.jpg)
+!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" licensed under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/)](penguin.jpg){width=50%}
 ```
@@
 ```{.md .code-overflow-wrap}
-!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/) from [`flic.kr/p/2pEKnUr`](https://flic.kr/p/2pEKnUr)](penguin.jpg)
+!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/) from [`flic.kr/p/2pEKnUr`](https://flic.kr/p/2pEKnUr)](penguin.jpg){width=50%}
 ```
@@
 In addition, we recommend creating a file called `LICENSE.txt`
-to describe the license of the image:
+to describe the license of the image.
+Click on _File_ > _New File_ > _Text File_,
+then on _File_ > _Save_, and choose `LICENSE.txt` as filename.
 

### 2025-02-10 data.qmd
Remove tinytable
tinytable is a lightweight package
but we don't even need it. The
data dictionary can be exported
to markdown using datawizard alone.
Also, we use datawizard::data_select
rather than base::subset to keep
the attributes of the data.frame.

@@
 In the following, we will demonstrate how to create a simple data dictionary
-using the R packages [`datawizard`][datawizard] and [`tinytable`][tinytable].
+using the R package [`datawizard`][datawizard].
 You can install them now using:
@@
 [datawizard]: https://easystats.github.io/datawizard/
-[tinytable]: https://vincentarelbundock.github.io/tinytable/
 
 ```{.r filename="Console"}
-renv::install(c(
-  "datawizard",
-  "tinytable"
-))
+renv::install("datawizard")
 ```
@@
 
-Next, you can use `datawizard` again to create
-the data dictionary containing the descriptions,
+Then, you can create the data dictionary containing the descriptions,
 but also some other information about each variable
-(e.g., the number of missing values).
-Finally, you can use `tinytable` to print the table.
-
-::: {.column-margin}
-![tinytable: Simple and Configurable Tables](images/tinytable.svg){width=250px}
-:::
+(e.g., the number of missing values) and print it.
 
@@
 #| column: "body-outset"
+#| classes: plain
 
-(dict <- datawizard::data_codebook(dat) |>
-  subset(select = -c(ID, .row_id)) |>
-  tinytable::tt() |>
-  tinytable::format_tt(replace = "")) # hide NAs
+datawizard::data_codebook(dat) |>
+  datawizard::data_select(exclude = c("ID", ".row_id")) |>
+  datawizard::data_filter(N != "") |>
+  datawizard::print_md()
 ```

### 2025-02-11 intro.qmd
Remove unnecessary anticipation
...of the term "research compendium"

@@
 documenting their usage,
-and linking results with their computations,
-leading to the creation _research compendia_.
+and linking results with their computations.
 

### 2025-02-11 copyright.qmd
Provide examples for boilerplate and free/open licenses

@@
 pre-formulated "boilerplate" licenses have been created.
+For example, you might have heard about ["Creative Commons" licenses][cc-licenses],
+which are used by Wikipedia.
 Recognizing them facilitates both sharing (on the side of the author)
@@
 
+[cc-licenses]: https://creativecommons.org/share-your-work/cclicenses/
 [sfconservancy]: https://sfconservancy.org/copyleft-compliance/
@@
 and that they can use it for commercial purposes.
-By the choice of license, authors can, however, demand that they are credited,
+Coming back to the example of Creative Commons licenses,
+not all of them are free/open.
+For example, some restrict commercial use and modification.
+The Creative Commons license "[Attribution-ShareAlike 4.0][cc-by-sa]"
+is an example of a free/open license, however,
+and we will cover several others below.
+
+[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/
+
+By the choice of license, authors can demand that they are credited,
 that the original license is indicated, that modifications are indicated,

### 2025-02-11 about.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.
Mention all authors and reviewers

@@
     orcid: "0000-0003-0374-5625"
+  - name: "Josephine Zerna"
+    orcid: "0000-0003-2892-884X"
+  - name: "Christoph Scheffel"
+    orcid: "0000-0001-5963-9229"
 bibliography: grateful-refs.bib
 ---
+
+---
+
+This tutorial was created by Florian Kohrt.
+It is based on previous work by Josephine Zerna and Christoph Scheffel.
+Comments on earlier versions were kindly provided by
+Malika Ihle, Florian Pargent, Jonas Hagenberg, Felix Schönbrodt,
+and Sarah von Grebmer zu Wolfsthurn.
 
@@
 
-Except where noted otherwise, the narrative text in this tutorial is licensed under [CC\ BY-SA\ 4.0](LICENSE-CC-BY-SA.txt); the code without any narrative text is also (at your option) available under [CC0\ 1.0](LICENSE-CC0.txt).
+Except where noted otherwise, the narrative text in this tutorial
+is licensed under [CC\ BY-SA\ 4.0](LICENSE-CC-BY-SA.txt);
+the code without any narrative text is also (at your option)
+available under [CC0\ 1.0](LICENSE-CC0.txt).
 
-The files `Manuscript.qmd` and `Bibliography.bib` are made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).
+The files `Manuscript.qmd` and `Bibliography.bib` are made available
+by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).
 
-The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are Copyright (C) 2024 Posit PBC and are available under the [GNU Affero General Public License v3](LICENSE-AGPL.txt).
+The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio)
+are Copyright (C) 2024 Posit PBC and are available
+under the [GNU Affero General Public License v3](LICENSE-AGPL.txt).
 The source code is available on [GitHub](https://github.com/rstudio/rstudio/tree/v2024.04.2%2B764).

### 2025-02-11 code.qmd
Tidy up the analysis code

@@
   knitr::kable()
-
 ```
@@
 
-# Create ggplot object
 ggplot2::ggplot(dat_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +

### 2025-02-12 choose_license.qmd
Follow-up from d5e2128
Don't cover data dictionary during
license choice. The data dictionary
is most likely only a derivative work
of the data and I would rather have
clear examples of own contributions
for this page.

@@
 For the purpose of this tutorial,
-we will consider the manuscript and the bibliography which you edited
-as well as the data dictionary you created.
+we will consider the manuscript and the bibliography which you edited.
 
@@
 under [AGPLv3][agpl] (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0][cc-by-sa],
-while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.
+while the bibliography is made available under CC0\ 1.0.
 

### 2025-02-12 choose_license.qmd
Fix location of footnotes
Place them after any punctuation

@@
 Either under a specified software license,
-or under a Creative Commons license[^media].
+or under a Creative Commons license.[^media]
 This is called multi-licensing and makes it easier

### 2025-02-12 code.qmd
Fix location of footnotes
Place them after any punctuation

@@
 and the $\TeX$ distribution we installed before.
-Then there are the individual R packages[^renv-copying],
+Then there are the individual R packages,[^renv-copying]
 Quarto extensions, and $\TeX$ packages.
@@
   Every dependency increases the risk of irreproducibility in the future.
-  Prefer packages that are well-maintained and light on dependencies[^pkg-deps].
+  Prefer packages that are well-maintained and light on dependencies.[^pkg-deps]
   We also recommend you to read "When should you take a dependency?"

### 2025-02-12 data.qmd
Be explicit about what renv::snapshot() does
Don't run code unnecessarily
Clarify code comments

@@
 
+# Store the description of variables
 vars <- c(
@@
 
+# Store the description of variable values
 vals <- list(
@@
 #| label: "data-citation"
-#| output: false
+#| eval: false
 
@@
 Having added the data and its documentation,
-one can view and record the utilized packages with `renv`:
+one can view and record the utilized packages with `renv`,
+thus bringing the project into a consistent state:
 

### 2025-02-12 make_readme.qmd
Clarify code comments

@@
 
-# Find all dependencies
+# Find all R package dependencies
 deps <- renv::dependencies()$Package |>
@@
 
-# Identify the system dependencies
+# Identify their system dependencies
 pak::pkg_sysreqs(deps)

### 2025-02-12 index.qmd
Clarify requirements

@@
 While the general principles taught in this tutorial apply to most researchers
-publishing code, we will make very specific suggestions what tools to use.
+publishing code, we will focus on the programming language R
+and the publishing system Quarto.
 
-In particular, we will assume you have installed
+To follow the tutorial you need to have installed
 [Git](https://git-scm.com/downloads),
-[R](https://cloud.r-project.org/),
-[RStudio](https://posit.co/download/rstudio-desktop/#:~:text=2%3A%20Install%20RStudio),
-and [Quarto](https://quarto.org/docs/download/).
-Windows users may additionally need
-[RTools](https://cloud.r-project.org/bin/windows/Rtools/).
+[R](https://cloud.r-project.org/), and
+[Quarto](https://quarto.org/docs/download/).
+Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).
+Some things will be explained using the [RStudio IDE](https://posit.co/download/rstudio-desktop/#:~:text=2%3A%20Install%20RStudio),
+but you are free to use any IDE you like.
 

### 2025-02-12 data.qmd
Don't remove .row_id
This was only necessary because we
previously used tinytable to export
the data dictionary. Now that we use
datawizard, we can leave them in,
making the code easier to read and
understand.
Provide examples more examples

@@
 datawizard::data_codebook(dat) |>
-  datawizard::data_select(exclude = c("ID", ".row_id")) |>
+  datawizard::data_select(exclude = ID) |>
   datawizard::data_filter(N != "") |>
@@
 Depending on the type of data, it may also be necessary
-to describe measurement instruments, sampling procedures, appropriate weighting,
-applied preprocessing steps, or contact information.
-In our case, as the data have already been published,
+to describe sampling procedures (e.g., selection criteria),
+measurement instruments (e.g., questionnaires),
+appropriate weighting,
+already applied preprocessing steps, or contact information.
+In our case, as the data has already been published,
 we only store a reference to its source.

### 2025-02-12 data.qmd
Add helpful code comments

@@
 ```{.r filename="Console"}
+# Store the description of variables
 vars <- c(

### 2025-03-05 copyright.qmd
Specify knitr as engine for pages
We rely on knitr-specific syntax
throughout the tutorial, even in
instances where no code is executed.

@@
 title: "Copyright"
+engine: knitr
 format:

### 2025-03-05 code.qmd
Specify knitr as engine for pages
We rely on knitr-specific syntax
throughout the tutorial, even in
instances where no code is executed.
Don't pass unncessary command line args
I think it's bad practice to skip them.
Minor improvements

@@
 title: "Add Code"
+engine: knitr
 ---
@@
 For the purpose of this tutorial, we suggest to conduct a simple analysis
-that explores whether the bill lengths of penguins vary between sex.
+that explores whether the bill lengths of penguins vary depending on their sex.
 You can write your own or use the following example:
@@
 ```{.bash filename="Terminal"}
-quarto add --no-prompt wjschne/apaquarto
+quarto add wjschne/apaquarto
 ```

### 2025-03-05 data.qmd
Minor improvements

@@
 Privacy laws and contractual obligations may require you
-to create a completely anonymized or synthetic data set[^anon-synth-tools] (if possible),
-or prohibit any sharing of data, in which case you should provide a reference
-to a data repository where they can be obtained from.
+to create an anonymized or synthetic data set[^anon-synth-tools],
+in which case you should provide a reference
+to a repository where the originally measured data can be obtained from.
 For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)"

### 2025-03-05 code.qmd
Rework README
Rework the instructions for creating
a README to better fit to papers/
manuscripts, rather than to software
projects.

@@
 If you already committed a file that contains credentials, you can follow @Chacon2024.
+:::
+
+::: {#wrn-dependencies .callout-warning}
+### Dealing with Dependencies
+
+Everything not included in the project folder that is required
+for running the project is called a dependency.
+Dependencies are possible breaking points in the future,
+therefore it's best to keep them at a minimum.
+If you cannot avoid taking a dependency,
+make sure that it's available to users of your project in the long term.
+In the following, we will discuss two common examples of dependencies:
+R packages and downloaded files.
+
+#### R Packages
+
+R packages that you use in your analysis are an obvious example of dependencies.
+Their version is recorded by `renv`,
+but you also need to ensure that they are available for download.
+First, identify the source from which you installed your packages:
+
+```{.r filename="Console"}
+# First, install {pak} and {sessioninfo}
+renv::install(c(
+  "pak",
+  "sessioninfo"
+))
+
+# Which R packages does the project directly depend on?
+deps <- renv::dependencies()$Package |>
+  unique()
+
+# Which R packages does the project indirectly depend on?
+deps <- deps |>
+  pak::pkg_deps(dependencies = NA) |>
+  getElement("package")
+
+# Get information about their source
+sessioninfo::package_info(deps)
+```
+
+If the `source` column only contains the entries `CRAN`, `RSPM`, or `Bioconductor`,
+they are already archived.
+If the `source` column instead mentions something else (e.g., `GitHub`),
+you need to make sure yourself that the package is available to users of your project.
+You can either archive the package,
+for example in the [Software Heritage archive](https://archive.softwareheritage.org/).
+Or you store a copy of the package in the project folder
+-- of course, you need to make sure that you are allowed to do so from a copyright perspective
+and comply with its license.
+
+#### Downloaded Files
+
+If your code interacts with the internet, for example, to download files,
+this is another common dependency with the risk of breaking in the future.
+If possible, store a copy of the downloaded file in your project folder.
+Alternatively, you can upload it to a permanent repository such as Zenodo
+(discussed [later](make_readme.qmd)).
 :::

### 2025-03-05 make_readme.qmd
Rework README
Rework the instructions for creating
a README to better fit to papers/
manuscripts, rather than to software
projects.

@@
 While you could just start writing along,
-most READMEs have some sections in common which we describe below.
+it is helpful to provide at least the following information in sections on their own.
 
-Name
+### Name and Description
 
-:   How is the project called?
+How is the project called?
+What is it about?
+Which files does the project folder contain?
+How are they organized?
 
-Badges/Project Status
+### Involved Data
 
-:   Badges typically report quick facts, like
-    [whether the project is actively maintained][repostatus],
-    [how many dependencies it has][tinyverse],
-    or [whether it has been published in a package repository][metacran].
-    Sometimes, the license, any associated DOIs,
-    or quality metrics like code coverage by unit tests
-    are communicated via badges as well.
+Are any (empirical) data involved (e.g., being analyzed or used as input)?
+From which sources can they be obtained?
+Are they already included in the project folder?
+Where is their data dictionary located?
+Which terms, usage restrictions, or licenses apply?
+If they are not publicly available,
+is an alternative, synthesized version provided?
 
-[repostatus]: https://www.repostatus.org/
-[tinyverse]: https://tinyverse.netlify.app/
-[metacran]: https://www.r-pkg.org/services#badges
+### Computational Requirements
 
-Description
+What software needs to be installed to run the analysis
+-- in other words, what are its dependencies?
+This also includes software that you have used for any manual steps.
+If the code has particular hardware requirements
+(e.g., in terms of processor or memory),
+these should be also noted.
+Finally, for steps that take more than a couple of seconds,
+the approximate runtime should be indicated.
 
-:   What is the project about? What are its features? Why was it created?
+@nte-dependencies provides more information
+on determining the dependencies of an R project.
 
-Visuals
+### Usage
 
-:   Is there anything you can show that demonstrates how the project can be used?
-    Screenshots or other visuals can make the README more appealing.
+How can one run the project -- is there a master script
+or a particular order in which any scripts need to be executed?
+Provide detailed instructions for running the full project.
 
-Installation/Dependencies
+### List of Results
 
-:   What steps need to be taken to run the project? What software needs to be installed?
+For every result (i.e., number, figure, or table) that is
+computed in the project and displayed in the manuscript,
+indicate where exactly it is computed.
+
+### Citation
+
+Is there a recommended way to cite this project?
+Is there a published article associated with it
+that you would like to have cited?
+
+### License
+
+Under which licenses are the works in this project folder available?
+
+## Create It!
+
+Create your README now as the file `README.md`.
+
+::: {#nte-dependencies .callout-note}
+### Identifying R Dependencies
+
 R itself and the R packages are already documented as this project uses `renv`.
@@
 as well as the version of Quarto.[^renv-quarto]
-    Also, don't forget to mention software that you have used for any manual steps.
-    See @sec-dependencies for additional information.
-    It may make sense to provide these information in the README
-    even if you already cite all programs and their version numbers in your manuscript.
 
@@
 has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).
-
-Usage
-
-:   Which files does the project folder contain? How are they organized?
-    How can one run the project -- is there a master script
-    or a particular order in which any scripts need to be executed?
-    How long does it take to run all scripts?
-    Is there additional documentation available?
-
-Support
-
-:   Do you offer support or help, for example, via GitHub discussions
-    or a mailing list?
-
-Contributing
-
-:   If the project is active: Can other people contribute? How?
-    Do you accept contributions? Do you review issues?
-    This section is sometimes outsourced into a file called `CONTRIBUTING.md`.
-
-Authors
-
-:   Who was involved in creating this project? This involves you,
-    your co-authors, and anybody you accepted contributions from.
-
-Citation
-
-:   Is there a recommended way to cite this project?
-    Is there a published article associated with it
-    that you would like to have cited?
-
-License
-
-:   Under which licenses are the works in this project folder available?
-
-## Installation/Dependencies {#sec-dependencies}
 
@@
 ```{.r filename="Console"}
-# First, install pak
-renv::install("pak")
-
 # Find all R package dependencies
@@
 Of course, all the system dependencies identified until now
-may have dependencies on their own. Use your own judgement to decide when not to dig deeper.
+may have dependencies on their own.
+Use your own judgement to decide when not to dig deeper.
+:::
 
-## Create It!
+If you feel stuck, you can have a look at the following examples:
 
-Create your README now as the file `README.md`. If you feel stuck,
-you can have a look at the following examples:
-
-::: {#tip-name .callout-tip collapse="true"}
-### Name
+::: {#tip-name-description .callout-tip collapse="true"}
+### Name and Description
 
@@
 
+This project contains the Quarto manuscript of our study on penguins ("Manuscript.qmd"). It is written in R and uses `renv` to track its dependencies.
+
+The most important file in this project folder is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:
+
+- `Bibliography.bib`: bibliographic references used in the manuscript
+- `data.csv`: a data set containing the simplified `palmerpenguins` data
+- `data_dictionary.html`: a dictionary to the data file,
+  created using `data_dictionary.qmd`
+
+The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.
 ```
@@
 
-::: {#tip-project-status .callout-tip collapse="true"}
-### Badges/Project Status
+::: {#tip-data .callout-tip collapse="true"}
+### Involved Data
 
 ```{.md .code-overflow-wrap filename="README.md"}
-[![Project Status: Unsupported – The project has reached a stable, usable state but the author(s) have ceased all work on it.
-A new maintainer may be desired.](https://www.repostatus.org/badges/latest/unsupported.svg)](https://www.repostatus.org/#unsupported)
+## Involved Data
+
+The manuscript analyzes the "palmerpenguins" data set available from <https://cran.r-project.org/package=palmerpenguins>.
+The data is stored as "data.csv" and documented in the file "data_dictionary.html".
+It is made available under CC0 1.0.
 ```
@@
 
-::: {#tip-description .callout-tip collapse="true"}
-### Description
-
-```{.md .code-overflow-wrap filename="README.md"}
-This project contains the Quarto manuscript of our study on penguins. It is written in R and uses `renv` to track its dependencies.
-```
-:::
-
-::: {#tip-installation .callout-tip collapse="true"}
-### Installation/Dependencies
+::: {#tip-computational-requirements .callout-tip collapse="true"}
+### Computational Requirements
 
 ``````{.md .code-overflow-wrap filename="README.md"}
-## Dependencies
+## Computational Requirements
 
@@
 
-The most important file in this project folder is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:
-
-- `Bibliography.bib`: bibliographic references used in the manuscript
-- `data.csv`: a data set containing the simplified `palmerpenguins` data
-- `data_dictionary.html`: a dictionary to the data file,
-  created using `data_dictionary.qmd`
-
-The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.
-
 The manuscript can be rendered to PDF using the following command:
@@
 ``````
+:::
+
+::: {#tip-list-of-results .callout-tip collapse="true"}
+### List of Results
+
+```{.md .code-overflow-wrap filename="README.md"}
+## List of Results
+
+- In-text numbers in the section "results": Calculated in the chunk "t-test" within "Manuscript.qmd"
+- Table 1: Calculated in the chunk "tbl-descriptive-statistics" within "Manuscript.qmd"
+- Figure 2: Calculated in the chunk "fig-bill-length-comparison" within "Manuscript.qmd"
+```
 :::

### 2025-03-05 copyright.qmd
Prefer images in high res

@@
 
-You can now download the image and move it to your project folder (maybe call it `penguin.jpg`).
+You can now download the image[^image-resolution]
+and move it to your project folder (maybe call it `penguin.jpg`).
 In your manuscript file, find the line that says "add image here", remove it,
 and add the image instead, including proper attribution.
+
+[^image-resolution]: If you are unsure which image resolution (size) is required,
+just download it in the largest possible resolution so it can be printed.
 

### 2025-07-09 choose_license.qmd
Reformat code and fix typos

@@
   at the end of the respective page,
-  or collectively on an separate page.
+  or collectively on a separate page.
 
@@
 Please apply the license of your choice now,
-using any of the methods mentioned above (except the `README`, which we will cover later).
+using any of the methods mentioned above (except the `README`,
+which we will [cover later](make_readme.qmd)).
 
@@
 
-Note. DRM = digital rights management
+_Note._ DRM = digital rights management
 
@@
 
+_Note._ DRM = digital rights management
+
 Advanced License Flowchart for Data(base)

### 2025-07-09 copyright.qmd
Reformat code and fix typos

@@
   This is also called multi-licensing.
-  For example, authors might allow to user a work under any version of a license.
+  For example, authors might allow to use a work under any version of a license.
 - In contrast, it may also be the case that one has to comply
@@
 -- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.
-Without permission you can check you national laws
+Without permission you can check your national laws
 for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_]
@@
 [cc-v4]: https://creativecommons.org/weblog/entry/40768
-[cc0-v1]: http://www.creativecommons.org/weblog/entry/13304
+[cc0-v1]: https://www.creativecommons.org/weblog/entry/13304
 [cc-versions]: https://wiki.creativecommons.org/wiki/License_Versions

### 2025-07-09 archive.qmd
Fix typos
Introduced in 9e90838
Also cite overall software deposit guidance

@@
 there are a few questions to ponder
-[excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0][cc-by]]:
+[excerpt from @Jackson2018Where, with footnotes removed, licensed under [CC\ BY\ 4.0][cc-by]]:
 
@@
 
-[re2data]: https://www.re3data.org/
+[re3data]: https://www.re3data.org/
 [sw-heritage]: https://www.softwareheritage.org/
@@
 For further information, you may be interested in
-[Software Deposit Guidance for Researchers](https://softwaresaved.github.io/software-deposit-guidance/).
+"Software Deposit Guidance for Researchers" [@Jackson2018Guidance].

### 2025-07-09 copyright.qmd
Cite and attribute educational video

@@
 
-The following video "The correct indication of a license"
+The following video by @Brinken2021
 embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy))
@@
   <summary>Load video</summary>
-  <div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
+  <p><div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
     <iframe loading="lazy" src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
-  </div>
+  </div></p>
+  <p>The video “The correct indication of a license: Aspects to consider when reusing the work of others” by Helene Brinken, Jonas Hauss, and Jessika Rücknagel is licensed under <a href="https://creativecommons.org/licenses/by/3.0/de/deed.de">CC&nbsp;BY&nbsp;3.0&nbsp;DE</a>.</p>
 </details>

### 2025-07-09 data.qmd
Also mention OpenDP

@@
 [ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools),
-or [Synthpop](https://www.synthpop.org.uk/).
+[Synthpop](https://www.synthpop.org.uk/), or [OpenDP](https://opendp.org/).
 

### 2025-07-09 data.qmd

@@
 ``````{cat}
-#| engine.opts = list(file = "_data_dictionary.qmd"),
-#| class.source = "md",
-#| filename = "data_dictionary.qmd"
+#| engine.opts: { file: "_data_dictionary.qmd" }
+#| class.source: "md"
+#| filename: "data_dictionary.qmd"
 
@@
 ``````{cat}
-#| engine.opts = list(file = "_data_dictionary.qmd", append = TRUE),
-#| class.source = "md",
-#| filename = "data_dictionary.qmd"
+#| engine.opts: { file: "_data_dictionary.qmd", append: TRUE }
+#| class.source: "md"
+#| filename: "data_dictionary.qmd"
 

### 2025-07-14 about.qmd
Replace chunk by shortcode
Reformat code and fix typos
Shorten chapter on copyright/ transfer to Appendix

@@
     orcid: "0000-0001-5963-9229"
+  - name: "Felix Schönbrodt"
+    orcid: "0000-0002-8282-3910"
 bibliography: grateful-refs.bib
@@
 and Sarah von Grebmer zu Wolfsthurn.
+Felix Schönbrodt created this short version of the [longer original tutorial](https://lmu-osc.github.io/code-publishing/).
 
@@
 
-```{r}
-#| output: "asis"
-#| echo: false
-
-cat(paste0("Quarto ", as.character(quarto::quarto_version()), " was utilized to render the files."))
-```
+Quarto {{< var version >}} was utilized to render the files.
 
@@
 
-This work was partly funded by the Berkeley Initiative for Transparency in the Social Sciences (BITSS), managed by the Center for Effective Global Action (CEGA).
+This work was partly funded by the Berkeley Initiative
+for Transparency in the Social Sciences (BITSS),
+managed by the Center for Effective Global Action (CEGA).

### 2025-07-14 copyright.qmd
Improve wording
Shorten chapter on copyright/ transfer to Appendix

--- 2025-03-05 copyright.qmd
@@
 ---
-title: "Copyright"
+title: "Sharing work of others: Copyright"
 engine: knitr
@@
 By default, everything you put into the project folder will be shared publicly.
-In many instances, this will also include works by others than yourself or your co-authors,
+This typically includes:
+
+- **works by others** than yourself or your co-authors,
   because you may use existing data, images or software. 
-In these cases, you should make sure to only share them to the extent you are allowed to.
-In the following, we will be focusing on copyright,
-but of course, there can also be other restrictions.
+  Make sure to only share those to the extent you are allowed to.
+- **work that was created by yourself** or your co-authors.
 
-::: {#nte-copyright-tldr .callout-note}
-## Copyright TL;DR
+Copyright and licensing can be complicated. 
+In this chapter, we focus on rather simple (but typical) situations.
+For an in-depth explanation of copyright and licenses, 
+see our supplementary in-depth material "[Introduction to Copyright and Licenses](in_depth_material/introduction_copyright.qmd)".
 
-Every expression of an idea, such as a literary or an artistic works,
-is automatically covered by copyright.
-Put simply, this means that nobody else than the copyright holder
-is allowed to copy, modify, or share it.
-Copyright licenses can grant others some of the necessary rights,
-but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
-Certain exceptions and limitations to copyright (such as fair use or the right to quote)
-may obviate the need for a license.
-
-With regard to copyright, free/open licenses can give permission to share a work by others.
-These licenses often require to attribute the authors,
-indicate whether changes were made,
-and provide the text of the license, among other things.
-
-The following video "The correct indication of a license"
-embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy))
-explains the correct attribution of works under Creative Commons licenses.
-
-```{=html}
-<details>
-  <summary>Load video</summary>
-  <div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
-    <iframe loading="lazy" src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
-  </div>
-</details>
-```
-:::
-
-## Introduction to Copyright and Licenses
+## Short Introduction to Copyright and Licenses
 
@@
 without the author's express permission. 
-This even applies if the author makes it available to others (e.g., on their website):
-First and foremost others are not allowed to copy, modify, or share it.
-Importantly, this legal default of "all rights reserved" was created to benefit publishers,
-not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
 Copyright licenses enable authors to free up their works for reuse by others.
-
-Note that copyright law [varies between countries][commons-rules]
-and the rule of territoriality (_lex loci protectionis_) applies:
-Applicable are the laws of the country where a work is used.
-Also, if the work was created as part of the author's job,
-it might be their employer who holds the copyright to the work,
-depending on the country and contract.
-
-[commons-rules]: https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory
 
@@
 can and cannot do with a copyright-protected work, and under which conditions.
-The license may be exclusive, such that it is only granted to one recipient,
-effectively limiting the author in dealing with their own work.
-It can also be non-exclusive, thus giving rights to others without the author losing them.
-To avoid that every author who would like to share their work writes their own license,
-pre-formulated "boilerplate" licenses have been created.
-For example, you might have heard about ["Creative Commons" licenses][cc-licenses],
-which are used by Wikipedia.
-Recognizing them facilitates both sharing (on the side of the author)
-and re-using (on the side of the recipient).^[In fact, authors should not
-create their own license texts if the purpose is to share their work freely
-and openly with others.
-By using a boilerplate license others know what to expect,
-also because it may have already been [tested][sfconservancy] in [courts][cc-db].]
-
-[cc-licenses]: https://creativecommons.org/share-your-work/cclicenses/
-[sfconservancy]: https://sfconservancy.org/copyleft-compliance/
-[cc-db]: https://legaldb.creativecommons.org/
+We strongly recommend to use *existing boilerplate licenses*
+(and not to craft your own license text).
 
 ::: {.column-margin}
-For the purpose of this tutorial, by _license_ we mean _copyright license_.
+For the purpose of this chapter, by _license_ we mean _copyright license_.
 :::
 
-An important class of boilerplate licenses are _free/open_ licenses.
-Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified,
+Works under a free/open license "can be freely studied, applied, copied and/or modified,
 by anyone, for any purpose" [@Moller2015].
 Importantly, this also means that others do not need to ask or notify the author
-and that they can use it for commercial purposes.
-Coming back to the example of Creative Commons licenses,
-not all of them are free/open.
-For example, some restrict commercial use and modification.
-The Creative Commons license "[Attribution-ShareAlike 4.0][cc-by-sa]"
-is an example of a free/open license, however,
-and we will cover several others below.
+and that they can use it also for commercial purposes. If you modified the work and put a sufficient amount of creativity into it,
+the result is called a _derivative work_.
 
-[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/
+Even for free licenses, authors can place certain demands on re-users of their material.
+For example, they can (optionally) ...
 
-By the choice of license, authors can demand that they are credited,
-that the original license is indicated, that modifications are indicated,
-that derivative works are only shared under the same license,
-and that no further restrictions are imposed on the work.
-Software licenses may additionally require to make the source code available
-to everybody the software is shared with
-and often require to display the full text of the license upon distribution.
+- demand that they are credited as the creators of a work,
+- that the original license must be indicated when a work is reused, 
+- that modifications are indicated,
+- that derivative works are only shared under the same license, or
+- that no further restrictions are imposed on the derivative work.
 
-::: {#nte-terminology .callout-note collapse="true"}
-### A Note on Terminology
+We explain below how to choose a license that brings the desired "settings".
 
-The terms _free_ and _open_, especially with regard to software, come with a history.
-The Free Software Foundation (FSF) was founded in 1985
-to protect "four essential freedoms" [-@FSF2024] of a program's user.
-These are the freedoms to use, study, share, and improve a program.
-Software whose users legally and practically have these freedoms
-(because, among other things, they have access to its source code) is considered _free_.
-The four freedoms are seen as vital for a society as a whole
-in the sense that they enable sharing, cooperation and ultimately freedom in general.
-The FSF maintains [a list of software licenses][fsf-list]
-that it considers to be protecting the four freedoms.
-Sometimes the term _libre_ (Spanish and French for _free_)
-is used to make a distinction from _gratis_ software.
-You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).
-
-[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses
-
-The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach.
-It is concerned with developing high-quality software,
-for which everyone's ability to obtain, modify and contribute back the source code
-is considered beneficial.
-Access to the source code is one out of multiple conditions for software
-to be considered _open source_ by the OSI [-@OSI2007],
-which equally maintains [a list of approved licenses][osi-list].
-
-[osi-list]: https://opensource.org/licenses
-
-Access to the source code is a necessary, but not sufficient requirement
-both for _free software_ and for _open source software_.
-Conversely, both can (and frequently are) sold for money,
-as their respective criteria only apply once one has access to the software.
-Throughout this tutorial, we write "free/open license"
-to mean a license that is approved by either the FSF or the OSI.
-Software which is neither but makes available its source code
-is sometimes referred to as _source-available software_.
-
-Apart from software, the term _open access_ has often been used for works
-that are available at no cost.
-For example, this is the commonality of bronze, green, hybrid, gold,
-and diamond/platinum open access _articles_,
-which otherwise vary in the rights that are granted to readers.
-In 2002, the Budapest Open Access Initiative declared
-that _open access_ additionally includes the right to use articles for any purpose,
-and in 2003, the Bethesda Statement and the Berlin Declaration added
-the right to make derivative works.
-
-Two other notable definitions include the _Open Definition_ [@OKFN2016],
-which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015],
-for which the open editing phase began in 2006.
-They are largely viewed as compatible with one another.
-
-:::: {.column-margin}
-![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
-::::
-:::
-
-### Can I Share That Work?
+## Can I Share the Work of Others?
 
@@
 and shared publicly, the first step is to search for its license.
-For example, this tutorial has an [about page](about.qmd) detailing its licensing,
+For example, this tutorial has an [About](about.qmd) page detailing its licensing,
 whereas other websites sometimes state the license in the footer.
-Other places to look include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
-If a project has been stored in a research repository,
-yet another means of communicating its copyright status might exist.
-For media from Wikimedia Commons, an [Attribution Generator][commons-gen] exists.
 
-[commons-gen]: https://lizenzhinweisgenerator.de/?lang=en
-
-::: {#tip-private-use .callout-tip}
-#### Consider Complicated Licensing Situations
-
-For works by one or only few authors, it is possible
-that one license consistently applies to the whole content.
-More often than not, however, the situation is a little more complicated:
-
-- Different parts of a work may be available under different licenses:
-  A document may incorporate direct quotes or images by others.
-  In this case, the license of the document may not cover those.
-  In the best case, the authors indicate which parts of the work are covered by which license.
-- A work may be available under different licenses at the choice of the (re-)user.
-  This is also called multi-licensing.
-  For example, authors might allow to user a work under any version of a license.
-- In contrast, it may also be the case that one has to comply
-  with the conditions of multiple licenses at the same time.
-  This is the case, for example, if a derivative work is published
-  under a different license than the original work.
-:::
+Other places to look for license statements include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
+If a project has been stored in a research repository, sometimes the license is stored in the metadata of the project and displayed somewhere on the project's page.
 
 If you cannot find any license, the work might be
-in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law.
-For example, the work may not cross the [threshold of originality][commons-threshold],
-it may not [qualify for copyright protection][commons-unprotected]
--- such as official works by the government --,
-or the copyright protection might have expired.
-According to the [Berne convention][berne-convention],
-copyright expires at earliest 50 years after the author's death,
-but in [many countries][commons-pd] it expires even later
--- you can use a [public domain calculator][pd-calc] to check.
-If no license notice is present and the work is not in the public domain,
-it is _in copyright_ and you are generally not allowed to share it with others.
+in the **public domain** (see details [here](in_depth_material/introduction_copyright.qmd#public-domain)). In this case, you can simply reuse it.
+
+If no license notice is present and the work is *not* in the public domain,
+it is by default  _in copyright_ and you are generally not allowed to share it with others.
 You can, however, ask the author to put the corresponding work under a free/open license
 -- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.
-Without permission you can check you national laws
-for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_]
-Examples include the right to quote, [trivial use][commons-minimis],
-and the US-American [fair use][wiki-fair-use].
 
-[commons-threshold]: https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality
-[commons-unprotected]: https://commons.wikimedia.org/wiki/Commons:Unprotected_works
-[berne-convention]: https://en.wikipedia.org/wiki/Berne_Convention
-[commons-pd]: https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country
-[pd-calc]: https://lawflow.org/public-domain-calculators/
 [commons-templates]: https://commons.wikimedia.org/wiki/Commons:Email_templates
-[commons-minimis]: https://commons.wikimedia.org/wiki/Commons:De_minimis
-[wiki-fair-use]: https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?
 
-Having found a license, you need to determine
-(1) whether it is a standard (boilerplate) license,
-(2) whether it is a free/open,
-and (3) whether it is appropriate for the particular type of work and use case.
+Once you found a license, you need to determine whether it is a standard (boilerplate) license and
+whether it is a free/open. Then you can share the reused material in your repository.
 
-1. __Standard licenses__ are usually stated by their name (or an abbreviation),
-   which may include a version number.
-   Most standard licenses were assigned
-   an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification,
-   though the identifier is often not indicated for a particular work
-   (which would be best practice).
-   Also, not all standard licenses have SPDX identifiers.
-   You can view a [list of all licenses with SPDX identifiers here](https://spdx.org/licenses/).
-2. __Free/open__: We have compiled lists of the most common free/open licenses below,
-   categorized by the type of work they apply to.
-   Be sure to look for an exact match,
-   as there are often many licenses with very similar names.
-   If a particular license is not listed below, we also provide links to other resources
-   which you can use to determine whether a license is free/open.
-3. __Type of work and use case__:
-   Many licenses were written with a particular type of work in mind,
-   along with a certain use case.
-   Software licenses, for example, may consider that the source code
-   is the preferred form for making modifications
-   and that software can be encumbered by patents,
-   while licenses for data can differentiate between the database
-   and any works produced from it.
-   Take caution if a license applies to a type work for which it was not originally created
-   and conflicts with your use case.
-   Sometimes, the license name hints at the intended type of work,
-   but when in doubt always read the full license text (and/or consult with a lawyer)
-   to determine whether the conditions are acceptable to you.
+Here are typical licenses that are free/open:
+
+[osi-list]: https://opensource.org/licenses
+[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses
 
@@
   <summary>GNU Affero General Public License v3.0 (`AGPL-3.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>GNU General Public License v3.0 (`GPL-3.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>GNU Lesser General Public License v3.0 (`LGPL-3.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Apache License 2.0 (`Apache-2.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Mozilla Public License 2.0 (`MPL-2.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>MIT License (`MIT`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>BSD 3-Clause "New" or "Revised" License (`BSD-3-Clause`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>BSD 2-Clause "Simplified" License (`BSD-2-Clause`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>CC0 1.0 (`CC0-1.0`)</summary>
-  __Major conditions when sharing works__: (none)
+  __You can share under these conditions__: (none)
 
@@
   <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Open Data Commons Public Domain Dedication & License 1.0 (`PDDL-1.0`)</summary>
-  __Major conditions when sharing works__: (none)
+  __You can share under these conditions__: (none)
 
@@
   <summary>Community Data License Agreement Permissive 2.0 (`CDLA-Permissive-2.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Community Data License Agreement Permissive 1.0 (`CDLA-Permissive-1.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Community Data License Agreement Sharing 1.0 (`CDLA-Sharing-1.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Open Data Commons Open Database License v1.0 (`ODbL-1.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
   <summary>Open Data Commons Attribution License v1.0 (`ODC-By-1.0`)</summary>
-  __Major conditions when sharing works__:
+  __You can share under these conditions__:
 
@@
 
-::: {#cau-license-versions .callout-caution}
-### License Versions Are Important
-
-You may have noticed that in many cases a version number is part of the license name.
-This is because the organizations that created the licenses sometimes publish updated versions
-to accommodate for developments in copyright law and the communities that use the licenses.
-For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002][cc-v1].
-Since then, the possibility to relicense under later ([v2.0][cc-v2], 2004)
-and compatible licenses ([v3.0][cc-v3], 2007) has been added,
-a 30-day window to correct license violations has been established to combat [copyleft trolls][commons-trolls],
-and _sui generis_ database rights have been covered explicitly ([v4.0][cc-v4], 2013).
-There are many more [subtle differences between license versions][cc-versions],
-therefore it is important to indicate which license version exactly one is referring to,
-as the license of a work does not "update" automatically
-(i.e., you have to follow the terms of that exact version).
-CC0, officially introduced in 2009, is still at version [1.0][cc0-v1].
-
-[cc-v1]: https://creativecommons.org/weblog/entry/3476
-[cc-v2]: https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders
-[cc-v3]: https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included
-[cc-v4]: https://creativecommons.org/weblog/entry/40768
-[cc0-v1]: http://www.creativecommons.org/weblog/entry/13304
-[cc-versions]: https://wiki.creativecommons.org/wiki/License_Versions
-[commons-trolls]: https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling
-
-For the GNU licenses it is even recommended to state whether a work is licensed
-_only_ under the indicated version of the license or, alternatively,
-also under newer versions of the license [@Stallman2022Version].
-:::
-
 ### Under Which Conditions?
@@
 and share it with others -- regardless of whether you modified it.
+
 In the following, we will explain the most important conditions of the licenses listed above.
@@
 The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com][cal-appendix],
-which you can use to get a general idea of a license
-(but which does not obviate reading the license text in full).
+which you can use to get a general idea of a license.
 Specifically for software licenses the _Open Source Automation Development Lab_
@@
 to the collective as a whole rather than being mentioned individually.
-For example, the flexible attribution requirements of Creative Commons licenses
-allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009].
-And software projects may use copyright notices that only contain the project name,
+Software projects may use copyright notices that only contain the project name,
 such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].
@@
 and to provide all required information. 
-Only linking to a page that provides the attribution (as with Wikipedia articles)
-most likely will not suffice in general.
+For images, for example, this can be a small text next to the image.
 
-#### State Changes
+#### Modifications of the work
 
-If a work has been modified, licenses may require that this is indicated.
-For example, if an image under a Creative Commons license other than CC0
-is reused in a cropped version, reusers could add "adapted from original"
-to indicate that changes have occurred.
+If a work has been modified, some licenses may require that this is indicated (e.g., with a note "adapted from original"). 
+Other licenses forbid that a work is modified.
+For example, the `NoDerivatives` clause of the `CC-BY-ND-4.0` licenses say:
+
+> If you remix, transform, or build upon the material, you may not distribute the modified material.
+
+Hence, in this case you are free to reuse and redistribute the material - as long as you don't change it.
+
+#### Commercial reuse
+
+The NC clause in the Creative Commons BY-NC (Attribution–NonCommercial) license stands for "NonCommercial", meaning that others are allowed to use, share, adapt, and build upon the work -- but not for commercial purposes.
+This restriction prohibits any use that is primarily intended for or directed toward commercial advantage or monetary compensation.
+For example, using the work in an advertisement, selling it, or incorporating it into a product for sale would not be permitted without additional permission from the copyright holder.
+However, non-profit educational use, personal projects, or sharing within a non-commercial community is generally allowed, as long as proper attribution is given to the original creator.
 
@@
 or they even require to copy the full license text along with the used work ("Add license text").
-Of course, the latter may not be feasible
-if the work is printed and smaller than the license text.
-For example, if the license of an image requires to print the full license text,
-the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].
-
-#### Disclose Code
-
-There are various situations in which one interacts with software
-without having access to the underlying source code.
-For example, software can be distributed as a binary ("executable") file
-or it can be used over the internet.
-In both cases, licenses may require that the source code is provided to its users.
-
-#### Anti-DRM
-
-You may have encountered images or videos which you could not download due to a copy protection.
-Such a protection is also called **d**igital **r**ights **m**anagement (DRM).
-For works published under a license with an anti-DRM clause,
-reusers may not apply DRM to the work or must provide a DRM-free version in parallel
-(depending on the exact license).
 
@@
 
-If you modified the work and put a sufficient amount of creativity into it,
-the result is called a _derivative work_.
-It is (again) protected by copyright and even if you share it,
-nobody except you is allowed to copy, modify, or distribute it further.
-You can put it under a free/open license but don't have to.[^still-abide]
-The only exception is if the original work has been made available under a "copyleft" license.
-In that case, if you share the work with others
-you must put it under the same license as the original work[^or-compatible]
--- or you commit a copyright violation.[^not-automatic]
+If the original work has been made available under a "copyleft" license, 
+any derivative work that is shared with others
+must be put under the same license as the original work[^or-compatible]
+-- or you commit a copyright violation.
+
 This explains the origin of the term _copyleft_,
-as it "turns copyright around" to make works permanently free: ©\ →\ 🄯.
-In the realm of Creative Commons licenses, copyleft is also called _share alike_.
-The opposite of a copyleft license is sometimes also called a _permissive_ license.
-
-[^still-abide]: Of course, you still need to abide
-by the license conditions of the original work, such as attribution.
+as it "turns copyright around" to make works permanently free.
+In the realm of Creative Commons licenses, copyleft is also called _share alike_ (e.g. in the `CC-BY-SA-4.0` license).
 
 [^or-compatible]: or sometimes also under a later version or under a compatible license
-
-[^not-automatic]: Note, however, that [copyleft licenses are not automatic][commons-automatic]:
-A derivative work of a copylefted work is not "implicitly" under the same license
--- others would commit a copyright violation if they just took it.
-Only if the derivative author explicitly makes it available under the same license,
-others may use the derivative work under the same terms.
-
-[commons-automatic]: https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic
 
@@
 Also note that the copyleft licenses we discuss here do not mandate sharing.
-Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015].
+Copyleft (and attribution) clauses are only triggered *if* the work is shared [@CC2015].
 This means that if you only use a work internally, you do not need to share your derivative works.
-
-::: {#nte-weak-strong-copyleft .callout-note collapse="true"}
-##### Weak and Strong Copyleft
-
-For software, copyleft licenses come in two flavors:
-
-- __Weak copyleft licenses__ only require that modifications to the software itself
-  are licensed under the same or a compatible license if shared.
-  For example, if you create and publish software under a weak copyleft license,
-  others who modify it and put their version on the internet have to apply the same license.
-  However, people who merely _use_ your software in their own work which they make publicly available
-  can choose any license for it.
-- __Strong copyleft licenses,__ on the other side, insist that any larger works
-  that use the copyleft-licensed software must also be licensed
-  under the same or a compatible license if shared.
-  For example, if your software were to be put under a strong copyleft license,
-  everybody publishing software that uses your software would need to put it under the same license.
-  Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024GPL].
-:::
-
-### Further Remarks
-
-The description of the conditions above is a simplified summary
--- read the full text of the license before sharing the respective work with others.
-For example, Creative Commons licenses other than CC0 also require that
-
-- you retain a notice about the disclaimer of warranties,
-- you retain an indication of any previous modifications,
-- the attribution does not imply endorsement by the original author, and
-- you remove the attribution again upon request by the original author.
 
@@
 
-- [ ] Author name
+- [ ] Original author name
 - [ ] Copyright notice
@@
 it is better to either follow it or to ask the author for clarification or explicit permission.
-This is especially important given the existence of [_copyleft trolls_][commons-trolls]:
-People who distribute works under a free/open license
-but send out statutory damages claims after minor violations
-without providing reusers an opportunity to fix their error.
-In addition, in many European countries attribution may be necessary regardless
-due to the author's moral rights.
-For these reasons, we recommend that you provide attribution
-for all works under a free/open license
+
+We recommend that you always provide attribution for all works under a free/open license
 -- even if the license does not strictly require it (e.g., CC0).
-For this reason, you need to make note of every foreign work
-included in your project folder and record its license.
+This also follows from the rules of good scientific practice (and not only from copyright law).
 
@@
 Also, their commercial use may require the consent of the depicted person.
-Finally, note that even if attribution is not a requirement of the license,
-good scientific practice demands that appropriate citations are made.
 :::
@@
 
-## Adding an Image
+## Practical Exercise: Adding an Image
 
@@
 2. Is it free/open?
-3. Is it appropriate for the type of work and use case?
-4. Under which conditions may the work be shared?
+3. Under which conditions may the work be shared?
 
@@
 2. The license is free/open
-3. The license is appropriate for images
-4. The license contains no (major) conditions
+3. The license contains no (major) conditions
 :::
@@
 
-You can now download the image[^image-resolution]
+You can now download the image
 and move it to your project folder (maybe call it `penguin.jpg`).
@@
 and add the image instead, including proper attribution.
-
-[^image-resolution]: If you are unsure which image resolution (size) is required,
-just download it in the largest possible resolution so it can be printed.
 
@@
 
+- Our own [in-depth material on copyright and licenses](in_depth_material/introduction_copyright.qmd) (which is an extension of this chapter)
 - "Open Content -- A Practical Guide to Using Creative Commons Licences"

### 2025-07-14 in_depth_material/introduction_copyright.qmd
Shorten chapter on copyright/ transfer to Appendix

---
title: "Introduction to Copyright and Licensing"
engine: knitr
format:
  html:
    include-in-header:
      - text: |
          <style>
          .unstyled-list > ul {
              list-style-type: none;
              padding-left: 1rem;
          }
          .unstyled-list summary {
              cursor: pointer;
          }
          </style>
---

By default, everything you put into the project folder will be shared publicly.
In many instances, this will also include works by others than yourself or your co-authors,
because you may use existing data, images or software.
In these cases, you should make sure to only share them to the extent you are allowed to.
In the following, we will be focusing on copyright,
but of course, there can also be other restrictions.

::: {#nte-copyright-tldr .callout-note}
## Copyright TL;DR

Every expression of an idea, such as a literary or an artistic works,
is automatically covered by copyright.
Put simply, this means that nobody else than the copyright holder
is allowed to copy, modify, or share it.
Copyright licenses can grant others some of the necessary rights,
but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
Certain exceptions and limitations to copyright (such as fair use or the right to quote)
may obviate the need for a license.

With regard to copyright, free/open licenses can give permission to share a work by others.
These licenses often require to attribute the authors,
indicate whether changes were made,
and provide the text of the license, among other things.

The following video by @Brinken2021
embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy))
explains the correct attribution of works under Creative Commons licenses.

```{=html}
<details>
  <summary>Load video</summary>
  <p><div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
    <iframe loading="lazy" src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
  </div></p>
  <p>The video “The correct indication of a license: Aspects to consider when reusing the work of others” by Helene Brinken, Jonas Hauss, and Jessika Rücknagel is licensed under <a href="https://creativecommons.org/licenses/by/3.0/de/deed.de">CC&nbsp;BY&nbsp;3.0&nbsp;DE</a>.</p>
</details>
```
:::

## Introduction to Copyright and Licenses

Whenever authors create a literary or artistic work
(such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_]
in most countries limits other people from copying, modifying, and sharing it
without the author's express permission.
This even applies if the author makes it available to others (e.g., on their website):
First and foremost others are not allowed to copy, modify, or share it.
Importantly, this legal default of "all rights reserved" was created to benefit publishers,
not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
Copyright licenses enable authors to free up their works for reuse by others.

Note that copyright law [varies between countries][commons-rules]
and the rule of territoriality (_lex loci protectionis_) applies:
Applicable are the laws of the country where a work is used.
Also, if the work was created as part of the author's job,
it might be their employer who holds the copyright to the work,
depending on the country and contract.

[commons-rules]: https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory

A license is a legal document that regulates what other people than the author
can and cannot do with a copyright-protected work, and under which conditions.
The license may be exclusive, such that it is only granted to one recipient,
effectively limiting the author in dealing with their own work.
It can also be non-exclusive, thus giving rights to others without the author losing them.
To avoid that every author who would like to share their work writes their own license,
pre-formulated "boilerplate" licenses have been created.
For example, you might have heard about ["Creative Commons" licenses][cc-licenses],
which are used by Wikipedia.
Recognizing them facilitates both sharing (on the side of the author)
and re-using (on the side of the recipient).^[In fact, authors should not
create their own license texts if the purpose is to share their work freely
and openly with others.
By using a boilerplate license others know what to expect,
also because it may have already been [tested][sfconservancy] in [courts][cc-db].]

[cc-licenses]: https://creativecommons.org/share-your-work/cclicenses/
[sfconservancy]: https://sfconservancy.org/copyleft-compliance/
[cc-db]: https://legaldb.creativecommons.org/

::: {.column-margin}
For the purpose of this tutorial, by _license_ we mean _copyright license_.
:::

An important class of boilerplate licenses are _free/open_ licenses.
Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified,
by anyone, for any purpose" [@Moller2015].
Importantly, this also means that others do not need to ask or notify the author
and that they can use it for commercial purposes.
Coming back to the example of Creative Commons licenses,
not all of them are free/open.
For example, some restrict commercial use and modification.
The Creative Commons license "[Attribution-ShareAlike 4.0][cc-by-sa]"
is an example of a free/open license, however,
and we will cover several others below.

[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/

By the choice of license, authors can demand that they are credited,
that the original license is indicated, that modifications are indicated,
that derivative works are only shared under the same license,
and that no further restrictions are imposed on the work.
Software licenses may additionally require to make the source code available
to everybody the software is shared with
and often require to display the full text of the license upon distribution.

::: {#nte-terminology .callout-note collapse="true"}
### A Note on Terminology

The terms _free_ and _open_, especially with regard to software, come with a history.
The Free Software Foundation (FSF) was founded in 1985
to protect "four essential freedoms" [-@FSF2024] of a program's user.
These are the freedoms to use, study, share, and improve a program.
Software whose users legally and practically have these freedoms
(because, among other things, they have access to its source code) is considered _free_.
The four freedoms are seen as vital for a society as a whole
in the sense that they enable sharing, cooperation and ultimately freedom in general.
The FSF maintains [a list of software licenses][fsf-list]
that it considers to be protecting the four freedoms.
Sometimes the term _libre_ (Spanish and French for _free_)
is used to make a distinction from _gratis_ software.
You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).

[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses

The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach.
It is concerned with developing high-quality software,
for which everyone's ability to obtain, modify and contribute back the source code
is considered beneficial.
Access to the source code is one out of multiple conditions for software
to be considered _open source_ by the OSI [-@OSI2007],
which equally maintains [a list of approved licenses][osi-list].

[osi-list]: https://opensource.org/licenses

Access to the source code is a necessary, but not sufficient requirement
both for _free software_ and for _open source software_.
Conversely, both can (and frequently are) sold for money,
as their respective criteria only apply once one has access to the software.
Throughout this tutorial, we write "free/open license"
to mean a license that is approved by either the FSF or the OSI.
Software which is neither but makes available its source code
is sometimes referred to as _source-available software_.

Apart from software, the term _open access_ has often been used for works
that are available at no cost.
For example, this is the commonality of bronze, green, hybrid, gold,
and diamond/platinum open access _articles_,
which otherwise vary in the rights that are granted to readers.
In 2002, the Budapest Open Access Initiative declared
that _open access_ additionally includes the right to use articles for any purpose,
and in 2003, the Bethesda Statement and the Berlin Declaration added
the right to make derivative works.

Two other notable definitions include the _Open Definition_ [@OKFN2016],
which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015],
for which the open editing phase began in 2006.
They are largely viewed as compatible with one another.

:::: {.column-margin}
![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
::::
:::

### Can I Share That Work?

To determine whether a particular work can be included in the project folder
and shared publicly, the first step is to search for its license.
For example, this tutorial has an [about page](../about.qmd) detailing its licensing,
whereas other websites sometimes state the license in the footer.
Other places to look include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
If a project has been stored in a research repository,
yet another means of communicating its copyright status might exist.
For media from Wikimedia Commons, an [Attribution Generator][commons-gen] exists.

[commons-gen]: https://lizenzhinweisgenerator.de/?lang=en

::: {#tip-private-use .callout-tip}
#### Consider Complicated Licensing Situations

For works by one or only few authors, it is possible
that one license consistently applies to the whole content.
More often than not, however, the situation is a little more complicated:

- Different parts of a work may be available under different licenses:
  A document may incorporate direct quotes or images by others.
  In this case, the license of the document may not cover those.
  In the best case, the authors indicate which parts of the work are covered by which license.
- A work may be available under different licenses at the choice of the (re-)user.
  This is also called multi-licensing.
  For example, authors might allow to use a work under any version of a license.
- In contrast, it may also be the case that one has to comply
  with the conditions of multiple licenses at the same time.
  This is the case, for example, if a derivative work is published
  under a different license than the original work.
:::

#### Public domain

If you cannot find any license, the work might be
in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law.
For example, the work may not cross the [threshold of originality][commons-threshold],
it may not [qualify for copyright protection][commons-unprotected]
-- such as official works by the government --,
or the copyright protection might have expired.
According to the [Berne convention][berne-convention],
copyright expires at earliest 50 years after the author's death,
but in [many countries][commons-pd] it expires even later
-- you can use a [public domain calculator][pd-calc] to check.
If no license notice is present and the work is not in the public domain,
it is _in copyright_ and you are generally not allowed to share it with others.
You can, however, ask the author to put the corresponding work under a free/open license
-- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.
Without permission you can check your national laws
for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_]
Examples include the right to quote, [trivial use][commons-minimis],
and the US-American [fair use][wiki-fair-use].

[commons-threshold]: https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality
[commons-unprotected]: https://commons.wikimedia.org/wiki/Commons:Unprotected_works
[berne-convention]: https://en.wikipedia.org/wiki/Berne_Convention
[commons-pd]: https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country
[pd-calc]: https://lawflow.org/public-domain-calculators/
[commons-templates]: https://commons.wikimedia.org/wiki/Commons:Email_templates
[commons-minimis]: https://commons.wikimedia.org/wiki/Commons:De_minimis
[wiki-fair-use]: https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?

Having found a license, you need to determine
(1) whether it is a standard (boilerplate) license,
(2) whether it is a free/open,
and (3) whether it is appropriate for the particular type of work and use case.

1. __Standard licenses__ are usually stated by their name (or an abbreviation),
   which may include a version number.
   Most standard licenses were assigned
   an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification,
   though the identifier is often not indicated for a particular work
   (which would be best practice).
   Also, not all standard licenses have SPDX identifiers.
   You can view a [list of all licenses with SPDX identifiers here](https://spdx.org/licenses/).
2. __Free/open__: We have compiled lists of the most common free/open licenses below,
   categorized by the type of work they apply to.
   Be sure to look for an exact match,
   as there are often many licenses with very similar names.
   If a particular license is not listed below, we also provide links to other resources
   which you can use to determine whether a license is free/open.
3. __Type of work and use case__:
   Many licenses were written with a particular type of work in mind,
   along with a certain use case.
   Software licenses, for example, may consider that the source code
   is the preferred form for making modifications
   and that software can be encumbered by patents,
   while licenses for data can differentiate between the database
   and any works produced from it.
   Take caution if a license applies to a type work for which it was not originally created
   and conflicts with your use case.
   Sometimes, the license name hints at the intended type of work,
   but when in doubt always read the full license text (and/or consult with a lawyer)
   to determine whether the conditions are acceptable to you.

::: panel-tabset
### Software

For a software license to be considered free/open,
it must be approved by either the [Open Source Initiative][osi-list] (OSI)
or the [Free Software Foundation][fsf-list] (FSF).
Some of the most common free/open software licenses are listed below:

:::: unstyled-list
- <details>
  <summary>GNU Affero General Public License v3.0 (`AGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/agpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU General Public License v3.0 (`GPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/gpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU Lesser General Public License v3.0 (`LGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.gnu.org/licenses/lgpl-3.0.html#license-text)
  </details>
- <details>
  <summary>Apache License 2.0 (`Apache-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes

  [Link to full text](https://www.apache.org/licenses/LICENSE-2.0)
  </details>
- <details>
  <summary>Mozilla Public License 2.0 (`MPL-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.mozilla.org/en-US/MPL/2.0/)
  </details>
- <details>
  <summary>MIT License (`MIT`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/mit/#license-text)
  </details>
- <details>
  <summary>BSD 3-Clause "New" or "Revised" License (`BSD-3-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-3-clause/#license-text)
  </details>
- <details>
  <summary>BSD 2-Clause "Simplified" License (`BSD-2-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-2-clause/#license-text)
  </details>
::::

If you did not find a particular software license in this list
and consult the resources by the OSI and/or FSF linked above,
be cautious because they also approve special-purpose licenses
that are unfit for software in the general case.
For example, the license CC0\ 1.0 expressively does not grant patent rights
and one should only use software under that license if it is most likely not covered by patents.

### Writing, Images, Audio, Video

Many authors of text and media put their work under a license created by Creative Commons.
Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use
or the distribution of modified works (such as sharing a translated version).
Therefore, we recommend to only use works made available to you under CC0
or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`),
but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`).
In their most recent version this includes the following licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

[cc0-text]: https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title
[cc-by-text]: https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title
[cc-by-sa-text]: https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title

Of course, if you encounter a Creative Commons license with a previous version
that satisfies the description above, it can be also considered free/open.
If you encounter a license not from Creative Commons and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for text and media.
For example, licenses for software or for documentation commonly require
to distribute a work together with the full license text,
which is impractical for usage in print media.

[commons-tags]: https://commons.wikimedia.org/wiki/Commons:Copyright_tags

### Data

Purely factual data such as measurements are usually not copyrightable,
but literary or artistic works that cross the threshold of originality are.
Additionally, in some jurisdictions, collections of data (i.e., databases)
can be subject to _sui generis_ database rights
which prevent extracting substantial parts of a database.

As a consequence, a database may be licensed differently than its content.
For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists,
but the license for the database as a whole might have been chosen by the curator,
including the selection of entries, field names and any _sui generis_ database rights.
Watch out for any signs that the license for the database and its content are not the same.
In the following, we list some popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

Note that Creative Commons licenses prior to version 4.0 are not suitable for databases.
You might also encounter the following, less popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>Open Data Commons Public Domain Dedication & License 1.0 (`PDDL-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text](https://opendatacommons.org/licenses/pddl/1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 2.0 (`CDLA-Permissive-2.0`)</summary>
  __Major conditions when sharing works__:

  - Add license text

  [Link to full text](https://cdla.dev/permissive-2-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 1.0 (`CDLA-Permissive-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://cdla.dev/permissive-1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Sharing 1.0 (`CDLA-Sharing-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Copyleft

  [Link to full text](https://cdla.dev/sharing-1-0/)
  </details>
::::

Finally, you should know about the following two free/open licenses because they only cover the database,
but not its individual contents or the works produced from it.
If any of these licenses is used, the license for the content must be stated separately.

:::: unstyled-list
- <details>
  <summary>Open Data Commons Open Database License v1.0 (`ODbL-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Anti-DRM
  - Copyleft

  [Link to full text](https://opendatacommons.org/licenses/odbl/1-0/)
  </details>
- <details>
  <summary>Open Data Commons Attribution License v1.0 (`ODC-By-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://opendatacommons.org/licenses/by/1-0/)
  </details>
::::

If you encounter a license not listed here and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for data,
because they do not mention _sui generis_ database rights.
:::

::: {#cau-license-versions .callout-caution}
### License Versions Are Important

You may have noticed that in many cases a version number is part of the license name.
This is because the organizations that created the licenses sometimes publish updated versions
to accommodate for developments in copyright law and the communities that use the licenses.
For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002][cc-v1].
Since then, the possibility to relicense under later ([v2.0][cc-v2], 2004)
and compatible licenses ([v3.0][cc-v3], 2007) has been added,
a 30-day window to correct license violations has been established to combat [copyleft trolls][commons-trolls],
and _sui generis_ database rights have been covered explicitly ([v4.0][cc-v4], 2013).
There are many more [subtle differences between license versions][cc-versions],
therefore it is important to indicate which license version exactly one is referring to,
as the license of a work does not "update" automatically
(i.e., you have to follow the terms of that exact version).
CC0, officially introduced in 2009, is still at version [1.0][cc0-v1].

[cc-v1]: https://creativecommons.org/weblog/entry/3476
[cc-v2]: https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders
[cc-v3]: https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included
[cc-v4]: https://creativecommons.org/weblog/entry/40768
[cc0-v1]: https://www.creativecommons.org/weblog/entry/13304
[cc-versions]: https://wiki.creativecommons.org/wiki/License_Versions
[commons-trolls]: https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling

For the GNU licenses it is even recommended to state whether a work is licensed
_only_ under the indicated version of the license or, alternatively,
also under newer versions of the license [@Stallman2022Version].
:::

### Under Which Conditions?

Finally, if you verified that the work has been made available to you under a free/open license,
you need to understand the conditions under which you are allowed to include the work in your project folder
and share it with others -- regardless of whether you modified it.
In the following, we will explain the most important conditions of the licenses listed above.
For other licenses we recommend reading their text in full and/or consulting with a lawyer.
The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com][cal-appendix],
which you can use to get a general idea of a license
(but which does not obviate reading the license text in full).
Specifically for software licenses the _Open Source Automation Development Lab_
provides [useful checklists detailing the respective obligations][osadl].[^osadl-note]

[cal-appendix]: https://choosealicense.com/appendix/
[osadl]: https://www.osadl.org/Checklists/
[^osadl-note]: Accessing the checklists requires a (free) account on their website.

::: {#nte-private-use .callout-note}
#### Private Use

In the following, we only discuss the conditions under which works may be _shared_.
This is because free/open licenses put no conditions on private use,
where the work is not made available to others.
:::

#### Attribution

Licenses may require that you provide attribution to the author(s)[^others]
and to retain any copyright notices,[^example-notice]
although the specifics vary between licenses.
For example, Creative Commons licenses other than CC0 require
that the name of the author, the title of the work, and a link to the work are provided.

[^others]: or others whose names are provided with the original work
[^example-notice]: Such as `Copyright (c) 2025 Jane Doe`

Projects with many authors often explicitly allow for an attribution
to the collective as a whole rather than being mentioned individually.
For example, the flexible attribution requirements of Creative Commons licenses
allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009].
And software projects may use copyright notices that only contain the project name,
such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].

Where to put the attribution depends on the type of the work,
but in general it is best to keep it close to where the work is used
and to provide all required information.
Only linking to a page that provides the attribution (as with Wikipedia articles)
most likely will not suffice in general.

#### State Changes

If a work has been modified, licenses may require that this is indicated.
For example, if an image under a Creative Commons license other than CC0
is reused in a cropped version, reusers could add "adapted from original"
to indicate that changes have occurred.

#### Indicate License / Add License Text

Licenses may either require that one mentions the applicable license
by name and possibly link ("Indicate license")
or they even require to copy the full license text along with the used work ("Add license text").
Of course, the latter may not be feasible
if the work is printed and smaller than the license text.
For example, if the license of an image requires to print the full license text,
the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].

#### Disclose Code

There are various situations in which one interacts with software
without having access to the underlying source code.
For example, software can be distributed as a binary ("executable") file
or it can be used over the internet.
In both cases, licenses may require that the source code is provided to its users.

#### Anti-DRM

You may have encountered images or videos which you could not download due to a copy protection.
Such a protection is also called **d**igital **r**ights **m**anagement (DRM).
For works published under a license with an anti-DRM clause,
reusers may not apply DRM to the work or must provide a DRM-free version in parallel
(depending on the exact license).

#### Copyleft

If you modified the work and put a sufficient amount of creativity into it,
the result is called a _derivative work_.
It is (again) protected by copyright and even if you share it,
nobody except you is allowed to copy, modify, or distribute it further.
You can put it under a free/open license but don't have to.[^still-abide]
The only exception is if the original work has been made available under a "copyleft" license.
In that case, if you share the work with others
you must put it under the same license as the original work[^or-compatible]
-- or you commit a copyright violation.[^not-automatic]
This explains the origin of the term _copyleft_,
as it "turns copyright around" to make works permanently free: ©\ →\ 🄯.
In the realm of Creative Commons licenses, copyleft is also called _share alike_.
The opposite of a copyleft license is sometimes also called a _permissive_ license.

[^still-abide]: Of course, you still need to abide
by the license conditions of the original work, such as attribution.

[^or-compatible]: or sometimes also under a later version or under a compatible license

[^not-automatic]: Note, however, that [copyleft licenses are not automatic][commons-automatic]:
A derivative work of a copylefted work is not "implicitly" under the same license
-- others would commit a copyright violation if they just took it.
Only if the derivative author explicitly makes it available under the same license,
others may use the derivative work under the same terms.

[commons-automatic]: https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic

What is considered creation of a derivative work varies between jurisdictions,
but the following are typical examples:
translating a text into a different language,
taking a picture of an artwork,
adding a song to a video, adapting a computer program to fit own use cases,
or taking a [screenshot of a computer program][commons-screenshots].
In all of these cases, if the original work is under a copyleft license,
then sharing of the resulting derivative work is only allowed
if it is put under the same license.

[commons-screenshots]: https://commons.wikimedia.org/wiki/Commons:Screenshots

The question is then: Which uses of a work do not create a derivative work?
In many cases, if an unaltered excerpt "is used to illuminate an idea
or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made.
The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p. 44]
and not be "merged with other material into a new and larger work" [@Kreutzer2024, p. 45].
For example, embedding images in a document and clearly separating them from the rest
or assembling works as part of a collection typically would not create a derivative work.
However, due to a lack of case law, there is only little guidance
on what constitutes a derivative work in general.

Also note that the copyleft licenses we discuss here do not mandate sharing.
Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015].
This means that if you only use a work internally, you do not need to share your derivative works.

::: {#nte-weak-strong-copyleft .callout-note collapse="true"}
##### Weak and Strong Copyleft

For software, copyleft licenses come in two flavors:

- __Weak copyleft licenses__ only require that modifications to the software itself
  are licensed under the same or a compatible license if shared.
  For example, if you create and publish software under a weak copyleft license,
  others who modify it and put their version on the internet have to apply the same license.
  However, people who merely _use_ your software in their own work which they make publicly available
  can choose any license for it.
- __Strong copyleft licenses,__ on the other side, insist that any larger works
  that use the copyleft-licensed software must also be licensed
  under the same or a compatible license if shared.
  For example, if your software were to be put under a strong copyleft license,
  everybody publishing software that uses your software would need to put it under the same license.
  Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024GPL].
:::

### Further Remarks

The description of the conditions above is a simplified summary
-- read the full text of the license before sharing the respective work with others.
For example, Creative Commons licenses other than CC0 also require that

- you retain a notice about the disclaimer of warranties,
- you retain an indication of any previous modifications,
- the attribution does not imply endorsement by the original author, and
- you remove the attribution again upon request by the original author.

::: {#tip-cc-licenses-checklist .callout-tip}
#### Required Information under Creative Commons Licenses

For most Creative Commons licenses,
providing the following information is required (if available):

- [ ] Author name
- [ ] Copyright notice
- [ ] Title of the work
  - [ ] Link to the work
- [ ] License name
  - [ ] Link to the license
- [ ] Link to warranty disclaimer
- [ ] Indicate modification
:::

Obviously, if you are in doubt whether a condition of the license applies in your use case,
it is better to either follow it or to ask the author for clarification or explicit permission.
This is especially important given the existence of [_copyleft trolls_][commons-trolls]:
People who distribute works under a free/open license
but send out statutory damages claims after minor violations
without providing reusers an opportunity to fix their error.
In addition, in many European countries attribution may be necessary regardless
due to the author's moral rights.
For these reasons, we recommend that you provide attribution
for all works under a free/open license
-- even if the license does not strictly require it (e.g., CC0).
Also, you should take note of every foreign work
included in your project folder and record its license.

::: {#imp-non-copyright-restrictions .callout-important}
#### Non-copyright Restrictions

The free/open licenses we discuss here mostly grant copyrights
(with some of them explicitly granting patent rights and _sui generis_ database rights).
Therefore, you may lack other rights such as personality, privacy, moral, or trademark rights.
For example, [sharing photos that depict people][commons-photographs]
is not only a matter of copyright, but also of privacy rights.
Also, their commercial use may require the consent of the depicted person.
Finally, note that even if attribution is not a requirement of the license,
good scientific practice demands that appropriate citations are made.
:::

[commons-photographs]: https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people

## Wrap-up

If you would like to learn more about copyright and licenses
you might find the following resources interesting:

- "Open Content -- A Practical Guide to Using Creative Commons Licences"
  by @Kreutzer2024
- "Creative Commons Certificate for Educators, Academic Librarians,
  and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
- "Rechtsfragen bei Open Science: Ein Leitfaden" by @Kreutzer2021

### 2025-07-15 data.qmd
Shorten the data chapter

@@
 ---
-title: "Add Data"
+title: "Add Data and Data Dictionary"
 engine: knitr
@@
 stopifnot(packageDescription("palmerpenguins", fields = "License") == "CC0")
+
+# write both in the top folder and in the in-depth folder
+# (so that both scripts can reference the data without relative paths)
 write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
+write.csv(palmerpenguins::penguins, file = "in_depth_material/data.csv", row.names = FALSE)
 ```
@@
 
-[data.csv](data.csv){.btn .btn-lg .btn-info download=""}
+[data.csv](data.csv){.btn .btn-lg .btn-info download="data.csv"}
 
@@
 
-## Add Data Dictionary
+## Add a Data Dictionary
 
@@
 This is typically done with a _data dictionary_ (also called a _codebook_).
-In the following, we will demonstrate how to create a simple data dictionary
-using the R package [`datawizard`][datawizard].
-You can install them now using:
 
-[datawizard]: https://easystats.github.io/datawizard/
+The recommendation for data dictionaries vary between fields - 
+both in terms of the recommended content (i.e., what exactly should be documented) 
+and the technical implementation (i.e., which file formats should be used).
 
-```{.r filename="Console"}
-renv::install("datawizard")
-```
+*For the purpose of this exercise, we keep it easy and propose that you 
+manually create a file with the data dictionary
+(e.g., as a table in .xlsx, .docx, .ods, or as a Markdown table), 
+documenting only the bare minimum.*
 
-You can put the code that follows in a separate document.
-Create it by clicking on _File_ > _New File_  > _Quarto Document..._.
-Choose a title such as `Data Dictionary`,
-select _HTML_ as format,
-uncheck the use of the visual markdown editor, and click on _Create_.
-Remove everything except the YAML header (between the `---`).
-To make the HTML file self-contained,
-also set `embed-resources: true` such that the YAML header looks as follows:
+<!-- TODO: Add a link to the forthcoming data dictionary OSC tutorial -->
 
-```{.yml filename="data_dictionary.qmd"}
----
-title: "Data Dictionary"
-format:
-  html:
-    embed-resources: true
----
-```
+::: {#tip-in-depth .callout-tip collapse="false"}
+### Manual is too much work? Automatic generation of (machine-readable) data dictionaries
+Our in-depth supplementary material "[Automatic Generation of Data Dictionaries](in_depth_material/data_dic_generation.qmd)" explains how you can automatically create
+a data dictionary with an R package. The package reads the data set and extracts relevant information from it. 
+This approach is in particular useful if you need to document data sets with many variables.
 
-Then, save it as `data_dictionary.qmd` by clicking on _File_ > _Save_.
-
-To create the actual data dictionary, first write a description for all columns
-so others can understand what the variable names mean.
-Where necessary, also document their value
--- this is especially important if their meaning is non-obvious.
-In the following, we demonstrate this by storing the penguins' binomial name
-along with the English name.
-
-``````{cat}
-#| engine.opts: { file: "_data_dictionary.qmd" }
-#| class.source: "md"
-#| filename: "data_dictionary.qmd"
-
-```{r}
-#| echo: false
-
-# Store the description of variables
-vars <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-
-# Store the description of variable values
-vals <- list(
-  species = c(
-    Adelie = "Pygoscelis adeliae",
-    Gentoo = "Pygoscelis papua",
-    Chinstrap = "Pygoscelis antarcticus"
-  )
-)
-```
-``````
-
-Then, load the data and use `datawizard`
-to add the descriptions to the `data.frame`:[^not-permanent]
-
-[^not-permanent]: Note that the code provided does not alter the data file
--- no description will be added to `data.csv`.
-The descriptions are only added to a (temporary) copy of the data set within R
-to create the data dictionary.
-
-::: {.column-margin}
-![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
+This advanced chapter also contains a section on how to create machine-readable data dictionaries.
 :::
 
-``````{cat}
-#| engine.opts: { file: "_data_dictionary.qmd", append: TRUE }
-#| class.source: "md"
-#| filename: "data_dictionary.qmd"
+### A bare minimum data dictionary
 
-```{r}
-#| echo: false
+Most standards for data dictionaries require at least 
+this information for each variable in your data set:
 
-dat <- read.csv("data.csv")
+- `name`: The (machine‑readable) name of the variable  
+- `label`: A short, human‑readable title or label for the variable  
+- `type`: The data type of the variable (e.g., `integer`, `float`, `string`, `date`)  
+- `description`: A brief description of what the variable measures or represents  
+- `values` (for categorical variables): A mapping of codes to their meanings, for example:  1 = Male, 2 = Female, 9 = Missing)
+- `units` (for numeric measures): The units of measurement (e.g., `kg`, `USD`, `years`, or the scale of a survey response), 
+- `missing_codes`: Any special codes used to denote missing or non‑applicable values (e.g., `-99 = Not answered`)
 
-for (x in names(vars)) {
-  if (x %in% names(vals)) {
-    dat <- datawizard::assign_labels(
-      dat,
-      select = I(x),
-      variable = vars[[x]],
-      values = vals[[x]]
-    )
-  } else {
-    dat <- datawizard::assign_labels(
-      dat,
-      select = I(x),
-      variable = vars[[x]]
-    )
-  }
-}
-```
+Here's an example from a different data set, with variables in rows, and the dictionary in columns:
 
-``````
+| name              | label                     | type    | description                             | values                                                                                  | units       | missing_codes      |
+| ----------------- | ------------------------- | ------- | --------------------------------------- | --------------------------------------------------------------------------------------- | ----------- | ------------------ |
+| gender            | Gender                    | integer | self‑identified gender                  | 1 = Male; 2 = Female; 3 = Other; 9 = Missing                                            |             | 9 = Missing        |
+| age               | Age                       | integer | age in years                            |                                                                                         | years       | -99 = Not answered |
+| blood_pressure    | Blood Pressure (systolic) | integer | systolic blood pressure                 |                                                                                         | mmHg        | -99 = Not measured |
+| life_satisfaction | Life Satisfaction         | integer | "How satisfied are you with your life?" | 1 = Very dissatisfied; 2 = Dissatisfied; 3 = Neutral; 4 = Satisfied; 5 = Very satisfied | scale (1–5) | -99 = Not answered |
 
-Then, you can create the data dictionary containing the descriptions,
-but also some other information about each variable
-(e.g., the number of missing values) and print it.
+### ✍️ Practical Exercise: Add your own data dictionary
 
-``````{cat}
-#| engine.opts: { file: "_data_dictionary.qmd", append: TRUE }
-#| class.source: "md"
-#| filename: "data_dictionary.qmd"
+Now go ahead and create a data dictionary for the penguins data set, 
+in a software (text or spreadsheet) of your choice.
 
-```{r}
-#| echo: false
-#| column: "body-outset"
-#| classes: plain
+Save the data dictionary in the same folder as the actual data set file.
 
-datawizard::data_codebook(dat) |>
-  datawizard::data_select(exclude = ID) |>
-  datawizard::data_filter(N != "") |>
-  datawizard::print_md()
-```
+::: {#tip-data_dic .callout-tip collapse="true"}
+#### Data dictionary for penguins data set (Solution)
 
-``````
+| name              | label          | type    | description                          | values                    | units | missing_codes |
+| ----------------- | -------------- | ------- | ------------------------------------ | ------------------------- | ----- | ------------- |
+| species           | Species        | string  | Penguin species                      | Adelie; Gentoo; Chinstrap |       | NA = Missing  |
+| island            | Island         | string  | Island where individual was observed | Torgersen; Biscoe; Dream  |       | NA = Missing  |
+| bill_length_mm    | Bill length    | float   | Length of the bill (beak)            |                           | mm    | NA = Missing  |
+| bill_depth_mm     | Bill depth     | float   | Depth of the bill (beak)             |                           | mm    | NA = Missing  |
+| flipper_length_mm | Flipper length | integer | Length of the flipper                |                           | mm    | NA = Missing  |
+| body_mass_g       | Body mass      | integer | Body mass                            |                           | g     | NA = Missing  |
+| sex               | Sex            | string  | Sex of the penguin                   | male; female              |       | NA = Missing  |
+| year              | Year           | integer | Year of data collection              | 2007; 2008; 2009          |       |               |
 
-```{r}
-#| child: "_data_dictionary.qmd"
-
-```
-
-Depending on the type of data, it may also be necessary
-to describe sampling procedures (e.g., selection criteria),
-measurement instruments (e.g., questionnaires),
-appropriate weighting,
-already applied preprocessing steps, or contact information.
-In our case, as the data has already been published,
-we only store a reference to its source.
-
-The data set is from the R package `palmerpenguins`.
-If you had it installed
-you could use the function `citation()` to create such a reference:
-
-```{r}
-#| label: "data-citation"
-#| eval: false
-
-citation("palmerpenguins", auto = TRUE) |>
-  format(bibtex = FALSE, style = "text")
-```
-
-Without the package `palmerpenguins` installed,
-you can find a [suggested citation on its website][palmerpenguins-citation]
-and add that to your data dictionary:
-
-[palmerpenguins-citation]: https://allisonhorst.github.io/palmerpenguins/#citation
-
-```{r}
-#| ref.label = "data-citation",
-#| render = function(x, options) gsub("\\n", " ", x = x),
-#| echo = FALSE,
-#| class.output = "md code-overflow-wrap",
-#| attr.output = 'filename="data_dictionary.qmd"'
-
-# This chunk takes the output from the chunk "data-citation"
-# and renders it with all newlines replaced by whitespaces.
-```
-
-Finally, you can render the data dictionary by running the following:
-
-```{.bash filename="Terminal"}
-quarto render data_dictionary.qmd
-```
-
-This should create the file `data_dictionary.html`
-which you open and view in your web browser.
-
-One could go even further by making the information machine-readable in a standardized way.
-We provide an optional example of that in @nte-frictionless.
-If you want to learn more about the sharing of research data,
-have a look at the tutorial "[FAIR research data management][fair-tutorial]".
-
-[fair-tutorial]: https://lmu-osc.github.io/FAIR-Data-Management/
-
-::: {#nte-frictionless .callout-note collapse="true"}
-### Create Machine-Readable Variable Documentation
-
-This example demonstrates how the title and description of the data set,
-the description of the variables and their valid values are stored in a machine-readable way.
-We'll reuse the descriptions we already created[^value-labels] and add a few others.
-
-[^value-labels]: Unfortunately, the descriptions of values are not reused in this example,
-as they are [not supported][enum-labels] by the specification we are using.
-
-[enum-labels]: https://specs.frictionlessdata.io/patterns/#table-schema-enum-labels-and-ordering
-
-First, store the title and description of the data set as a whole:
-
-```{.r filename="Console"}
-table_info <- c(
-  title = "penguins data set",
-  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
-)
-```
-
-As before, also provide a reference to the source.
-
-```{r}
-#| echo: false
-#| class-output: "r code-overflow-wrap"
-#| attr-output: 'filename="Console"'
-
-# We have provided the data set as CSV file to the readers.
-# Therefore, we cannot assume or require that readers have
-# the R package palmerpenguins installed. Instead, we create
-# the citation on our end and hide how we obtained it.
-
-citation("palmerpenguins", auto = TRUE)$url |>
-  paste0("dat_source <- \"", ... = _, "\"") |>
-  cat()
-```
-
-Next, create a list of the categorical variables' valid values:
-
-```{.r filename="Console"}
-valid_vals <- list(
-  species = c("Adelie", "Gentoo", "Chinstrap"),
-  island = c("Torgersen", "Biscoe", "Dream"),
-  sex = c("male", "female"),
-  year = c(2007, 2008, 2009)
-)
-```
-
-Finally, store the descriptions of the variables we already created earlier:
-
-```{.r filename="Console"}
-# Store the description of variables
-vars <- c(
-  species = "a character string denoting penguin species",
-  island = "a character string denoting island in Palmer Archipelago, Antarctica",
-  bill_length_mm = "a number denoting bill length (millimeters)",
-  bill_depth_mm = "a number denoting bill depth (millimeters)",
-  flipper_length_mm = "an integer denoting flipper length (millimeters)",
-  body_mass_g = "an integer denoting body mass (grams)",
-  sex = "a character string denoting penguin sex",
-  year = "an integer denoting the study year"
-)
-```
-
-Generally, metadata are either stored embedded into the data or externally,
-for example, in a separate file.
-We will use the "[frictionless data](https://frictionlessdata.io/)" standard,
-where metadata are stored separately.
-Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).
-
-Specifically, one can use the R package [`frictionless`][frictionless]
-to create a _schema_ which describes the structure of the data.[^frictionless-note]
-For the purpose of the following code,
-it is just a nested list that we edit to include our own information.
-We also explicitly record in the schema
-that missing values are stored in the data file as `NA`
-and that the data are licensed under CC0\ 1.0.
-Finally, the package is used to create a metadata file that contains the schema.
-
-[frictionless]: https://docs.ropensci.org/frictionless/
-
-[^frictionless-note]: In June 2024, [version 2](https://datapackage.org/)
-of the frictionless data standard has been released.
-As of November 2024, the R package `frictionless` only supports the first version,
-though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).
-
-```{.r filename="Console"}
-# Install {frictionless} and the required dependency {stringi}
-renv::install(c(
-  "frictionless",
-  "stringi"
-))
-
-# Read data and create schema
-dat_filename <- "data.csv"
-dat <- read.csv(dat_filename)
-dat_schema <- frictionless::create_schema(dat)
-
-# Add descriptions to the fields
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  c(x, description = vars[[x$name]])
-})
-
-# Record valid values
-dat_schema$fields <- lapply(dat_schema$fields, \(x) {
-  if (x[["name"]] %in% names(valid_vals)) {
-    modifyList(x, list(constraints = list(enum = valid_vals[[x$name]])))
-  } else {
-    x
-  }
-})
-
-# Define missing values
-dat_schema$missingValues <- c("", "NA")
-
-# Create package with license info and write it
-dat_package <- frictionless::create_package() |>
-  frictionless::add_resource(
-    resource_name = "penguins",
-    data = dat_filename,
-    schema = dat_schema,
-    title = table_info[["title"]],
-    description = table_info[["description"]],
-    licenses = list(list(
-      name = "CC0-1.0",
-      path = "https://creativecommons.org/publicdomain/zero/1.0/",
-      title = "CC0 1.0 Universal"
-    )),
-    sources = list(list(
-      title = "CRAN",
-      path = dat_source
-    ))
-  )
-frictionless::write_package(dat_package, directory = ".")
-```
-
-This creates the metadata file `datapackage.json` in the current directory.
-Make sure it is located in the same folder as `data.csv`,
-as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
 :::
 
-Having added the data and its documentation,
-one can view and record the utilized packages with `renv`,
-thus bringing the project into a consistent state:
-
-```{.r filename="Console"}
-renv::status()
-renv::snapshot()
-```
-
-## Add Data Citation and Attribution
+## ✍️ Practical Exercise: Add Data Citation and Attribution
 
@@
 to allow for precise identification and access.
-From the "eight core principles of data citation" by @Starr2015,
-licensed under [CC0\ 1.0][cc0]:
+Now, it's your turn to add an appropriate citation for the data set to the manuscript.
 
-> **Principle 1 – Importance**: "Data should be considered legitimate,
-> citable products of research. Data citations should be accorded the same
-> importance in the scholarly record as citations of other research objects,
-> such as publications."
-> 
-> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a
-> claim relies upon data, the corresponding data should be cited."
-> 
-> **Principle 5 – Access**: "Data citations should facilitate access to the data
-> themselves and to such associated metadata, documentation, code, and other
-> materials, as are necessary for both humans and machines to make informed use
-> of the referenced data."
-> 
-> **Principle 7 – Specificity and Verifiability**: "Data citations should
-> facilitate identification of, access to, and verification of the specific data
-> that support a claim. Citations or citation metadata should include
-> information about provenance and fixity sufficient to facilitate verifying
-> that the specific time slice, version and/or granular portion of data
-> retrieved subsequently is the same as was originally cited."
+Hints:
 
-Now, it's your turn to add an appropriate citation for the data set to the manuscript.
-Does your citation adhere to the principles above?
+- You can find an appropriate BibTeX entry on [the package website](https://allisonhorst.github.io/palmerpenguins/#citation)
+  or with the function `citation()`^[Note that this function requires to have the respective package installed.].
+- Add the citation in the manuscript where it says "cite data here".
 
@@
 
-You can find an appropriate BibTeX entry on [the package website][palmerpenguins-citation]
-or with the function `citation()`:^[Note that this function requires
-to have the respective package installed.]
+Show the correct reference of the data set:
 
@@
 
-While citation happens in the manuscript for reasons of academic integrity and reproducibility,
-to comply with any licenses you also may need to provide attribution within your project folder.
-Even though the data file we use here does not require attribution,
-we recommend adding a short paragraph to `LICENSE.txt`:
+## Wrap up
 
-```{r}
-#| echo: false
-#| class-output: "txt code-overflow-wrap"
-#| attr-output: 'filename="LICENSE.txt"'
+Congrats! You documented your data set and cited it correctly.
 
-cat(paste0(
-  "The penguins data stored in \"data.csv\" by Allison Horst, Alison Hill, and Kristen Gorman available from <",
-  citation("palmerpenguins", auto = TRUE)$url,
-  "> are licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>"
-))
-```
-
-As before, if the license required adding the full license text,
-you would also need to copy it to the project folder (if not already in there).
-
-Finally, you can go through the commit routine:
+To finalize this step, you can go through the commit routine:
 

### 2025-07-16 make_readme.qmd
Also state where deps can be obtained
Shorten README Chapter; add new in-depth page

@@
 having forgotten most of the details of what you did exactly.
-What would be useful to know in order to quickly understand
+What would your future self want to know in order to quickly understand
 what is going on in the project?
-This is what needs to be described in the README.
-While you could just start writing along,
-it is helpful to provide at least the following information in sections on their own.
 
-### Name and Description
+This is what needs to be described in the `README` file.
+It should be the primary entry point into your project,
+the place where new users go first to get an orientation.
 
-How is the project called?
-What is it about?
-Which files does the project folder contain?
-How are they organized?
+There is no common standard how to structure a README file,
+but you should usually provide at least the following information,
+structured by sections:
 
-### Involved Data
+**Name and Description**
 
-Are any (empirical) data involved (e.g., being analyzed or used as input)?
-From which sources can they be obtained?
-Are they already included in the project folder?
-Where is their data dictionary located?
-Which terms, usage restrictions, or licenses apply?
-If they are not publicly available,
-is an alternative, synthesized version provided?
+- [ ] How is the project called?
+- [ ] What is it about?
+- [ ] Which files does the project folder contain?
+- [ ] How are they organized?
 
-### Computational Requirements
+**Data**
 
-What software needs to be installed to run the analysis
--- in other words, what are its dependencies?
-This also includes software that you have used for any manual steps.
+- [ ] Are any (empirical) data involved (e.g., being analyzed or used as input)?
+- [ ] Are they already included in the project folder? If not: From which sources can they be obtained?
+- [ ] Where is their data dictionary located?
+- [ ] Which terms, usage restrictions, or licenses apply?
+- [ ] (optionally) If they are not publicly available, is an alternative, synthesized version provided?
+
+**Computational Requirements / Dependencies**
+
+- [ ] What software needs to be installed to run the analysis?
+
+For every dependency (e.g., packages that you load in R), describe where it can be obtained from.
 If the code has particular hardware requirements
-(e.g., in terms of processor or memory),
-these should be also noted.
+(e.g., in terms of processor or memory), these should be also noted.
 Finally, for steps that take more than a couple of seconds,
-the approximate runtime should be indicated.
+indicate the approximate runtime.
 
-@nte-dependencies provides more information
-on determining the dependencies of an R project.
+For an R project, the most important information for enabling reproducibility are the specific version of R and the loaded packages.
+Ideally, you also list other system dependencies.
+The in-depth supplementary material [*Identify additional system dependencies for the README*](/in_depth_material/other_dependencies.qmd) provides more information on determining the dependencies of an R project.
+*For the purpose of this tutorial, however, you can skip this step.*
 
-### Usage
+**How to reproduce the results**
 
-How can one run the project -- is there a master script
-or a particular order in which any scripts need to be executed?
-Provide detailed instructions for running the full project.
+- [ ] Provide step-by-step instructions that are necessary to reproduce all reported results.
 
-### List of Results
+Is there a master script or a particular order in which any scripts need to be executed?
 
-For every result (i.e., number, figure, or table) that is
-computed in the project and displayed in the manuscript,
-indicate where exactly it is computed.
+**Citation**
 
-### Citation
+- [ ] Is there a recommended way to cite this project?
+- [ ] Is there a published article associated with it that you would like to have cited?
 
-Is there a recommended way to cite this project?
-Is there a published article associated with it
-that you would like to have cited?
+**License**
 
-### License
+- [ ] Under which licenses are the works in this project folder available?
 
-Under which licenses are the works in this project folder available?
+## Identifying R Dependencies
 
-## Create It!
+If you just write "We used R to analyze the data" in your manuscript or in the *Computational Requirements / Dependencies* section of your README file, this would be too underspecified.
+To enable someone else to really reproduce your results, you need specific information on ...
 
-Create your README now as the file `README.md`.
+- The exact version of R
+- All packages and their versions that you loaded in your project
+- (optionally) Further system dependencies
 
-::: {#nte-dependencies .callout-note}
-### Identifying R Dependencies
+### R & Package Dependencies
 
-R itself and the R packages are already documented as this project uses `renv`.
-Therefore you can focus on all other dependencies,
-such as the system dependencies of R packages
-as well as the version of Quarto.[^renv-quarto]
+R itself and the R packages are already documented as our tutorial project uses `renv`: The information is stored in the `renv.lock` file at the root of the project, which tracks each package’s version and source.
+In this case, you can simply refer to that in the README file and all is done:
 
-[^renv-quarto]: As of August 2024, a proposal for `renv` to record the version of Quarto
-has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).
+> The R and packages versions used to compute the results are stored in the `renv` environment (see `renv.lock` file).
 
-An overview over the system dependencies of R packages can be created
-using the function `pak::pkg_sysreqs()`.
-In combination with `renv`, we can obtain the system dependencies
-of all R packages the current project directly depends on:
+If you do *not* use `renv`, the easiest solution to get a list of all needed packages is the `sessionInfo()` command. *Call it **after** you loaded all necessary packages!*
 
-```{.r filename="Console"}
-# Find all R package dependencies
-deps <- renv::dependencies()$Package |>
-  unique() |>
-  pak::pkg_deps(dependencies = NA) |>
-  getElement("package")
+It lists all packages including their versions, and you can copy & paste that information into your README file. Here's an example for a project:
 
-# Identify their system dependencies
-pak::pkg_sysreqs(deps)
+```txt
+> sessionInfo()
+R version 4.4.1 (2024-06-14)
+Platform: aarch64-apple-darwin20
+Running under: macOS 15.5
+
+Matrix products: default
+BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
+LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0
+
+locale:
+[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
+
+time zone: Europe/Berlin
+tzcode source: internal
+
+attached base packages:
+[1] parallel  stats     graphics  grDevices utils     datasets  methods   base
+
+other attached packages:
+ [1] rio_1.2.2            doRNG_1.8.6          rngtools_1.5.2      
+ [4] doParallel_1.0.17    iterators_1.0.14     foreach_1.5.2       
+ [7] fitPCurve_0.0.0.9000 testthat_3.2.3       prettycode_1.1.0    
+[10] colorDF_0.1.7
+
+loaded via a namespace (and not attached):
+ [1] gtable_0.3.6       ggplot2_3.5.2      htmlwidgets_1.6.4  devtools_2.4.5    
+ [5] remotes_2.5.0      TruncExpFam_1.2.1  vctrs_0.6.5        tools_4.4.1       
+ [9] generics_0.1.4     tibble_3.2.1       pkgconfig_2.0.3    R.oo_1.26.0       
+[13] data.table_1.17.4  RColorBrewer_1.1-3 desc_1.4.3         lifecycle_1.0.4 
 ```
 
-The output may look like the following:
+## ✍️ Practical Exercise: Create your README!
 
-```txt
-── Install scripts ────────────────── Fedora 40 ── 
-dnf install -y make pandoc git
-
-── Packages and their system dependencies ────────
-fs        – make
-knitr     – pandoc
-remotes   – git
-rmarkdown – pandoc
-sass      – make
-```
-
-We can see that the programs `make`, `pandoc`,
-and `git` were identified as system dependencies.
-Often, one can obtain their version by running them with the `--version` argument:
-
-```{.bash filename="Terminal"}
-make --version
-pandoc --version
-git --version
-```
-
-However, this does not work for all system dependencies.
-Specifically, it does not work for libraries -- software that is not supposed to be run on its own.
-Identifying their version is beyond the scope of this tutorial.
-
-We also know that we need Quarto to create the PDF,
-so let's find out its version as well:
-
-```{.bash filename="Terminal"}
-quarto --version
-```
-
-If you installed `apaquarto` or any other Quarto extension,
-one can query their versions as follows:[^already-included]
-
-[^already-included]: Luckily, the extensions are included in the project folder,
-so technically their version is already recorded in the project's files.
-
-```{.bash filename="Terminal"}
-quarto list extensions
-```
-
-Finally, we know that we installed a $\TeX$ distribution to create the PDF,
-so let's find out its version by running:
-
-```{.bash filename="Terminal"}
-quarto check
-```
-
-The output is quite long and it might look slightly different for you,
-but the relevant sections are the following:
-
-```txt
-[✓] Checking tools....................OK
-      TinyTeX: v2024.09
-      Chromium: (not installed)
-
-[✓] Checking LaTeX....................OK
-      Using: TinyTex
-      Path: /home/r155953/.TinyTeX/bin/x86_64-linux
-      Version: 2024
-```
-
-Of course, all the system dependencies identified until now
-may have dependencies on their own.
-Use your own judgement to decide when not to dig deeper.
-:::
+Create your README now as the file `README.md`, located at the top level folder.
 
@@
 
-The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.
+The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF according to APA guidelines.
 ```
@@
 ::: {#tip-data .callout-tip collapse="true"}
-### Involved Data
+### Data
 
 ```{.md .code-overflow-wrap filename="README.md"}
-## Involved Data
+## Data
 
@@
 
-Next, one can open a new R session in the root folder of this project and run the following, which should install all required R packages at their recorded versions:
+Next, one can open a new R session in the root folder of this project and run the following command, which should install all required R packages at their recorded versions:
 
@@
 
-::: {#tip-usage .callout-tip collapse="true"}
-### Usage
+::: {#tip-how-to-reproduce .callout-tip collapse="true"}
+### How to reproduce the results
 
 `````{.md .code-overflow-wrap filename="README.md"}
-## Usage
+## How to reproduce the results
 
@@
 ``````
-:::
-
-::: {#tip-list-of-results .callout-tip collapse="true"}
-### List of Results
-
-```{.md .code-overflow-wrap filename="README.md"}
-## List of Results
-
-- In-text numbers in the section "results": Calculated in the chunk "t-test" within "Manuscript.qmd"
-- Table 1: Calculated in the chunk "tbl-descriptive-statistics" within "Manuscript.qmd"
-- Figure 2: Calculated in the chunk "fig-bill-length-comparison" within "Manuscript.qmd"
-```
 :::
@@
 
-The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) or (at your option) under the [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later).
+The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/).
 For further copyright information, see `LICENSE.txt`.
